### 2013-05-04 gall-manuscript.Rmd
lots of additions

```{r include=FALSE}
library(bipartite)
library(knitr)
library(knitcitations)
bibnet <- read.bibtex("~/Documents/BibTex/willow gall-ptoid networks.bib")
```

Plant genetic variation shapes the structure and function of herbivore-parasitoid networks (a work in progress)
-----------------------------------------------------------

Matthew A. Barbour, Jordi Bascompte, and Greg W. Crutsinger

### Objective

This paper examines how plant genetic variation influences herbivore-parasitoid interaction networks and how these networks shape selection pressures acting on herbivores.

Introduction
------------

If plant genotype has a significant effect on gall-parasitoid network structure

We tested the relative importance of plant genotype versus the environment in shaping gall-parasitoid networks.

### Research Questions

**(1)** How does plant genotype influence gall-parasitoid interaction networks?

**(2)** How do different interaction networks influence the form of natural selection acting on herbivore traits?

*Possible (3)* How does the simulated loss of plant genotypes influence gall-parasitoid interaction networks?

### Methods

**Quantifying gall-parasitoid interaction networks**

Following the method proposed by `r citet(bibnet["Poisot2012a"])`, we quantified the dissimilarity in quantitative gall-parasitoid networks across all willow tree sampled (n = 140).

To measure the interaction strengths between each herbivore and parasitoid species in each network, we calculated the number of observed interactions divided by the sum of observed interactions between that herbivore and all parasitoid species.  This represents a quantitative measure of the relative impact of a particular parasitoid species on a herbivore species. *need to think about whether this is the actual measure I want*

**Other Approaches at Community Detection**
Following Albert et al. (in review at Ecology Letters) and `r citet(bibnet["Karrer2008"])`, we could compare the module composition of multiple networks.  Note that this approach will only work, if the same nodes are present in each network.  So if I find differences in "interaction turnover" between networks using `r citet(bibnet["Poisot2012a"])` method, I could follow up by examining how this interaction turnover influences the modular composition of the network.

Also, consider using hierarchical random graphs in the igraph library (in R) to detect hierarchical structure in the networks.  This method appears to utilize clustering at multiple levels, which may be useful for my tripartite networks of genotype-herbivore-parasitoid.  I also may be able to add ecological correlates (e.g. plant traits, gall size) that will influence community structure.  See `r citet(bibnet[c("Newman2011","Clauset2008"])` for examples.

Also, consider an individual-based approach (sensu `r citet(bibnet["Gomez2011"])`) at the plant and herbivore level.

Also, remember Jordi's advice which presents a pretty clear study.  Search for variation in modular composition of the network across genotypes.  Then use partial mantel tests and path analysis to see how genetic distance and plant traits influence the modular composition of these networks.  Then, I can identify genotype-modules (using bipartite modularity measures focused on genotypes, `r citet(bibnet[c("Guimara2007","Fortuna2009"])).  And follow up by seeing how variation in module composition influences natural selection acting on herbivores (section below).  But before this, I should explore the more fine scale structure (as suggested by Miguel) by looking at dissimilarity in interaction networks between individual trees and making sure all nodes are present on each tree.

**Quantifying natural selection**
Depending on the appropriate the scale (each tree, each genotype, or multiple genotypes) identified from the community detection analysis, I will quantify how differences in community structure influence the form and strength of natural selection acting on galling insects.  I need to find the R package, but I should be able to use a cubic spline method, as advocated by Dolph Schluter and `r citet(bibnet["Brodie1995"])`.

Older ideas
-----------
#### Q1
Build a meta-web of herbivore-parasitoid interactions by pooling across all genotypes (try a both qualitative and quantitative version, hopefully both will yield the same results)

Run an appropriate modularity algorithm to identify modules on the meta-web.

***Figure out how to quantify the frequency of representation of the different modules.***

Run a chi-square test based on the frequencies of each module occurring on each genotype to see whether there is an association.

### Q2
Build a meta-web of herbivore-parasitoid interactions by pooling across all genotypes (try both qualitative and quantitative version, hopefully both will yield the same results)

Evaluate the structure of the meta-web (nested or modular or both).

Compute the appropriate network measure for each replicate of each genotype and see whether I can detect any correlations between network structure and genetic distance or traits.

Could also use the method proposed in Poisot 2012 (Dissimilarity of Interaction Networks), but it doesn't examine how network structure changes (this may or may not be important, depending on whether I see any noticeable changes in my graphs of the different networks)

Could also see how the functional role of species (Guimera and Amaral Functional Cartography paper in Nature) change across genotypes.  Test this using a chi-square analysis based on genotypes and the seven different functional roles (I'm thinking now I won't see a big difference...)

***What I think may be a promising method***: Look for dissimilarity in module participation, by creating a matrix of all possible interactions (e.g H1,P1 H1,P2 H2,P1 etc.) for a particular module on a tree (or genotype), then calculate a bray-curtis dissimilarity within modules across the trees/genotypes.  Then see if this is associated with any particular gall and/or plant traits.**this is a more likely area where we will see gall trait variation influencing the network**

***Conversation with Rudolf***: Look for mantel correlation between dissimilarity in herbivore traits (e.g. gall size, multilocularity, etc.) and dissimilarity in network topology (e.g. see Bersier et al. 2002, for quantitative versions of network measures).  And see whether this correlation varies across genotypes.  Or simply look to see whether the quantitative versions of these network metrics vary across genotypes (similar to Tylianakis et al. 2007, but with different genotypes instead of habitats)

### Q3
Option 1: Use partial mantel tests and path analysis to see how genetic distance and trait dissimilarity influence dissimilarity in module frequency (either all modules combined, or modules individually).  Consider at first using all plant traits and then looking at individual traits.

Also, look for patterns between genotypes and herbivore traits, as well as herbivore traits and network topology.  Could use a mantel test for genetic distance and herbivore trait dissimilarity and measures of network topology.

### 2013-05-04 Network 28 betalink functions.r

@@ -1,0 +1,394 @@
+##This file contains functions written by Poisot and colleagues to implement their analysis of dissimilarity of species interaction networks (Poisot et al 2012 Ecology Letters).
+##Rabie's adjustments are noted in the code
+
+library(vegan)
+
+B01 = function(pm) with(pm,{2*(a+b+c)/(2*a+b+c) - 1})
+#B02 = function(pm) B01(pm)
+#B03 = function(pm) with(pm,{(b+c)/2})
+#B04 = function(pm) with(pm,{(b+c)})
+#B05 = function(pm) with(pm,{(((a+b+c)^2)/((a+b+c)^2-(2*b*c)))-1})
+#B06 = function(pm) with(pm,{log(2*a+b+c)-((2*a*log(2))/(2*a+b+c))-(((a+b)*log(a+b)+(a+c)*log(a+c))/(2*a+b+c))})
+#B07 = function(pm) exp(B06(pm))-1
+#B08 = function(pm) with(pm,{(b+c)/(2*a+b+c)})
+#B09 = function(pm) B08(pm)
+#B10 = function(pm) with(pm,{a/(a+b+c)})
+#B11 = function(pm) with(pm,{(2*a)/(2*a+b+c)})
+#B12 = function(pm) with(pm,{(2*a+b+c)*(1-(a/(a+b+c)))})
+#B13 = function(pm) with(pm,{min(b,c)/(max(b,c)+a)})
+#B14 = function(pm) with(pm,{1-((a*(2*a+b+c))/(2*(a+b)*(a+c)))})
+#B15 = function(pm) with(pm,{(b+c)/(a+b+c)})
+#B16 = function(pm) B15(pm)
+#B17 = function(pm) with(pm,{min(b,c)/(a+b+c)})
+#B18 = function(pm) with(pm,{(b+c)/2})
+#B19 = function(pm) with(pm,{(b*c+1)/(((a+b+c)^2-(a+b+c))/2)})
+#B20 = function(pm) with(pm,{1-(2*a)/(2*a+b+c)/2})
+#B21 = function(pm) with(pm,{a/(a+c)})
+#B22 = function(pm) with(pm,{min(b,c)/(min(b,c)+a)})
+#B23 = function(pm) with(pm,{(2*aba(b-c))/(2*a+b+c)})
+#B24 = function(pm) with(pm,{1-(log((2*a+b+c)/(a+b+c))/log(2))})
+
[…]

### 2014-11-15 Rscripts/ms_willow_network.Rmd
lots of updates, too many to detail
need to do a better job of keeping track of the changes I make.

---
title: "Plant genetic variation structures herbivore-parasitoid networks"
author: "Matthew Barbour"
date: "20 August, 2014"
output: word_document
---

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2014-11-15 Rscripts/Allometry - Willow Garden - 2012.R

@@ -1,6 +1,8 @@
 #### Metadata
 # Data Collection Date: September 29, 2012 (end of growing season)
 
+library(dplyr)
+
 # setworking directions
 setwd('~/Documents/Genotype_Networks/data/')
 
@@ -27,10 +29,22 @@ hist(log(allom$Shoot.Count)) # distribution more normalized except for one outli
 
 # shoot allometry. Equation = log(Shoot.Count) = 3.7245*log(Stem.Diamter) - 8.7794
 plot(log(Shoot.Count)~log(Stem.Diamter), allom)
+plot(Shoot.Count ~ Stem.Diamter, allom)
+
+lm.shoot.notrans <- lm(Shoot.Count ~ Stem.Diamter, allom)
+summary(lm.shoot.notrans)
+abline(lm.shoot.notrans)
+plot(lm.shoot.notrans)
+
+plot(log(Shoot.Count) ~ log(Stem.Diamter), allom[-18,])
 lm.shoot <- lm(log(Shoot.Count)~log(Stem.Diamter), allom[-18,])
 summary(lm.shoot) # model explains 61.3% of the variance with the quadratic fit.
 abline(lm.shoot)
 plot(lm.shoot) # point 18 appears to be an outlier, but is also the smallest branch s
+
+library(visreg)
+lm.shoot.visreg <- visreg(lm.shoot, trans = exp)
+
 
 # leaf allometry.
 plot(log(Leaf.Estimate)~log(Stem.Diamter),allom)
@@ -42,12 +56,19 @@ plot(lm.leaf)
 
 # allometric relationships. Focusing on shoots because this relationship was stronger 
 
-# log(Shoot.Count) = 3.7245*log(Stem.Diamter) - 8.7794  # all data points. 61.3% of variance explained
-# log(Shoot.Count) = log(Stem.Diameter)*3.0371 - 6.1602  # data point #18 removed. 54.6% of variance explained
+# log(Shoot.Count) = 3.7245*log(Stem.Diamter) - 8.7794  # all data points.  61.3% of variance explained
+# log(Shoot.Count) = log(Stem.Diameter)*3.0371 - 6.1602  # datapoint 18 removed. 54.6% of variance explained. Although lower variance explained, it appears to be a better fit of the data by inspecting the residuals.
 
[…]

### 2014-11-15 Rscripts/best_QuaBiMo_function.R

@@ -1,0 +1,18 @@
+best_QuaBiMo <- function(web, QuaBiMo_reps = 100) {
+  require('bipartite')
+  
+  observed_web_repeats <- list()
+  for(i in 1:QuaBiMo_reps){ 
+    observed_web_repeats[[i]] <- web
+  }
+  
+  QuaBiMo_repeats <- sapply(observed_web_repeats, computeModules)
+  best_observed_value <- max(sapply(QuaBiMo_repeats, function(x) x@likelihood)) # takes the maximum out of a user specified number of replications, since this is an optimization algorithm and we want to avoid reaching a local optimum.
+  
+  best_reps <- which.max(sapply(QuaBiMo_repeats, function(x) x@likelihood)) # identify best partition of modularity algorithm
+  
+  output <- list(best_observed_value = best_observed_value,
+                 best_reps = best_reps,
+                 best_module_info = QuaBiMo_repeats[best_reps])
+}
+

### 2014-11-15 Rscripts/betalink_analysis_genotype_level.R

@@ -1,0 +1,369 @@
+
+library(reshape)
+library(reshape2)
+library(plyr)
+library(dplyr)
+library(vegan)
+
+
+# upload molten gall network data
+gall_net_melt <- read.csv("~/Documents/Genotype_Networks/data/gall_network_data.csv")
+gall_net_melt <- tbl_df(gall_net_melt)
+
+##### Does the number of shoots sampled vary among genotypes?
+
+# data frame
+shootEsts_df <- gall_net_melt %>%
+  mutate(plant.position = as.factor(plant.position)) %>%
+  group_by(Gender, Genotype, plant.position) %>%
+  summarise(shootEst.no18 = mean(shootEst.no18), shootEst.all = mean(shootEst.all), galls.found = mean(Galls.found)) # needed to take the mean because these shoot estimates are duplicated throughout the molten dataframe
+shootEsts_df <- mutate(shootEsts_df, plant.position = as.character(plant.position))
+
+shootEsts_genotype_df <- shootEsts_df %>%
+  group_by(Genotype) %>%
+  summarise(n = n(), shootEst.no18 = sum(shootEst.no18), shootEst.all = sum(shootEst.all))
+
+
+##### How dissimilar are gall-parasitoid interaction networks among genotypes?
+
+### Genotype-level networks
+levels(gall_net_melt$gall_contents)
+
+genotype_net_filter <- gall_net_melt %>%
+  filter(gall_contents %in% c("Eulo.fem", "Eulo.mal", "Eury.fem", "Eury.mal", "Lathro.fem", "Lathro.mal", "Lestodip", "Mesopol", "Mymarid", "Platy", "Ptero.2", "Tory.fem", "Tory.mal")) # removed survival data for gall species
+
+genotype_net_add <- mutate(genotype_net_filter, gall_contents_collapse = revalue(gall_contents, c("Eulo.fem" = "Eulo", "Eulo.mal" = "Eulo", "Eury.fem" = "Eury", "Eury.mal" = "Eury", "Lathro.fem" = "Lathro", "Lathro.mal" = "Lathro", "Ptero.2" = "Mesopol", "Tory.fem" = "Tory", "Tory.mal" = "Tory") ))
+
+
+genotype_net <- cast(genotype_net_add, gall.sp ~ gall_contents_collapse | Genotype, sum)
+genotype_net <- genotype_net[-21] # never any interactions for Genotype U
+
[…]

### 2014-11-15 Rscripts/betalink_plot_testing.R

@@ -1,0 +1,131 @@
+source('~/Documents/ggnet/bipartite_plot_info.R')
+source('~/Documents/ggnet/ggnet_bipartite.R')
+
+betalink_bipartite_plot_info <- function(m1, m2, order.type = "cca") {
+  m1.info <- bipartite_plot_info(m1, order.type = order.type)
+  m2.info <- bipartite_plot_info(m2, order.type = order.type)
+  
+  m1.m2.bind <- rbind(m1.info[[2]], m2.info[[2]])
+  m1.m2.bind$webID <- factor(c(rep(1, sum(dim(m1))), 
+                               rep(2, sum(dim(m2)))))
+  
+  up.1 = rownames(m1)
+  up.2 = rownames(m2)
+  lo.1 = colnames(m1)
+  lo.2 = colnames(m2)
+  up.shared = up.1[up.1%in%up.2]
+  lo.shared = lo.1[lo.1%in%lo.2]
+  up.un.1 = up.1[!(up.1%in%up.shared)]
+  up.un.2 = up.2[!(up.2%in%up.shared)]
+  lo.un.1 = lo.1[!(lo.1%in%lo.shared)]
+  lo.un.2 = lo.2[!(lo.2%in%lo.shared)]
+  
+  shared = c(up.shared, lo.shared)
+  unshared = c(up.un.1, up.un.2, lo.un.1, lo.un.2)
+  
+  m1.m2.bind$species.turnover.colors <- rep(NA, dim(m1.m2.bind)[1])
+  m1.m2.bind$species.turnover.colors[which(m1.m2.bind$vertex.names %in% shared)] <- "shared"
+  m1.m2.bind$species.turnover.colors[which(m1.m2.bind$vertex.names %in% unshared)] <- "unshared"
+  out <- list(m1.info, m2.info, m1.m2.bind)
+}
+
+comp1 <- betalink_bipartite_plot_info(m1 = genotype_net_matrices[[1]], m2 = genotype_net_matrices[[3]], order.type = "normal")
+
+m1.plot <- ggnet_bipartite(comp1[[1]]) + 
+  geom_point(data = filter(comp1[[3]], webID == 1), 
+             aes(x = x, y = y, fill = species.turnover.colors), size = 12, shape = 21) + 
+  scale_fill_manual(values = c("gray", "steelblue"), guide = "none") + 
+  geom_text(data = filter(comp1[[3]], webID == 1), 
[…]

### 2014-11-15 Rscripts/betalink_practice.R

@@ -1,46 +1,76 @@
-library(devtools)
+#library(devtools)
 
-install_github("tpoisot/betalink")
+#install_github("tpoisot/betalink")
 
 library(betalink)
+library(reshape2)
+library(plyr)
+library(dplyr)
 
-w1 = matrix(c(1, 0, 0, 1), ncol=2)
-w2 = matrix(c(1, 1, 0, 1), ncol=2)
-w3 = matrix(c(1, 0, 1, 1), ncol = 2)
-colnames(w1) = c('p1', 'p2')
-rownames(w1) = c('h1', 'h2')
+# Comparing 2 networks whose dissimilarity is entirely driven by "rewiring" of species interactions
+w1 = matrix(c(1, 0, 0, 
+              0, 1, 0, 
+              0, 0, 1), ncol=3)
+w2 = matrix(c(0, 1, 1, 
+              1, 0, 1,
+              1, 1, 0), ncol=3)
+
+colnames(w1) = c('p1', 'p2', 'p3')
+rownames(w1) = c('h1', 'h2', 'h3')
 colnames(w2) = colnames(w1)
 rownames(w2) = rownames(w1)
-colnames(w3) = colnames(w1)
-rownames(w3) = rownames(w1)
 
-realizations = list(w1, w2, w3)
+realizations = list(w1, w2)
 
 metaweb_1 = metaweb(W = realizations)
 
 betalink(w1, w2, bf = B01)
 betalink.plot(m1 = w1, m2 = w2, by.unique = TRUE)
-beta
 
-betalink.dist(realizations, bf = B01)
+w1_trans <- melt(w1)
+w1_trans <- mutate(w1_trans, site = rep("w1", dim(w1_trans)[1]))
+
+w2_trans <- melt(w2)
+w2_trans <- mutate(w2_trans, site = rep("w2", dim(w2_trans)[1]))
+
+w1_w2_comp <- rbind.data.frame(w1_trans, w2_trans)
+w1_w2_comp_cast <- dcast(w1_w2_comp, site ~ Var1 + Var2, sum)
+
+library(vegan)
+betadiver(w1_w2_comp_cast[ ,-1], method = "w")
+
+# Comparing 2 networks that differ solely due to species turnover. Note that there has to be at least 1 overlapping link for betalink to calculate species turnover...does that make sense!?!?!
+w3 = matrix(c(1, 0, 0, 
+              0, 1, 0, 
[…]

### 2014-11-15 Rscripts/discriminant_function_analysis_genotype_gall_network_modules.R

@@ -1,0 +1,60 @@
+#### Discriminant function analysis of modules
+
+library(plyr)
+library(dplyr)
+
+genotype_gall_parasitoid_network <- read.csv('~/Documents/Genotype_Networks/data/genotype_gall_parasitoid_network_data.csv')
+genotype_gall_parasitoid_network$Genotype <- genotype_gall_parasitoid_network$X
+genotype_gall_parasitoid_network <- select(genotype_gall_parasitoid_network, -X)
+
+module.info <- read.csv('~/Documents/Genotype_Networks/data/module_info_genotype_gall_parasitoid_network.csv')
+module.info.genotype = module.info %>%
+  select(Genotype = Vertex, Module.ID, Trophic) %>%
+  filter(Trophic == "Genotype")
+
+gall_density_genotype_summary_df <- read.csv('~/Documents/Genotype_Networks/data/gall_density_genotype_summary_df.csv')
+gall_density_genotype_summary_df <- select(gall_density_genotype_summary_df, -X)
+
+gall_size_genotype_summary_df <- read.csv('~/Documents/Genotype_Networks/data/gall_size_genotype_summary_df.csv')
+gall_size_genotype_summary_df <- select(gall_size_genotype_summary_df, -X)
+
+dfa_data <- join_all(list(genotype_gall_parasitoid_network, module.info.genotype, gall_density_genotype_summary_df, gall_size_genotype_summary_df))
+dfa_data_other <- dfa_data
+
+### Perhaps finding the regressions for some of the key players, rather than trying to see which ones predict the modules the best is the key...
+
+cor.test(dfa_data_other$vLG_volume_mean, dfa_data_other$vLG_abund, na.action = na.omit)
+plot(vLG_Platy ~ vLG_volume_mean, dfa_data_other)
+vLG_Platy_glm <- glm(rsLG_Eury ~ rsLG_abund, dfa_data_other, family = "poisson")
+summary(vLG_Platy_glm)
+plot(vLG_Platy_glm)
+
+library(visreg)
+visreg(vLG_Platy_glm, scale = "response")
+
+dfa_data <- dfa_data %>%
+  select(Genotype, Module.ID, aSG_abund:rsLG_abund, vLG_volume_mean, rG_volume_mean) %>%
+  mutate(Module.ID = as.factor(Module.ID))
+
[…]

### 2014-11-15 Rscripts/food_web_overall.R

@@ -1,0 +1,75 @@
+############# Overall Gall-Parasitoid Network Analysis
+
+# load libraries
+library(reshape2)
+library(reshape)
+library(dplyr)
+library(bipartite)
+library(RColorBrewer)
+
+#source('~/Documents/ggnet/bipartite_plot_info.R')
+#source('~/Documents/ggnet/ggnet_bipartite.R')
+
+# upload molten gall network data and manage it.
+gall_net_melt <- read.csv("~/Documents/Genotype_Networks/data/gall_network_data.csv")
+gall_net_melt <- tbl_df(gall_net_melt)
+
+total_gall_parasitoid_network <- gall_net_melt %>%
+  filter(gall_contents %in% c("aSG.larv", "Pont.ad", "Pont.prep", "rG.larv", "SG.larv", "vLG.pupa", "Eulo.fem", "Eulo.mal", "Eury.fem", "Eury.mal", "Lathro.fem", "Lathro.mal", "Lestodip", "Mesopol", "Mymarid", "Platy", "Ptero.2", "Tory.fem", "Tory.mal")) 
+
+total_gall_parasitoid_network <- mutate(total_gall_parasitoid_network, gall_contents_collapse = revalue(gall_contents, c("Pont.ad" = "Pont.surv", "Pont.prep" = "Pont.surv", "Eulo.fem" = "Eulo", "Eulo.mal" = "Eulo", "Eury.fem" = "Eury", "Eury.mal" = "Eury", "Lathro.fem" = "Lathro", "Lathro.mal" = "Lathro", "Ptero.2" = "Mesopol", "Tory.fem" = "Tory", "Tory.mal" = "Tory") ))
+
+gall.ptoid.net <- dcast(total_gall_parasitoid_network, gall.sp ~ gall_contents_collapse, sum)
+gall.ptoid.net <- gall.ptoid.net %>%
+  mutate(survive = aSG.larv + rG.larv + SG.larv + vLG.pupa + Pont.surv) %>%
+  select(gall.sp, Eulo, Eury, Lathro, Lestodip, Mesopol, Mymarid, Platy, Tory, survive)
+rownames(gall.ptoid.net) <- gall.ptoid.net$gall.sp
+gall.ptoid.net$total <- rowSums(select(gall.ptoid.net, -gall.sp))
+
+total.interactions <- sum(rowSums(select(gall.ptoid.net, Eulo:Tory)))
[…]

### 2014-11-15 Rscripts/gall_net_highlights.R

@@ -1,0 +1,368 @@
+source('~/Documents/Genotype_Networks/Rscripts/matrix_highlights.R')
+source('~/Documents/Genotype_Networks/Rscripts/plotweb2_highlights.R')
+source('~/Documents/ggnet/bipartite_plot_info.R')
+source('~/Documents/ggnet/tripartite_plot_info.R')
+source('~/Documents/ggnet/ggnet_bipartite.R')
+
+library(reshape)
+library(reshape2)
+library(bipartite)
+library(dplyr)
+library(igraph)
+
+
+########### upload datasets
+
+###  upload general molten gall network data
+gall_net_melt <- read.csv("~/Documents/Genotype_Networks/data/gall_network_data.csv")
+gall_net_melt <- tbl_df(gall_net_melt)
+
+total_gall_parasitoid_network <- gall_net_melt %>%
+  filter(gall_contents %in% c("aSG.larv", "Pont.ad", "Pont.prep", "rG.larv", "SG.larv", "vLG.pupa", "Eulo.fem", "Eulo.mal", "Eury.fem", "Eury.mal", "Lathro.fem", "Lathro.mal", "Lestodip", "Mesopol", "Mymarid", "Platy", "Ptero.2", "Tory.fem", "Tory.mal")) 
+
+total_gall_parasitoid_network <- mutate(total_gall_parasitoid_network, gall_contents_collapse = revalue(gall_contents, c("Pont.ad" = "Pont.surv", "Pont.prep" = "Pont.surv", "Eulo.fem" = "Eulo", "Eulo.mal" = "Eulo", "Eury.fem" = "Eury", "Eury.mal" = "Eury", "Lathro.fem" = "Lathro", "Lathro.mal" = "Lathro", "Ptero.2" = "Mesopol", "Tory.fem" = "Tory", "Tory.mal" = "Tory") ))
+
+# shoot estimates on a genotype level
+shootEsts_df <- gall_net_melt %>%
+  mutate(plant.position = as.factor(plant.position)) %>%
+  group_by(Gender, Genotype, plant.position) %>%
+  summarise(shootEst.no18 = mean(shootEst.no18), shootEst.all = mean(shootEst.all), galls.found = mean(Galls.found)) # needed to take the mean because these shoot estimates are duplicated throughout the molten dataframe
+shootEsts_df <- mutate(shootEsts_df, plant.position = as.character(plant.position))
+
+shootEsts_genotype_df <- shootEsts_df %>%
+  group_by(Genotype) %>%
[…]

### 2014-11-15 Rscripts/gall_net_pies.R

@@ -1,0 +1,198 @@
+# source in functions
+source('~/Documents/ggnet/bipartite_plot_info.R')
+source('~/Documents/ggnet/ggnet_bipartite.R')
+
+
+#### load data
+
+#  upload general molten gall network data
+gall_net_melt <- read.csv("~/Documents/Genotype_Networks/data/gall_network_data.csv")
+gall_net_melt <- tbl_df(gall_net_melt)
+
+total_gall_parasitoid_network <- gall_net_melt %>%
+  filter(gall_contents %in% c("aSG.larv", "Pont.ad", "Pont.prep", "rG.larv", "SG.larv", "vLG.pupa", "Eulo.fem", "Eulo.mal", "Eury.fem", "Eury.mal", "Lathro.fem", "Lathro.mal", "Lestodip", "Mesopol", "Mymarid", "Platy", "Ptero.2", "Tory.fem", "Tory.mal")) 
+
+total_gall_parasitoid_network <- mutate(total_gall_parasitoid_network, gall_contents_collapse = revalue(gall_contents, c("Pont.ad" = "Pont.surv", "Pont.prep" = "Pont.surv", "Eulo.fem" = "Eulo", "Eulo.mal" = "Eulo", "Eury.fem" = "Eury", "Eury.mal" = "Eury", "Lathro.fem" = "Lathro", "Lathro.mal" = "Lathro", "Ptero.2" = "Mesopol", "Tory.fem" = "Tory", "Tory.mal" = "Tory") ))
+
+# focus on gall-parasitoids
+total_gall_parasitoid_network <- dcast(total_gall_parasitoid_network, gall.sp ~ gall_contents_collapse, sum)
+rownames(total_gall_parasitoid_network) <- total_gall_parasitoid_network$gall.sp
+total_gall_parasitoid_network <- select(total_gall_parasitoid_network, -gall.sp)
+interaxns_focus <- select(total_gall_parasitoid_network, -c(aSG.larv, Pont.surv, rG.larv, SG.larv, vLG.pupa))
+interaxns_focus_density <- interaxns_focus/shootEsts_genotype_df$shootEst.no18*1000
+
+interaxns_focus_qual <- interaxns_focus_density
+interaxns_focus_qual[interaxns_focus_qual > 0] <- 1
+
+gall_ptoid_qual <- bipartite_plot_info(interaxns_focus_qual, order.type = "cca")
[…]

### 2014-11-15 Rscripts/gall_network_analysis_modularity_nestedness.R

@@ -1,0 +1,331 @@
+source('~/Documents/vegan/R/nestednodf.R')
+source('~/Documents/Genotype_Networks/Rscripts/null_model_analysis_WNODF_QuaBiMo.R')
+source('~/Documents/Genotype_Networks/Rscripts/best_QuaBiMo_function.R')
+source('~/Documents/Genotype_Willow_Community/datasets_&_Rscripts/beta.div.R')
+
+library(dplyr)
+library(bipartite)
+library(GGally)
+library(ggplot2)
+
+# possibly required packages for plotting
+
+
+genotype_gall_parasitoid_network <- read.csv('~/Documents/Genotype_Networks/data/genotype_gall_parasitoid_network_data.csv')
+rownames(genotype_gall_parasitoid_network) <- genotype_gall_parasitoid_network$X 
+genotype_gall_parasitoid_network <- select(genotype_gall_parasitoid_network, -X)
+
+
+###### Plots of Contributions to network structure
+
+### Genotypes
+geno.contrib.df <- data.frame(contrib = rowSums(genotype_gall_parasitoid_network),
+                              percent.contrib = rowSums(genotype_gall_parasitoid_network)/sum(rowSums(genotype_gall_parasitoid_network))*100,
+                              genotypes = rownames(genotype_gall_parasitoid_network))
+geno.contrib.df <- mutate(geno.contrib.df, geno.ord = reorder(genotypes, percent.contrib))
+mean.percent.contrib <- with(geno.contrib.df, mean(percent.contrib))
+
+hist(log(geno.contrib.df$contrib)) # log-normal distribution?
+
+ggplot(geno.contrib.df, aes(x = geno.ord, y = percent.contrib)) +
+  geom_bar(stat = "identity") +
+  xlab("Genotype") +
+  ylab("Percent Contribution to Network Structure") +
+  coord_flip() +
+  theme_classic() +
+  geom_hline(yintercept = mean.percent.contrib, linetype = "dashed")
+
+# Gall-parasitoid interactions
+gall.ptoid.contrib.df <- data.frame(contrib = colSums(genotype_gall_parasitoid_network),
+                              percent.contrib = colSums(genotype_gall_parasitoid_network)/sum(colSums(genotype_gall_parasitoid_network))*100,
+                              interactions = colnames(genotype_gall_parasitoid_network))
[…]

### 2014-11-15 Rscripts/gall_size_analysis.R

@@ -1,0 +1,398 @@
+
+
+library(dplyr)
+library(reshape2)
+library(visreg)
+library(mgcv)
+library(ggplot2)
+
+# function for calculating adjusted repeatability for nested random effect model. 
+genotype_nested_model <- function(data, y){
+  require(nlme)
+  
+  y <<- y
+  # calculate broad-sense heritability of genotype
+  random_geno_lme <- lme(y ~ 1, 
+                         data, 
+                         random = ~ 1 | Genotype/plant.position, 
+                         method = "REML", 
+                         na.action = na.omit) # use lme function because it is much easier to extract variance of random effects
+
+  gen_var <- as.numeric(VarCorr(random_geno_lme)[2])
+  tot_var <-  gen_var +  as.numeric(VarCorr(random_geno_lme)[5]) # not taking account into the variation explained by plant position. Therefore, this is an estimate of the adjusted repeatability or broad-sense heritability
+  her <- gen_var / tot_var
+  names(her) <- "Broad sense Heritability"
+  print(her)
+  
+  # examining significance of random effect
+  no_geno <- lm(y ~ 1, data) # no random effect to assess the effect of plant genotype. I tested this against a null model instead of a model with just plant.position as a random effect, because the variation due to plant position is primarily a result of plant genotype.
+  rand_anova <- anova(random_geno_lme, no_geno)
+  rand.Lratio <- rand_anova$L.Ratio[2]
+  names(rand.Lratio) <- "Rand eff chi.sq"
+  rand.pval <- rand_anova$"p-value"[2]
+  names(rand.pval) <- "Rand eff p-val"
+  print(rand_anova)
+  
+  # Return all objects in a list
+  output = c(rand.Lratio, rand.pval, her)
+}
+
+
+
+#library(arm)
+#rpt.remlLMM.adj(formula = g.height ~ 1 + (1 | Genotype/plant.position), 
+ #               data = vLG.size.df, 
+  #              grname = "Genotype",
+   #             nboot = 10, 
[…]

### 2014-11-15 Rscripts/genotype_gall-parasitoid_network_tapered-intensity-curved_edges.R

@@ -1,0 +1,47 @@
+# source in functions
+source('~/Documents/ggnet/bipartite_plot_info.R')
+source('~/Documents/ggnet/ggnet_bipartite.R')
+
+
+# load data
+genotype_gall_parasitoid_network <- read.csv('~/Documents/Genotype_Networks/data/genotype_gall_parasitoid_network_data.csv')
+rownames(genotype_gall_parasitoid_network) <- genotype_gall_parasitoid_network$X 
+genotype_gall_parasitoid_network <- select(genotype_gall_parasitoid_network, -X)
+
+module.info <- read.csv('~/Documents/Genotype_Networks/data/module_info_genotype_gall_parasitoid_network.csv')
+module.info <- arrange(module.info, Trophic, Module.ID, Vertex) # reorder for easy plotting
+genotypes <- filter(module.info, Trophic == "Genotype")
+gall.parasitoid <- filter(module.info, Trophic == "Gall-Parasitoid")
+
+# reorder web based on modules
+web <- genotype_gall_parasitoid_network
+web <- web[as.character(genotypes$Vertex), # reorder within web
+           as.character(gall.parasitoid$Vertex)] # reorder within web
+
+gall.net.info <- bipartite_plot_info(web = web, order.type = "normal")
+
+# add some important node attributes
+gall.net.info[[2]]$Module.ID <- factor(c(genotypes$Module.ID, gall.parasitoid$Module.ID))
+gall.net.info[[2]]$Trophic <- factor(c(genotypes$Trophic, gall.parasitoid$Trophic))
+
+# color blind friednly color palette
+cb_palette <- c("#56B4E9", "#999999", "#E69F00", "#009E73", "#CC79A7")
+col.mod1 <- "#56B4E9"
+col.mod2 <- "#999999"
+col.mod3 <- "#E69F00"
+col.mod4 <- "#009E73"
+col.mod5 <- "#CC79A7"
+
+# create initial plot
+gall.net.plot <- ggnet_bipartite(gall.net.info)
+
+# make plot pretty. Save as .png, minimum width aspect ratio of 1100
+gall.net.plot + 
+  geom_point(data = gall.net.info[[2]], aes(x = x, y = y, fill = Module.ID, shape = Trophic), size = 12) +
+  scale_fill_manual(values = cb_palette, guide = "none") +
+  scale_shape_manual(values = c(23, 21), guide = "none")
+  
[…]

### 2014-11-15 Rscripts/genotype_gall_ptoid_modules_figure.R

@@ -1,0 +1,436 @@
+##### Setup
+
+# source in required functions
+source('~/Documents/Genotype_Networks/Rscripts/matrix_highlights.R')
+source('~/Documents/Genotype_Networks/Rscripts/plotweb2_highlights.R')
+source('~/Documents/ggnet/bipartite_plot_info.R')
+source('~/Documents/ggnet/tripartite_plot_info.R')
+source('~/Documents/ggnet/ggnet_bipartite.R')
+
+library(reshape)
+library(reshape2)
+library(bipartite)
+library(plyr)
+library(dplyr)
+library(igraph)
+
+#  upload general molten gall network data
+gall_net_melt <- read.csv("~/Documents/Genotype_Networks/data/gall_network_data.csv")
+gall_net_melt <- tbl_df(gall_net_melt)
+
+total_gall_parasitoid_network <- gall_net_melt %>%
+  filter(gall_contents %in% c("aSG.larv", "Pont.ad", "Pont.prep", "rG.larv", "SG.larv", "vLG.pupa", "Eulo.fem", "Eulo.mal", "Eury.fem", "Eury.mal", "Lathro.fem", "Lathro.mal", "Lestodip", "Mesopol", "Mymarid", "Platy", "Ptero.2", "Tory.fem", "Tory.mal")) 
+
+total_gall_parasitoid_network <- mutate(total_gall_parasitoid_network, gall_contents_collapse = revalue(gall_contents, c("Pont.ad" = "Pont.surv", "Pont.prep" = "Pont.surv", "Eulo.fem" = "Eulo", "Eulo.mal" = "Eulo", "Eury.fem" = "Eury", "Eury.mal" = "Eury", "Lathro.fem" = "Lathro", "Lathro.mal" = "Lathro", "Ptero.2" = "Mesopol", "Tory.fem" = "Tory", "Tory.mal" = "Tory") ))
+
+# shoot estimates on a genotype level
+shootEsts_df <- gall_net_melt %>%
+  mutate(plant.position = as.factor(plant.position)) %>%
+  group_by(Gender, Genotype, plant.position) %>%
+  summarise(shootEst.no18 = mean(shootEst.no18), shootEst.all = mean(shootEst.all), galls.found = mean(Galls.found)) # needed to take the mean because these shoot estimates are duplicated throughout the molten dataframe
+shootEsts_df <- mutate(shootEsts_df, plant.position = as.character(plant.position))
+
+shootEsts_genotype_df <- shootEsts_df %>%
+  group_by(Genotype) %>%
[…]

### 2014-11-15 Rscripts/genotype_level_associations_gall_net.R

@@ -1,0 +1,116 @@
+# Genotype level trait-gall and gall-ptoid associations
+library(dplyr)
+library(psych)
+library(ggplot2)
+library(visreg)
+
+# load data
+plant.traits <- read.csv('~/Documents/Genotype_Networks/data/plant.trait.galls.2011.genotype.df.csv')
+plant.traits <- select(plant.traits, -X)
+
+genotype_level_df <- read.csv('~/Documents/Genotype_Networks/data/genotype_level.df.csv')
+genotype_level_df <- select(genotype_level_df, -X)
+
+module.info <- read.csv('~/Documents/Genotype_Networks/data/module_info_genotype_gall_parasitoid_network.csv')
+module.info <- arrange(module.info, Trophic, Module.ID, Vertex) # reorder for easy plotting
+genotypes <- filter(module.info, Trophic == "Genotype")
+genotypes <- select(genotypes, Genotype = Vertex, Module.ID)
+
+#galldens <- read.csv("~/Documents/Genotype_Networks/data/gall_density_genotype_summary_df.csv")
+#galldens <- select(galldens, -X)
+
+vLG_gallsize <- read.csv("~/Documents/Genotype_Networks/data/vLG_gall_size_geno.df.csv")
+vLG_gallsize <- select(vLG_gallsize, -X)
+#gallsize <- read.csv("~/Documents/Genotype_Networks/data/gall_size_genotype_summary_df.csv")
+#gallsize <- select(gallsize, -X)
+
+#galldens.size <- left_join(galldens, gallsize)
+#galldens.size.noU <- filter(galldens.size, Genotype != "U")
+
+geno.df <- plyr::join_all(list(genotype_level_df, plant.traits, genotypes, vLG_gallsize), type = "left")
+
+# vLG_Platy
+plot(vLG_Platy ~ vLG_density, geno.df)
+plot(vLG_Platy ~ g.height.vLG.mean, geno.df)
+plot(vLG_Platy ~ prop.small.vLG, geno.df)
+
+plot(vLG_Mesopol ~ vLG_density, geno.df)
+plot(vLG_Mesopol ~ g.height.vLG.mean, geno.df)
+plot(vLG_Mesopol ~ prop.small.vLG, geno.df)
+
+small.vLG.density <- with(geno.df, small.vLG.count/shootEst.no18)
+
[…]

### 2014-11-15 Rscripts/matrix_highlights.R

@@ -1,0 +1,61 @@
+# matrix highlighter function for plotweb2_highlights
+
+matrix_highlighter <- function(web, high.names = "", low.names = "") {
+  
+  web <- as.matrix(web)
+  
+  grab.colnames <- which(colnames(web) %in% high.names)
+  grab.rownames <- which(rownames(web) %in% low.names)
+  
+  web.sub <- web
+  web.sub[ ,-grab.colnames] <- 0
+  web.sub[-grab.rownames, ] <- 0
+  
+  out <- web.sub
+}
+
+
+matrix_highlighter_2 <- function(web, interaction.names = "", replace.values = NULL) {
+  require(dplyr)
+  
+  if(!is.null(replace.values)){
+    replace.df <- data.frame(interaction.names = interaction.names, replace.values = replace.values)
+  }
+  
+  web <- data.frame(web)
+  web$low.nodes <- factor(rownames(web))
+  
+  web.melt <- melt(web)
+  web.melt <- mutate(web.melt, interactions = paste(low.nodes, variable, sep = "_"))
+  
+  web.melt.sub <- web.melt
+  web.melt.sub[-which(web.melt.sub$interactions %in% interaction.names), 3] <- 0
+  
+  if(!is.null(replace.values)){
+    interaction.order <- web.melt.sub$interactions[which(web.melt.sub$interactions %in% interaction.names)]
+    replace.df <- mutate(replace.df, interaction.order = factor(interaction.order, ordered = TRUE))
+    test <- match(replace.df$interaction.names, replace.df$interaction.order)
+    #replace.df.sort <- sort_df(replace.df, "interaction.order")
+    test2 <- web.melt.sub[which(web.melt.sub$interactions %in% interaction.names), ]
+    test3 <- replace.df$replace.values[test]
+    
+    web.melt.sub <- web.melt.sub %>% 
+      left_join(replace.df %>% select(interactions = interaction.names, replace.values),
+                missing = 0) %>%
+      mutate(value = ifelse(is.na(replace.values),yes = 0,no = replace.values))
+  }
+    #web.melt.sub[which(web.melt.sub$interactions %in% interaction.names), 3] <- replace.df$replace.values[test]
+  #}
+
+  web.melt.sub.matrix <- dcast(web.melt.sub, low.nodes ~ variable, value.var = "value")
[…]

### 2014-11-15 Rscripts/network_management_genotype_level.R

@@ -1,0 +1,31 @@
+library(dplyr)
+
+tree_level_interaxn_all_plants <- read.csv('~/Documents/Genotype_Networks/data/tree_level_interaxn_all_plants.csv')
+
+### Genotype level data including larva survival and shoot estimates as separate variables
+genotype_level.df <- tree_level_interaxn_all_plants %>%
+  group_by(Genotype) %>%
+  select(shootEst.no18, shootEst.all, aSG_aSG.larv:gall_survive_abund) %>%
+  summarise_each(funs(sum)) %>%
+  ungroup()
+
+write.csv(genotype_level.df, "~/Documents/Genotype_Networks/data/genotype_level.df.csv")
+
+
+### Modularity and nestedness analysis. 
+modularity_df <- tree_level_interaxn_all_plants %>%
+  group_by(Genotype) %>%
+  select(aSG_Tory:rG_Platy, rG_Tory:rsLG_Lathro, SG_Platy, vLG_Eulo:vLG_Tory, shootEst.no18, shootEst.all) %>%
+  filter(Genotype != "U") %>% # removing because there were no observed interactions
+  summarise_each(funs(sum)) %>%
+  ungroup()
+
+rownames(modularity_df) <- as.character(modularity_df$Genotype)
+modularity_df_interaxns <- modularity_df %>%
+  select(-Genotype) %>%
+  apply(MARGIN = 2, FUN = function(x) round(x/modularity_df$shootEst.no18*1000)) # standardize the frequency of all interactions by dividing by the number of shoots sampled, mulitplying it by 1000, then rounding to the nearest whole integer. Therefore, this represents the frequency of each interaction per 1000 shoots sampled. I chose 1000 shoots because it was near the minimum number of shoots sampled per genotype.
+modularity_df_interaxns <- as.data.frame(modularity_df_interaxns)
+modularity_df_interaxns <- select(modularity_df_interaxns, aSG_Tory:vLG_Tory)
+
+write.csv(modularity_df_interaxns, "~/Documents/Genotype_Networks/data/genotype_gall_parasitoid_network_data.csv")
+

### 2014-11-15 Rscripts/network_management_tree_level.R

@@ -1,0 +1,104 @@
+######## Tree level Network, Community, and Species level analysis
+
+#### upload necessary packages
+
+# analysis
+library(mvabund)
+library(vegan)
+library(visreg)
+library(bipartite)
+
+# data manipulation. order of libraries is important. always load dplyr last.
+library(reshape2)
+library(reshape)
+library(dplyr)
+
+# plotting
+library(ggplot2)
+
+
+# upload molten gall network data
+gall_net_melt <- read.csv("~/Documents/Genotype_Networks/data/gall_network_data.csv")
+gall_net_melt <- tbl_df(gall_net_melt) # 1,511 galls
+sum(gall_net_melt$value[1:1511]) # 1740 total samples
+sum(filter(gall_net_melt, gall_contents == "nothing")$value) # 530 "nothing" samples
+sum(filter(gall_net_melt, gall_contents == "exit.hole")$value) # 136
+
+##### Does the number of shoots sampled vary among genotypes?
+
+# data frame
+shootEsts_df <- gall_net_melt %>%
+  mutate(plant.position = as.factor(plant.position)) %>%
+  group_by(Gender, Genotype, plant.position) %>%
+  summarise(shootEst.no18 = mean(shootEst.no18), shootEst.all = mean(shootEst.all), galls.found = mean(Galls.found)) # needed to take the mean because these shoot estimates are duplicated throughout the molten dataframe
+shootEsts_df <- mutate(shootEsts_df, plant.position = as.character(plant.position))
+
+shootEsts_genotype_df <- shootEsts_df %>%
+  group_by(Genotype) %>%
+  summarise(n = n(), shootEst.no18 = sum(shootEst.no18), shootEst.all = sum(shootEst.all))
+
+
+# create the tree-level data
+
+tree_interaxns_filter <- gall_net_melt %>%
+  filter(gall_contents %in% c("aSG.larv", "Pont.ad", "Pont.prep", "rG.larv", "SG.larv", "vLG.pupa", "Eulo.fem", "Eulo.mal", "Eury.fem", "Eury.mal", "Lathro.fem", "Lathro.mal", "Lestodip", "Mesopol", "Mymarid", "Platy", "Ptero.2", "Tory.fem", "Tory.mal")) # 764 galls
+sum(tree_interaxns_filter$value) # 991 individuals
+
[…]

### 2014-11-15 Rscripts/network_mechanism_models.R

@@ -1,0 +1,1716 @@
+#####################################################################
+# Random and Mixed effect models of plant genotype correlations with gall density and gall size, as well as how these factors shape gall survival.
+#####################################################################
+
+##### Upload required packages and functions for analysis
+
+### packages
+
+# visualization
+library(visreg)
+library(car)
+library(ggplot2)
+
+# analysis
+library(psych)
+library(RLRsim)
+library(MASS)
+library(mgcv)
+library(gamm4)
+library(nlme)
+library(lme4)
+library(pbkrtest)
+library(vegan)
+
+# data manipulation
+library(reshape)
+library(reshape2) 
+library(plyr)
+library(dplyr)
+
+### functions
+
+source('~/Documents/Genotype_Willow_Community/datasets_&_Rscripts/functions_ms_willow_community.R') # used for testing for random effects of plant genotype on gall densities
+
+overdisp_fun <- function(model) {
+  ## number of variance parameters in 
+  ##   an n-by-n variance-covariance matrix
+  vpars <- function(m) {
+    nrow(m)*(nrow(m)+1)/2
+  }
+  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
+  rdf <- nrow(model.frame(model))-model.df
+  rp <- residuals(model,type="pearson")
+  Pearson.chisq <- sum(rp^2)
+  prat <- Pearson.chisq/rdf
+  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
+  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
+}
+
+# function for evaluating normality of random effects assumption.
+ggQQ_ranef <- function(ranef) # argument: vector of random effects
+{
+  y <- quantile(ranef, c(0.25, 0.75))
+  x <- qnorm(c(0.25, 0.75))
+  slope <- diff(y)/diff(x)
+  int <- y[1L] - slope * x[1L]
+  p <- ggplot(data.frame(ranef = ranef), aes(sample=ranef)) +
+    stat_qq(alpha = 0.5) +
+    geom_abline(slope = slope, intercept = int, color="blue")
+  
+  return(p)
+} 
+
+# function for calculating adjusted repeatability for nested random effect model. 
+genotype_nested_model <- function(data, y){
+  require(nlme)
+  
+  y <<- y
+  # calculate broad-sense heritability of genotype
+  random_geno_lme <- lme(y ~ 1, 
+                         data, 
[…]

### 2014-11-15 Rscripts/null_model_analysis_WNODF_QuaBiMo.R

@@ -1,0 +1,35 @@
+null_model_analysis_WNODF_QuaBiMo <- function(web, type, observed_value, N_null_webs = 10000, null_model = "swap.web") {
+  require('bipartite')
+  source('~/Documents/vegan/R/nestednodf.R')
+  
+  # generate null models
+  null_web_list <- nullmodel(web, N = N_null_webs, method = null_model)
+  
+  # WNODF analysis
+  if(type == "WNODF"){
+    
+    nulls <- sapply(null_web_list, FUN = function(x) nestednodf(x, order = TRUE, weighted = TRUE))
+    
+    null_values <- vector()
+    for(i in 1:dim(nulls)[2]){
+      null_values[i] <- nulls[ ,i]$statistic[3]
+    }
+  }
+  
+  # QuaBiMo analysis
+  if(type == "QuaBiMo"){
+    
+    nulls <- sapply(null_web_list, computeModules)
+    
+    null_values <- sapply(nulls, function(x) x@likelihood)  
+  }
+  
+  mean_null_values <- mean(null_values)
+  sd_null_values <- sd(null_values)
+  
+  z_score <- (observed_value - mean_null_values)/sd_null_values
+  names(z_score) <- "z-score"
+  names(mean_null_values) <- "mean of null values"
+  names(sd_null_values) <- "SD of null values"
+  return(c(z_score, mean_null_values, sd_null_values))
+}

### 2014-11-15 Rscripts/plant.trait.df.R

@@ -1,0 +1,36 @@
+library(dplyr)
+library(visreg)
+library(ggplot2)
+
+# upload and select plant trait data
+plant.trait.data <- read.csv('~/Documents/Genotype_Willow_Community/datasets_&_Rscripts/arthropods_&_traits/complete.data.csv')
+plant.trait.data <- tbl_df(plant.trait.data)
+
+plant.trait.data <- plant.trait.data %>%
+  select(Genotype, plant.position = Plant.Position, vLG, rsLG, SG, PG, Total_Area:flavanonOLES.PC1) # includes 2011 gall data (except for rG)
+
+write.csv(plant.trait.data, "~/Documents/Genotype_Networks/data/plant.trait.galls.2011.tree.df.csv")
+
+vLG_2011 <- MASS::glm.nb(vLG ~ offset(log(Total_Area)) + Genotype, plant.trait.data)
+summary(vLG_2011)
+vLG_2011_fit <- visreg(vLG_2011, "Genotype", scale = "response", cond = list(Total_Area = 2), ylim = c(0,60))
+vLG_2011_density <- vLG_2011_fit$Genotype$y$fit
+
+plant.trait.geno.df <- plant.trait.data %>%
+  group_by(Genotype) %>%
+  select(-Genotype, -plant.position) %>%
+  summarise_each(funs(mean.na.rm = mean(., na.rm = TRUE))) %>%
+  mutate(vLG_2011_glm.nb_density = vLG_2011_density)
+
+write.csv(plant.trait.geno.df, '~/Documents/Genotype_Networks/data/plant.trait.galls.2011.genotype.df.csv')
+
+#plant.trait.geno.noQ <- filter(plant.trait.geno.df, Genotype != "Q")
+
+ggplot(plant.trait.geno.df, aes(x = log(C_N_ratio), y = log(vLG_2011_density+1))) + geom_point() + stat_smooth(method = "lm")
+vLG_2011_CN <- lm(log(vLG_2011_density+1) ~ log(C_N_ratio), plant.trait.geno.df)
+summary(vLG_2011_CN)
+plot(vLG_2011_CN)
+
+vLG_2011_height <- lm(vLG_2011_density ~ Height, plant.trait.geno.df)
+summary(vLG_2011_height)
+visreg(vLG_2011_height)

### 2014-11-15 Rscripts/plotweb2_highlights.R

@@ -1,50 +1,52 @@
-# ideas, add two separate food web matrices  that I want to "highlight"
 
-# load data
-genotype_gall_parasitoid_network <- read.csv('~/Documents/Genotype_Networks/data/genotype_gall_parasitoid_network_data.csv')
-rownames(genotype_gall_parasitoid_network) <- genotype_gall_parasitoid_network$X 
-genotype_gall_parasitoid_network <- select(genotype_gall_parasitoid_network, -X)
-
-gall_ptoid <- genotype_gall_parasitoid_network
-
-rowSums(select(gall_ptoid, "vLG_Platy"))
-grab.geno <- which(rownames(gall_ptoid) %in% c("B","D","X"))
-grab.interaction <- which(colnames(gall_ptoid) %in% "vLG_Platy")
-vLG_Platy_BDX <- sum(gall_ptoid[grab.geno, grab.interaction])
-interaxns_vLG_Platy_sub <- interaxns_focus_density
-interaxns_vLG_Platy_sub["vLG","Platy"] <- vLG_Platy_BDX
-interaxns_vLG_Platy_sub[-5,] <- 0
-interaxns_vLG_Platy_sub[,-7] <- 0
-
-debug(plotweb2_highlights)
-plotweb2_highlights(genotype_gall_network_density, interaxns_focus_density, low.abun.col2 = "white", highlight.web = low.web.high, highlight.web.high = interaxns_vLG_Platy_sub, node.highlight = c("vLG","X","B","D","Platy"))
-
-grab.prey <- which(rownames(genotype_gall_network_density) %in% c("B","D","X"))
-grab.pred <- which(colnames(genotype_gall_network_density) %in% "vLG")
-
-low.web.high <- genotype_gall_network_density
-low.web.high[-grab.prey, ] <- 0
-low.web.high[ ,-grab.pred] <- 0
-
-
-plotweb2_highlights <- function (web, web2, method = "cca", empty = FALSE, labsize = 1, 
-                                 ybig = 1, y_width = 0.1, spacing = 0.05, arrow = "no", col.interaction = "grey80", 
-                                 col.pred = "grey10", col.prey = "grey10",
+plotweb2_highlights <- function (web, 
+                                 web2, 
+                                 highlight.web = NULL,
+                                 highlight.web.high = NULL,
+                                 method = "cca", 
+                                 empty = FALSE, # when FALSE, method changes to "normal"
+                                 labsize = 1, 
+                                 ybig = 1, 
+                                 y_width = 0.1,
[…]

### 2014-11-15 Rscripts/proportion_parasitized_contingency_analysis.R

@@ -1,0 +1,82 @@
+library(reshape2)
+library(reshape)
+library(dplyr)
+
+# upload molten gall network data
+gall_net_melt <- read.csv("~/Documents/Genotype_Networks/data/gall_network_data.csv")
+gall_net_melt <- tbl_df(gall_net_melt)
+
+##### Does the proportion of parasitized vLG galls vary among genotypes?
+
+### contingency table analyses for each gall species separately. 
+
+# Iteomyia salicisverruca
+vLG_geno_parasitism <- gall_net_melt %>%
+  dcast(Genotype ~ gall.sp + gall_contents, sum) %>%
+  select(Genotype, starts_with("vLG")) %>%
+  mutate(vLG_ptoid_attack = vLG_Eulo.fem + vLG_Eulo.mal + vLG_Mesopol + vLG_Mymarid + vLG_Platy + vLG_Ptero.2 + vLG_Tory.fem + vLG_Tory.mal, vLG_prop_ptoid_attack = vLG_ptoid_attack/(vLG_vLG.pupa + vLG_ptoid_attack)) %>% # omitting exit hole because of potential overlap with other ectoparasitoids. Need to probably resolve this by not using "unk" gall.id
+  select(Genotype, vLG_ptoid_attack, vLG_vLG.pupa, vLG_prop_ptoid_attack)
+
+vLG_for_chi <- vLG_geno_parasitism %>%
+  filter(vLG_ptoid_attack > 0 & vLG_vLG.pupa > 0) %>% # remove all observations with zeros in either column
+  select(vLG_ptoid_attack, vLG_vLG.pupa)
+
+chisq.test(vLG_for_chi)
+chisq.test(vLG_for_chi, simulate.p.value = TRUE, B = 10000) # same results as above
+
+# Rabdophaga salicisbrassicoides
+rG_geno_parasitism <- gall_net_melt %>%
+  dcast(Genotype ~ gall.sp + gall_contents, sum) %>%
+  select(Genotype, starts_with("rG")) %>%
[…]

### 2014-11-15 Rscripts/survey_2_network_management.R

@@ -13,16 +13,26 @@ library(dplyr)
 
 #### Data management
 
-# upload gall network data (survey #2)
-Gall_Network_Data_Survey_2_2012 <- read.csv("~/Documents/Genotype_Networks/data/Gall_Network_Data_Survey_2_raw.csv", skip=1, stringsAsFactors = FALSE, strip.white = TRUE)
-Gall_Network_Data_Survey_2_2012 <- tbl_df(Gall_Network_Data_Survey_2_2012)
-Gall_Network_Data_Survey_2_2012[which(Gall_Network_Data_Survey_2_2012$plant.position == 49.8),"plant.position"] <- rep(167, 7) # 49.8 label came from the stem diameter for this plant, but we also knew it was genotype "O". I also noticed I was missing plant 167 from the dataset and plant was Genotype "O" and had a stem diameter of 49.8, therefore I input 167 as the plant position.
+# Shoot count estimates as well as plant positions where zero galls were collected
+shoot.countEst <- read.csv("~/Documents/Genotype_Networks/data/survey_2_stem_diams_shootEsts.csv")
+shoot.count_df <- tbl_df(shoot.countEst)
+shoot.count_df <- shoot.count_df %>%
+  select(plant.position = Plant.Position, Galls.found, shootEst.all, shootEst.no18)
 
 # upload plant position data with details on plant identity to eventually merge with gall network data
 plant.position.info <- read.csv("~/Documents/Genotype_Willow_Community/datasets_&_Rscripts/Willow Garden Positions.csv")
 plant.position.info <- tbl_df(plant.position.info)
 plant.position.info <- plant.position.info %>%
   select(Genotype, Gender, Row = Row.., plant.position = Plant.Position)
+
+# merge plant position info with shoot estimate data
+plant.info_df <- left_join(shoot.count_df, plant.position.info) %>%
+  select(Gender, Genotype, Row, plant.position:shootEst.no18)
+
+# upload gall network data (survey #2)
+Gall_Network_Data_Survey_2_2012 <- read.csv("~/Documents/Genotype_Networks/data/Gall_Network_Data_Survey_2_raw.csv", skip=1, stringsAsFactors = FALSE, strip.white = TRUE)
+Gall_Network_Data_Survey_2_2012 <- tbl_df(Gall_Network_Data_Survey_2_2012)
[…]

### 2014-11-15 Rscripts/tapered-intensity-curved_edges.R

@@ -1,0 +1,62 @@
+doInstall <- TRUE  # Change to FALSE if you don't want packages installed.
+toInstall <- c("sna", "ggplot2", "Hmisc", "reshape2")
+if(doInstall){install.packages(toInstall, repos = "http://cran.r-project.org")}
+lapply(toInstall, library, character.only = TRUE)
+
+# Empty ggplot2 theme
+new_theme_empty <- theme_bw()
+new_theme_empty$line <- element_blank()
+new_theme_empty$rect <- element_blank()
+new_theme_empty$strip.text <- element_blank()
+new_theme_empty$axis.text <- element_blank()
+new_theme_empty$plot.title <- element_blank()
+new_theme_empty$axis.title <- element_blank()
+new_theme_empty$plot.margin <- structure(c(0, 0, -1, -1), unit = "lines",
+                                         valid.unit = 3L, class = "unit")
+
+data(coleman)  # Load a high school friendship network
+adjacencyMatrix <- coleman[1, , ]  # Fall semester
+layoutCoordinates <- gplot(adjacencyMatrix)  # Get graph layout coordinates
+
+adjacencyList <- melt(adjacencyMatrix)  # Convert to list of ties only
+adjacencyList <- adjacencyList[adjacencyList$value > 0, ]
+
+# Function to generate paths between each connected node
+edgeMaker <- function(whichRow, len = 100, curved = TRUE){
+  fromC <- layoutCoordinates[adjacencyList[whichRow, 1], ]  # Origin
+  toC <- layoutCoordinates[adjacencyList[whichRow, 2], ]  # Terminus
+
+  # Add curve:
+  graphCenter <- colMeans(layoutCoordinates)  # Center of the overall graph
+  bezierMid <- c(fromC[1], toC[2])  # A midpoint, for bended edges
+  distance1 <- sum((graphCenter - bezierMid)^2)
+  if(distance1 < sum((graphCenter - c(toC[1], fromC[2]))^2)){
+    bezierMid <- c(toC[1], fromC[2])
+    }  # To select the best Bezier midpoint
+  bezierMid <- (fromC + toC + bezierMid) / 3  # Moderate the Bezier midpoint
+  if(curved == FALSE){bezierMid <- (fromC + toC) / 2}  # Remove the curve
+
+  edge <- data.frame(bezier(c(fromC[1], bezierMid[1], toC[1]),  # Generate
+                            c(fromC[2], bezierMid[2], toC[2]),  # X & y
[…]

### 2014-11-15 Rscripts/test_bipartite_networks.R

@@ -1,0 +1,147 @@
+# Bipartite network with tapered, intensity curved edges
+
+# load libraries
+library(sna)
+library(ggplot2)
+library(Hmisc)
+library(reshape2)
+require(plyr)
+require(dplyr)
+require(vegan)
+
+
+
+# load data
+genotype_gall_parasitoid_network <- read.csv('~/Documents/Genotype_Networks/data/genotype_gall_parasitoid_network_data.csv')
+rownames(genotype_gall_parasitoid_network) <- genotype_gall_parasitoid_network$X 
+genotype_gall_parasitoid_network <- select(genotype_gall_parasitoid_network, -X)
+
+module.info <- read.csv('~/Documents/Genotype_Networks/data/module_info_genotype_gall_parasitoid_network.csv')
+module.info <- arrange(module.info, Trophic, Module.ID, Vertex) # reorder for easy plotting
+genotypes <- filter(module.info, Trophic == "Genotype")
+gall.parasitoid <- filter(module.info, Trophic == "Gall-Parasitoid")
+
+web <- genotype_gall_parasitoid_network
+web <- web[as.character(genotypes$Vertex), # reorder within web
+           as.character(gall.parasitoid$Vertex)] # reorder within web
+
+web <- as.matrix(web)
+attr(web, "modules") <- list(genotypes$Module.ID,
+                             gall.parasitoid$Module.ID)
+
+# Requires an ordered matrix with a "modules" attributes giving the module ID for each node of the bipartite network
+
+bipartite_info <- function(web){
+
+  # layout for bipartite network. Evenly spaces out trophic level with the fewest number of nodes
+  layout.bipartite <- matrix(nrow = sum(dim(web)), ncol = 2)
+  layout.bipartite[ ,1] <- c(1:max(dim(web)), 
+                             seq(1, max(dim(web)), length.out = min(dim(web)))) 
+  layout.bipartite[ ,2] <- c(rep(which.max(dim(web)), max(dim(web))), rep(which.min(dim(web)), min(dim(web))))
+  
+  # name a replicate of the web "adjacencyMatrix" with unique, numbered values for each node. Necessary for accurate plotting.
+  adjacencyMatrix <- web
+  rownames(adjacencyMatrix) <- 1:dim(web)[1] 
+  colnames(adjacencyMatrix) <- (dim(web)[1]+1):sum(dim(web)) 
+
+  # get layout coordinates for graph
+  layoutCoordinates <- gplot(adjacencyMatrix, 
[…]

### 2014-11-15 Rscripts/total_gall_parasitoid_networks.R

@@ -1,0 +1,117 @@
+# Total Gall-Parasitoid Network Analysis
+library(reshape2)
+library(reshape)
+library(dplyr)
+library(bipartite)
+
+source('~/Documents/ggnet/bipartite_plot_info.R')
+source('~/Documents/ggnet/ggnet_bipartite.R')
+
+# upload molten gall network data
+gall_net_melt <- read.csv("~/Documents/Genotype_Networks/data/gall_network_data.csv")
+gall_net_melt <- tbl_df(gall_net_melt)
+
+total_gall_parasitoid_network <- gall_net_melt %>%
+  filter(gall_contents %in% c("aSG.larv", "Pont.ad", "Pont.prep", "rG.larv", "SG.larv", "vLG.pupa", "Eulo.fem", "Eulo.mal", "Eury.fem", "Eury.mal", "Lathro.fem", "Lathro.mal", "Lestodip", "Mesopol", "Mymarid", "Platy", "Ptero.2", "Tory.fem", "Tory.mal")) 
+
+total_gall_parasitoid_network <- mutate(total_gall_parasitoid_network, gall_contents_collapse = revalue(gall_contents, c("Pont.ad" = "Pont.surv", "Pont.prep" = "Pont.surv", "Eulo.fem" = "Eulo", "Eulo.mal" = "Eulo", "Eury.fem" = "Eury", "Eury.mal" = "Eury", "Lathro.fem" = "Lathro", "Lathro.mal" = "Lathro", "Ptero.2" = "Mesopol", "Tory.fem" = "Tory", "Tory.mal" = "Tory") ))
+
+# shoot estimates on a genotype level
+shootEsts_df <- gall_net_melt %>%
+  mutate(plant.position = as.factor(plant.position)) %>%
+  group_by(Gender, Genotype, plant.position) %>%
+  summarise(shootEst.no18 = mean(shootEst.no18), shootEst.all = mean(shootEst.all), galls.found = mean(Galls.found)) # needed to take the mean because these shoot estimates are duplicated throughout the molten dataframe
+shootEsts_df <- mutate(shootEsts_df, plant.position = as.character(plant.position))
+
+shootEsts_genotype_df <- shootEsts_df %>%
+  group_by(Genotype) %>%
+  summarise(n = n(), shootEst.no18 = sum(shootEst.no18), shootEst.all = sum(shootEst.all))
+
[…]

### 2014-11-15 Rscripts/trait_network_analysis.R

@@ -1,0 +1,108 @@
+
+##### load require libraries
+library(psych)
+library(plyr)
+library(dplyr)
+
+##### upload data
+#genotype_gall_parasitoid_network <- read.csv('~/Documents/Genotype_Networks/data/genotype_gall_parasitoid_network_data.csv')
+#rownames(genotype_gall_parasitoid_network) <- genotype_gall_parasitoid_network$X 
+#genotype_gall_parasitoid_network <- select(genotype_gall_parasitoid_network, -X)
+glevel.net <- read.csv('~/Documents/Genotype_Networks/data/genotype_level.df.csv')
+
+glevel.traits <- read.csv("~/Documents/Genotype_Networks/data/plant.trait.galls.2011.genotype.df.csv")
+glevel.traits <- select(glevel.traits, -X)
+
+glevel.gall.densities <- read.csv("~/Documents/Genotype_Networks/data/gall_density_estimates_genotype_level.csv")
+glevel.gall.densities <- select(glevel.gall.densities, -X)
+
+# module ID
+modules <- read.csv("~/Documents/Genotype_Networks/data/module_info_genotype_gall_parasitoid_network.csv")
+modules <- modules %>%
+  filter(Trophic == "Genotype") %>%
+  mutate(Genotype = as.character(Vertex), Module.ID = factor(Module.ID)) %>%
+  select(Genotype, Module.ID)
+
+# vLG gall size
+vLG_size <- read.csv("~/Documents/Genotype_Networks/data/vLG_gall_size_geno.df.csv")
+vLG_size <- select(vLG_size, -X, Genotype, gall.height.vLG.mean = g.height.vLG.mean, large.vLG.count, small.vLG.count)
+
+
+df <- join_all(list(glevel.net, glevel.traits, modules, vLG_size, glevel.gall.densities), by = "Genotype")
+df <- mutate(df, vLG_ptoid_attack = vLG_Platy + vLG_Mesopol + vLG_Tory + vLG_Mymarid + vLG_Eulo,
+             prop.small.galls = small.vLG.count/(large.vLG.count + small.vLG.count))
+
+#### Correlations
+df.numeric <- select(df, -X, -Genotype, -Module.ID)
[…]

### 2014-11-15 Rscripts/tree_level_networks_analysis.R

@@ -2,721 +2,850 @@
 
 #### upload necessary packages
 
+source('~/Documents/Genotype_Willow_Community/datasets_&_Rscripts/functions_ms_willow_community.R') # used for testing for random effects of plant genotype on gall densities
+
+# function for evaluating normality of random effects assumption.
+ggQQ_ranef <- function(ranef) # argument: vector of random effects
+{
+  y <- quantile(ranef, c(0.25, 0.75))
+  x <- qnorm(c(0.25, 0.75))
+  slope <- diff(y)/diff(x)
+  int <- y[1L] - slope * x[1L]
+  p <- ggplot(data.frame(ranef = ranef), aes(sample=ranef)) +
+    stat_qq(alpha = 0.5) +
+    geom_abline(slope = slope, intercept = int, color="blue")
+  
+  return(p)
+} 
+
+
+# analysis
+#library(mvabund)
+#library(vegan)
+library(visreg)
+#library(bipartite)
+#library(pscl)
+library(car)
+library(lme4)
+library(MASS)
+
 # data manipulation. order of libraries is important. always load dplyr last.
 library(reshape2)
 library(reshape)
+library(plyr)
 library(dplyr)
 
 # plotting
 library(ggplot2)
 
-# analysis
-library(mvabund)
-library(vegan)
+##### upload data
 
-# upload molten gall network data
-gall_net_melt <- read.csv("~/Documents/Genotype_Networks/data/gall_network_data.csv")
-gall_net_melt <- tbl_df(gall_net_melt)
-gall_net_melt <- filter(gall_net_melt, plant.position != 9999) # removed unknown plant.position
+# interaction data
+tree_level_interaxn_all_plants <- read.csv("~/Documents/Genotype_Networks/data/tree_level_interaxn_all_plants.csv")
 
-### NEED TO ADD PLANTS WITH NO GALLS
+table(tree_level_interaxn_all_plants$Genotype) # sample sizes
+sum(table(tree_level_interaxn_all_plants$Genotype))
 
-##### Does the proportion of parasitized vLG galls vary among genotypes?
+# plant trait data
+tree_level_traits <- read.csv("~/Documents/Genotype_Networks/data/plant.trait.galls.2011.tree.df.csv")
[…]

### 2014-11-15 rptR/R/pqlglmm.pois.model.R

@@ -1,0 +1,18 @@
+pqlglmm.pois.model <- function(y, groups, link, returnR=TRUE) {
+  mod     <-  glmmPQL(y ~ 1,random=~1|groups,  family=quasipoisson(link=eval(link)), verbose=FALSE) 
+  VarComp <- nlme::VarCorr(mod)
+  beta0   <- as.numeric(mod$coefficients$fixed)
+  omega   <- (as.numeric(VarComp[2,1]))
+  var.a   <- (as.numeric(VarComp[1,1]))
+  if (link=="log") {
+    R.link  <- var.a/(var.a + omega*log(1/exp(beta0)+1))
+    EY 		<- exp(beta0+var.a/2)
+    R.org 	<- EY*(exp(var.a)-1)/(EY*(exp(var.a)-1)+omega) 
+  }
+  if (link=="sqrt") {
+    R.link  <- var.a/(var.a + omega*0.25)
+    R.org 	<- NA 
+  }	
+  if(returnR) return(list(R.link=R.link, R.org=R.org))
+  else return(list(beta0=beta0, omega=omega, var.a=var.a))
+}

### 2014-11-15 rptR/R/pqlglmm.pois.model.offset.R

@@ -1,0 +1,18 @@
+pqlglmm.pois.model.offset <- function(y, offset.trans, groups, link, returnR=TRUE) {
+  mod     <-  glmmPQL(y ~ 1 + offset(offset.trans),random=~1|groups,  family=quasipoisson(link=eval(link)), verbose=FALSE, control = lmeControl(opt = "optim")) 
+  VarComp <- nlme::VarCorr(mod)
+  beta0   <- as.numeric(mod$coefficients$fixed)
+  omega   <- (as.numeric(VarComp[2,1]))
+  var.a   <- (as.numeric(VarComp[1,1]))
+  if (link=="log") {
+    R.link  <- var.a/(var.a + omega*log(1/exp(beta0)+1))
+    EY   	<- exp(beta0+var.a/2)
+    R.org 	<- EY*(exp(var.a)-1)/(EY*(exp(var.a)-1)+omega) 
+  }
+  if (link=="sqrt") {
+    R.link  <- var.a/(var.a + omega*0.25)
+    R.org 	<- NA 
+  }	
+  if(returnR) return(list(R.link=R.link, R.org=R.org))
+  else return(list(beta0=beta0, omega=omega, var.a=var.a))
+}

### 2014-11-15 rptR/R/rpt.poisGLMM.multi.offset.MAB.R

@@ -1,0 +1,93 @@
+# Matt Barbour has added an offset parameter to allow for estimation of density.
+
+rpt.poisGLMM.multi.offset = function(y, offset.trans, groups, link=c("log", "sqrt"), CI=0.95, nboot=1000, npermut=1000) {
+  # initial checks
+  if(length(y) != length(groups)) 
+    stop("y and group hav to be of equal length")
+  if(nboot < 0)   nboot <- 0
+  if(npermut < 1) npermut <- 1
+  if(length(link) > 1)
+    link   <- link[1]
+  if(link != "log" &  link != "sqrt") 
+    stop("inappropriate link (has to be 'log' or 'sqrt')")
+  if(any(is.na(y))) {
+    warning("missing values in y are removed")
+    groups <- groups[!is.na(y)]
+    y      <- y[!is.na(y)]
+  }
+  # preparation
+  groups <- factor(groups)
+  N <- length(y)
+  k <- length(levels(groups))
+  # functions
+  pqlglmm.pois.model <- function(y, offset.trans, groups, link, returnR=TRUE) {
+    mod     <-  glmmPQL(y ~ 1 + offset(offset.trans),random=~1|groups,  family=quasipoisson(link=eval(link)), verbose=FALSE) 
+    VarComp <- nlme::VarCorr(mod)
+    beta0   <- as.numeric(mod$coefficients$fixed)
+    omega   <- (as.numeric(VarComp[2,1]))
+    var.a   <- (as.numeric(VarComp[1,1]))
+    if (link=="log") {
+      R.link  <- var.a/(var.a + omega*log(1/exp(beta0)+1))
+      EY 		<- exp(beta0+var.a/2)
+      R.org 	<- EY*(exp(var.a)-1)/(EY*(exp(var.a)-1)+omega) 
+    }
+    if (link=="sqrt") {
+      R.link  <- var.a/(var.a + omega*0.25)
+      R.org 	<- NA 
+    }	
+    if(returnR) return(list(R.link=R.link, R.org=R.org))
+    else return(list(beta0=beta0, omega=omega, var.a=var.a))
+  }
+  # point estimation according to model 17 equations 18-20
+  R   <- pqlglmm.pois.model(y, offset.trans, groups, link)
+  
+  #browser()
+  # confidence interval estimation by parametric bootstrapping
[…]

### 2014-11-15 rptR/R/rptR.functions.R

@@ -1,0 +1,73 @@
+#### Functions taken from rptR package
+
+# general count model function
+pqlglmm.pois.model <- function(y, # response variable
+                               groups, # grouping variable
+                               link, # link function
+                               returnR=TRUE) {
+  mod     <-  glmmPQL(y ~ 1,random=~1|groups,  family=quasipoisson(link=eval(link)), verbose=FALSE) 
+  VarComp <- nlme::VarCorr(mod)
+  beta0   <- as.numeric(mod$coefficients$fixed)
+  omega   <- (as.numeric(VarComp[2,1]))
+  var.a   <- (as.numeric(VarComp[1,1]))
+  if (link=="log") {
+    R.link  <- var.a/(var.a + omega*log(1/exp(beta0)+1))
+    EY   	<- exp(beta0+var.a/2)
+    R.org 	<- EY*(exp(var.a)-1)/(EY*(exp(var.a)-1)+omega) 
+  }
+  if (link=="sqrt") {
+    R.link  <- var.a/(var.a + omega*0.25)
+    R.org 	<- NA 
+  }	
+  if(returnR) return(list(R.link=R.link, R.org=R.org))
+  else return(list(beta0=beta0, omega=omega, var.a=var.a))
+}
+
+# general count model with offset option. Most parameters same as for general count model
+pqlglmm.pois.model.offset <- function(y, 
+                                      offset.trans, # variable to use for offsetting that has already been transformed according to the link function used.
+                                      groups, 
+                                      link, 
+                                      returnR=TRUE) {
+  mod     <-  glmmPQL(y ~ 1 + offset(offset.trans),random=~1|groups,  family=quasipoisson(link=eval(link)), verbose=FALSE) 
+  VarComp <- nlme::VarCorr(mod)
+  beta0   <- as.numeric(mod$coefficients$fixed)
+  omega   <- (as.numeric(VarComp[2,1]))
+  var.a   <- (as.numeric(VarComp[1,1]))
+  if (link=="log") {
+    R.link  <- var.a/(var.a + omega*log(1/exp(beta0)+1))
+    EY     <- exp(beta0+var.a/2)
+    R.org 	<- EY*(exp(var.a)-1)/(EY*(exp(var.a)-1)+omega) 
+  }
+  if (link=="sqrt") {
+    R.link  <- var.a/(var.a + omega*0.25)
+    R.org 	<- NA 
+  }	
[…]

### 2014-11-15 Rscripts/gall_network_analysis_modularity_nestedness.R
removed contributions to network structure
Initially tried plotting the contribution of different interactions and
genotypes to network structure in terms of the frequency of
interactions. I think I have identified a better way to do this which
accounts for the roles different nodes play in the network.

@@ -16,52 +16,6 @@ rownames(genotype_gall_parasitoid_network) <- genotype_gall_parasitoid_network$X
 genotype_gall_parasitoid_network <- select(genotype_gall_parasitoid_network, -X)
 
 
-###### Plots of Contributions to network structure
-
-### Genotypes
-geno.contrib.df <- data.frame(contrib = rowSums(genotype_gall_parasitoid_network),
-                              percent.contrib = rowSums(genotype_gall_parasitoid_network)/sum(rowSums(genotype_gall_parasitoid_network))*100,
-                              genotypes = rownames(genotype_gall_parasitoid_network))
-geno.contrib.df <- mutate(geno.contrib.df, geno.ord = reorder(genotypes, percent.contrib))
-mean.percent.contrib <- with(geno.contrib.df, mean(percent.contrib))
-
-hist(log(geno.contrib.df$contrib)) # log-normal distribution?
-
-ggplot(geno.contrib.df, aes(x = geno.ord, y = percent.contrib)) +
-  geom_bar(stat = "identity") +
-  xlab("Genotype") +
-  ylab("Percent Contribution to Network Structure") +
-  coord_flip() +
-  theme_classic() +
-  geom_hline(yintercept = mean.percent.contrib, linetype = "dashed")
-
-# Gall-parasitoid interactions
-gall.ptoid.contrib.df <- data.frame(contrib = colSums(genotype_gall_parasitoid_network),
-                              percent.contrib = colSums(genotype_gall_parasitoid_network)/sum(colSums(genotype_gall_parasitoid_network))*100,
-                              interactions = colnames(genotype_gall_parasitoid_network))
-gall.ptoid.contrib.df <- mutate(gall.ptoid.contrib.df, interaction.ord = reorder(interactions, percent.contrib))
-mean.gp.percent.contrib <- with(gall.ptoid.contrib.df, mean(percent.contrib))
-
-hist(log(gall.ptoid.contrib.df$contrib)) # log-normal distribution?
-
-ggplot(gall.ptoid.contrib.df, aes(x = interaction.ord, y = percent.contrib)) +
-  geom_bar(stat = "identity") +
-  xlab("Interactions") +
-  ylab("Percent Contribution to Network Structure") +
-  coord_flip() +
-  theme_classic() +
-  geom_hline(yintercept = mean.gp.percent.contrib, linetype = "dashed")
-
-# betadiversity of genotype-gall-parasitoid network
[…]

### 2015-01-28 ms_willow_network.Rmd
made a ton of changes.
I made this commit so I could begin tracking my changes for the network
manuscript in R markdown.

---
title: "Plant genetic variation drives beta-diversity of herbivore-parasitoid interaction networks"
author: "Matthew A. Barbour"
date: "January 19, 2015"
output: word_document
---

### Abstract
Interactions between individual organisms give rise to the arrangement and strength of mutualistic and antagonistic interactions among species within a community. While we know this ecological network structure can have a profound influence on community dynamics, we still have a limited understanding of how different genotypes and phenotypes of individual organisms contribute to overall network structure. We address this knowledge gap by using a common garden experiment to examine how genetic and phenotypic variation within a dominant host-plant species influences an insect herbivore-parasitoid interaction network. We found that the plant genotype-herbivore-parasitoid network exhibited a modular structure, due to distinct genotype-herbivore-parasitoid associations. In particular, we found that genetic variation in plant traits influenced both the density and size of herbivores, which in turn, affected attack rates from parasitoids. Taken together, our results indicate that genetic variation can play a key role in structuring ecological networks.. Furthermore, our results highlight the potential for microevolutionary processes to shape both the structure and dynamics of ecological networks.

### Introduction
Network theory has emerged as a powerful and general framework for quantifying the arrangement and strength of species interactions within communities and makes clear predictions for how this structure should influence community dynamics (Strogatz 2001). For example, mutualistic interaction networks (e.g., plant-pollinator) tend to have a nested structure, where specialist species interactions are a subset of generalist species interactions (Bascompte et al., 2003; Olesen et al. 2007; Thebault and Fontaine 2011). In contrast, antagonistic interaction networks (e.g., food webs) tend to exhibit a modular structure, where groups of species frequently interact with each other compared to other species (Krause et al., 2003; Rezende et al., 2009; Thebault and Fontaine 2011). These nested and modular structures, for mutualistic and antagonistic networks respectively, determine the robustness of these networks to perturbations (Thebault and Fontaine 2011; Stouffer and Bascompte, 2011; Rohr et al., 2014). However, studies quantifying the structure of ecological networks have been resolved predominantly to the species level (Ings et al., 2011); neglecting the considerable phenotypic variation that can exist within species and the fact that interactions occur between individuals -- not species -- within a community. By ignoring how individual differences scale up to affect species interactions, we risk overlooking fundamentally important aspects of ecological network structure.

Intraspecific genetic variation is a key driver of individual differences and there are a growing number of examples demonstrating its cascading effects on community assembly (Antonovics 1992; Whitham et al. 2003, 2012). For example, genetic variation within host-plant species can influence the abundance and diversity of herbivores and, in turn, alter interactions with predators and parasitoids at higher trophic levels (Bailey et al., 2006; Johnson 2008; Abdala-Roberts and Mooney 2014). Similarly, genetic and phenotypic variation within predators has been found to mediate the strength of trophic cascades (Post et al. 2008; Harmon et al., 2009; Bassar et al., 2010). While it is clear that microevolutionary processes within a single species can influence community assembly, these studies have failed to examine the ecological interactions (i.e., genotype-species and species-species) that determine the assembly process. Consequently, we lack a detailed understanding of the structure of ecological interactions that shape the genetic-basis to community assembly.

Incorporating intraspecific genetic variation into network theory has the potential to reveal the underlying structures that mediate the ecological and evolutionary dynamics of communities. For instance, if certain species interact more frequently with particular genotypes as a result of their phenotype, then we hypothesize that this genetic specificity would result in a modular network structure. This modular structure may have two-fold consequences on the community. Ecologically, a modular structure has been shown to buffer the cascading effects of extinction on other species in antagonistic networks (Stouffer and Bascompte, 2011). Therefore, genetic variation may enhance the persistence of communities by creating a modular structure that is nested within a species. Evolutionarily, if we know the phenotypes associated with modules of genotype-species interactions, we can begin to predict how trait evolution will affect the eco-evolutionary dynamics of communities (Guimeras et al., 2011; Nuismer et al., 2012). Despite these promising avenues for future research, we currently lack empirical tests of genetic variation affecting the structure of ecological networks.

Here, we test the hypothesis that intraspecific genetic variation results in a modular network structure using the insect gall-parasitoid network (5 gall species, 8 parasitoid species) associated with the coastal willow Salix hookeriana. This willow-gall-parasitoid system was ideal for testing our hypothesis for several reasons. First, S. hookeriana, displays considerable genetic and phenotypic variation that corresponds to variation in herbivory by galling insects (Barbour et al. 2014 in press). Second, the closed nature of galls restricts their parasitoid community to a small number of species that can successfully locate and parasitize galls. Third, the sedentary nature of galling insects make them amenable to building quantitative interaction networks due to the fact that galls can be collected and reared in the lab to determine survival and the extent of parasitism from different species (van Veen et al. 2006). Finally, prior studies in a different willow species have shown that genetic and phenotypic variation can influence both the density and size of a common stem-galling sawfly, which in turn, determines the likelihood of parasitoid attack (Price 1988; Craig, Itami, & Price, 1990; Craig 1994).

If genetic variation within species influences ecological network structure, we would predict the following two things. First, we predict that the willow-gall-parasitoid network would exhibit a modular structure, due to variation in the frequency of particular gall-parasitoid interactions among willow genotypes. Second, we predict that genetic variation in plant traits would be associated with variation in gall density and gall size, which in turn, would mediate the strength of gall-parasitoid interactions.  Taken together, our study seeks to understand how genetic variation shapes the structure of ecological networks. And since genetic variation provides the raw material for natural selection, establishing a genetic-basis to network structure will begin to inform the eco-evolutionary dynamics of communities.

### Methods

## Common Garden
To isolate the effects of genetic variation within Salix hookeriana (hereafter ‘willow’) on gall-parasitoid network structure, we used a common garden experiment consisting of 26 different genotypes of S. hookeriana (13 males; 13 females), located at Humboldt Bay National Wildlife Refuge (HBNWR) (40°40'53"N, 124°12'4"W) near Loleta, California, USA. Willow genotypes were collected from a single population of willows growing around Humboldt Bay. This common garden was planted in February 2009 with 25 clonal replicates (i.e., stem cuttings) of each willow genotype in a completely randomized design in two hectares of a former cattle pasture at HBNWR. Willows in our garden began flowering in February and reach their peak growth in early August. During this study, willows had reached 2-4 m in height. Further details on the genotyping and planting of the common garden are available in Barbour et al. (2014, in press).

## Beta-diversity of gall-parasitoid networks among willow genotypes
To build a quantitative gall-parasitoid network for each willow genotype, we collected all galls occurring on one haphazardly selected basal branch from about 5 randomly chosen replicates of each genotype (N = 146 trees, range = 4-9 trees per genotype). To control for differences in branch size, we estimated the number of shoots on each branch based on an allometric equation using the stem diameter of the sampled basal branch (mean ± SD shoot count = 280 ± 124; details in supplementary materials). We collected all galls in September 2012 when gall larva were in late instars of their development or had already spun cocoons within the gall. All galls were placed into 30 mL plastic transport vials (loosely capped at the end) and maintained at room temperature in the lab for four months. We then opened galls under a dissecting scope and determined whether the gall survived or was parasitized, and if parasitized, the identity of the parasitoid species. We omitted galls for which we could not reliably determine the cause of mortality from further analysis.

We used a null model to test our prediction that the willow genotype-gall-parasitoid network exhibits a modular structure. To do this, we first pooled all observed gall-parasitoid interactions for each willow genotype. This resulted in a weighted, bipartite (i.e., two sets of nodes) network, where each willow genotype and each unique gall-parasitoid interaction comprised the two sets of nodes, respectively. The weights between nodes corresponded to the frequency of each link (willow genotype-gall-parasitoid interaction). To identify different modules of willow genotype-gall-parasitoid associations, we used an algorithm (QuaBiMo) that seeks to find the optimal partition of the network into groups, or modules, of nodes that frequently interact with each other compared to other nodes (Dorman and Strauss 2014). This module-finding algorithm seeks to maximize the modularity value Q, which is calculated as,  
￼
 where m is half of the total number of observed links in the network, Aij is the weighted, bipartite network and Kij is the network of expected weights (Dormann and Strauss 2013). The module to which a node i or j is assigned is ci,cj. The indicator function δ(ci,cj) = 1 if ci = cj and 0 if ci ≠ cj. Q ranges from 0, which means the community has no more links within modules than expected by random chance, to a maximum value of 1, which indicates increasing support for the division of a network into modules. Since Q is determined by an optimality function and is susceptible to being trapped at local optima, we repeated this calculation 100 times and used the iteration with the highest Q value to identify the modules within our network. Importantly, the magnitude of Q will be influenced by the number of nodes, the number of links between nodes, and the total number of links observed in the network (Dormann & Strauss, 2014; Thébault, 2012). Therefore, we used a conservative null model to examine whether the degree of modularity (Q) we observed was significantly different from what we would expect by chance. Specifically, we used a swapping algorithm that randomized the observed weights in the bipartite network while preserving both the total weights for each genotype and each gall-parasitoid interaction, as well as the number of unique links from the original network. We generated 1000 null bipartite networks and calculated Q for each one. We then calculated the Z-score, which is equal to the difference between the observed modularity and the mean of the null Q’s divided by the standard deviation of the null Q’s. Assuming that Q is normally distributed, a Z-score > 1.96 corresponds to a significantly modular network structure (P < 0.05; Whitlock and Schluter, 2009). Both the modularity and null model analysis were conducted with the bipartite package in R (Dormann and Strauss 2014; R Development Core Team 2014).

## Genetic variation in plant traits influences gall density and size
To identify the plant traits mediating the resistance of willow genotypes to herbivory from galling insects, we measured 40 different traits associated with variation in leaf quality (36 traits) and plant architecture (4 traits). Details on how these willow traits were sampled and quantified are given in Barbour et al. (2014, in press), but we summarize which traits were sampled here. Leaf quality traits included: phenolic chemistry (7 classes of compounds, 31 individual metabolites), trichome density, specific leaf area (SLA), water content, and percent Carbon and Nitrogen (converted to C:N). Plant architecture traits included: plant size, fractal dimension (index of architectural complexity), height, and foliage density. We calculated the broad-sense heritability (H2) of plant traits using the equation: H2 = VG / VP, where VG is the total genotypic variance among clones, and VP is the total phenotypic variance, calculated as the sum of the residual and genetic variance (Lynch & Walsh 1998). Broad-sense heritability values range between 0-1, where values close to zero indicate low heritability (i.e., the trait is strongly influenced by the environment), and values close to 1 indicate high heritability (i.e., the trait is strongly controlled by underlying genetic variation). Each of these traits exhibited significant broad-sense heritable variation among willow genotypes (mean leaf quality H2 = 0.72; mean architecture H2 = 0.27; range of H2 = 0.15 - 0.97; Barbour et al., 2014 in press).

We calculated both gall density and gall size size (measured as the maximum diameter perpendicular to direction of plant tissue growth, to the nearest 0.01 mm) of each species from our gall collections on each replicate willow tree (N = 146). We then conducted two types of analyses to test whether heritable variation in plant traits affects the density and size of galling insects. First, we calculated separate restricted maximum likelihood (REML) random-effect models and performed restricted likelihood-ratio tests to examine whether willow genotypes exhibited variation in the density and size of each gall species. Gall density and gall size were transformed as needed to improve normality and reduce heteroscedasticity of model residuals. As with each plant trait, we calculated the broad-sense heritability of gall density and gall size for each species. If we identified heritable variation in gall density or gall size among willow genotypes, we then used multiple regression with forward model selection to identify the key plant traits mediating resistance to galling insects. We mitigated multicollinearity using three different methods: principal components analysis, residual and sequential regression (Graham 2003), and omitting highly correlated traits (details in Barbour et al. 2014, in press). We used the forward model selection approach advocated by Blanchet et al. (2008), which prevents inclusion of spurious variables (i.e., inflated Type 1 error) and overestimation of explained variance (i.e., R2).

## Variation in gall density and size affects the strength of gall-parasitoid interactions
To test this prediction, we analyzed generalized linear and additive mixed-effect models (GLMM and GAMM, respectively). We modeled gall density, gall size and their interaction as fixed-effects, and included each replicate willow tree as a random effect. We modeled willow tree as a random effect because gall density was measured at the scale of the entire willow tree, but gall size was measured for individual galls collected from each willow tree. Since parasitoids often exhibit a non-linear response to variation in gall size (e.g., Hezewijk and Roland, 2003), we first analyzed GAMMs and only switched to simpler GLMMs if the difference in aikaike’s information criteria (AIC) values between the models was less than two (i.e., we chose the most parsimonious model). For GAMMs, we used the chi-squared statistic to test for non-linear effects of gall density, gall size, and their interaction. For GLMMs, we used parametric bootstrapping to test the effect of gall density, gall size, and their interaction. We always started with the most complex model and removed non-significant predictors (P > 0.05) until we identified the most parsimonious model. All GLMMs and GAMMs were analyzed in R using the lme4, gamm4, and pbkrtest packages (R core team 2013).

### Results

## Beta-diversity of gall-parasitoid networks among willow genotypes
In concordance with our prediction, we found that willow genotypes were associated with particular gall-parasitoid interactions, resulting in a modular food web structure (Q = 0.33, Z = 2.41, P = 0.008; Fig. 1). In particular, we detected five distinct modules. Three of these modules (Fig. 1; blue, purple, and red) were primarily determined by variation in the frequency of parasitism from three different parasitoid species on the most abundant gall former Iteomyia salicisverruca (63% of total observed gall-parasitoid interactions). Specifically, the frequency of parasitism from the egg, endoparasitoid Platygaster on Iteomyia varied 34.9-fold among willow genotypes (H2 = 0.31, RLRT = 21.61, P < 0.001), while parasitism from the larval, ectoparasitoids Mesopolobus (H2 = 0.11, RLRT = 3.77, P = 0.024) and Torymus (H2 = 0.25, RLRT = 14.75, P < 0.001) varied 10.5- and 5.7-fold among willow genotypes, respectively. Another gall former, R. salicisbrassicoides experienced a similar shift in its source of parasitism, although the strength of these interactions were weaker than the interactions with Iteomyia. In contrast, the three other gall species in our study system each participated in a single, but not necessarily distinct, compartment. The leaf galling sawfly Pontania had a distinct parasitoid community from the four gall midges. For the gall formers R. salicisbattatus and Cecidomyiid sp. A, we detected only a single associated parasitoid species each (Platygaster sp. and Torymus sp., respectively), but this may simply be a reflection of their relatively low abundances of these galls (5% and 4% of total galls, respectively).

##Genetic variation in plant traits influences gall density and size
We found that willows displayed heritable variation in resistance to galling insects in terms of both the density and size of galls. Specifically, the density for four of the five gall species varied between 22.8- and 70.2-fold among willow genotypes (Fig. 2A; range of H2 = 0.12 - 0.36), but was the most pronounced for the most common gall former, Iteomyia. Iteomyia was also the only gall species that varied in size (2.3-fold) among willow genotypes (Fig. 2B; H2 = 0.13, RLRT = 3.68, P = 0.022).

We found that variation in both the density and size of galls was explained by both leaf quality and plant architecture traits. For example, the density of Iteomyia galls was higher on shorter willows with higher leaf C:N (R2 = 0.17, F2,119 = 12.14, P < 0.001); however, the size of Iteomyia galls was larger on willows with higher concentrations of salicylates and flavones in their leaves (R2 = 0.14, F2,75 = 5.88, P = 0.004). As with Iteomyia, the density of R. salicisbrassicoides galls was higher on shorter willows and higher leaf C:N (R2 = 0.15, F2,120 = 10.97, P < 0.001). The density of Pontania galls was higher on smaller willows with low leaf trichome density, but higher concentrations of flavones (R2 = 0.17, F3,106 = 7.38, P < 0.001). The density of Cecidomyiid sp. A galls was higher on willows with higher concentrations of flavanones and flavanonols (R2 = 0.10, F1,131 = 15.21, P < 0.001).

##Variation in gall density and size affects the strength of gall-parasitoid interactions
We found that the dominant interactions in the gall-parasitoid network were shaped by an interaction between gall size and gall density. Specifically, the probability of Platygaster parasitizing Iteomyia decreased 18-fold over the range of gall sizes (Fig. 3A). For small-to-intermediate size galls though, the probability of Platygaster parasitizing Iteomyia increased 4.5-fold over the range of gall densities (Fig. 3C; 𝛘2 = 20.61, P < 0.001). In contrast to Platygaster, the probability of Mesopolobus parasitizing Iteomyia was highest for intermediate sized galls (Fig. 3B) and decreased 2.8-fold over the range of gall densities (Fig. 3D; 𝛘2 = 18.10, P = 0.003). The probability of Torymus parasitizing Iteomyia was not influenced by gall size, but similar to Mesopolobus, parasitism decreased 19-fold over the range of gall densities (Fig. 3E; 𝛘2 = 10.82, P = 0.001).

## Network Dissimilarity

## Linking heritable plant traits to herbivore community/traits to network dissimilarity
Genotype and phenotype level. Need to consider removing a couple of the outlying trees...(at least for vLG and rG)

Mantel tests and path analysis?
First, do mantel tests for each one. Genotype and phenotype level.
Download package for looking at dissimilarity in size-structure of a community.

May be able to perform mantel tests with subset of plant traits that actually influence herbivores. And same thing with herbivores influencing network dissimilarity...

Consider including individual species responses in a supplement. I think these results are important for teasing apart the density-independent, -dependent, and size-dependent parasitoid responses.

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2015-01-28 Rscripts/betalink_tree_net.R

@@ -1,0 +1,500 @@
+# NOTE THAT I HAVE NOT PERFORMED THESE ANALYSES AFTER REMOVING PP 373 (GENOTYPE S: OUTLYING DATAPOINT FOR VLG AND RG)
+
+## upload source code for managing network data at the tree-level
+source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
+
+## upload source code for analyzing beta-diversity of interaction networks. Note that this code is necessary for partitioning the components of network beta-diversity, but unnecessary for calculating the total dissimilarity of interaction networks (species turnover + interaction turnover)
+source('~/Documents/betalink/R/betalink.R')
+source('~/Documents/betalink/R/vec2data.frame.R')
+source('~/Documents/betalink/R/measures.r')
+source('~/Documents/betalink/R/betalink.b.r')
+source('~/Documents/betalink/R/betalink.dist.r')
+source('~/Documents/betalink/R/betalink.q.R')
+
+## upload required libraries for analysis
+library(vegan)
+
+## main dataset is 'tree_level_interaxn_all_plants'
+
+#### Analysis of dissimilarity of gall-parasitoid interaction networks among genotypes
+abund_interaxns_gallsurv_noPont <- rowSums(tree_level_interaxn_all_plants_traits_size[ ,interaxns_gallsurv_noPont] 
+
+abund_interaxns_noPont <- rowSums(tree_level_interaxn_all_plants_traits_size[ ,interaxns_noPont])
+
+table(filter(tree_level_interaxn_all_plants_traits_size, abund_interaxns_noPont > 0)$Genotype)
+## when 'net_df <- filter(tree_level_interaxn_all_plants_traits_size, abund_interaxns_noPont > 0'. The replicate samples of many of the Genotypes is reduced to a very low level. To see whether removing these genotypes affects the outcome of the analysis, I have created the following subsets. 
+geno_2plus <- c("*","A","B","D","E","F","I","K","L","Q","S","T","V","W","X","Y","Z") # 17
+geno_3plus <- c("*","B","D","I","K","L","Q","S","V","W","X","Y","Z") # 13
+geno_4plus <- c("*","B","I","K","L","S","V","X","Z") # 9 (if Pontania galls were included, then I would have retained Genotype D as well.)
+
+## dataset to use for PERMANOVA
[…]

### 2015-01-28 Rscripts/boxplots genotypic variation.R

@@ -1,0 +1,84 @@
+## source in required data
+source('~/Documents/Genotype_Networks/Rscripts/network_management_genotype_level.R')
+
+## load required libraries for plotting
+library(ggplot2)
+library(grid)
+library(gridExtra)
+
+#### Generate boxplots
+
+## create data frame for plotting
+plot.df <- genotype_level_interaxn_traits_size %>%
+  melt(id.vars = "Genotype", measure.vars = c("vLG_abund", "vLG.height.mean_mean.na.rm", "rG_abund", "rG.height.mean_mean.na.rm", "vLG_egg", "vLG_ecto", "rG_ecto")) %>%
+  mutate(species = factor(c(rep("Iteomyia", 2*26), rep("Rabdophaga (bud)", 2*26), 
+                     rep("Iteomyia - Egg parasitoids", 26), rep("Iteomyia - Larval parasitoids", 26),
+                     rep("Rabdophaga (bud) - Larval parasitoids", 26))),
+         variable.replace = c(rep("mean_gall.density", 26), rep("mean_height", 26), 
+                              rep("mean_gall.density", 26), rep("mean_height", 26),
+                              rep("mean_interaxn.density", 26*3)))
+levels(plot.df$species) <- gsub("- ", "-\n", levels(plot.df$species))
+#levels(plot.df$species) <- rev(levels(plot.df$species))
+plot.df$species <- factor(plot.df$species, levels = rev(levels(plot.df$species)))
+
+## create theme for boxplots
+theme_boxplot <- theme_classic() + theme(axis.title = element_text(size = 14, vjust = -0.5),
+                                         axis.text = element_text(size = 12))
+
+## boxplots
+gall.density.plot <- ggplot(data = filter(plot.df, variable.replace == "mean_gall.density"), aes(x = species, y = value)) +
+  geom_boxplot() + ylab("Gall density (count per 100 shoots)") + xlab("") + theme_boxplot + coord_flip()
+
+gall.size.plot <- ggplot(data = filter(plot.df, variable.replace == "mean_height"), aes(x = species, y = value)) +
+  geom_boxplot() + ylab("Gall diameter (mm)") + xlab("") + theme_boxplot + coord_flip()
+
+interaction.density.plot <- ggplot(data = filter(plot.df, variable.replace == "mean_interaxn.density"), 
[…]

### 2015-01-28 Rscripts/czvalues_error.R

@@ -1,0 +1,28 @@
+czvalues_error <- function(web, QuaBiMo_reps = 100) {
+  require('bipartite')
+  
+  observed_web_repeats <- list()
+  for(i in 1:QuaBiMo_reps){ 
+    observed_web_repeats[[i]] <- web
+  }
+  
+  QuaBiMo_repeats <- sapply(observed_web_repeats, computeModules)
+  Q_values <- sapply(QuaBiMo_repeats, function(x) x@likelihood)
+  cz_higher_repeats <- sapply(QuaBiMo_repeats, function(x) czvalues(x, weighted = TRUE, level = "higher"))
+  cz_lower_repeats <- sapply(QuaBiMo_repeats, function(x) czvalues(x, weighted = TRUE, level = "lower"))
+  
+  c_higher_df <- ldply(cz_higher_repeats[1, ])
+  z_higher_df <- ldply(cz_higher_repeats[2, ])
+  
+  c_lower_df <- ldply(cz_lower_repeats[1, ])
+  z_lower_df <- ldply(cz_lower_repeats[2, ])
+  
+  # make a list
+  list_Qcz.values <- list(Q_values,
+                         c_higher_df, 
+                         z_higher_df,
+                         c_lower_df,
+                         z_lower_df)
+  names(list_Qcz.values) <- c("Q_values", "c_higher", "z_higher", "c_lower", "z_lower")
+  return(list_Qcz.values)
+}

### 2015-01-28 Rscripts/discriminant_function_analysis_genotype_gall_network_modules.R

@@ -12,6 +12,9 @@ module.info.genotype = module.info %>%
   select(Genotype = Vertex, Module.ID, Trophic) %>%
   filter(Trophic == "Genotype")
 
+#testing
+module.info.genotype <- cbind.data.frame(module.info.genotype, cz.values.df[1:25,c("Nodes","z.values")])
+
 gall_density_genotype_summary_df <- read.csv('~/Documents/Genotype_Networks/data/gall_density_genotype_summary_df.csv')
 gall_density_genotype_summary_df <- select(gall_density_genotype_summary_df, -X)
 
@@ -33,8 +36,14 @@ library(visreg)
 visreg(vLG_Platy_glm, scale = "response")
 
 dfa_data <- dfa_data %>%
-  select(Genotype, Module.ID, aSG_abund:rsLG_abund, vLG_volume_mean, rG_volume_mean) %>%
+  select(Genotype, Module.ID, z.values, aSG_abund:rsLG_abund, vLG_volume_mean, rG_volume_mean) %>%
   mutate(Module.ID = as.factor(Module.ID))
+
+summary(lm(sqrt(vLG_abund) ~ Module.ID, data = subset(dfa_data, Module.ID %in% c(1,3,4,5)), weights = z.values))
+summary(lm(sqrt(rsLG_abund) ~ Module.ID, data = subset(dfa_data, Module.ID %in% c(1,3,5)), weights = z.values))
+summary(lm(sqrt(vLG_volume_mean) ~ Module.ID, data = subset(dfa_data, Module.ID %in% c(1,3,5)), weights = z.values))
+plot(lm(sqrt(rG_volume_mean) ~ Module.ID, data = subset(dfa_data, Module.ID %in% c(1,3,4)), weights = z.values))
+
 
 plot(vLG_abund ~ Module.ID, dfa_data)
 plot(vLG_volume_mean ~ Module.ID, dfa_data)

### 2015-01-28 Rscripts/gall density and size models.R

@@ -1,0 +1,307 @@
+# script borrowed heavily from older network_mechanism_models' data sheet
+
+## source required script
+source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
+source('~/Documents/Genotype_Willow_Community/datasets_&_Rscripts/functions_ms_willow_community.R') # used for testing for random effects of plant genotype on gall densities
+
+## load required libraries
+library(RLRsim) # for testing random effect. Using exactRLRT function gives more accurate P-values.
+library(lme4) # for random effect models
+library(ggplot2)
+
+## function for evaluating normality of random effects assumption.
+ggQQ_ranef <- function(ranef) # argument: vector of random effects
+{
+  y <- quantile(ranef, c(0.25, 0.75))
+  x <- qnorm(c(0.25, 0.75))
+  slope <- diff(y)/diff(x)
+  int <- y[1L] - slope * x[1L]
+  p <- ggplot(data.frame(ranef = ranef), aes(sample=ranef)) +
+    stat_qq(alpha = 0.5) +
+    geom_abline(slope = slope, intercept = int, color="blue")
+  
+  return(p)
+} 
+
+## focused data frame to use for analyses in this script
+abund.df <- tree_level_interaxn_all_plants_traits_size %>%
+  select(Genotype, shootEst.no18, aSG_aSG.larv:gall_total_abund, vLG_egg, vLG_ecto, vLG.height.mean, rG.height.mean) 
+
+geno_3plus <- c("*","B","D","I","K","L","Q","S","V","W","X","Y","Z") # Genotypes with 3 plus replicates after removing trees with zero gall-parasitoid interactions
+#abund.df <- filter(abund.df, Genotype %in% geno_3plus)
+
+## vLG_egg = Iteomyia egg parasitoid interactions (predominantly Platygaster)
+abund.df.noX <- filter(abund.df, Genotype != "X")
+vLG_egg.lmer <- lmer(sqrt(vLG_egg/shootEst.no18) ~ (1 | Genotype), abund.df)
+summary(vLG_egg.lmer)
+exactRLRT(vLG_egg.lmer)
[…]

### 2015-01-28 Rscripts/gall_network_analysis_modularity_nestedness.R

@@ -15,6 +15,47 @@ genotype_gall_parasitoid_network <- read.csv('~/Documents/Genotype_Networks/data
 rownames(genotype_gall_parasitoid_network) <- genotype_gall_parasitoid_network$X 
 genotype_gall_parasitoid_network <- select(genotype_gall_parasitoid_network, -X)
 
+# node contribution
+node.Q.contribs <- node.Q.contribution(web = genotype_gall_parasitoid_network, n_null_webs = 100, Null_model = "swap.web")
+rownames(node.Q.contribs) <- colnames(genotype_gall_parasitoid_network)
+
+write.csv(node.Q.contribs, "~/Documents/Genotype_Networks/data/node.Q.contribs.csv")
+
+tree_level_interaxn_all_plants <- read.csv("~/Documents/Genotype_Networks/data/tree_level_interaxn_all_plants.csv")
+tree_level_interaxn_all_plants <- mutate(tree_level_interaxn_all_plants, plant.position = factor(plant.position))
+
+genotype_gall_web <- tree_level_interaxn_all_plants %>%
+  group_by(Genotype) %>%
+  select(vLG_abund, rG_abund, rsLG_abund, aSG_abund, SG_abund, shootEst.no18) %>%
+  summarise_each(funs(sum)) %>%
+  mutate(vLG_dens = round(vLG_abund/shootEst.no18*1000),
+         rG_dens = round(rG_abund/shootEst.no18*1000),
+         rsLG_dens = round(rsLG_abund/shootEst.no18*1000),
+         aSG_dens = round(aSG_abund/shootEst.no18*1000),
+         SG_dens = round(SG_abund/shootEst.no18*1000)) %>%
+  select(Genotype, vLG_dens:SG_dens)
+rownames(genotype_gall_web) <- genotype_gall_web[ ,1]
+
+visweb(genotype_gall_web[ ,-1]) # not nested, even after null model
+visweb(genotype_gall_web[ ,-1], type = "diagonal")
+nestednodf(comm = genotype_gall_web[ ,-1], order = T, weighted = T) # 38.48
+g.gall.web.best <- best_QuaBiMo(web = genotype_gall_web[ ,-1], QuaBiMo_reps = 10)
+plotModuleWeb(g.gall.web.best$best_module_info[[1]]) # interesting, doesn't strongly correspond to modules identified from tri-trophic interaction.
+czvalues(g.gall.web.best$best_module_info[[1]], weighted = TRUE)
+g.gall.web.best$best_observed_value # 0.199335
[…]

### 2015-01-28 Rscripts/genotype gall parasitoid network analysis.R

@@ -1,0 +1,178 @@
+####################################################################################
+# This R script analyzes the willow genotype-gall-parasitoid network
+# Coded by: Matt Barbour
+####################################################################################
+
+##### Upload source files
+source('~/Documents/vegan/R/nestednodf.R')
+source('~/Documents/Genotype_Networks/Rscripts/null_model_analysis_WNODF_QuaBiMo.R')
+source('~/Documents/Genotype_Networks/Rscripts/best_QuaBiMo_function.R')
+source("~/Documents/Genotype_Networks/Rscripts/module.contribution.R")
+source("~/Documents/Genotype_Networks/Rscripts/null_model_analysis_czvalues.R")
+source("~/Documents/Genotype_Networks/Rscripts/czvalues_error.R")
+source("~/Documents/Genotype_Networks/Rscripts/teammates.R")
+source("~/Documents/Genotype_Networks/Rscripts/teammates_table.R")
+source("~/Documents/miscellaneous_R/ggplot_themes.R")
+
+##### Upload required libraries
+library(plyr)
+library(dplyr)
+library(bipartite)
+library(ggplot2)
+library(gridExtra)
+
+##### Upload required datasets
+genotype_gall_parasitoid_network <- read.csv('~/Documents/Genotype_Networks/data/genotype_gall_parasitoid_network_data.csv')
+rownames(genotype_gall_parasitoid_network) <- genotype_gall_parasitoid_network$X 
+genotype_gall_parasitoid_network <- select(genotype_gall_parasitoid_network, -X)
+
+#### Summary statistics
+colSums(genotype_gall_parasitoid_network)/sum(colSums(genotype_gall_parasitoid_network)) # aSG_Tory, rG_Eulo, rG_Lestodip, rG_Mesopol, rG_Platy, SG_Platy, and vLG_Mymarid each make up less than 5% of the total interactions.
+
+rowSums(genotype_gall_parasitoid_network)/sum(rowSums(genotype_gall_parasitoid_network)) # Genotypes G, P, H, Q, O, N, E, M, A, and J each make up less than 2% of the total interactions in the network
+
+##### Null Model Analysis. Used 1000 reps of the 'swap_web' algorithm for both WNODF and QuaBiMo.
+
+### Nestedness: weighted NODF (Almeida-Neto & Ulrich, 2010, Environmental Modelling Software).
+# WNODF of network = 23.7
[…]

### 2015-01-28 Rscripts/metaweb plot.R

@@ -1,0 +1,73 @@
+## metaweb
+library(vegan) # for cca function
+library(igraph)
+library(ggplot2)
+library(grid)
+
+## source in required files
+source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
+source('~/Documents/ggnet/bipartite_plot_info.R')
+source('~/Documents/ggnet/tripartite_plot_info.R')
+source('~/Documents/ggnet/ggnet_bipartite.R')
+
+## create metaweb data
+metaweb <- tree_interaxns_filter_add %>%
+  filter(gall_contents_collapse %in% c("aSG.larv", "rG.larv", "SG.larv", "vLG.pupa", "Eulo", "Lestodip", "Mesopol", "Mymarid", "Platy", "Tory")) %>%
+  dcast(gall.sp ~ gall_contents_collapse, sum, value.var = "value") %>%
+  mutate(gall.surv = aSG.larv + rG.larv + SG.larv + vLG.pupa) %>%
+  select(gall.sp, Eulo:Platy, Tory, gall.surv)
+rownames(metaweb) <- metaweb$gall.sp
+
+# get plot info
+metaweb.info <- bipartite_plot_info(select(metaweb, Eulo:Tory), order.type = "cca")
+
+interaction.df <- metaweb.info[[1]] %>%
+  filter(Sequence == 500 | Sequence == 1) %>%
+  reshape(idvar = "Group", timevar = "Sequence", direction = "wide") %>%
+  mutate(weight.trans = Weight.1) # unscaled weights
+
+nodeinfo.df <- metaweb.info[[2]] %>%
+  mutate(guild = factor(c(rep("gall", 4),
+                   "predator",
+                   rep("Larval parasitoid", 3),
+                   rep("Egg parasitoid", 2))),
+         names = c("Cecidomyiid",
+                   "Rab. (bud)",
+                   "Iteomyia",
+                   "Rab. (stem)",
+                   "Lestodiplosis", "Torymus", "Eulophid", "Mesopolobus", "Platygaster", "Mymarid"))
+
+new_theme_empty <- theme_bw()
+new_theme_empty$line <- element_blank()
+new_theme_empty$rect <- element_blank()
+new_theme_empty$strip.text <- element_blank()
+new_theme_empty$axis.text <- element_blank()
+new_theme_empty$plot.title <- element_blank()
+new_theme_empty$axis.title <- element_blank()
[…]

### 2015-01-28 Rscripts/module.contribution.R

@@ -1,0 +1,64 @@
+module.contribution <- function (moduleWebObject, weighted = FALSE, level = "higher") 
+{
+  # this function was modified from the czvalues function in the bipartite package. Anything with comments represents a modification from the czvalues function. I kept all naming the same as in the czvalues function just to avoid errors. I did not modify the c-value calculation, but I changed the z-value calculation to estimate the percent contribution of a node to the total number of links in the module. I then changed the output to 'module.percent.contribution' to reflect what it is actually calculating.
+  
+  require(bipartite) # added because I don't know how to put this into the 'bipartite' environment
+  
+  #if (!isCorrectModuleWebObject(moduleWebObject)) # I omitted this because it seems that I need to be in the 'bipartite' environment to reference this function
+   # stop("This function cannot be applied to this type of object!") 
+  zvalues <- function(web, weighted = FALSE) {
+    if (weighted) {
+      depL <- web/matrix(rowSums(web), nrow = NROW(web), 
+                         ncol = NCOL(web), byrow = FALSE)
+      k.is <- colSums(depL)
+    }
+    else {
+      k.is <- colSums(web > 0)
+    }
+    out <- k.is/sum(k.is) # altered this to calculate the % contribution of each node to the number of links in the module. Note that this was only changed for the weighted version.
+    out
+  }
+  modInfo <- listModuleInformation(moduleWebObject)
+  modules <- modInfo[[2]]
+  nModules <- length(modInfo[[2]])
+  if (nModules < 2) 
+    stop("This web has no modules.")
+  web <- moduleWebObject@originalWeb
+  if (level == "lower") 
+    web <- t(web)
+  if (weighted) {
+    depL <- web/matrix(rowSums(web), nrow = NROW(web), ncol = NCOL(web), 
+                       byrow = FALSE)
+    k.i <- colSums(depL)
+  }
+  else {
+    k.i <- colSums(web > 0)
+  }
+  z.values.all <- web[1, ]
+  k.it.for.t <- matrix(NA, ncol = NCOL(web), nrow = nModules)
+  for (t in 1:nModules) {
+    h.level.species.names <- if (level == "higher") 
+      modules[[t]][[2]]
+    else modules[[t]][[1]]
+    l.level.species.names <- if (level == "higher") 
+      modules[[t]][[1]]
+    else modules[[t]][[2]]
+    if (weighted) {
+      depL.mod1 <- web[l.level.species.names, , drop = FALSE]/matrix(rowSums(web[l.level.species.names, 
[…]

### 2015-01-28 Rscripts/network denddrograms.R

@@ -1,0 +1,28 @@
+source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
+source('~/Documents/Genotype_Networks/Rscripts/network_management_genotype_level.R')
+
+library(ggplot2)
+library(ggdendro)
+
+geno.level.net <- genotype_level_interaxn_traits_size[ ,interaxns_noPont]
+
+geno.members <- table(tree_level_interaxn_all_plants_traits_size$Genotype)
+
+plot(hclust(d = dist(geno.level.net)^2, method = "cen", members = geno.members), 
+     labels = genotype_level_interaxn_traits_size$Genotype, hang = -1)
+
+plot(hclust(d = dist(geno.level.net), method = "ave", members = geno.members), 
+     labels = genotype_level_interaxn_traits_size$Genotype, hang = -1)
+
+t <- hclust(d = dist(geno.level.net), method = "ward", members = geno.members)#, 
+     #labels = genotype_level_interaxn_traits_size$Genotype, hang = -1, leaflab = "perpendicular")
+t <- as.dendrogram(t)
+
+hc <- hclust(d = dist(geno.level.net), method = "ward", members = geno.members)
+dhc <- as.dendrogram(hc)
+ddata <- dendro_data(dhc)
+
+ggplot(data = segment(ddata)) +
+  geom_segment(aes(x = x, y = y, xend = xend, y = yend)) #+ 
+  #coord_flip() + 
+  #scale_y_reverse(expand = c(0.2, 0))

### 2015-01-28 Rscripts/network mixed effect models.R

@@ -1,0 +1,73 @@
+## network mixed-effect models
+
+## source in required data
+source('~/Documents/Genotype_Networks/Rscripts/network_management_gall_level.R')
+
+## source in required libraries
+library(lme4)
+library(visreg)
+library(gamm4)
+
+gall_mech.df
+
+## vLG mechanisms df
+vLG_mech.df <- filter(gall_mech.df, vLG_total > 0) %>%
+  mutate(sc.gall.height = scale(gall.height, center = TRUE, scale = TRUE),
+         sc.vLG_density = scale(vLG_density, center = TRUE, scale = TRUE),
+         vLG_egg = vLG_Platy + vLG_Mymarid,
+         vLG_ecto = vLG_Eulo + vLG_Mesopol + vLG_Tory)
+
+
+## vLG_egg
+vLG_egg.glmer <- glmer(cbind(vLG_egg, vLG_total - vLG_egg) ~ sc.vLG_density*sc.gall.height + 
+                         (1 | Genotype.x/plant.position), data = vLG_mech.df,
+                       family = binomial(link = "logit"))
+summary(vLG_egg.glmer)
+AIC(vLG_egg.glmer)
+
+vLG_egg.glmer.vis <- glmer(cbind(vLG_egg, vLG_total - vLG_egg) ~ vLG_density*gall.height + 
+                         (1 | plant.position), data = vLG_mech.df,
+                       family = binomial(link = "logit"))
+summary(vLG_egg.glmer.vis) # same overall picture as standardized model that includes Genotype as a random effect.
+AIC(vLG_egg.glmer.vis)
+
+## vLG_ecto
+vLG_ecto.glmer <- glmer(cbind(vLG_ecto, vLG_total - vLG_ecto) ~ sc.vLG_density + sc.gall.height +
+                         (1 | Genotype.x/plant.position), data = vLG_mech.df, 
+                        family = binomial(link = "logit"))
+summary(vLG_ecto.glmer)
+AIC(vLG_ecto.glmer)
+
+vLG_ecto.glmer.vis <- glmer(cbind(vLG_ecto, vLG_total - vLG_ecto) ~ vLG_density + gall.height + 
+                             (1 | plant.position), data = vLG_mech.df,
+                           family = binomial(link = "logit"))
+summary(vLG_ecto.glmer.vis) # same overall picture as standardized model that includes Genotype as a random effect.
+AIC(vLG_ecto.glmer.vis)
+
[…]

### 2015-01-28 Rscripts/network_management_gall_level.R

@@ -1,0 +1,42 @@
+### prepare data that is resolved to the individual gall level
+source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
+
+gall.size.df <- tree_interaxns_filter_add %>%
+  group_by(Genotype, plant.position, gall.id.nest, gall.sp) %>% # this makes sure I don't double count galls with multiple larvae, and is also why I take the "average" below, which is just the same number.
+  summarise(gall.height = mean(g.height)) %>%
+  ungroup()
+
+gall.contents.df <- tree_interaxns_filter_add %>%
+  dcast(gall.id.nest ~ gall.sp + gall_contents_collapse, sum) %>%
+  tbl_df()
+
+## merge data sets together and make some new columns of data
+gall.size.contents.df <- left_join(gall.size.df, gall.contents.df) 
+gall.size.contents.df <- mutate(gall.size.contents.df, plant.position = factor(plant.position))
+gall.size.contents.df$vLG_total <- rowSums(select(gall.size.contents.df, starts_with("vLG")))
+gall.size.contents.df$rG_total <- rowSums(select(gall.size.contents.df, starts_with("rG")))
+gall.size.contents.df$rsLG_total <- rowSums(select(gall.size.contents.df, starts_with("rsLG")))
+gall.size.contents.df$aSG_total <- rowSums(select(gall.size.contents.df, starts_with("aSG")))
+gall.size.contents.df$SG_total <- rowSums(select(gall.size.contents.df, starts_with("SG")))
+
+
+## create gall density data and plant architecture and trichome density. These were the only plant traits that I expected a priori to affects parasitoids. I would expect the other mechanisms to act indirectly by changing gall densities.
+gall_density.df <- with(tree_level_interaxn_all_plants_traits_size, 
+                        data.frame(plant.position = plant.position, 
+                                   Genotype = Genotype, 
+                                   Gender = Gender, 
+                                   vLG_density = vLG_abund/shootEst.no18*100,
+                                   rG_density = rG_abund/shootEst.no18*100,
+                                   aSG_density = aSG_abund/shootEst.no18*100,
+                                   rsLG_density = rsLG_abund/shootEst.no18*100,
+                                   SG_density = SG_abund/shootEst.no18*100,
+                                   Total_Area = Total_Area,
+                                   Density = Density,
+                                   Height = Height,
+                                   D = D_mean_smoothed,
[…]

### 2015-01-28 Rscripts/network_management_genotype_level.R

@@ -1,4 +1,39 @@
-library(dplyr)
+## This script manages network data at genotype level
+
+## source in required scripts and data
+source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
+source('~/Documents/miscellaneous_R/length2_function.R')
+
+## create genotype-level interaction data. I accounted for density by dividing by the number of estimated shoots, because that was the most straightforward way to account for differences in sampling among trees. RIGHT NOW I HAVEN'T CALCULATE RICHNESS RESPONSES AND PARASITISM YET. OR 2011 GALL DATA. I SHOULD NOT CONTROL FOR RICHNESS OR PARASITISM IN TERMS OF SHOOT COUNTS, BUT I SHOULD DIVIDE 2011 GALL DATA BY TOTAL AREA.
+tree_level_interaxn_density <- select(tree_level_interaxn_all_plants_traits_size, aSG_aSG.larv:link_abund, vLG_ecto:rG_ecto, vLG_egg)/tree_level_interaxn_all_plants$shootEst.no18*100 # scale to a per 100 shoot basis. 
+
+geno_level_interaxn_density <- aggregate(tree_level_interaxn_density, 
+                                         by = list(Genotype = tree_level_interaxn_all_plants_traits_size$Genotype), 
+                                         FUN = mean) %>%
+  tbl_df()
+
+## create genotype-level trait data.
+geno_level_traits <- tree_level_traits %>%
+  group_by(Genotype) %>%
+  select(Total_Area:flavanonOLES.PC1) %>%
+  summarise_each(funs(mean.na.rm = mean(., na.rm = TRUE))) 
+
+## create genotype-level gall size data
+geno_level_gall.size.count.join <- select(tree_level_gall.size.count.join, 
+                                          Genotype:vLG.height.mean) %>% # counts per tree not needed
+  group_by(Genotype) %>% 
+  select(-plant.position) %>%
+  summarise_each(funs(mean.na.rm = mean(., na.rm = TRUE), 
+                      length2.na.rm = length2(., na.rm = TRUE))) # length2 enables me to calculate the number of plant replicates that were used to estimate mean gall size.
+
+## merge interaction and genotype-trait data
+genotype_level_interaxn_traits_size <- join_all(list(geno_level_interaxn_density, geno_level_traits, geno_level_gall.size.count.join), by = "Genotype") %>%
+  tbl_df()
+
[…]

### 2015-01-28 Rscripts/network_management_tree_level.R

@@ -1,52 +1,55 @@
-######## Tree level Network, Community, and Species level analysis
+## This R script organizes the data for gall-parasitoid networks on individual trees of different willow genotypes.
+## Author: Matt Barbour
+## Email: barbour@zoology.ubc.ca
 
-#### upload necessary packages
-
-# analysis
-library(mvabund)
-library(vegan)
-library(visreg)
-library(bipartite)
-
-# data manipulation. order of libraries is important. always load dplyr last.
+## upload packages for data manipulation. order of libraries is important. always load dplyr last.
 library(reshape2)
 library(reshape)
+library(plyr)
 library(dplyr)
 
-# plotting
-library(ggplot2)
-
-
-# upload molten gall network data
+## upload molten gall network data and calculate some basic summary data
 gall_net_melt <- read.csv("~/Documents/Genotype_Networks/data/gall_network_data.csv")
-gall_net_melt <- tbl_df(gall_net_melt) # 1,511 galls
-sum(gall_net_melt$value[1:1511]) # 1740 total samples
-sum(filter(gall_net_melt, gall_contents == "nothing")$value) # 530 "nothing" samples
-sum(filter(gall_net_melt, gall_contents == "exit.hole")$value) # 136
+gall_net_melt <- tbl_df(gall_net_melt) # 1,495 galls
+sum(gall_net_melt$value, na.rm = TRUE) # 1709 total samples
+sum(filter(gall_net_melt, gall_contents == "nothing")$value) # 518 "nothing" samples
+sum(filter(gall_net_melt, gall_contents == "exit.hole")$value) # 136 exit holes. These indicate ectoparasitism, but I was unable to reliably determine attack by the particular parasitoid.
 
-##### Does the number of shoots sampled vary among genotypes?
-
-# data frame
+## create a data frame for the estimated number of shoots surveyed on each willow.
 shootEsts_df <- gall_net_melt %>%
   mutate(plant.position = as.factor(plant.position)) %>%
   group_by(Gender, Genotype, plant.position) %>%
   summarise(shootEst.no18 = mean(shootEst.no18), shootEst.all = mean(shootEst.all), galls.found = mean(Galls.found)) # needed to take the mean because these shoot estimates are duplicated throughout the molten dataframe
 shootEsts_df <- mutate(shootEsts_df, plant.position = as.character(plant.position))
 
-shootEsts_genotype_df <- shootEsts_df %>%
-  group_by(Genotype) %>%
[…]

### 2015-01-28 Rscripts/network_mechanism_models.R

@@ -101,6 +101,15 @@ genotype_nested_model <- function(data, y){
 tree_level_interaxn_all_plants <- read.csv("~/Documents/Genotype_Networks/data/tree_level_interaxn_all_plants.csv")
 tree_level_interaxn_all_plants <- mutate(tree_level_interaxn_all_plants, plant.position = factor(plant.position))
 
+# test analysis. don't know how promising the tree-level approach is going to be
+sub.tree <- select(tree_level_interaxn_all_plants, plant.position, vLG_Tory, vLG_Platy, vLG_Mesopol)#aSG_aSG.larv:vLG_vLG.pupa, -aSG_aSG.larv, -rsLG_Pont.surv, -SG_SG.larv, - vLG_vLG.pupa, -rG_rG.larv)
+row.names(sub.tree) <- sub.tree$plant.position
+visweb(sub.tree[ ,-1], "diagonal")
+library(bipartite)
+sub.tree.mod <- computeModules(sub.tree[ ,-1])
+plotModuleWeb(sub.tree.mod)
+printoutModuleInformation(sub.tree.mod)
+
 # plant trait data
 tree_level_traits <- read.csv("~/Documents/Genotype_Networks/data/plant.trait.galls.2011.tree.df.csv")
 tree_level_traits <- select(tree_level_traits, -X, -Genotype, plant.position, vLG.2011 = vLG, rsLG.2011 = rsLG, Total_Area:flavanonOLES.PC1)
@@ -111,6 +120,9 @@ geno_level_traits <- select(geno_level_traits, -X)
 
 # merge interaction and tree-trait data
 tree_level_interaxn_all_plants <- left_join(tree_level_interaxn_all_plants, tree_level_traits, by = "plant.position")
+
+module.info <- read.csv('~/Documents/Genotype_Networks/data/module_info_genotype_gall_parasitoid_network.csv')
+
 
 ### prepare data that is resolved to the individual gall level
 

### 2015-01-28 Rscripts/node.Q.contribution.R

@@ -1,0 +1,25 @@
+
+# need to calculate z-score of removing each interaction. It is not sufficient, to just remove a node and recalculate modularity to see how it changes, since it is sensitive to the number of nodes, number of links in the web, and the total weights in the web.
+
+node.Q.contribution <- function(web, n_null_webs = 1, Null_model = "swap.web"){
+  require('bipartite')
+  require('plyr')
+  require('dplyr')
+  
+  upper.node.exclusion.webs <- list()
+  for(i in 1:dim(web)[2]){
+    upper.node.exclusion.webs[[i]] <- web[ ,-i]
+  }
+  
+  upper.node.excl.modules <- sapply(upper.node.exclusion.webs, computeModules)
+  upper.node.Q.values <- sapply(upper.node.excl.modules, function(x) x@likelihood)
+  
+  upper.node.mod.contribs <- list()
+  for(i in 1:dim(web)[2]){
+    upper.node.mod.contribs[[i]] <- null_model_analysis_WNODF_QuaBiMo(web = upper.node.exclusion.webs[[i]], type = "QuaBiMo", observed_value = upper.node.Q.values[i], N_null_webs = n_null_webs, null_model = Null_model)
+  }
+  
+  upper.node.mod.contribs.df <- ldply(upper.node.mod.contribs)
+  return(upper.node.mod.contribs.df)
+}
+

### 2015-01-28 Rscripts/null_model_analysis_czvalues.R

@@ -1,42 +1,68 @@
-null_model_analysis_czvalues <- function(web, N_null_webs = 10, null_model = "swap.web", cz_level = "higher") {
+null_model_analysis_czvalues <- function(web, N_null_webs = 10, null_model = "swap.web", ColNames = colnames(web), RowNames = rownames(web)) {
   require('bipartite')
+  require('plyr')
   
   # generate null models
   null_web_list <- nullmodel(web, N = N_null_webs, method = null_model)
   
-  # add row and column names to each null model
-  t <- sapply(null_web_list, provideDimnames)
-  
-  ### ON THE RIGHT TRACK. NEED TO ADD DIMNAMES TO ALL NULL MODELS, THEN I SHOULD BE ABLE TO CALCULATE THE CZ VALUES.
-  browser()
-  # need to extract cz values
-  #nulls <- sapply(null_web_list, computeModules)
-
-  null_cz.values <- list()
-  for(i in 1:length(null_web_list)){
-    tmp <- computeModules(null_web_list[[i]])
-    browser()
-    nulls[[i]] <- czvalues(tmp, weighted = TRUE, level = cz_level)
+  # add row and column names to each null model. Necessary to run czvalues function on moduleWebObject
+  for(i in 1:N_null_webs){
+    colnames(null_web_list[[i]]) <- ColNames
+    rownames(null_web_list[[i]]) <- RowNames
   }
-  #browser()
-
   
-  #null_cz.values <- sapply(nulls, FUN = function(x) czvalues(x, weighted = TRUE, level = cz_level))
+  # compute modules for each null model
+  nulls <- sapply(null_web_list, computeModules)
+
+  # compute "higher" trophic level cz values for each null model
+  high.nulls_cz.values <- sapply(nulls, function(x) module.contribution(x, weighted = TRUE, level = "higher"))
+  
   
  
-  
-  null_values <- vector()
-  for(i in 1:dim(nulls)[2]){
-    null_values[i] <- nulls[ ,i]$statistic[3] 
+  # extract c values
+  high.null_c.values <- list()
+  for(i in 1:N_null_webs){
+    high.null_c.values[[i]] <- high.nulls_cz.values[ ,i]$c 
   }
   
-  mean_null_values <- mean(null_values)
-  sd_null_values <- sd(null_values)
+  # extract z values
+  high.null_z.values <- list()
+  for(i in 1:N_null_webs){
[…]

### 2015-01-28 Rscripts/survey_2_network_management.R

@@ -234,7 +234,18 @@ gall_net_melt_with_no_gall_data <- rbind(gall_net_melt, no_gall_df)
 gall_net_melt_plant_info <- left_join(gall_net_melt_with_no_gall_data, plant.info_df) %>%
   select(Gender:shootEst.no18, plant.position:value)
 
-#### Data has been checked and appears to be error free.
+### After performing preliminary analyses. I have decided to remove plant position 373 (Genotype S) from further analysis. It had very large outliers for both vLG and rG density. rG density at the Genotype level was especially skewed by this outlier. This plant position also had the lowest number of estimated shoots sampled (between 23 & 12), whereas the next lowest at was between 66 & 50. Plant position 390 had a very large number of shoots sampled (between 993 & 1189), but did not appear to be an outlier for gall densities, therefore I kept it.
+gall_net_melt_plant_info <- filter(gall_net_melt_plant_info, plant.position != "373")
 
+## although plant position 77 was a "mystery" (see notes above) I'm assigning it positive values for "Galls.found" to avoid confusion in downstream analyses, because I have decided to retain this data.
+pp77s <- which(gall_net_melt_plant_info$plant.position == "77")
+gall_net_melt_plant_info$Galls.found[pp77s] <- c(rep(1, length(pp77s))) 
+
+## reassign this tree to have zero galls found. We did have a rsLG-moth.ad3 associated with it, but I already decided above to remove this from the data set, so it should be involved in the density estimates.
+pp521 <- with(gall_net_melt_plant_info, which(Galls.found == "1" & plant.position == "521")) 
+gall_net_melt_plant_info$Galls.found[pp521] <- 0
+
+
+#### Data has been checked and appears to be error free.
 write.csv(gall_net_melt_plant_info,"~/Documents/Genotype_Networks/data/gall_network_data.csv")
 

### 2015-01-28 Rscripts/teammates.R

@@ -1,0 +1,14 @@
+#### Andrew MacDonald wrote this code for me.
+
+teammates <- function(modelrun, focal = "vLG_Platy", node.location = "column"){
+  node.set <- ifelse(node.location == "column", 2, 1)
+  
+  interact <- sapply(modelrun[[2]], function(x) x[[node.set]])
+  
+  has_focal <- sapply(interact, function(x) focal %in% x)
+  
+  in_mod <- interact[[which(has_focal)]]
+  
+  in_mod[!(in_mod %in% focal)]
+}
+

### 2015-01-28 Rscripts/teammates_table.R

@@ -1,0 +1,21 @@
+source('~/Documents/Genotype_Networks/Rscripts/teammates.R')
+
+teammates_table <- function(web, focal = "vLG_Platy", node.location = "column", QuaBiMo_reps = 100) { # focal
+  require('bipartite')
+  
+  # run modularity algorithm and create a list of the module information
+  observed_web_repeats <- list()
+  for(i in 1:QuaBiMo_reps){ 
+    observed_web_repeats[[i]] <- web
+  }
+  
+  QuaBiMo_repeats <- sapply(observed_web_repeats, computeModules)  
+  modlist <- lapply(QuaBiMo_repeats, listModuleInformation)
+  
+  run_teammates <- lapply(modlist, function(x) teammates(x, focal = focal, node.location = node.location))
+  
+  total <- do.call(c, run_teammates)
+  
+  output <- table(total)
+}
+  

### 2015-01-28 Rscripts/trait_network_analysis.R

@@ -1,3 +1,143 @@
+## This script analyzes the association between plant traits and gall-parasitoid networks.
+
+
+plot(log(vLG.2011+1) ~ log(Total_Area) + C_N_imputed, tree_level_interaxn_all_plants_traits_size)
+test <- with(tree_level_interaxn_all_plants_traits_size, vLG.2011/Total_Area)
+visreg(lm(test ~ C_N_imputed, tree_level_interaxn_all_plants_traits_size))
+summary(lm(vLG.2011 ~ poly(Total_Area,2) + C_N_imputed, tree_level_interaxn_all_plants_traits_size))
+
+summary(MASS::glm.nb(vLG.2011 ~ offset(log(Total_Area)) + C_N_imputed + Height, tree_level_interaxn_all_plants_traits_size))
+
+test <- MASS::glm.nb(vLG_abund ~ offset(log(shootEst.no18)) + C_N_imputed + Height, tree_level_interaxn_all_plants_traits_size)
+summary(test)
+library(AER)
+dispersiontest(test)
+
+test.df <- filter(tree_level_interaxn_all_plants_traits_size, vLG_abund > 0)
+summary(lm(log(vLG_abund) ~ log(shootEst.no18) + C_N_imputed, test.df))
+summary(lm(log(vLG_abund) ~ offset(log(shootEst.no18)) + C_N_imputed, test.df))
+summary(lm(log(vLG_abund/shootEst.no18) ~ C_N_imputed, test.df))
+
+shoothigh <- which(tree_level_interaxn_all_plants_traits_size$shootEst.no18 > 800)
+## upload source code for managing network data at the tree-level
+source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
+source('~/Documents/miscellaneous_R/length2_function.R')
+
+## upload required libraries
+library(visreg) # regression visualization
+
+## get genotype-level interaction data. I accounted for density by dividing by the number of estimated shoots, because that was the most straightforward way to account for differences in sampling among trees.
+tree_level_interaxn_density <- select(tree_level_interaxn_all_plants, aSG_aSG.larv:gall_survive_abund)/tree_level_interaxn_all_plants$shootEst.no18*100 # scale to a per 100 shoot basis
+
+geno_level_interaxn_density <- aggregate(tree_level_interaxn_density, 
[…]

### 2015-01-28 ms_willow_network.Rmd
editing network ms

@@
 
-### Abstract
-Interactions between individual organisms give rise to the arrangement and strength of mutualistic and antagonistic interactions among species within a community.
-While we know this ecological network structure can have a profound influence on community dynamics, we still have a limited understanding of how different genotypes and phenotypes of individual organisms contribute to overall network structure.
-We address this knowledge gap by using a common garden experiment to examine how genetic and phenotypic variation within a dominant host-plant species influences an insect herbivore-parasitoid interaction network.
-We found that the plant genotype-herbivore-parasitoid network exhibited a modular structure, due to distinct genotype-herbivore-parasitoid associations.
-In particular, we found that genetic variation in plant traits influenced both the density and size of herbivores, which in turn, affected attack rates from parasitoids.
-Taken together, our results indicate that genetic variation can play a key role in structuring ecological networks..
+### Abstract. Re-read and revise later.
+The structure of ecological networks is determined by interactions between individual organisms -- not species -- within a community.
+While we know this network structure can have a profound influence on community dynamics, we still have a limited understanding of how different genotypes and phenotypes of individual organisms contribute to overall network structure.
+We address this knowledge gap by using a common garden experiment to examine how genetic and phenotypic variation within a dominant plant species influences an insect herbivore-parasitoid interaction network.
+We found considerable beta-diversity in herbivore-parasitoid networks, which was primarily due to species turnover.
+In other words, herbivores were susceptible to different parasitoids depending on which plant genotype they were found on.
+In particular, we found that genetic variation in plant traits influenced both the density and size of herbivores, which in turn affected attack rates from parasitoids.
+Taken together, our results indicate that genetic variation can play a key role in structuring ecological networks.
 Furthermore, our results highlight the potential for microevolutionary processes to shape both the structure and dynamics of ecological networks.
 
-### Introduction
-Network theory has emerged as a powerful and general framework for quantifying the arrangement and strength of species interactions within communities and makes clear predictions for how this structure should influence community dynamics (Strogatz 2001).
-For example, mutualistic interaction networks (e.g., plant-pollinator) tend to have a nested structure, where specialist species interactions are a subset of generalist species interactions (Bascompte et al., 2003; Olesen et al.
-2007; Thebault and Fontaine 2011).
-In contrast, antagonistic interaction networks (e.g., food webs) tend to exhibit a modular structure, where groups of species frequently interact with each other compared to other species (Krause et al., 2003; Rezende et al., 2009; Thebault and Fontaine 2011).
-These nested and modular structures, for mutualistic and antagonistic networks respectively, determine the robustness of these networks to perturbations (Thebault and Fontaine 2011; Stouffer and Bascompte, 2011; Rohr et al., 2014).
+### Introduction.  Re-read and revise later.
+Network theory has emerged as a powerful framework for quantifying the arrangement of species interactions within communities and makes clear predictions for how this structure should influence community dynamics (Thebault2011, Stouffer2011, Rohr2014).
 However, studies quantifying the structure of ecological networks have been resolved predominantly to the species level (Ings et al., 2011); neglecting the considerable phenotypic variation that can exist within species and the fact that interactions occur between individuals -- not species -- within a community.
@@
 2003, 2012).
-For example, genetic variation within host-plant species can influence the abundance and diversity of herbivores and, in turn, alter interactions with predators and parasitoids at higher trophic levels (Bailey et al., 2006; Johnson 2008; Abdala-Roberts and Mooney 2014).
+For example, genetic variation within host-plant species can influence the abundance and diversity of herbivores and, in turn, alter interactions with higher trophic levels (Bailey et al., 2006; Johnson 2008; Abdala-Roberts and Mooney 2014).
 Similarly, genetic and phenotypic variation within predators has been found to mediate the strength of trophic cascades (Post et al.
@@
 While it is clear that microevolutionary processes within a single species can influence community assembly, these studies have failed to examine the ecological interactions (i.e., genotype-species and species-species) that determine the assembly process.
-Consequently, we lack a detailed understanding of the structure of ecological interactions that shape the genetic-basis to community assembly.
+Consequently, we lack a detailed understanding of the structure of ecological interactions that shape the genetic basis to community assembly.
 
-Incorporating intraspecific genetic variation into network theory has the potential to reveal the underlying structures that mediate the ecological and evolutionary dynamics of communities.
+Incorporating genetic variation into network theory has the potential to reveal the underlying structures that mediate the ecological and evolutionary dynamics of communities.
 For instance, if certain species interact more frequently with particular genotypes as a result of their phenotype, then we hypothesize that this genetic specificity would result in a modular network structure.
@@
 Here, we test the hypothesis that intraspecific genetic variation results in a modular network structure using the insect gall-parasitoid network (5 gall species, 8 parasitoid species) associated with the coastal willow Salix hookeriana.
-This willow-gall-parasitoid system was ideal for testing our hypothesis for several reasons.
-First, S.
-hookeriana, displays considerable genetic and phenotypic variation that corresponds to variation in herbivory by galling insects (Barbour et al.
-2014 in press).
+This willow-gall-parasitoid system was ideal for testing our hypothesis for two reasons.
+Firstly, *S.
+hookeriana* displays considerable genetic and phenotypic variation that results in variation in herbivory by galling insects (Barbour et al.
+2015 in press).
 Second, the closed nature of galls restricts their parasitoid community to a small number of species that can successfully locate and parasitize galls.
-Third, the sedentary nature of galling insects make them amenable to building quantitative interaction networks due to the fact that galls can be collected and reared in the lab to determine survival and the extent of parasitism from different species (van Veen et al.
+Consequently, it is easy to identify and quantify the sources of mortality, and ultimately, build quantitative interaction networks (van Veen et al.
 2006).
-Finally, prior studies in a different willow species have shown that genetic and phenotypic variation can influence both the density and size of a common stem-galling sawfly, which in turn, determines the likelihood of parasitoid attack (Price 1988; Craig, Itami, & Price, 1990; Craig 1994).
 
@@
 Taken together, our study seeks to understand how genetic variation shapes the structure of ecological networks.
-And since genetic variation provides the raw material for natural selection, establishing a genetic-basis to network structure will begin to inform the eco-evolutionary dynamics of communities.
+And since genetic variation provides the raw material for natural selection, establishing a genetic-basis to network structure will begin to inform the eco-evolutionary dynamics of ecological networks.
 
@@
 ## Beta-diversity of gall-parasitoid networks among willow genotypes
-To build a quantitative gall-parasitoid network for each willow genotype, we collected all galls occurring on one haphazardly selected basal branch from about 5 randomly chosen replicates of each genotype (N = 146 trees, range = 4-9 trees per genotype).
-To control for differences in branch size, we estimated the number of shoots on each branch based on an allometric equation using the stem diameter of the sampled basal branch (mean ± SD shoot count = 280 ± 124; details in supplementary materials).
-We collected all galls in September 2012 when gall larva were in late instars of their development or had already spun cocoons within the gall.
-All galls were placed into 30 mL plastic transport vials (loosely capped at the end) and maintained at room temperature in the lab for four months.
+**HOW DATA WAS COLLECTED** To build quantitative gall-parasitoid networks, we collected galls from about 5 randomly chosen replicates of each genotype in September 2012 (N = 146 trees, range = 4-9 trees per genotype).
+For each replicate willow, we collected all galls occurring on one randomly selected basal branch.
+For each branch, we estimated the number of shoots based on an allometric equation using the stem diameter of the sampled basal branch (mean ± SD shoot count = 280 ± 124; details in supplementary materials).
+We then used these shoot estimates to quantify the density of galls and gall-parasitoid interactions on each replicate willow.
+To quantify gall-parasitoid interactions, we placed collected galls into 30 mL plastic transport vials (loosely capped at the end), which we maintained at room temperature in the lab for four months.
 We then opened galls under a dissecting scope and determined whether the gall survived or was parasitized, and if parasitized, the identity of the parasitoid species.
-We omitted galls for which we could not reliably determine the cause of mortality from further analysis.
+We omitted from analyses those galls for which we could not reliably determine the cause of mortality.
 
-We used a null model to test our prediction that the willow genotype-gall-parasitoid network exhibits a modular structure.
-To do this, we first pooled all observed gall-parasitoid interactions for each willow genotype.
-This resulted in a weighted, bipartite (i.e., two sets of nodes) network, where each willow genotype and each unique gall-parasitoid interaction comprised the two sets of nodes, respectively.
-The weights between nodes corresponded to the frequency of each link (willow genotype-gall-parasitoid interaction).
-To identify different modules of willow genotype-gall-parasitoid associations, we used an algorithm (QuaBiMo) that seeks to find the optimal partition of the network into groups, or modules, of nodes that frequently interact with each other compared to other nodes (Dorman and Strauss 2014).
-This module-finding algorithm seeks to maximize the modularity value Q, which is calculated as,
-￼
- where m is half of the total number of observed links in the network, Aij is the weighted, bipartite network and Kij is the network of expected weights (Dormann and Strauss 2013).
-The module to which a node i or j is assigned is ci,cj.
-The indicator function δ(ci,cj) = 1 if ci = cj and 0 if ci ≠ cj.
-Q ranges from 0, which means the community has no more links within modules than expected by random chance, to a maximum value of 1, which indicates increasing support for the division of a network into modules.
-Since Q is determined by an optimality function and is susceptible to being trapped at local optima, we repeated this calculation 100 times and used the iteration with the highest Q value to identify the modules within our network.
-Importantly, the magnitude of Q will be influenced by the number of nodes, the number of links between nodes, and the total number of links observed in the network (Dormann & Strauss, 2014; Thébault, 2012).
-Therefore, we used a conservative null model to examine whether the degree of modularity (Q) we observed was significantly different from what we would expect by chance.
-Specifically, we used a swapping algorithm that randomized the observed weights in the bipartite network while preserving both the total weights for each genotype and each gall-parasitoid interaction, as well as the number of unique links from the original network.
-We generated 1000 null bipartite networks and calculated Q for each one.
-We then calculated the Z-score, which is equal to the difference between the observed modularity and the mean of the null Q’s divided by the standard deviation of the null Q’s.
-Assuming that Q is normally distributed, a Z-score > 1.96 corresponds to a significantly modular network structure (P < 0.05; Whitlock and Schluter, 2009).
-Both the modularity and null model analysis were conducted with the bipartite package in R (Dormann and Strauss 2014; R Development Core Team 2014).
+**FRAMEWORK FOR MEASURING NETWORK BETA-DIVERSITY** We measured the beta-diversity of gall-parasitoid networks among willows by analyzing both qualitative and quantitaive differences in the composition of their links (i.e., species interactions) (Poisot et al.
+2012).
+These differences in link composition can then be partitioned into species turnover and interaction components.
+Differences due to species turnover arise from the gain/loss of species (either gall or parasitoid) altering link composition, whereas differences due to interactions occur when species switch with whom they are interacting, despite having the same species composition.
+For qualitative differences, we transformed the quantitative network into presence/absence data and calculated the pairwise Euclidean distances between sites.
+We used Euclidean distance because this treats joint absences of interactions between sites as meaningful, THIS IS USUALLY JUSTIFIED WHEN LOOKING AT TURNOVER IN A RELATIVELY FOCUSED COMMUNITY, BUT WHEN YOU ARE LOOKING AT BETA-DIVERSITY ACROSS LARGER GRADIENTS IT MAY NOT BE USEFUL BECAUSE TWO SITES MAY NOT BE MORE SIMILAR BECAUSE THEY DON'T HAVE THE SAME SPECIES.
+WE ALSO USED THE JACCARD AND WHITTAKER INDEX WITH A SUBSET OF DATA, WHERE WE HAD ENOUGH REPLICATE GENOTYPES WITHOUT JOINT ABSENCES (N = 58 OR SOMETHING).
+THE RESULTS ARE QUALITATIVELY THE SAME AS THE EUCLIDEAN DISTANCE SO WE ONLY REPORT THEM IN THE SUPPLEMENTARY INFO.
+Given that beta-diversity may also be influenced by differences in alpha-diversity between sites, we used the Raup-Crick null model to partition whether beta-diversity was due to simple changes in alpha-diversity among sites.
+We also looked at the euclidean distance of presence/absence networks.
+We even used subsets of the data where we had sufficients samples for each genotype.
+We therefore report these results in the supplementary info.
+FOR QUANTITATIVE DIFFERENCES, WE ALSO USED EUCLIDEAN DISTANCE. AS WITH THE QUALITATIVE DATA, OTHER DISSIMILARITY INDICES GAVE QUALITATIVELY THE SAME RESULTS, WHICH WE REPORT IN THE SUPPLEMENTARY INFO.
+
+**ANALYZING NETWORK BETA-DIVERSITY** To test whether willow genetic variation drove beta-diversity of gall-parasitoid networks, we used permutational ANOVA (PERMANOVA, 'adonis' function with *vegan* package in R) on pairwise Euclidean distances between link composition, with willow genotype modelled as a random effect.
+We then partitioned the relative contribution of species turnover and interaction components (with betalink package in R) to understand the processes driving network beta-diversity among genotypes.
+To estimate the proportion of variance in network dissimilarity explained by willow genotype, we calculated broad-sense heritability (*H^2*) using the equation: H2 = VG / VP, where VG is the total genotypic variance among clones, and VP is the total phenotypic variance, calculated as the sum of the residual (Vr) and genetic variance (Lynch & Walsh 1998).
+Following Lynch and Walsh (1998), we estimated Vg and Vp from out PERMANOVA table.
+Broad-sense heritability values range between 0-1, where values close to zero indicate low heritability (i.e., gall resistance is strongly influenced by the environment), and values close to 1 indicate high heritability (i.e., gall resistance is strongly controlled by underlying genetic variation).
+
+**PAIRWISE SPECIES-INTERACTIONS AMONG WILLOW GENOTYPES* FOR THE QUANTITATIVE DATA, WE WERE ALSO INTERESTED IN IDENTIFYING WHICH GALL-PARASITOID INTERACTIONS WERE DRIVING NETWORK BETA-DIVERSITY.
+TO DO THIS Given the quantitative nature of our data, we were interested in also identifying which gall-parasitoid interactions was driving network beta-diversity.
+To test whether the density of gall-parasitoid interactions varied among willow genotypes, we analyzed separate random effect models.
+Since we only had one response variable in each model, we used restricted maximum likelihood (REML) to estimate the variance due to willow genotype (VG) and unexplained residuals (Vr), which we used for calculating the broad-sense heritability of each interaction (METHODS PAPER).
+We then performed restricted likelihood-ratio tests to examine whether willow genotype explained a significant proportion of the variance in each gall-parasitoid interaction ('RLRTsim' function in *RLRTsim* package in R).
+
+To get at a more interpretable understanding of how species turnover could affect galling insects, we also divided up the natural enemy community into three different guilds.
+Egg parasitoids, larval parasitoids, and predators.
+These guilds may impose different functional constraints on galling insects.
+In particular, gall size.
+For example, for Cecidomyiid midges, larva initiate gall development, therefore we would not expect any characteristics of the gall to impose a functional constraint (e.g.
+gall size) on the likelihood of parasitoid attack, and is therefore causing selection on this stage of their life history.
+In contrast, larval parasitoids attack after the gall is formed and therefore, we would expect them to be influenced by the size of the gall and potentiall constrained from being able to oviposit in large galls.
+In our system, one of the species is a predatory midge which presumably attacks the larval stage as well, but giving their distant evolutionary relationship to parasitoid wasps, may be constrained by other factors, which is why we retained it as a separate guild.
+Therefore, we also analyzed differences in link composition when parasitoid species were collapsed into different guilds.
 
 ## Genetic variation in plant traits influences gall density and size
-To identify the plant traits mediating the resistance of willow genotypes to herbivory from galling insects, we measured 40 different traits associated with variation in leaf quality (36 traits) and plant architecture (4 traits).
+To identify the plant traits mediating the resistance of willow genotypes to herbivory from galling insects, we first measured 40 different traits associated with variation in leaf quality (36 traits) and plant architecture (4 traits).
 Details on how these willow traits were sampled and quantified are given in Barbour et al.
@@
 Plant architecture traits included: plant size, fractal dimension (index of architectural complexity), height, and foliage density.
-We calculated the broad-sense heritability (H2) of plant traits using the equation: H2 = VG / VP, where VG is the total genotypic variance among clones, and VP is the total phenotypic variance, calculated as the sum of the residual and genetic variance (Lynch & Walsh 1998).
+We calculated the broad-sense heritability (*H^2*) of plant traits using the equation: H2 = VG / VP, where VG is the total genotypic variance among clones, and VP is the total phenotypic variance, calculated as the sum of the residual and genetic variance (Lynch & Walsh 1998).
 Broad-sense heritability values range between 0-1, where values close to zero indicate low heritability (i.e., the trait is strongly influenced by the environment), and values close to 1 indicate high heritability (i.e., the trait is strongly controlled by underlying genetic variation).
-Each of these traits exhibited significant broad-sense heritable variation among willow genotypes (mean leaf quality H2 = 0.72; mean architecture H2 = 0.27; range of H2 = 0.15 - 0.97; Barbour et al., 2014 in press).
+Each of these 40 traits exhibited significant broad-sense heritable variation among willow genotypes (mean leaf quality H2 = 0.72; mean architecture H2 = 0.27; range of H2 = 0.15 - 0.97; Barbour et al., 2014 in press).
 
-We calculated both gall density and gall size size (measured as the maximum diameter perpendicular to direction of plant tissue growth, to the nearest 0.01 mm) of each species from our gall collections on each replicate willow tree (N = 146).
-We then conducted two types of analyses to test whether heritable variation in plant traits affects the density and size of galling insects.
-First, we calculated separate restricted maximum likelihood (REML) random-effect models and performed restricted likelihood-ratio tests to examine whether willow genotypes exhibited variation in the density and size of each gall species.
+We calculated both gall density and gall size (measured as the maximum diameter perpendicular to direction of plant tissue growth, to the nearest 0.01 mm) of each species from our gall collections on each replicate willow tree (N = 146).
+
+To test whether willows exhibited genetic variation in resistance to galling insects, we analyzed separate random effect models for the density and size of each gall species.
+We used restricted maximum likelihood (REML) to estimate the variance due to willow genotype (VG) and unexplained residuals (Vr).
+We then performed restricted likelihood-ratio tests to examine whether willow genotype explained a significant portion of the variance in the density and size of each gall species ('RLRTsim' function in *RLRTsim* package in R).
+To estimate the proportion of variance in gall density and gall size explained by willow genotype, we calculated broad-sense heritability (*H^2*) using the equation: H2 = VG / VP, where VG is the total genotypic variance among clones, and VP is the total phenotypic variance, calculated as the sum of the residual (Vr) and genetic variance (Lynch & Walsh 1998).
+Broad-sense heritability values range between 0-1, where values close to zero indicate low heritability (i.e., gall resistance is strongly influenced by the environment), and values close to 1 indicate high heritability (i.e., gall resistance is strongly controlled by underlying genetic variation).
+
+We only conducted our analysis though for gall species that were associated with genetic variation in gall-parasitoid interactions.
 Gall density and gall size were transformed as needed to improve normality and reduce heteroscedasticity of model residuals.
+
 As with each plant trait, we calculated the broad-sense heritability of gall density and gall size for each species.
@@
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+### Figure Legends
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r}
-summary(cars)
-```
-
-You can also embed plots, for example:
-
-```{r, echo=FALSE}
-plot(cars)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+![Figure 1](figures_network_ms_portrait.001.png)

### 2015-03-06 ms_willow_network_supplement.Rmd
lots of changes. updating manuscript

---
title: "ms_willow_network_supplement"
author: "Matthew Barbour"
date: "January 31, 2015"
output: word_document
---

The code below replicates the analyses for the qualitative and quantitative network data.

```{r prepare, echo=FALSE}
## source in required datasets and functions. 
source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
source('~/Documents/miscellaneous_R/model_diagnostic_functions.R')
source('~/Documents/miscellaneous_R/jensen_magnitude_function.R')

## load required libraries
#library(vegan) # for adonis analysis
#library(knitr) # for making tables
library(ggplot2) # for plotting
theme_set(theme_classic()) # customize ggplot for prettier default graphs
library(gridExtra) # for grid.arrange
library(mvabund) # multivariate analysis
library(HH) # for variance inflation factor analysis (vif) in multiple regression

## create datasets for analysis
full.df <- tree_level_interaxn_all_plants_traits_size 
interaxns_noPont # all gall-parasitoid interactions used for analysis
```

```{r mv abund galls}
gall.mvabund <- mvabund(full.df[ ,c("vLG_abund","rG_abund","aSG_abund","SG_abund")])
meanvar.plot(gall.mvabund ~ full.df$Genotype)
poisson_curve(from = 0.1, to = 10)
quasipoisson_curve(from = 0.1, to = 10, quasi.scalar = 2)
neg.binomial_curve(from = 0.1, to = 10, theta.negbin = 0.7) # negative binomial provides the best fit to the data.
```

Assessed negative binomial assumptions
```{r gall abund negative binomial assumptions}
manyglm.gall <- manyglm(gall.mvabund ~ Genotype,
                        data = full.df,
                        family = "negative.binomial")
plot(manyglm.gall, which = 1:3) # residuals look really good. No heteroscadisticity or non-normality. Note that replotting the residuals gives qualitatively the same picture (it's important to replot them because the residuals involve random number generation, see ?plot.manyglm)
```

Tested whether genotype affected gall community composition
```{r gall abund test}
anova.gall <- anova.manyglm(manyglm.gall, p.uni = "adjusted")
anova.gall # vLG_abund, rG_abund and aSG_abund are driving the community response.
```

Boxplots of gall abundance among the different genotypes.
```{r gall abund across genotypes, echo = FALSE}
vLG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = vLG_abund)) + 
  geom_boxplot() + ylab("Iteomyia salicisverruca density (#/branch)")
rG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = rG_abund)) + 
  geom_boxplot() + ylab("Rabdophaga salicisbrassicoides density (#/branch)")
aSG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = aSG_abund)) + 
  geom_boxplot() + ylab("Cecidomyid sp. A density (#/branch)")
SG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = SG_abund)) + 
  geom_boxplot() + ylab("Rabdophaga salicisbattatus density (#/branch)")

grid.arrange(vLG.abund.plot, rG.abund.plot, aSG.abund.plot, SG.abund.plot)
```

```{r mvabund gall-trait associations}
# selects galls and traits for analysis
galls.traits.df <- full.df %>%
  select(aSG_abund:SG_abund,
         vLG.height.mean, vLG.gall.count,
         Trichome.No., Total_Area, Height, Density,
         C_N_imputed, water_content, specific_leaf_area, sal_tannin.PC1:flavanonOLES.PC1)

# create some new variables to reduce multicollinearity among non-principal component traits
gall.density.df <- galls.traits.df %>%
  mutate(log_size = log(Total_Area),
         log_trichomes = log(Trichome.No.+1)) %>%
  select(aSG_abund:SG_abund, Height:log_trichomes) %>%
  na.omit()
gall.density.df$Height_resid = resid(lm(Height ~ log_size, gall.density.df))
gall.density.df$Density_resid = resid(lm(Density ~ log_size, gall.density.df))
gall.density.df$SLA_resid = resid(lm(specific_leaf_area ~ water_content, gall.density.df))

# subset the final datasets
galls_galls.traits <- mvabund(select(gall.density.df, aSG_abund:SG_abund))
traits_galls.traits <- select(gall.density.df, C_N_imputed, water_content, sal_tannin.PC1:SLA_resid)

mvabund.galls.traits <- manyglm(galls_galls.traits ~ C_N_imputed + water_content + sal_tannin.PC1 + cinn.PC1 + cinn.PC2 + flavonOLES.PC1 + flavonOLES.PC2 + flavanonOLES.PC1 + log_size + log_trichomes + Height_resid + Density_resid + SLA_resid, 
                                data = traits_galls.traits, 
                                family = "negative.binomial")
plot(mvabund.galls.traits, which = 1:3)
```

```{r stepAIC galls-traits}
## use stepAIC to identify traits that reduce AIC the most and remove them in a stepwise fashion from analysis.
drop1(mvabund.galls.traits)
mod2 <- update(mvabund.galls.traits, .~. - sal_tannin.PC1); drop1(mod2)
mod3 <- update(mod2, .~. - flavonOLES.PC1); drop1(mod3)
mod4 <- update(mod3, .~. -water_content); drop1(mod4)
mod5 <- update(mod4, .~. - cinn.PC2); drop1(mod5)
mod6 <- update(mod5, .~. - cinn.PC1); drop1(mod6)
mod7 <- update(mod6, .~. -SLA_resid); drop1(mod7)
mod8 <- update(mod7, .~. -Height_resid); drop1(mod8)
mod9 <- update(mod8, .~. -Density_resid); drop1(mod9)
mod10 <- update(mod9, .~. -flavonOLES.PC2); drop1(mod10)
mod11 <- update(mod10, .~. -log_trichomes); drop1(mod11)
mod12 <- update(mod11, .~. -C_N_imputed); drop1(mod12)
mod13 <- update(mod12, .~. -log_size); drop1(mod13)
mod14.null <- update(mod13, .~. -flavanonOLES.PC1)

## use AIC to compare models
AIC.models.galls.traits <- data.frame(Model = 1:14, # progressive order in which models were fit
           Formula = c(paste(formula(mvabund.galls.traits)[3]), 
                       paste(formula(mod2)[3]),
                       paste(formula(mod3)[3]), 
                       paste(formula(mod4)[3]),
                       paste(formula(mod5)[3]), 
                       paste(formula(mod6)[3]),
                       paste(formula(mod7)[3]),
                       paste(formula(mod8)[3]), 
                       paste(formula(mod9)[3]),
                       paste(formula(mod10)[3]), 
                       paste(formula(mod11)[3]),
                       paste(formula(mod12)[3]),
                       paste(formula(mod13)[3]),
                       paste(formula(mod14.null)[3])),
           AIC = c(sum(AIC(mvabund.galls.traits)), 
                   sum(AIC(mod2)), 
                   sum(AIC(mod3)), 
                   sum(AIC(mod4)),
                   sum(AIC(mod5)), 
                   sum(AIC(mod6)),
                   sum(AIC(mod7)), 
                   sum(AIC(mod8)), 
                   sum(AIC(mod9)),
                   sum(AIC(mod10)), 
                   sum(AIC(mod11)),
                   sum(AIC(mod12)), 
                   sum(AIC(mod13)), 
                   sum(AIC(mod14.null))))
arrange(AIC.models.galls.traits, AIC)

## Likelihood ratio tests suggest that all of the models with lower AIC than the null model provide a better fit to the data than the null.
anova.manyglm(mod14.null, mod13)
anova.manyglm(mod14.null, mod12)
anova.manyglm(mod14.null, mod11)
anova.manyglm(mod14.null, mod10)

# identify gall-trait associations. C_N_imputed has a marginally significant effect on Iteomyia and R. salicisbrassicoides abundance. FlavononOLES.PC1 has a signficiant effect on Cecidomyiid sp. A abundance, and plant size as a significant effect on R. salicisbrassicoides and a marginal effect on R. salicisbattatus abundance.
uni.galls.traits <-anova.manyglm(mod11, p.uni = "unadjusted")

coef.df <- data.frame(coef(mod11))

range.flavanonOLES.PC1 <- with(gall.density.df, max(flavanonOLES.PC1) - min(flavanonOLES.PC1))
exp(coef.df["flavanonOLES.PC1","aSG_abund"]*range.flavanonOLES.PC1) # over the range of flavanonOLES.PC1, Cecidomyiid sp. A abundance increased 15-fold

range.C_N_imputed <- with(gall.density.df, max(C_N_imputed) - min(C_N_imputed))
exp(coef.df["C_N_imputed", "vLG_abund"]*range.C_N_imputed) # over the range of C:N ratio, Iteomyia abundance increased 3-fold (2.9)

exp(coef.df["C_N_imputed", "rG_abund"]*range.C_N_imputed) # over the range of C:N ratio, R. salicisbrassicoides abundance increased 8-fold (7.6)

1 - 1.10^coef.df["log_size","rG_abund"] # for every 10% increase in plant size, R. salicisbrassicoides density decreased 9%

1 - 1.10^coef.df["log_size","SG_abund"] # for every 10% increase in plant size, R. salicisbattatus density decreasted by 37%
```

```{r visreg univariate gall-trait associations}
library(visreg)
# include model with all of the traits, but only examine one of the variables
vLG.trait <- MASS::glm.nb(vLG_abund ~ C_N_imputed + log_size + flavanonOLES.PC1, gall.density.df)
summary(vLG.trait) 
visreg(vLG.trait, xvar = "C_N_imputed", scale = "response")

aSG.trait <- MASS::glm.nb(aSG_abund ~ C_N_imputed + log_size + flavanonOLES.PC1, gall.density.df)
summary(aSG.trait)
visreg(aSG.trait, xvar = "flavanonOLES.PC1", scale = "response") # may or may not be too important

rG.trait <- MASS::glm.nb(rG_abund ~ C_N_imputed + log_size + flavanonOLES.PC1, gall.density.df)
summary(rG.trait)
visreg(rG.trait, xvar = "log_size", scale = "response")
visreg(rG.trait, xvar = "C_N_imputed", scale = "response")

SG.trait <- MASS::glm.nb(SG_abund ~ C_N_imputed + log_size + flavanonOLES.PC1, gall.density.df)
summary(SG.trait) # some warnings about lack of convergence
visreg(SG.trait, xvar = "log_size", scale = "response") # not a great fit. Outlier seems to be driving most of this relationship.
```

```{r}
## gall size
gall.size.df <- galls.traits.df %>%
  mutate(log_size = log(Total_Area),
         log_trichomes = log(Trichome.No.+1)) %>%
  select(vLG.height.mean, vLG.gall.count, Height:log_trichomes) %>%
  na.omit()
gall.size.df$Height_resid = resid(lm(Height ~ log_size, gall.size.df))
gall.size.df$Density_resid = resid(lm(Density ~ log_size, gall.size.df))
gall.size.df$SLA_resid = resid(lm(specific_leaf_area ~ water_content, gall.size.df))
gall.size.df <- select(gall.size.df, vLG.height.mean, vLG.gall.count, 
                       C_N_imputed, water_content,
                       sal_tannin.PC1:SLA_resid)

vLG.size.lm <- lm(vLG.height.mean ~ ., select(gall.size.df, -vLG.gall.count), weights = gall.size.df$vLG.gall.count)

drop1(vLG.size.lm)
size2 <- update(vLG.size.lm, .~. -C_N_imputed); drop1(size2)
size3 <- update(size2, .~. -Density_resid); drop1(size3)
size4 <- update(size3, .~. -flavonOLES.PC2); drop1(size4)
size5 <- update(size4, .~. -flavanonOLES.PC1); drop1(size5)
size6 <- update(size5, .~. -log_trichomes); drop1(size6)
size7 <- update(size6, .~. -SLA_resid); drop1(size7)
size8 <- update(size7, .~. -Height_resid); drop1(size8)
size9 <- update(size8, .~. -water_content); drop1(size9)
size10 <- update(size9, .~. -log_size); drop1(size10)
size11 <- update(size10, .~. -cinn.PC1); drop1(size11)
size12 <- update(size11, .~. -cinn.PC2); drop1(size12)
size13 <- update(size12, .~. -flavonOLES.PC1); drop1(size13)
size14.null <- update(size13, .~. -sal_tannin.PC1)

## use AIC to compare models
AIC.vLG.size.traits <- data.frame(Models = 1:14, # progressive order in which sizeels were fit
           Formula = c(paste(formula(vLG.size.lm)[3]), 
                       paste(formula(size2)[3]),
                       paste(formula(size3)[3]), 
                       paste(formula(size4)[3]),
                       paste(formula(size5)[3]), 
                       paste(formula(size6)[3]),
                       paste(formula(size7)[3]),
                       paste(formula(size8)[3]), 
                       paste(formula(size9)[3]),
                       paste(formula(size10)[3]), 
                       paste(formula(size11)[3]),
                       paste(formula(size12)[3]),
                       paste(formula(size13)[3]),
                       paste(formula(size14.null)[3])),
           AIC = c(sum(AIC(vLG.size.lm)), 
                   sum(AIC(size2)), 
                   sum(AIC(size3)), 
                   sum(AIC(size4)),
                   sum(AIC(size5)), 
                   sum(AIC(size6)),
                   sum(AIC(size7)), 
                   sum(AIC(size8)), 
                   sum(AIC(size9)),
                   sum(AIC(size10)), 
                   sum(AIC(size11)),
                   sum(AIC(size12)), 
                   sum(AIC(size13)), 
                   sum(AIC(size14.null))))
arrange(AIC.vLG.size.traits, AIC)

# all of these models give the same picture. Sal_tannin.PC1 and flavonOLES.PC1 are the best predictors.
summary(size11)
summary(size10)
summary(size12)
summary(size9)

plot(size12) # residuals looks pretty good
vLG.size.predict <- visreg(size12)

# Iteomyia size decreased 1.2 fold over the range in salicylate/tannin chemistry
max(vLG.size.predict$sal_tannin.PC1$y$fit)/min(vLG.size.predict$sal_tannin.PC1$y$fit)

# Iteomyia size decreased 1.2 fold over the range in flavonid chemistry
max(vLG.size.predict$flavonOLES.PC1$y$fit)/min(vLG.size.predict$flavonOLES.PC1$y$fit)

```

```{r gall abund predict}
gall.predict <- predict(manyglm.gall, 
                        newdata = data.frame(Genotype = levels(full.df$Genotype)), 
                        type = "response")
gall.predict.df <- cbind.data.frame(gall.predict, Genotype = levels(full.df$Genotype))
gall.predict.melt <- melt(gall.predict.df)

ggplot(gall.predict.melt, aes(x = Genotype, y = value, fill = variable)) + geom_bar(stat = "identity")

colMeans(gall.predict)
library(vegan)
diversity(colMeans(gall.predict))

gall.predict.X.R <- predict(manyglm.gall, 
                            newdata = data.frame(Genotype = c("X","R")),
                            type = "response")
colMeans(gall.predict.X.R)
diversity(colMeans(gall.predict.X.R))

sample(levels(full.df$Genotype), 5)

names(manyglm.gall$data)
predict.allcombos <- function(manyglm.object, factor.index, predict.type = "response"){
  require(plyr)
  require(dplyr)
  factor.levels <- levels(manyglm.object$data[ ,factor.index])

  predict.list <- list()
  for(i in 2:length(factor.levels)){
    tmp.data <- data.frame(sample(factor.levels, i))
    colnames(tmp.data) <- names(manyglm.object$data)[factor.index]
    predict.tmp <- predict(manyglm.object, newdata = tmp.data, type = predict.type)
    predict.list[[i]] <- colMeans(predict.tmp)
  }
  predict.df <- ldply(predict.list) %>%
    mutate(Number.of.Genotypes = 2:length(factor.levels))
}

N <- 50
gall.div.list <- list()
for(i in 1:N){
  gall.div.list[[i]] <- round(predict.allcombos(manyglm.gall, factor.index = 2), 1) # rounding to avoid pseudoaccuracy
} # takes a couple minutes with 50 reps

gall.div.df <- ldply(gall.div.list) %>% tbl_df()
gall.div.df$diversity <- diversity(select(gall.div.df, vLG_abund:SG_abund))

boxplot(diversity ~ factor(Number.of.Genotypes), gall.div.df, ylab = "Gall diversity", xlab = "Number of Genotypes")

N <- 50
link.div.list <- list()
for(i in 1:N){
  link.div.list[[i]] <- round(predict.allcombos(manyglm.full, factor.index = 2), 1)
} # takes a couple minutes with 50 reps

link.div.df <- ldply(link.div.list) %>% tbl_df()
link.div.df$diversity <- diversity(select(link.div.df, aSG_Tory:vLG_Tory))
link.div.df$repID <- 1:dim(link.div.df)[1]

boxplot(diversity ~ factor(Number.of.Genotypes), link.div.df, ylab = "Interaction Diversity", xlab = "Number of Genotypes")

link.melt <- melt(select(link.div.df, -diversity), id.vars = c("repID", "Number.of.Genotypes"))

link.split <- colsplit(link.melt$variable, "_", names = c("gall","parasitoid"))
link.melt.df <- cbind(link.melt, link.split)

web.list <- cast(link.melt.df, gall ~ parasitoid | repID + Number.of.Genotypes, mean)

library(bipartite)
web.list.measures <- c()
genotype.number <- c()
for(i in 1:dim(web.list)){
  genotype.number[i] <- names(web.list[[i]][1])
  tmp.web <- web.list[[i]][[1]]
  rownames(tmp.web) <- tmp.web$gall
  tmp.web <- as.matrix.data.frame(tmp.web[ ,-1])
  web.list.measures[i] <- networklevel(tmp.web, index = "linkage density", weighted = TRUE)
}
web.measures.df <- cbind.data.frame(Number.of.Genotypes = as.numeric(genotype.number),
                                    Weighted.Linkage.Density = web.list.measures)

boxplot(Weighted.Linkage.Density ~ factor(Number.of.Genotypes), web.measures.df, ylab = "Weighted Linkage Density", xlab = "Number of Genotypes")

web.measures.summary <- web.measures.df %>%
  group_by(Number.of.Genotypes) %>%
  summarise(median.link.density = median(Weighted.Linkage.Density))
with(web.measures.summary, max(median.link.density)/min(median.link.density))

library(mgcv)
web.sum.lm <- lm(log(median.link.density) ~ log(Number.of.Genotypes), data = web.measures.summary)
summary(web.sum.lm) # Need to understand interpretation here.
visreg(web.sum.lm) # not quite there
```

We were unable to use the mvabund framework for this analysis, so we conducted separate linear models to test for these effects.
```{r gall size}
plot(vLG.height.mean ~ Genotype, full.df)
vLG.size.lm <- lm(log(vLG.height.mean) ~ Genotype, filter(full.df, Genotype != "J"))
#plot(vLG.size.lm)
summary(vLG.size.lm)

vLG.size.lmer <- lmer(log(vLG.height.mean) ~ (1|Genotype), full.df)
heritability(vLG.size.lmer) # significant. Note though, that I originally valuate gall size using nested random effect models to account for individual measurements I had for each gall.

plot(rG.height.mean ~ Genotype, full.df)
rG.size.lm <- lm(log(rG.height.mean) ~ Genotype, full.df)
#plot(rG.size.lm)
summary(rG.size.lm)

plot(aSG.height.mean ~ Genotype, full.df) # too few data points to reliably evaluate
aSG.size.lm <- lm(log(aSG.height.mean) ~ Genotype, full.df)
#plot(aSG.size.lm)
summary(aSG.size.lm)

plot(SG.height.mean ~ Genotype, full.df) # too few data points to reliably evaluate
SG.size.lm <- lm(log(SG.height.mean) ~ Genotype, full.df)
plot(SG.size.lm)
summary(SG.size.lm)
```

```{r gall size variation among willow genotypes}
vLG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = vLG.height.mean)) + 
  geom_boxplot() + ylab("Iteomyia salicisverruca diameter (mm)")
rG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = rG.height.mean)) + 
  geom_boxplot() + ylab("Rabdophaga salicisbrassicoides diameter (mm)")
aSG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = aSG.height.mean)) + 
  geom_boxplot() + ylab("Cecidomyid sp. A diameter (mm)")
SG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = SG.height.mean)) + 
  geom_boxplot() + ylab("Rabdophaga salicisbattatus diameter (mm)")

grid.arrange(vLG.height.mean.plot, rG.height.mean.plot, aSG.height.mean.plot, SG.height.mean.plot)
```

Here, we evaluate the assumptions of mvabund and see which error distribution is appropriate. Specifically, we first look at a plot of the mean-variance relationship of our response variables. It is easy to see that the negative binomial model (black line) provides the best fit to this data, suggesting that we should specify this as the error distribution in our model
```{r mv abund interactions explore}
full.mvabund <- mvabund(full.df[ ,interaxns_noPont]) # create mvabund object for easy analysis
meanvar.plot(full.mvabund ~ full.df$Genotype) # easy to see that the data do not have a poisson distribution, therefore, a negative binomial may be a good fit.
poisson_curve(from = 0.1, to = 3.0, color = "blue")
quasipoisson_curve(from = 0.1, to = 3.0, quasi.scalar = 2, color = "red")
neg.binomial_curve(from = 0.1, to = 3.0, theta.negbin = 0.7, color = "black") # negative binomial appears to provide the best fit to this relationship
```

```{r genotype level}
geno.level.interactions.df <- full.df %>%
  select(Genotype, aSG_Tory:rG_Platy, rG_Tory, SG_Platy, vLG_Eulo:vLG_Tory, aSG_abund:vLG_abund, vLG.height.mean) %>%
  group_by(Genotype) %>%
  summarise_each(funs(mean.na.rm = mean(., na.rm = TRUE))) %>%
  #na.omit() %>%
  mutate(log.vLG_abund = log(vLG_abund),
         log.vLG.height.mean = log(vLG.height.mean),
         log.1.rG_abund = log(rG_abund+1),
         log.1.aSG_abund = log(aSG_abund+1))

library(psych)
corr.test(geno.level.interactions.df[ ,c("vLG_abund","rG_abund","aSG_abund","vLG.height.mean")])

car::scatterplotMatrix(select(geno.level.interactions.df, vLG_abund, rG_abund, aSG_abund, vLG.height.mean))
corr.test(geno.level.interactions.df[ ,c("log.vLG_abund","log.1.rG_abund","log.1.aSG_abund","log.vLG.height.mean")])

# correlations are rather weak
corr.test(select(geno.level.interactions.df, aSG_Tory:vLG_Tory))
corr.test(select(geno.level.interactions.df, vLG_Platy, vLG_Mesopol, vLG_Tory, rG_Tory))
car::scatterplotMatrix(select(geno.level.interactions.df, vLG_Platy, vLG_Mesopol, vLG_Tory, rG_Tory))

plot(rG_Eulo ~ vLG_Mymarid, geno.level.interactions.df)

plot(log.1.aSG_abund ~ log.vLG_abund, geno.level.interactions.df)

geno.level.mvabund <- mvabund(geno.level.interactions.df[ ,interaxns_noPont])

geno.level.manyglm <- manylm(geno.level.mvabund ~ vLG_abund*vLG.height.mean + rG_abund + aSG_abund, data = geno.level.interactions.df)

plot(geno.level.manyglm, which = 1:3)
anova.manylm(geno.level.manyglm, p.uni = "unadjusted")

hist(geno.level.interactions.df$vLG_Platy)
plot(vLG_Mesopol ~ vLG.height.mean, geno.level.interactions.df)
vLG_Platy.lm <- lm(vLG_Mesopol ~ vLG_abund+vLG.height.mean, geno.level.interactions.df)
summary(vLG_Platy.lm)
plot(vLG_Platy.lm)
AIC(vLG_Platy.lm)
library(visreg)
visreg(vLG_Platy.lm, scale = "response")
```

We then fit a model and used residual plots to diagnose the model fit.
```{r mv abund negative binomial assumptions}
manyglm.full <- manyglm(full.mvabund ~ Genotype,
                          data = full.df,
                          family = "negative.binomial")
plot(manyglm.full, which = 1:3) # residuals aren't quite normally distributed, but there doesn't seem to be any heteroscedasticity in the model fit. Note that replotting the residuals gives qualitatively the same picture (it's important to replot them because the residuals involve random number generation, see ?plot.manyglm)
```

```{r}
link.predict <- predict(manyglm.full, 
                        newdata = data.frame(Genotype = levels(full.df$Genotype)), 
                        type = "response")
link.predict.df <- cbind.data.frame(link.predict, Genotype = levels(full.df$Genotype))
library(vegan)
mean(vegdist(select(link.predict.df, -Genotype), method = "horn"))
sd(vegdist(select(link.predict.df, -Genotype), method = "horn"))
hist(vegdist(select(link.predict.df, -Genotype), method = "horn"))

link.predict.melt <- melt(link.predict.df)

ggplot(link.predict.melt, aes(x = Genotype, y = value, fill = variable)) + geom_bar(stat = "identity")
```

Given that a negative binomial distribution seems to provide a good fit to the data, we tested whether the composition of gall-parasitoid interaction varied among willow genotypes. To further diagnose which interactions were driving this response, we conducted univariate analyses on each predictor, but adjusted for multiple comparisons. P-values were adjusted for multiple testing using a step-down resampling procedure. This methods provides strong control of family-wise error rates and makes use of resampling to ensure inferences take into account correlation between variables (Westfall & Young 1993).

From the table, it is clear to say that genotype has a strong effect on the composition of links in the network. Moreover, differences in community composition are driven primarily by 3 interactions: vLG_Platy, vLG_Tory, and vLG_Mesopol.
```{r}
anova.full <- anova.manyglm(manyglm.full, p.uni = "unadjusted") # Takes about 1 min and 30 sec to run.
anova.full # vLG_Platy, vLG_Tory, and vLG_Mesopol are driving the community response. rG_Tory is marginally significant and vLG_Eulo is close to marginal as well.
```

Now we examine how variation in gall densities and gall size (for Iteomyia) affects the network.

First we created a dataset that contained complete observations of the network and predictor variables
```{r mv abund interaction predictors}
full.predictors <- full.df %>%
  select(aSG_Tory:rG_Platy, rG_Tory, SG_Platy, vLG_Eulo:vLG_Tory, # interactions
         aSG_abund:vLG_abund, # gall counts. Exclusing SG_abund because it didn't vary among genotypes
         vLG.height.mean) %>% # Iteomyia gall size
# we transformed some of the variables so we could examine their vif 
  mutate(log.vLG_abund = log(vLG_abund),
         log.1.vLG_abund = log(vLG_abund + 1),
         log.1.rG_abund = log(rG_abund + 1),
         log.1.aSG_abund = log(aSG_abund + 1),
         log.vLG.height.mean = log(vLG.height.mean)) %>%
  na.omit() # removes all instances that don't have vLG.height.mean
dim(full.predictors)[1] # 81 data points
```

We then looked for evidence of variance inflation among the predictor variables. But found little evidence for it.
```{r vif}
vif(full.predictors[ ,c("log.1.aSG_abund","log.1.rG_abund","log.vLG_abund","log.vLG.height.mean")]) # still little evidence of variance inflation among predictor variables.
```

We log transformed all predictor variables because it provided a much better fit to the data as determined by AIC. We then used AIC to compare our most complex model to our least complex. Instead of exploring all possible combinations, we started with the most complex model and used AIC to drop predictor variables with the lowest AIC values (i.e. least impact on removal). Using AIC, we identified 3 equivalenet models (difference in AIC < 2.1 amongst models). 
```{r AIC model assessment}
net.trait <- mvabund(full.predictors[ ,interaxns_noPont])
net.trait.sub <- mvabund(full.predictors[ ,c("vLG_Platy","vLG_Mesopol","vLG_Tory","rG_Tory")])
# even though manyglm doesn't appear to test interaction terms appropriately (sequential vs. dropping different terms), there is not interactive effect of vLG height and vLG abundance in univariate negative binomial models for any vLG_parasitoid interaction, so I believe this final model is appropriate for the data.
net.mvabund.1.full <- manyglm(net.trait ~ log.vLG_abund*vLG.height.mean + 
                         log.1.rG_abund + log.1.aSG_abund,
                       data = full.predictors,
                       family = "negative.binomial")
MASS::stepAIC(net.mvabund.1.full, direction = "both")

net.mvabund.2 <- update(net.mvabund.1.full, .~. - log.vLG_abund:vLG.height.mean)

drop1(net.mvabund.2)
net.mvabund.3 <- update(net.mvabund.2, .~. - vLG.height.mean)

drop1(net.mvabund.3)
net.mvabund.4 <- update(net.mvabund.3, .~. - log.1.aSG_abund)

drop1(net.mvabund.4)
net.mvabund.5 <- update(net.mvabund.4, .~. - log.vLG_abund)

drop1(net.mvabund.5)
net.mvabund.6.null <- update(net.mvabund.5, .~. -log.1.rG_abund)

AIC.models <- data.frame(Model = 1:6, # progressive order in which models were fit
           Formula = c(paste(formula(net.mvabund.1.full)[3]), 
                       paste(formula(net.mvabund.2)[3]),
                       paste(formula(net.mvabund.3)[3]), 
                       paste(formula(net.mvabund.4)[3]),
                       paste(formula(net.mvabund.5)[3]), 
                       paste(formula(net.mvabund.6.null)[3])),
           AIC = c(sum(AIC(net.mvabund.1.full)), 
                   sum(AIC(net.mvabund.2)), 
                   sum(AIC(net.mvabund.3)), 
                   sum(AIC(net.mvabund.4)),
                   sum(AIC(net.mvabund.5)), 
                   sum(AIC(net.mvabund.6.null))))
arrange(AIC.models, AIC)

anova.manyglm(net.mvabund.1.full, net.mvabund.2, net.mvabund.3, net.mvabund.4, net.mvabund.5, net.mvabund.6.null) # takes about 3 min to run.
```

After deciding on the model with all main effects and no interactions, we examined the residuals and everything looked pretty good.
```{r}
plot(net.mvabund.2, which = 1:3) # residuals look pretty good.
```

We then examined the which processes underlied variation in the gall-parasitoid interactions.
```{r}
anova.net <- anova.manyglm(net.mvabund.2, p.uni = "unadjusted") # takes about 2.5 min to run.
anova.net # not that since the P-values are determined by a resampling procedure, they may differ slightly between runs. Therefore, we retain all P-values < 0.10 (during at least one run) for the coefficient summary below.

sig.coef.df.2 <- mutate(melt(coef(net.mvabund.2)),
               predictor_response = paste(X1, X2, sep = "_")) %>%
 # select(predictor = X1, response = X2, predictor_response, value) %>%
  subset(predictor_response %in% c("log.1.aSG_abund_aSG_Tory", 
                                 "log.1.rG_abund_rG_Eulo",
                                 "log.1.rG_abund_rG_Platy",
                                 "log.1.rG_abund_rG_Lestodip", # marginal
                                 "log.vLG_abund_SG_Platy", # marginal
                                 "log.vLG_abund_vLG_Mymarid",
                                 "log.vLG.height.mean_vLG_Eulo", # marginal
                                 "log.vLG.height.mean_rG_Tory", "log.1.rG_abund_rG_Tory",
                                 "log.vLG_abund_vLG_Mesopol", "log.vLG.height.mean_vLG_Mesopol", 
                                 "log.vLG.height.mean_vLG_Platy", "log.vLG_abund_vLG_Platy",
                                 "log.vLG.height.mean_vLG_Tory", "log.vLG_abund_vLG_Tory")) %>% # vLG.height.mean is marginal for vLG_Tory
  select(predictor = X1, response = X2, coefficient = value)
arrange(sig.coef.df.2, predictor, response)
```

```{r Interaction predictions}
# calculate the mean values, on the original scale, of all predictor variables
no.var.data <- with(full.predictors,
                    data.frame(vLG_abund = mean(vLG_abund),
                               rG_abund = mean(rG_abund),
                               aSG_abund = mean(aSG_abund),
                               vLG.height.mean = mean(vLG.height.mean)))

# http://www.ats.ucla.edu/stat/mult_pkg/faq/general/log_transformed_regression.htm Use this web page for interpreting log and non-transformed coefficients in regression.
library(visreg)
par(mfrow = c(1,2))
vLG_Platy.glm <- MASS::glm.nb(vLG_Platy ~ log(vLG_abund) + vLG.height.mean, full.predictors); summary(vLG_Platy.glm); jensen_magnitude(vLG_Platy.glm, no.var.data)
visreg(vLG_Platy.glm, scale = "response")
#plot(vLG_Platy.glm)

vLG_Mesopol.glm <- MASS::glm.nb(vLG_Mesopol ~ log(vLG_abund) + vLG.height.mean, full.predictors); summary(vLG_Mesopol.glm); jensen_magnitude(vLG_Mesopol.glm, no.var.data)
#plot(vLG_Mesopol.glm)
visreg(vLG_Mesopol.glm, scale = "response")

vLG_Tory.glm <- MASS::glm.nb(vLG_Tory ~ vLG_abund+vLG.height.mean, full.predictors); summary(vLG_Tory.glm); jensen_magnitude(vLG_Tory.glm, no.var.data) # gall height is marginally significant, but the AIC is slightly better than the model with only vLG_abund.
visreg(vLG_Tory.glm, scale = "response")

rG_Tory.glm <- glm(rG_Tory ~ rG_abund + vLG.height.mean, full.predictors, family = quasipoisson); summary(rG_Tory.glm); jensen_magnitude(rG_Tory.glm, no.var.data) # underdispersed, which is why I used quasipoisson.
#plot(rG_Tory.glm)
visreg(rG_Tory.glm, scale = "response")

rG_Platy.glm <- glm(rG_Platy ~ rG_abund, full.predictors, family = quasipoisson); summary(rG_Platy.glm); jensen_magnitude(rG_Platy.glm, no.var.data)  # underdispersed, which is why I used quasipoisson.

rG_Eulo.glm <- glm(rG_Eulo ~ rG_abund, full.predictors, family = quasipoisson); summary(rG_Eulo.glm); jensen_magnitude(rG_Eulo.glm, no.var.data) # underdispersed, which is why I used quasipoisson

aSG_Tory.glm <- glm(aSG_Tory ~ aSG_abund, full.predictors, family = quasipoisson); summary(aSG_Tory.glm); jensen_magnitude(aSG_Tory.glm, no.var.data) # underdispersed, so I used quasipoisson.

rG_Lestodip.glm <- glm(rG_Lestodip ~ log(rG_abund + 1), full.predictors, family = quasipoisson)
summary(rG_Lestodip.glm); jensen_magnitude(rG_Lestodip.glm, no.var.data) # rG_abund is marginally significant.

## intercept only models.
vLG_Eulo.glm <- glm(vLG_Eulo ~ 1, full.predictors, family = poisson); summary(vLG_Eulo.glm); jensen_magnitude(vLG_Eulo.glm, no.var.data) # not a lot of support for including gall height as a predictor, althought it was marginally significant in multivariate model, so I dropped it from the model

SG_Platy.glm <- glm(SG_Platy ~ 1, full.predictors, family = poisson); summary(SG_Platy.glm); jensen_magnitude(SG_Platy.glm, no.var.data) # no evidence of vLG_abund being a clear predictor, so we dropped it from the model

rG_Mesopol.glm <- glm(rG_Mesopol ~ 1, full.predictors, family = poisson); summary(rG_Mesopol.glm); jensen_magnitude(rG_Mesopol.glm, no.var.data)

vLG_Mymarid.glm <- glm(vLG_Mymarid ~ 1, full.predictors, family = poisson); summary(vLG_Mymarid.glm); jensen_magnitude(vLG_Mymarid.glm, no.var.data) # vLG_abund is a significant predictor, but the visreg predictions look so wonky that I don't really trust it, so I just modelled it with an intercept
```

```{r visreg plots}
# Platygaster interactions
layout(matrix(c(1, 2, 3, 4), nrow = 2, ncol = 2, byrow = FALSE))
visreg(vLG_Platy.glm, scale = "response", ask = F)
plot(x = range(full.predictors$vLG_abund), y = c(0, 1), type = "n", ylab = "SG_Platy", xlab = "vLG_abund"); abline(h = exp(coef(SG_Platy.glm)), col = "steelblue", lwd = 3) # intercept only models
visreg(rG_Platy.glm, scale = "response", ask = F)

# Mesopolobus interactions
layout(matrix(c(1, 2, 3, 0), nrow = 2, ncol = 2, byrow = FALSE))
visreg(vLG_Mesopol.glm, scale = "response", ask = F)
plot(x = range(full.predictors$rG_abund), y = c(0, 1), type = "n", ylab = "rG_Mesopol", xlab = "rG_abund"); abline(h = exp(coef(rG_Mesopol.glm)), col = "steelblue", lwd = 3) # intercept only models

# Torymus interactions
layout(matrix(c(1, 2, 3, 4, 5, 0), nrow = 2, ncol = 3, byrow = FALSE))
visreg(vLG_Tory.glm, scale = "response", ask = F)
visreg(rG_Tory.glm, scale = "response", ask = F)
visreg(aSG_Tory.glm, scale = "response", ask = F)

# Eulophid interactions and other less abundant ones
layout(matrix(c(1, 2, 3, 4), nrow = 2, ncol = 2, byrow = FALSE))
plot(x = range(full.predictors$vLG_abund), y = c(0, 1), type = "n", ylab = "vLG_Eulo", xlab = "vLG_abund"); abline(h = exp(coef(vLG_Eulo.glm)), col = "steelblue", lwd = 3)
visreg(rG_Eulo.glm, scale = "response", ask = F)
plot(x = range(full.predictors$vLG_abund), y = c(0, 1), type = "n", ylab = "vLG_Mymarid", xlab = "vLG_abund"); abline(h = exp(coef(vLG_Mymarid.glm)), col = "steelblue", lwd = 3)
visreg(rG_Lestodip.glm, scale = "response", ask = F)

par(mfrow = c(1,1))
```

```{r table of jensen effects}
jensen.df <- rbind_all(list(jensen_magnitude(vLG_Platy.glm, no.var.data),
               jensen_magnitude(vLG_Mesopol.glm, no.var.data),
               jensen_magnitude(vLG_Tory.glm, no.var.data),
               jensen_magnitude(vLG_Eulo.glm, no.var.data),
               jensen_magnitude(vLG_Mymarid.glm, no.var.data),
               jensen_magnitude(rG_Tory.glm, no.var.data),
               jensen_magnitude(rG_Mesopol.glm, no.var.data),
               jensen_magnitude(rG_Platy.glm, no.var.data),
               jensen_magnitude(rG_Mesopol.glm, no.var.data),
               jensen_magnitude(rG_Lestodip.glm, no.var.data),
               jensen_magnitude(aSG_Tory.glm, no.var.data),
               jensen_magnitude(SG_Platy.glm, no.var.data)))
jensen.df <- mutate(jensen.df, Interactions = c("vLG_Platy","vLG_Mesopol","vLG_Tory","vLG_Eulo","vLG_Mymarid","rG_Tory","rG_Mesopol","rG_Platy","rG_Mesopol","rG_Lestodip","aSG_Tory","SG_Platy"))
plot(log(Magnitude.Difference) ~ log(mean.no.var), filter(jensen.df, Magnitude.Difference != "1"))
visreg(lm(log(Magnitude.Difference) ~ log(mean.no.var), filter(jensen.df, Magnitude.Difference != "1")), scale = "response", xtrans = log, xlab = "log(mean.no.var)")

interactions.split <- colsplit(jensen.df$Interactions, "_", names = c("gall","parasitoid"))
jensen.df <- cbind(jensen.df, interactions.split)

var.web <- cast(jensen.df, gall ~ parasitoid, mean, value = "mean.with.var")
rownames(var.web) <- var.web$gall
var.web <- as.matrix.data.frame(var.web[ ,-1])

no.var.web <- cast(jensen.df, gall ~ parasitoid, mean, value = "mean.no.var")
rownames(no.var.web) <- no.var.web$gall
no.var.web <- as.matrix.data.frame(no.var.web[ ,-1])

library(bipartite)
plotweb(var.web, method = "normal")
plotweb(no.var.web, method = "normal")
```

```{r remove weak interactions}
colSums(full.predictors[ ,interaxns_noPont]) # remove rG_Lestodip, rG_Mesopol, vLG_Mymarid, and SG_Platy. SG_Platy had 9 interactions, but all of these interactions came from a single, multichambered gall.

net.trait.noweak <- mvabund(full.predictors[ ,c("aSG_Tory", "rG_Eulo", "rG_Tory", "vLG_Eulo", "vLG_Mesopol","vLG_Platy", "vLG_Tory")])

net.mvabund.2.noweak <- manyglm(net.trait.noweak ~ log.vLG_abund + log.vLG.height.mean + 
                         log.1.rG_abund + log.1.aSG_abund,
                       data = full.predictors,
                       family = "negative.binomial")
```

```{r}
## Identify the effect of phenotypic variation (gall density and gall size) on food web structure

## create data with the same mean, but NO VARIANCE in predictor variables. For some reason, the predict.manyglm needs a dataset that is the same dimension as the original one.
row.count <- dim(full.predictors)[1]
no.var.data <- with(full.predictors, 
                  data.frame(log.vLG_abund = rep(log(mean(vLG_abund)), row.count),
                             log.1.rG_abund = rep(log(mean(rG_abund)+1), row.count),
                             log.1.aSG_abund = rep(log(mean(aSG_abund)+1), row.count),
                             log.vLG.height.mean = rep(log(mean(vLG.height.mean)), row.count)))
# increase mean by one standard deviation. With this type of code, I should be able to make predictions for how evolution will affect the food web.
no.var.data.vLG.size.Increase <- with(full.predictors, 
                  data.frame(log.vLG_abund = rep(log(mean(vLG_abund)), row.count),
                             log.1.rG_abund = rep(log(mean(rG_abund)+1), row.count),
                             log.1.aSG_abund = rep(log(mean(aSG_abund)+1), row.count),
                             log.vLG.height.mean = rep(log(mean(vLG.height.mean)+sd(vLG.height.mean)), row.count)))

jensen.net.mvabund.2 <- jensen_magnitude_manyglm(net.mvabund.2, original.data = full.predictors,
                                                 comparison.data = no.var.data)
jensen.net.mvabund.2.noweak <- jensen_magnitude_manyglm(net.mvabund.2.noweak, original.data = full.predictors, comparison.data = no.var.data)
jensen.net.mvabund.2.meanIncrease <- jensen_magnitude_manyglm(net.mvabund.2, original.data = no.var.data, comparison.data = no.var.data.vLG.size.Increase)

# used a wilcoxon signed rank test because the differences were not normally distributed
with(jensen.net.mvabund.2, hist(var.predict - no.var.predict, breaks = seq(0,2,0.01)))
with(jensen.net.mvabund.2, wilcox.test(var.predict, no.var.predict, 
                                       paired = TRUE, conf.int = TRUE))
magnitude.lm <- lm(log(Percent.Difference) ~ log(no.var.predict), jensen.net.mvabund.2)
summary(magnitude.lm); visreg(magnitude.lm, scale = "response", xtrans = log)

# I get the same qualitative results even after removing the weakest interactions from the dataset
with(jensen.net.mvabund.2.noweak, wilcox.test(var.predict, no.var.predict, 
                                              paired = TRUE, conf.int = TRUE)) 
magnitude.lm.noweak <- lm(log(Percent.Difference) ~ log(no.var.predict), 
                          jensen.net.mvabund.2.noweak)
summary(magnitude.lm.noweak); visreg(magnitude.lm.noweak, scale = "response", xtrans = log)
```

```{r effects of variation on food web structure}
library(bipartite)
jensen.to.use <- jensen.net.mvabund.2.noweak
interactions.split <- colsplit(jensen.to.use$response.variables, 
                               "_", names = c("gall","parasitoid"))
jensen.to.use <- cbind(jensen.to.use, interactions.split)

var.web <- cast(jensen.to.use, gall ~ parasitoid, mean, value = "var.predict")
rownames(var.web) <- var.web$gall
var.web <- as.matrix.data.frame(var.web[ ,-1])

no.var.web <- cast(jensen.to.use, gall ~ parasitoid, mean, value = "no.var.predict")
rownames(no.var.web) <- no.var.web$gall
no.var.web <- as.matrix.data.frame(no.var.web[ ,-1])

par(mfrow = c(2,1))
plotweb(var.web, method = "normal")
plotweb(no.var.web, method = "normal")

cbind(networklevel(var.web, 
                   index = c("linkage density", "interaction evenness")),
      networklevel(no.var.web,
                   index = c("linkage density", "interaction evenness")))
```

```{r}
vLG.ptized.glm <- glm(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~ vLG.height.mean, full.df, family = binomial)
summary(vLG.ptized.glm)
visreg(vLG.ptized.glm, scale = "response")
vLG.ptized.no.var.data <- with(full.predictors, data.frame(vLG.height.mean = mean(vLG.height.mean)))
jensen_magnitude(vLG.ptized.glm, no.var.data = vLG.ptized.no.var.data)
#plot(vLG.ptized.glm)

vLG.Platy.glm <- glm(cbind(vLG_Platy, vLG_abund - vLG_Platy) ~ vLG.height.mean, full.df, family = binomial)
summary(vLG.Platy.glm)
visreg(vLG.Platy.glm, scale = "response")
vLG.Platy.no.var.data <- with(full.predictors, data.frame(vLG.height.mean = mean(vLG.height.mean)))
jensen_magnitude(vLG.Platy.glm, no.var.data = vLG.Platy.no.var.data)

vLG.Tory.glm <- glm(cbind(vLG_Tory, vLG_abund - vLG_Tory) ~ vLG.height.mean, full.df, family = binomial)
summary(vLG.Tory.glm)
visreg(vLG.Tory.glm, scale = "response")
vLG.Tory.no.var.data <- with(full.predictors, data.frame(vLG.height.mean = mean(vLG.height.mean)))
jensen_magnitude(vLG.Tory.glm, no.var.data = vLG.Tory.no.var.data)

#plot(vLG_Platy ~ Genotype, full.df)
#plot(vLG_Tory ~ Genotype, full.df)

vLG.ptized.glmer <- glmer(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~ log(vLG.height.mean) + (1|Genotype), full.df, family = binomial)
summary(vLG.ptized.glmer)
visreg(vLG.ptized.glmer, scale = "response")
mean(predict(vLG.ptized.glmer, type = "response"))
#mean(predict(vLG.ptized.glmer, newdata = vLG.ptized.no.var.data, type = "response"))

vLG_Platy.glmer <- glmer(vLG_Platy ~ log(vLG_abund) + log(vLG.height.mean) + (1|Genotype), full.df, family = poisson)
summary(vLG_Platy.glmer)
overdisp_fun(vLG_Platy.glmer) # no overdispersion when I model genotype as a randome effect.
visreg(vLG_Platy.glmer, scale = "response")
```

```{r}
## Examine whether the proportion of galls parasitized varies among willow genotypes

## vLG parasitism
vLG.ptized.df <- full.df %>%
  filter(vLG_abund > 0)
table(vLG.ptized.df$Genotype)
vLG.ptized.3plus.genos <- c("*","B","D","E","F","H","I","K","L","O","S","V","X","Y","Z")
colSums(vLG.ptized.df[ ,c("vLG_parasitized","vLG_abund","vLG_Tory","vLG_Eulo","vLG_Platy",
                         "vLG_Mesopol","vLG_Mymarid")])

# interesting, only marginally significant with quasibinomial
vLG.ptized.bin <- glm(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~ Genotype,
                  data = filter(vLG.ptized.df, Genotype %in% vLG.ptized.3plus.genos), 
                  family = 'binomial')
anova(glm.binomial.disp(vLG.ptized.bin), test = "LR") # just significant, when I account for overdispersion using this method...

vLG.ptized.bin.glmer <- glmer(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~ (1|Genotype),
                  data = filter(vLG.ptized.df, Genotype %in% vLG.ptized.3plus.genos), 
                  family = 'binomial')
overdisp_fun(vLG.ptized.bin.glmer)

test.2 <- rpt.binomGLMM.add(y = with(vLG.ptized.df, cbind(vLG_parasitized, vLG_abund - vLG_parasitized)), groups = vLG.ptized.df$Genotype) 
test.2
library(emdbook)
library(MASS)
vLG.ptized.df <- mutate(vLG.ptized.df, 
                        plant.position = factor(plant.position))
library(lme4)
sc.abund <- scale(vLG.ptized.df$vLG_abund)
sc.height <- scale(vLG.ptized.df$vLG.height.mean)
test.4 <- glmer(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~  vLG.height.mean + (1|Genotype),
                  data = vLG.ptized.df,
                  family = 'binomial')
summary(test.4)
AIC(test.4)
confint(test.4, method = "profile") # variance > 0!
overdisp_fun(test.4) # no clear evidence of overdispersion
ggQQ_ranef(ranef(test.4)$Genotype$'(Intercept)') # and the residuals look great!
visreg(test.4, scale = "response")
R.link  <- .6411 / (.6411 +pi^2 /3) # heritability is reasonable as well
R.link.low  <- .1153673 / (.1153673 +pi^2 /3) 
R.link.high  <- 1.44 / (1.44 +pi^2 /3) 
profile(test.4, which = 1)
summary(test.4)
test.3 <- rpt.binomGLMM.multi(y = with(vLG.ptized.df, cbind(vLG_parasitized, vLG_abund - vLG_parasitized)), groups = vLG.ptized.df$Genotype)

library(RLRsim)

#plot(vLG.ptized.bin)
summary(vLG.ptized.bin)
dfun(vLG.ptized.bin)
anova(vLG.ptized.bin, test = "LR") # still significant after retaining Genotypes with 3 plus replicates.

vLG_ptized.bin.pred <- gam(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~ s(vLG.height.mean),
                  data = full.df, 
                  family = 'binomial')
summary(vLG_ptized.bin.pred)
AIC(vLG_ptized.bin.pred)
visreg(vLG_ptized.bin.pred, scale = "response")
#jensen.single(vLG_ptized.bin.pred)

# vLG_Platy test
vLG_Platy.bin <- glm(cbind(vLG_Platy, vLG_abund - vLG_Platy) ~ Genotype,
                  data = filter(full.df, Genotype %in% vLG.ptized.3plus.genos), 
                  family = 'quasibinomial')
plot(vLG_Platy.bin)
summary(vLG_Platy.bin)
anova(vLG_Platy.bin, test = "F")

library(dispmod)
vLG_Platy.bin.pred <- glm(cbind(vLG_Platy, vLG_abund - vLG_Platy) ~ vLG.height.mean,
                  data = full.predictors, 
                  family = 'binomial')
dfun(vLG_Platy.bin.pred)
summary(vLG_Platy.bin.pred)
glm.binomial.disp(vLG_Platy.bin.pred)
AIC(vLG_Platy.bin.pred)
plot(vLG_Platy.bin.pred)
visreg(vLG_Platy.bin.pred, scale = "response")
jensen.single(vLG_Platy.bin.pred)
glm.binomial.disp(vLG_Platy.bin.pred)
qAIC(vLG_Platy.bin.pred, dispersion = dfun(vLG_Platy.bin.pred))

dfun(vLG_Platy.bin.pred)

vLG_Platy.bin.pred.glmer <- glmer(cbind(vLG_Platy, vLG_abund - vLG_Platy) ~ vLG.height.mean + (1|Genotype),
                  data = full.df, 
                  family = 'binomial')
summary(vLG_Platy.bin.pred.glmer)
overdisp_fun(vLG_Platy.bin.pred.glmer)
AIC(vLG_Platy.bin.pred.glmer)
plot(vLG_Platy.bin.pred.glmer)
ggQQ_ranef(ranef(vLG_Platy.bin.pred.glmer)$Genotype$'(Intercept)') # not too bad.

# AIC gam interaction 207.6398, AIC = 204 with main effect model in gam. 213 with main model

## Identify the effect of phenotypic variation (gall density and gall size) on food web structure

jensen.single(final.vLG_Platy)
jensen.single(vLG_Mesopol.bin.pred)
jensen.single(vLG_Tory.bin.pred)

# vLG_Mesopol test
vLG_Mesopol.bin <- glm(cbind(vLG_Mesopol, vLG_abund - vLG_Mesopol) ~ Genotype,
                  data = filter(full.df, Genotype %in% vLG.ptized.3plus.genos), 
                  family = 'binomial')
plot(vLG_Mesopol.bin)
summary(vLG_Mesopol.bin)
dfun(vLG_Mesopol.bin)
anova(vLG_Mesopol.bin, test = "LR")

vLG_Mesopol.bin.pred <- gam(cbind(vLG_Mesopol, vLG_abund - vLG_Mesopol) ~ vLG.height.mean,
                  data = full.df, 
                  family = 'quasibinomial')
summary(vLG_Mesopol.bin.pred)
visreg(vLG_Mesopol.bin.pred, xvar = "vLG.height.mean", scale = "response")
dfun(vLG_Mesopol.bin.pred)

jensen.single(vLG_Mesopol.bin.pred)

# vLG_Tory test
vLG_Tory.bin <- glm(cbind(vLG_Tory, vLG_abund - vLG_Tory) ~ Genotype,
                  data = filter(full.df, Genotype %in% vLG.ptized.3plus.genos), 
                  family = 'binomial')
plot(vLG_Tory.bin)
summary(vLG_Tory.bin)
anova(vLG_Tory.bin, test = "LR")

vLG_Tory.bin.pred <- glm(cbind(vLG_Tory, vLG_abund - vLG_Tory) ~ vLG_abund + vLG.height.mean,
                  data = full.df, 
                  family = 'binomial')
summary(vLG_Tory.bin.pred)
AIC(vLG_Tory.bin.pred) # 136.1538 is the lowest score. Main effect model
visreg(vLG_Tory.bin.pred, scale = "response")

# vLG_Eulo test
vLG_Eulo.bin <- glm(cbind(vLG_Eulo, vLG_abund - vLG_Eulo) ~ Genotype,
                  data = filter(full.df, Genotype %in% vLG.ptized.3plus.genos), 
                  family = 'binomial')
#plot(vLG_Eulo.bin)
summary(vLG_Eulo.bin)
anova(vLG_Eulo.bin, test = "LR") # not significant

# not testing vLG_Mymarid because it is in such low abundance
predict.df <- data.frame(Genotype = vLG.ptized.3plus.genos)
vLG.genotype.predictions <- cbind.data.frame(predict.df, 
                                         vLG.Proportion.Ptized = predict(vLG.ptized.bin,
                                                                               newdata = predict.df,
                                                                               type = "response"))
summary(vLG.genotype.predictions)

ggplot(filter(vLG.ptized.df, Genotype %in% vLG.ptized.3plus.genos),
       aes(x = Genotype, y = vLG_parasitized/vLG_abund)) +
  geom_point(aes(size = vLG_abund), color = "grey", shape = 1,
             position = position_jitter(height = 0, width = 0.1)) + 
  scale_size(range = c(3,15)) + 
  geom_point(data = vLG.genotype.predictions, 
             aes(x = Genotype, y = vLG.Proportion.Ptized),
             color = "steelblue", size = 30, shape = "--") +
  theme_classic() +
  ylab("Proportion of galls parasitized")
range(filter(vLG.ptized.df, Genotype %in% vLG.ptized.3plus.genos)$vLG_abund)

## rG parasitism
rG.ptized.df <- full.df %>%
  filter(rG_abund > 0)
table(rG.ptized.df$Genotype)
rG.ptized.3plus.genos <- c("*","A","D", "F", "G", "H","I","K","O","Q", "R", "S","X", "Y", "Z") 
colSums(rG.ptized.df[ ,c("rG_parasitized","rG_abund","rG_Tory","rG_Eulo","rG_Platy",
                         "rG_Mesopol","rG_Lestodip")]) # too few rG_eulo, rG_Platy, rG_Mesopol, and rG_Lestodip to test for differences in parasitism among genotypes.

rG.ptized.bin <- glm(cbind(rG_parasitized, rG_abund - rG_parasitized) ~ Genotype,
                  data = filter(full.df, Genotype %in% rG.ptized.3plus.genos), 
                  family = 'binomial')
#plot(rG.ptized.bin) # residuals not great
summary(rG.ptized.bin)
anova(rG.ptized.bin, test = "LR") # marginally significant after retaining Genotypes with 3 plus replicates.

# rG_Tory test
rG_Tory.bin <- glm(cbind(rG_Tory, rG_abund - rG_Tory) ~ Genotype,
                  data = filter(full.df, Genotype %in% rG.ptized.3plus.genos), 
                  family = 'binomial')
plot(rG_Tory.bin)
summary(rG_Tory.bin)
anova(rG_Tory.bin, test = "LR") # not significant

rG_Eulo.bin <- glm(cbind(rG_Eulo, rG_abund - rG_Eulo) ~ Genotype,
                  data = filter(full.df, Genotype %in% rG.ptized.3plus.genos), 
                  family = 'binomial')
plot(rG_Eulo.bin)
summary(rG_Eulo.bin)
anova(rG_Eulo.bin, test = "LR") # not significant

predict.df <- data.frame(Genotype = rG.ptized.3plus.genos)
rG.genotype.predictions <- cbind.data.frame(predict.df, 
                                         rG.Proportion.Ptized = predict(rG.ptized.bin,
                                                                               newdata = predict.df,
                                                                               type = "response"))
summary(rG.genotype.predictions)

ggplot(filter(rG.ptized.df, Genotype %in% rG.ptized.3plus.genos),
       aes(x = Genotype, y = rG_parasitized/rG_abund)) +
  geom_point(aes(size = rG_abund), color = "grey", shape = 1,
             position = position_jitter(height = 0, width = 0.1)) + 
  scale_size(range = c(3,15)) + 
  geom_point(data = rG.genotype.predictions, 
             aes(x = Genotype, y = rG.Proportion.Ptized),
             color = "steelblue", size = 30, shape = "--") +
  theme_classic() +
  ylab("Proportion of galls parasitized")
range(filter(rG.ptized.df, Genotype %in% rG.ptized.3plus.genos)$rG_abund)

## aSG parasitism
aSG.ptized.df <- full.df %>%
  filter(aSG_abund > 0)
table(aSG.ptized.df$Genotype)
aSG.ptized.3plus.genos <- c("G","I","K","Q")

aSG.ptized.bin <- glm(cbind(aSG_parasitized, aSG_abund - aSG_parasitized) ~ Genotype,
                  data = filter(full.df, Genotype %in% aSG.ptized.3plus.genos), 
                  family = 'binomial')
plot(aSG.ptized.bin) # residuals not great
summary(aSG.ptized.bin)
anova(aSG.ptized.bin, test = "LR") # marginally significant after retaining Genotypes with 3 plus replicates.
predict.df <- data.frame(Genotype = aSG.ptized.3plus.genos)
aSG.genotype.predictions <- cbind.data.frame(predict.df, 
                                         aSG.Proportion.Ptized = predict(aSG.ptized.bin,
                                                                               newdata = predict.df,
                                                                               type = "response"))
summary(aSG.genotype.predictions)

ggplot(filter(aSG.ptized.df, Genotype %in% aSG.ptized.3plus.genos),
       aes(x = Genotype, y = aSG_parasitized/aSG_abund)) +
  geom_point(aes(size = aSG_abund), color = "grey", shape = 1,
             position = position_jitter(height = 0, width = 0.1)) + 
  scale_size(range = c(3,15)) + 
  geom_point(data = aSG.genotype.predictions, 
             aes(x = Genotype, y = aSG.Proportion.Ptized),
             color = "steelblue", size = 30, shape = "--") +
  theme_classic() +
  ylab("Proportion of galls parasitized")
range(filter(aSG.ptized.df, Genotype %in% aSG.ptized.3plus.genos)$aSG_abund)

## SG parasitism
SG.ptized.df <- full.df %>%
  filter(SG_abund > 0)
table(SG.ptized.df$Genotype) # no Genotypes with 3+ replicates, so I'm unable to evaluate whether SG parasitism defers among willow genotypes.

##
uni.check.bin <- glm(vLG_Platy/vLG_abund ~ vLG_abund*vLG.height.mean, full.predictors,
                     family = binomial,
                     weights = vLG_abund)
AIC(uni.check.bin)
summary(uni.check.bin)
visreg(uni.check.bin, xvar = "vLG_abund", by = "vLG.height.mean", scale = "response")

# NEED TO THINK ABOUT HOW I SCALE THIS UP TO DIFFERENT NUMBERS OF SHOOTS SAMPLED.

#### messing around with different ideas.
## Create mvabund model based on probability of observing an interaction
prob.df <- full.predictors %>%
  mutate(succ.vLG_Platy.succ = vLG_Platy,
         fail.vLG_Platy.fail = vLG_abund - vLG_Platy,
         succ.vLG_Tory.succ = vLG_Tory,
         fail.vLG_Tory.fail = vLG_abund - vLG_Tory,
         succ.vLG_Mesopol.succ = vLG_Mesopol,
         fail.vLG_Mesopol.fail = vLG_abund - vLG_Mesopol,
         succ.vLG_Eulo.succ = vLG_Eulo,
         fail.vLG_Eulo.fail = vLG_abund - vLG_Eulo,
         succ.rG_Platy.succ = rG_Platy,
         fail.rG_Platy.fail = rG_abund - rG_Platy,
         succ.rG_Tory.succ = rG_Tory,
         fail.rG_Tory.fail = rG_abund - rG_Tory,
         succ.rG_Mesopol.succ = rG_Mesopol,
         fail.rG_Mesopol.fail = rG_abund - rG_Mesopol,
         succ.rG_Eulo.succ = rG_Eulo,
         fail.rG_Eulo.fail = rG_abund - rG_Eulo,
         succ.SG_Platy.succ = SG_Platy,
         fail.SG_Platy.fail = SG_abund - SG_Platy,
         succ.aSG_Tory.succ = aSG_Tory,
         fail.aSG_Tory.fail = aSG_abund - aSG_Tory) %>%
  select(succ.vLG_Platy.succ:fail.aSG_Tory.fail) %>%
  mvabund()

vLG.interaction.df <- mvabund(select(full.predictors, vLG_Eulo, vLG_Platy, vLG_Mymarid, vLG_Tory, vLG_Mesopol))
vLG.interaction.df[vLG.interaction.df > 0] <- 1

test <- manyglm(vLG.interaction.df ~ vLG_density*vLG.height.mean, data = full.predictors, family = "binomial")
plot(test)
anova(test, p.uni = "adjusted")
length(prob.df[prob.df > 1] > 0)

test <- MASS::glm.nb(vLG_Platy ~ offset(log(vLG_abund)) + vLG.height.mean, data = full.predictors) #
summary(test)
coef(test)
AIC(test)
#exp(sum(-4.5 + .34*5 -0.25*8))
exp(sum(coef(test)[c(1:3)]*c(1,0.02,8))) # 8 mm gall at a density of 11 per 100 shoots. Would be 6.11 parasitoids per 100 shoots.
predict(test, type = "response")*10
visreg(test, scale = "response")
```

```{r, functions, echo=FALSE}
kable.adonis <- function(adonis.object){
  require(knitr)
  kable(adonis.object$aov.tab[ ,-5], # remove R2 calculation
        col.names = c("df","SS","MS","F","P"), 
        digits = c(0,2,2,2,3))
}

heritability.adonis <- function(adonis.object, data){
  gen.row <- which(rownames(adonis.object$aov.tab) == "Genotype")
  res.row <- which(rownames(adonis.object$aov.tab) == "Residuals")
  gen.MS <- adonis.object$aov.tab$MeanSqs[gen.row]
  res.MS <- adonis.object$aov.tab$MeanSqs[res.row]
  mean.sample.size <- mean(table(data$Genotype))
  gen.variance <- (gen.MS - res.MS)/mean.sample.size
  heritability <- gen.variance/(gen.variance + res.MS) # note that residual MS is equal to the residual variance
  print(heritability)
}

heritability <- function(merMod.object){
  require(lme4)
  require(RLRsim)

  var.comps <- as.data.frame(VarCorr(merMod.object))
  genetic.var <- var.comps$vcov[1]
  residual.var <- var.comps$vcov[2]
  heritability <- genetic.var/(genetic.var + residual.var)
  restricted.likelihood.ratio.test <- exactRLRT(merMod.object)
  data.frame(#Model = formula(merMod.object),
             Heritability = round(heritability, 2),
             RLRT_statistic = round(restricted.likelihood.ratio.test$statistic, 2),
             P = round(restricted.likelihood.ratio.test$p.value, 3))
}

jensen.single <- function(model.object, predict.type = "response"){
  model <- model.object

  var.predict <- predict(model.object, type = predict.type)

  predictor.names <- names(model.object$model)[-1] # omits response variable
  data <- data.frame(model.object$model[ ,predictor.names])

  no.var.data <- list()
  for(i in 1:length(data)){
    no.var.data[[i]] <- mean(data[ ,i])
  }
  no.var.data <- as.data.frame(no.var.data)
  names(no.var.data) <- predictor.names

  no.var.predict <- predict(model.object, newdata = no.var.data, type = predict.type)

  data.frame(mean.with.var = mean(var.predict, na.rm = TRUE),
             mean.no.var = mean(no.var.predict, na.rm = TRUE))
}

jensen.single <- function(model.object, no.var.data, predict.type = "response"){
  model <- model.object

  var.predict <- predict(model.object, type = predict.type)

  predictor.names <- names(model.object$model)[-1] # omits response variable
  data <- data.frame(model.object$model[ ,predictor.names])

  #no.var.data <- list()
  #for(i in 1:length(data)){
   # no.var.data[[i]] <- mean(data[ ,i])
  #}
  #no.var.data <- as.data.frame(no.var.data)
  #names(no.var.data) <- predictor.names

  no.var.predict <- predict(model.object, newdata = no.var.data, type = predict.type)

  data.frame(mean.with.var = mean(var.predict, na.rm = TRUE),
             mean.no.var = mean(no.var.predict, na.rm = TRUE))
}

```

```{r}
library(psych)
gall.df.cors <- full.df %>%
  mutate(vLG_density = log(vLG_abund+1) - log(shootEst.no18),
         vLG_density2 = log(vLG_abund/shootEst.no18+0.01),
         vLG_density3 = sqrt(vLG_abund/shootEst.no18),
         rG_density = log(vLG_abund+1) - log(shootEst.no18),
         rG_density2 = log(rG_abund/shootEst.no18+0.01),
         rG_density3 = sqrt(rG_abund/shootEst.no18)) %>%
  select(vLG.height.mean, vLG_density:rG_density3)
trait.df.cors <- full.df %>%
  select(Total_Area:flavanonOLES.PC1)

gall.trait.cors <- corr.test(x = trait.df.cors, y = gall.df.cors)
round(gall.trait.cors$r,2)
# further examine all correlations greater than 0.20
# vLG.height.mean: salicortin__A270nm, luteolin.glucoside.5.glu.__A320nm, quercetyin.der1, luteolin.der1, luteolin der2, termulacin, Carbon, sal_tannin.PC1, flavonoOLES.PC1
# vLG_density: Total_Area, Height, D_mean_smoothed, ampelopsin.der__A320nm, myricitrin__A320nm, eriodictyol.7.glucoside__A270nm, luteolin.der1__A320nm, C_N_imputed (sal_tannin.PC1? a bit low). Note though the Nitrogen and C_N_ratio are not good predictors for the "un_imputed" data. 
# rG_density: total_area, height, D_mean_smoothed, ampelopsin.der__A320nm, eriodictyol.7.glucoside__A270nm, luteolin.der1__A320nm, C_N_imputed

library(car)
# need to work on code below.
scatterplotMatrix(cbind.data.frame(gall.df.cors$vLG.height.mean, 
                                   trait.df.cors[ ,"salicortin__A270nm", 
                                                 "luteolin.glucoside.5.glu.__A320nm",
                                                 "quercetin.der1__A320nm", 
                                                 "luteolin.der1__A320nm", 
                                                 "luteolin der2__A320nm", 
                                                 "termulacin__A320nm",
                                                 "Carbon", "sal_tannin.PC1", 
                                                 "flavonoOLES.PC1"]))
```

```{r}
interaction_abund <- rowSums(full.df[ ,interaxns_noPont])
interaction_richness <- rowSums(full.df[ ,interaxns_noPont] > 0)
ggplot(full.df, aes(x = Genotype, y = interaction_abund)) +
  geom_boxplot()

ggplot(full.df, aes(x = Genotype, y = interaction_abund/shootEst.no18)) +
  geom_boxplot() # G, H, M, N, O, P, R, U all have 75% of their observations at zero for interaction density

ggplot(full.df, aes(x = Genotype, y = interaction_richness)) +
  geom_boxplot() # G, H, M, N, O, P, R, U all have 75% of their observations at zero for interaction richness

non.zero.genos <- c("*","A","B","D","E","F","I","J","K","L","Q","S","T","V","W","X","Y","Z")

df.non.zero.genos <- filter(full.df, Genotype %in% non.zero.genos)
hist(log(rowSums(df.non.zero.genos[ ,interaxns_noPont]) +1))
interaction_abund.non.zero.genos <- rowSums(df.non.zero.genos[ ,interaxns_noPont])
library(lme4)
interaction.abund.lmer <- lmer(log(interaction_abund+1) ~ offset(log(shootEst.no18)) + (1|Genotype), full.df)
heritability(interaction.abund.lmer) # ~0.28 with non.zero.genos. 0.33 with full dataset, but the residuals look pretty good actually. Actually, setting an offset and log(x+1) transforming the response variable seems to do a pretty good job of making the random effect normally distributed.
ggQQ_ranef(ranef(interaction.abund.lmer)$Genotype$'(Intercept)')
hist(ranef(interaction.abund.lmer)$Genotype$'(Intercept)')
```

Table 1: Quantitative full data set with euclidean distance
```{r, results='asis', echo=FALSE}
full.adonis <- adonis(log(full.df.quant+1) ~ log(shootEst.no18) + Genotype, 
                      data = full.df, method = "euclidean")
heritability.adonis(full.adonis, full.df)
kable.adonis(full.adonis)
```

Table S2: Qualitative full data set with euclidean distance
```{r, results='asis', echo=FALSE}
full.adonis.qual <- adonis(full.df.qual ~ log(shootEst.no18) + Genotype, 
                      data = full.df, method = "euclidean")
heritability.adonis(full.adonis.qual, full.df)
kable.adonis(full.adonis.qual)
```

Table S3: Qualitative full data set with Raup-crick null model
```{r, results='asis', echo=FALSE}
rc.dist.full <- raupcrick(full.df.qual) # calculate raupcrick dissimilarity for teasing apart differences in alpha vs. beta-diversity
rc.dist.full.adonis <- adonis(rc.dist.full ~ log(shootEst.no18) + Genotype, 
                      data = full.df)
heritability.adonis(rc.dist.full.adonis, full.df)
kable.adonis(rc.dist.full.adonis)
```

Table S3: Qualitative sub data set with Raup-crick null model
```{r, results='asis', echo=FALSE}
rc.dist.sub <- raupcrick(sub.df.qual) # calculate raupcrick dissimilarity for teasing apart differences in alpha vs. beta-diversity
rc.dist.sub.adonis <- adonis(rc.dist.sub ~ log(shootEst.no18) + Genotype, 
                      data = sub.df)
heritability.adonis(rc.dist.sub.adonis, sub.df)
kable.adonis(rc.dist.sub.adonis)
```

Table S2: Quantitative data subset, euclidean
```{r, results='asis', echo=FALSE}
sub.adonis.euclid <- adonis(log(sub.df.quant+1) ~ log(shootEst.no18) + Genotype, 
                            data = sub.df, method = "euclidean")
kable.adonis(sub.adonis.euclid)
```

Table S3: Quantitative data subset, with bray-curtis dissimilarity
```{r, results='asis', echo=FALSE}
sub.adonis.bray <- adonis(sub.df.quant ~ log(shootEst.no18) + Genotype, 
                          data = sub.df, method = "bray")
kable.adonis(sub.adonis.bray)
```

Table S3: Qualitative data subset, with Jaccard dissimilarity
```{r, results='asis', echo=FALSE}
sub.adonis.jac <- adonis(sub.df.qual ~ log(shootEst.no18) + Genotype, 
                          data = sub.df, method = "jaccard")
heritability.adonis(adonis.object = sub.adonis.jac, data = sub.df)
kable.adonis(adonis.object = sub.adonis.jac)
```

Table S4: Qualitative data subset, with euclidean dissimilarity
```{r, results='asis', echo=FALSE}
sub.adonis.euc <- adonis(sub.df.qual ~ log(shootEst.no18) + Genotype, 
                          data = sub.df, method = "euclidean")
kable.adonis(adonis.object = sub.adonis.euc)
```

Heritability of focal gall-parasitoid guild interactions
```{r gall-ptoid heritability, echo=FALSE}
## Variation among genotypes
gall.ptoid.variation <- cbind(Genotype = full.df$Genotype, full.df[ ,c("vLG_egg","vLG_ecto","rG_ecto")]/full.df$shootEst.no18) %>%
  group_by(Genotype) %>%
  select(vLG_egg, vLG_ecto, rG_ecto) %>%
  summarise_each(funs(mean))

max(filter(gall.ptoid.variation, vLG_egg > 0)$vLG_egg)/min(filter(gall.ptoid.variation, vLG_egg > 0)$vLG_egg) # 35-fold variation in vLG_egg parasitism among genotypes that hosted this interaction

max(filter(gall.ptoid.variation, vLG_ecto > 0)$vLG_ecto)/min(filter(gall.ptoid.variation, vLG_ecto > 0)$vLG_ecto) # 20-fold variation in vLG_ecto parasitism among genotypes that hosted this interaction

max(filter(gall.ptoid.variation, rG_ecto > 0)$rG_ecto)/min(filter(gall.ptoid.variation, rG_ecto > 0)$rG_ecto) # 9-fold variation in vLG_ecto parasitism among genotypes that hosted this interaction

## Iteomyia (vLG) - Egg parasitoid 
vLG_egg.lmer.full <- lmer(log(vLG_egg + 1) ~ log(shootEst.no18) + (1|Genotype), 
                          data = full.df)
heritability(vLG_egg.lmer.full)
ggQQ_ranef(ranef(vLG_egg.lmer.full)$Genotype$"(Intercept)") # non-normal random effect

vLG_egg.lmer.sub <- lmer(log(vLG_egg + 1) ~ log(shootEst.no18) + (1|Genotype), 
                         data = sub.df)
heritability(vLG_egg.lmer.sub)
ggQQ_ranef(ranef(vLG_egg.lmer.sub)$Genotype$"(Intercept)") # normal random effect

## Iteomyia (vLG) - Larval parasitoid (ecto)
vLG_ecto.lmer.full <- lmer(log(vLG_ecto + 1) ~ log(shootEst.no18) + (1|Genotype), 
                          data = full.df)
heritability(vLG_ecto.lmer.full)
ggQQ_ranef(ranef(vLG_ecto.lmer.full)$Genotype$"(Intercept)") # close to normal random effect

vLG_ecto.lmer.sub <- lmer(log(vLG_ecto + 1) ~ log(shootEst.no18) + (1|Genotype), 
                         data = sub.df)
heritability(vLG_ecto.lmer.sub) # marginally significant
ggQQ_ranef(ranef(vLG_ecto.lmer.sub)$Genotype$"(Intercept)") # normal random effect

## Rabdophaga salicisbrassicoides (rG) - Larval parasitoid (ecto)
rG_ecto.lmer.full <- lmer(log(rG_ecto + 1) ~ log(shootEst.no18) + (1|Genotype), 
                          data = full.df)
heritability(rG_ecto.lmer.full)
ggQQ_ranef(ranef(rG_ecto.lmer.full)$Genotype$"(Intercept)") # non-normal random effect

rG_ecto.lmer.sub <- lmer(log(rG_ecto + 1) ~ log(shootEst.no18) + (1|Genotype), 
                         data = sub.df)
heritability(rG_ecto.lmer.sub)
ggQQ_ranef(ranef(rG_ecto.lmer.sub)$Genotype$"(Intercept)") # close to normal random effect
```

Gall density and gall size variation among willow genotypes
```{r}
gall.density.variation <- cbind(Genotype = full.df$Genotype, full.df[ ,c("vLG_abund","rG_abund")]/full.df$shootEst.no18) %>%
  group_by(Genotype) %>%
  select(vLG_abund, rG_abund) %>%
  summarise_each(funs(mean))

max(filter(gall.density.variation, vLG_abund > 0)$vLG_abund)/min(filter(gall.density.variation, vLG_abund > 0)$vLG_abund) # 67-fold variation in Iteomyia among willow genotypes

max(filter(gall.density.variation, rG_abund > 0)$rG_abund)/min(filter(gall.density.variation, rG_abund > 0)$rG_abund) # 62-fold variation in Rabdophaga-bud among willow genotypes.

vLG.lmer.full <- lmer(log(vLG_abund + 1) ~ log(shootEst.no18) + (1|Genotype),
                      data = full.df)
heritability(merMod.object = vLG.lmer.full)

rG.lmer.full <- lmer(log(rG_abund + 1) ~ log(shootEst.no18) + (1|Genotype),
                      data = full.df)
heritability(merMod.object = rG.lmer.full)

## vLG size. To test this, I took advantage of the full gall data set and used a nested random effect model.
vLG.size.df <- filter(gall.size.df, gall.sp == "vLG")
table(vLG.size.df$Genotype) # large heterogeneity in sample sizes
table(vLG.size.df$plant.position) # large heterogeneity in sample sizes

vLG.size.lmer <- lmer(gall.height ~ 1 + (1 | Genotype) + (1 |plant.position:Genotype),
                      data = vLG.size.df)
vLG.size.lmer.gen <- update(vLG.size.lmer, .~. - (1 | plant.position:Genotype)) # only Genotype as random effect
vLG.size.lmer.pp <- update(vLG.size.lmer, .~. - (1 | Genotype)) # only plant.position as random effect

vLG.size.lmer.sum <- summary(vLG.size.lmer)
summary(vLG.size.lmer.gen)
summary(vLG.size.lmer.pp)

gen.var <- vLG.size.lmer.sum$varcor$Genotype[1] # Variance due to Genotype
pp.var <- vLG.size.lmer.sum$varcor$'plant.position:Genotype'[1] # variance due to plant position
res.var <- 4.1133 # residual variance - don't know how to extract this from summary object.

H2.vLG.size <- gen.var/(gen.var + pp.var + res.var) # 0.15
exactRLRT(m = vLG.size.lmer.gen, mA = vLG.size.lmer, m0 = vLG.size.lmer.pp) # this model tests whether

## rG size. To test this, I took advantage of the full gall data set and used a nested random effect model.
rG.size.df <- filter(gall.size.df, gall.sp == "rG")
table(rG.size.df$Genotype) # large heterogeneity in sample sizes
table(rG.size.df$plant.position) # large heterogeneity in sample sizes

rG.size.lmer <- lmer(gall.height ~ 1 + (1 | Genotype) + (1 |plant.position:Genotype),
                      data = rG.size.df)
rG.size.lmer.gen <- update(rG.size.lmer, .~. - (1 | plant.position:Genotype)) # only Genotype as random effect
rG.size.lmer.pp <- update(rG.size.lmer, .~. - (1 | Genotype)) # only plant.position as random effect

rG.size.lmer.sum <- summary(rG.size.lmer)
summary(rG.size.lmer.gen)
summary(rG.size.lmer.pp)

gen.var <- rG.size.lmer.sum$varcor$Genotype[1] # Variance due to Genotype
pp.var <- rG.size.lmer.sum$varcor$'plant.position:Genotype'[1] # variance due to plant position
res.var <- 1.80518 # residual variance - don't know how to extract this from summary object.

H2.rG.size <- gen.var/(gen.var + pp.var + res.var) # 0.0418
exactRLRT(m = rG.size.lmer.gen, mA = rG.size.lmer, m0 = rG.size.lmer.pp) # this model tests whether 
```

RDA analysis of community response to Iteomyia density and size as well as R. salicisbrassicoides density
```{r RDA, echo=FALSE}
full.rda.df <- full.df %>%
  filter(vLG.height.mean > 0) %>% 
  mutate(vLG_density = vLG_abund/shootEst.no18,
         rG_density = rG_abund/shootEst.no18)
full.rda.df.interaxn <- full.rda.df[ ,interaxns_noPont]
full.rda.df.interaxn.qual <- ifelse(full.rda.df[ ,interaxns_noPont] > 0, 1, 0)

full.rda.df.interaxn.log1 <- log(full.rda.df.interaxn+1)
full.rda.df.interaxn.log1.offset <- full.rda.df.interaxn.log1 - log(full.rda.df$shootEst.no18) # this is the same as applying an "offset" to all of the data. In other words, to interpret it in terms of density

rda.full <- rda(full.rda.df.interaxn.log1.offset ~  
                  vLG_density*vLG.height.mean + rG_density, data = full.rda.df)
RsquareAdj(rda.full)
anova(rda.full)
anova(rda.full, by = "margin")

rda.traits.df <- na.omit(data.frame(full.rda.df.interaxn, 
                                  shootEst.no18 = full.rda.df$shootEst.no18,
                                  C_N_imputed = full.rda.df$C_N_imputed, 
                                  Height = full.rda.df$Height, 
                                  sal_tannin.PC1 = full.rda.df$sal_tannin.PC1, 
                                  flavonOLES.PC1 = full.rda.df$flavonOLES.PC1))
rda.full.traits <- rda(log(rda.traits.df[ ,interaxns_noPont]+1) ~ 
                         Condition(log(shootEst.no18)) + 
                         sal_tannin.PC1, data = rda.traits.df)
RsquareAdj(rda.full.traits)
anova(rda.full.traits)
anova(rda.full.traits, by = "margin")
vif.cca(rda.full.traits)
plot(rda.full.traits)

rda.full.qual <- rda(full.rda.df.interaxn.qual ~  
                  vLG_density + vLG.height.mean + rG_density, data = full.rda.df)
RsquareAdj(rda.full.qual)
anova(rda.full.qual)
anova(rda.full.qual, by = "margin")

pos.interactions <- which(rowSums(full.rda.df.interaxn.qual) > 0)
rda.full.qual.jac <- capscale(full.rda.df.interaxn.qual[pos.interactions, ] ~  
                  vLG_density+vLG.height.mean + rG_density, 
                  data = full.rda.df[pos.interactions, ], distance = "jaccard")
RsquareAdj(rda.full.qual.jac)
plot(rda.full.qual.jac)
anova(rda.full.qual.jac)
anova(rda.full.qual.jac, by = "margin")
vif.cca(rda.full.qual.jac)

## vLG-egg parasitoid response
vLG_egg.lm <- lm(log(vLG_egg + 1) ~ offset(log(shootEst.no18)) + 
                   vLG_density*vLG.height.mean,
                 data = full.rda.df)
summary(vLG_egg.lm)
plot(vLG_egg.lm)
library(visreg)
visreg(vLG_egg.lm, xvar = "vLG_density", by = "vLG.height.mean")

## vLG-ecto parasitoid response
vLG_ecto.lm <- lm(log(vLG_ecto+1) ~ offset(log(shootEst.no18)) + vLG_density + vLG.height.mean,
                 data = full.rda.df)
summary(vLG_ecto.lm)
plot(vLG_ecto.lm)

## rG-ecto parasitoid response
rG_ecto.lm <- lm(log(rG_ecto+1) ~ offset(log(shootEst.no18)) + rG_density,
                 data = full.rda.df)
summary(rG_ecto.lm)

```

```{r functional responses}
library(ggplot2)
full.rda.df.small.galls <- filter(full.rda.df, vLG.height.mean < 8)
ggplot(full.rda.df.small.galls, aes(x = vLG_density, y = vLG_egg/vLG_abund)) +
  geom_point(aes(size = vLG_abund)) +
  stat_smooth(method = "glm", family = binomial, aes(weight = full.rda.df.small.galls$vLG_abund))

vLG_egg.bin <- glm(cbind(vLG_egg, vLG_abund - vLG_egg) ~ vLG_density*vLG.height.mean, 
                   data = full.rda.df, family = binomial)
summary(vLG_egg.bin)
anova(update(vLG_egg.bin, .~. - vLG_density*vLG.height.mean), vLG_egg.bin, test = "Chi")

plot(vLG_egg.bin)
visreg(vLG_egg.bin, xvar = "vLG_density", by = "vLG.height.mean", scale = "response")
visreg2d(vLG_egg.bin, x = "vLG.height.mean", y = "vLG_density", scale = "response")

vLG_ecto.bin <- glm(cbind(vLG_ecto, vLG_abund - vLG_ecto) ~ vLG_density + vLG.height.mean,
                    data = full.rda.df, family = binomial)
summary(vLG_ecto.bin)
anova(update(vLG_ecto.bin, .~. - vLG_density - vLG.height.mean), vLG_ecto.bin, test = "Chi")
plot(vLG_ecto.bin)
visreg(vLG_ecto.bin, scale = "response")
visreg2d(vLG_ecto.bin, x = "vLG.height.mean", y = "vLG_density", scale = "response")

rG_ecto.bin <- glm(cbind(rG_ecto, rG_abund - rG_ecto) ~ I(rG_abund/shootEst.no18),
                    data = full.df, family = binomial)
summary(rG_ecto.bin)
anova(update(rG_ecto.bin, .~. - I(rG_abund/shootEst.no18)), rG_ecto.bin, test = "Chi")
plot(rG_ecto.bin)
```

```{r}
# try mantel test
data.for.mantel <- na.omit(select(full.df, plant.position, shootEst.no18,
                                  aSG_Tory:rG_Platy, rG_Tory, SG_Platy, vLG_Eulo:vLG_Tory, # all interactions
                                  Total_Area:HCH.tremulacin.__A220nm, water_content,
                                  specific_leaf_area, C_N_imputed))
traits.stand <- decostand(select(data.for.mantel, Total_Area:C_N_imputed), method = "standardize")
traits.dist <- vegdist(traits.stand, method = "euclidean")
network.stand <- (select(data.for.mantel, aSG_Tory:vLG_Tory)/data.for.mantel$shootEst.no18)
network.dist <- vegdist(network.stand, method = "euclidean")

network.for.dist <- mantel(xdis = traits.dist, ydis = network.dist)

```

Gall density mechanisms
```{r}
traits <- full.df %>%
  select(plant.position,
         Total_Area, Height, Density, # architecture traits
         Trichome.No., specific_leaf_area, water_content, C_N_imputed, # other leaf quality traits
         sal_tannin.PC1, cinn.PC1, cinn.PC2, # salicylates/tannins and phenolic acids
         flavonOLES.PC1, flavonOLES.PC2, flavanonOLES.PC1) %>% # flavonoids
  mutate(log_size = log(Total_Area),
         log_trichomes = log(Trichome.No.+1))
traits <- na.omit(traits)
height_resid <- residuals(lm(Height ~ log_size, traits))
density_resid <- residuals(lm(Density ~ log_size, traits))
sla_resid <- residuals(lm(specific_leaf_area ~ water_content, traits))

traits.df <- cbind.data.frame(select(traits, plant.position, water_content:log_trichomes), 
                              height_resid, density_resid, sla_resid)

trait.names <- colnames(traits.df)[-1]

gall.density.size.df <- full.df %>% 
  select(vLG_abund, rG_abund, shootEst.no18, plant.position, vLG.height.mean) %>%
  mutate(vLG_density = vLG_abund/shootEst.no18,
         rG_density = rG_abund/shootEst.no18) %>%
  select(plant.position:rG_density)

gall.density.size.traits.df <- left_join(gall.density.size.df, traits.df)
vLG.df <- na.omit(select(gall.density.size.traits.df, vLG_density, water_content:sla_resid))
vLG.size.df <- na.omit(select(gall.density.size.traits.df, vLG.height.mean, water_content:sla_resid))
rG.df <- na.omit(select(gall.density.size.traits.df, rG_density, water_content:sla_resid))
#focal.df <- na.omit(gall.density.size.traits.df)
#focal.df.traits <- focal.df[ ,trait.names]

# Iteomyia salicisverruca (vLG) abundance model
vLG.dens.glm <- glm(vLG_abund > 0 ~ offset(log(shootEst.no18)) + C_N_imputed + Total_Area + sal_tannin.PC1, full.df, family = binomial)
summary(vLG.dens.glm)
plot(vLG.dens.glm)

vLG.dens.lm <- lm(log(vLG_abund) ~ offset(log(shootEst.no18)) + C_N_imputed + Total_Area + sal_tannin.PC1, filter(full.df, vLG_abund > 0))
summary(vLG.dens.lm)
plot(vLG.dens.lm)
visreg(vLG.dens.lm)

plot(log(vLG_abund/shootEst.no18) ~ sal_tannin.PC1, filter(full.df, vLG_abund > 0))
library(MuMIn)

dredge(vLG.full)
vLG.null <- lm(vLG.df$vLG_density ~ 1, vLG.df[ ,trait.names])
vLG.full <- lm(vLG.df$vLG_density ~ ., vLG.df[ ,trait.names])
summary(vLG.full) # adj. R2 = 0.03 DOESN'T PASS THE TEST!!!
vLG.step <- step(vLG.null, scope = formula(vLG.full), direction = "forward")
vLG.update <- update(vLG.null, .~. + sal_tannin.PC1 + log_size + height_resid) 
summary(vLG.update)
#plot(vLG.update)

vLG.size.null <- lm(vLG.size.df$vLG.height.mean ~ 1, vLG.size.df[ ,trait.names])
vLG.size.full <- lm(vLG.size.df$vLG.height.mean ~ ., vLG.size.df[ ,trait.names])
summary(vLG.size.full) # adj. R2 = 0.03 DOESN'T PASS THE TEST!!!
vLG.size.step <- step(vLG.size.null, scope = formula(vLG.size.full), direction = "forward")
vLG.size.update <- update(vLG.size.null, .~. + flavonOLES.PC1) 
summary(vLG.size.update)

rG.null <- lm(rG.df$rG_density ~ 1, rG.df[ ,trait.names])
rG.full <- lm(rG.df$rG_density ~ ., rG.df[ ,trait.names])
summary(rG.full) # adj. R2 = 0.12 DOESN'T PASS THE TEST!!!
rG.step <- step(rG.null, scope = formula(rG.full), direction = "forward")
rG.update <- update(rG.null, .~. + log_size  + height_resid + 
    density_resid) 
summary(rG.update)

```

### 2015-03-06 Rscripts/betalink_tree_net.R

@@ -30,6 +30,7 @@ geno_4plus <- c("*","B","I","K","L","S","V","X","Z") # 9 (if Pontania galls were
 ## dataset to use for PERMANOVA
 net_df <- tree_level_interaxn_all_plants_traits_size#filter(tree_level_interaxn_all_plants_traits_size, abund_interaxns_noPont > 0, Genotype %in% geno_3plus)
 
+
 ## Permutational multivariate analysis of variance (PERMANOVA)
 net_qual <- ifelse(net_df[ ,interaxns_noPont] > 0, 1, 0)
 
@@ -311,7 +312,8 @@ X.df <- X_info[[1]] %>%
   reshape(idvar = "Group", timevar = "Sequence", direction = "wide") %>%
   mutate(weight.trans = Weight.1) # unscaled weights
 X.nodes <- X_info[[2]] %>%
-  mutate(colrs = c(rab.bud.col, iteo.col, eulo.col, meso.col, platy.col, tory.col)) %>%
+  mutate(colrs = c(rab.bud.col, iteo.col, eulo.col, meso.col, platy.col, tory.col),
+         guilds = c("gall","gall","ecto","ecto","egg","ecto")) %>%
   arrange(vertex.names)
 
 metaplot.col <- brewer.pal(11, "Spectral")
@@ -332,8 +334,10 @@ X.plot <- ggplot(X.df) +
                size = X.df$weight.trans/max.seg.wt*15,
                alpha = 0.75) +
   new_theme_empty + 
-  geom_point(data = filter(X.nodes, y > 1), aes(x = x, y = y, fill = vertex.names), 
-             shape = 25, size = 30) +
+  geom_point(data = filter(X.nodes, y > 1), aes(x = x, y = y, fill = vertex.names, shape = guilds), 
+             #shape = 25, 
+             size = 30) +
+  scale_shape_manual(values = c(22, 25), guide = "none") + 
   geom_point(data = filter(X.nodes, y == 1), aes(x = x, y = y, fill = vertex.names), 
              shape = 21, size = 30) +
   scale_fill_manual(values = X.nodes$colrs, guide = "none")
@@ -351,7 +355,8 @@ F.df <- F_info[[1]] %>%
   reshape(idvar = "Group", timevar = "Sequence", direction = "wide") %>%
   mutate(weight.trans = Weight.1) # unscaled weights
 F.nodes <- F_info[[2]] %>%
[…]

### 2015-03-06 Rscripts/metaweb plot.R

@@ -28,9 +28,9 @@ interaction.df <- metaweb.info[[1]] %>%
 
 nodeinfo.df <- metaweb.info[[2]] %>%
   mutate(guild = factor(c(rep("gall", 4),
-                   "predator",
-                   rep("Larval parasitoid", 3),
-                   rep("Egg parasitoid", 2))),
+                   "Predator",
+                   rep("Larval\nparasitoid", 3),
+                   rep("Egg\nparasitoid", 2))),
          names = c("Cecidomyiid",
                    "Rab. (bud)",
                    "Iteomyia",
@@ -54,17 +54,20 @@ metaweb.plot <- ggplot(interaction.df) +
                size = interaction.df$weight.trans/max(interaction.df$weight.trans)*25,
                alpha = 0.75)  +
   new_theme_empty + 
-  geom_point(data = filter(nodeinfo.df, y > 1), aes(x = x, y = y, fill = vertex.names),
+  geom_point(data = filter(nodeinfo.df, y > 1), aes(x = x, y = y, fill = vertex.names, shape = guild),
              color = "black",
-             shape = 25,
+             #shape = 25,
              size = 30)  + 
-  geom_text(data = filter(nodeinfo.df, y > 1), aes(x = x, y = y + 0.15, label = names), size = 8) +  
+  scale_shape_manual(values = c(25, 22, 23), name = "Natural enemy guild") + 
+  geom_text(data = filter(nodeinfo.df, y > 1), aes(x = x, y = y + 0.15, label = names), size = 6) +  
   geom_point(data = filter(nodeinfo.df, y == 1), aes(x = x, y = y, fill = vertex.names),
              color = "black",
              shape = 21,
              size = 30) + 
   scale_fill_brewer(palette = "Spectral", guide = "none") +
-  geom_text(data = filter(nodeinfo.df, y == 1), aes(x = x, y = y - 0.15, label = names), size = 8)
+  geom_text(data = filter(nodeinfo.df, y == 1), aes(x = x, y = y - 0.15, label = names), size = 8) +
+  theme(legend.text = element_text(size = 14),
+        legend.title = element_text(size = 16))
 
 #Code to override clipping
 metaweb.plot

### 2015-03-06 Rscripts/morgue_ms_willow_network_supplement.R

@@ -1,0 +1,79 @@
+# subset of full data set. Only contains Genotype with 3+ replicates in the subset of data that only contains willows with at least 1 gall-parasitoid interaction found on them.
+abund_interaxns_noPont <- rowSums(full.df[ ,interaxns_noPont]) # used to filter willow replicates with at least 1 gall-parasitoid interaction found on them.
+table(filter(full.df, abund_interaxns_noPont > 0)$Genotype) # number of replicates for all genotypes in reduced dataset.
+geno_3plus <- c("*","B","D","I","K","L","Q","S","V","W","X","Y","Z") # Genotypes with 3+ replicates
+sub.df <- filter(full.df, abund_interaxns_noPont > 0, Genotype %in% geno_3plus)
+
+# quantitative interaction data
+full.df.quant <- full.df[ ,interaxns_noPont]
+sub.df.quant <- sub.df[ ,interaxns_noPont]
+
+# qualitative interaction data
+full.df.qual <- ifelse(full.df[ ,interaxns_noPont] > 0, 1, 0)
+sub.df.qual <- ifelse(sub.df[ ,interaxns_noPont] > 0, 1, 0)
+
+library(lme4)
+library(RLRsim)
+plot(vLG_abund ~ Genotype, full.df)
+vLG.lmer <- lmer(log(vLG_abund+1) ~ (1|Genotype), full.df)
+heritability(vLG.lmer)
+
+with(full.df, table(vLG_abund, Genotype))
+with(full.df, table(rG_abund, Genotype))
+hist(full.df$rG_abund)
+hist(log(full.df$rG_abund+1))
+
+library(glmmADMB)
+hist(subset(full.df, vLG_abund > 0)$vLG_abund)
+fit1 <- glmmadmb(vLG_abund ~ (1|Genotype), subset(full.df, vLG_abund > 0), family="truncpoiss" )
+summary(fit1)
+t <- ranef(fit1)$Genotype
+ggQQ_ranef(t)
+
+fit2 <- glmmadmb(vLG_abund >0 ~ (1|Genotype), subset(full.df, vLG_abund > 0), family="binomial" )
+summary(fit1)
+
+vLG.glmer <- glmer(vLG_abund>0 ~ (1|Genotype), full.df, family = binomial)
+summary(vLG.glmer)
+plot(vLG.glmer)
[…]

### 2015-03-06 Rscripts/network mixed effect models.R

@@ -10,26 +10,145 @@ library(gamm4)
 
 gall_mech.df
 
+## testing for overdispersion
+overdisp_fun <- function(model) {
+  ## number of variance parameters in 
+  ##   an n-by-n variance-covariance matrix
+  vpars <- function(m) {
+    nrow(m)*(nrow(m)+1)/2
+  }
+  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
+  rdf <- nrow(model.frame(model))-model.df
+  rp <- residuals(model,type="pearson")
+  Pearson.chisq <- sum(rp^2)
+  prat <- Pearson.chisq/rdf
+  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
+  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
+}
+
 ## vLG mechanisms df
-vLG_mech.df <- filter(gall_mech.df, vLG_total > 0) %>%
+vLG_mech.df <- filter(gall_mech.df, vLG_total > 0 & gall.height > 0) %>%
   mutate(sc.gall.height = scale(gall.height, center = TRUE, scale = TRUE),
          sc.vLG_density = scale(vLG_density, center = TRUE, scale = TRUE),
+         sc.vLG_abund = scale(vLG_abund, center = TRUE, scale = TRUE),
          vLG_egg = vLG_Platy + vLG_Mymarid,
-         vLG_ecto = vLG_Eulo + vLG_Mesopol + vLG_Tory)
+         vLG_ecto = vLG_Eulo + vLG_Mesopol + vLG_Tory,
+         vLG_parasitized = vLG_egg + vLG_ecto,
+         gall.id.nest = factor(gall.id.nest))
+str(vLG_mech.df)
 
+## vLG_parasitized
+library(MCMCglmm)
+t <- rpt.binomGLMM.add(y = with(vLG_mech.df, cbind(vLG_parasitized, vLG_total-vLG_parasitized)), groups = vLG_mech.df$Genotype.x)
+t
+library(embd)
+q <- rpt.binomGLMM.multi(y = with(vLG_mech.df, cbind(vLG_parasitized, vLG_total-vLG_parasitized)), groups = vLG_mech.df$Genotype.x)
+test <- glmmPQL(cbind(vLG_parasitized, vLG_total-vLG_parasitized) ~ 1,
+                random = ~ 1|Genotype.x/plant.position,
+                data = vLG_mech.df,
+                family = binomial)
+summary(test)
[…]

### 2015-03-06 Rscripts/network_management_gall_level.R

@@ -25,6 +25,10 @@ gall_density.df <- with(tree_level_interaxn_all_plants_traits_size,
                         data.frame(plant.position = plant.position, 
                                    Genotype = Genotype, 
                                    Gender = Gender, 
+                                   vLG_abund = vLG_abund,
+                                   rG_abund = rG_abund,
+                                   aSG_abund = aSG_abund,
+                                   SG_abund = SG_abund,
                                    vLG_density = vLG_abund/shootEst.no18*100,
                                    rG_density = rG_abund/shootEst.no18*100,
                                    aSG_density = aSG_abund/shootEst.no18*100,

### 2015-03-06 Rscripts/numerical responses.R

@@ -1,0 +1,221 @@
+## numerical responses of natural enemies
+
+## source in required data
+source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
+
+library(visreg)
+library(lme4)
+library(RLRsim)
+library(vegan)
+#library(pbkrtest)
+
+#### Analyses at the tree-level
+
+## make focal data frame for analysis
+df <- tree_level_interaxn_all_plants_traits_size %>%
+  select(aSG_Tory, rG_Eulo, rG_Lestodip, rG_Mesopol, rG_Platy, rG_Tory, 
+         SG_Platy, vLG_Eulo, vLG_Mesopol, vLG_Mymarid, vLG_Platy, vLG_Tory, 
+         vLG_egg, vLG_ecto, rG_ecto, # all gall-parasitoid interactions for both pairwise and dissimilarity analyses
+         shootEst.no18,
+         Genotype, 
+         vLG_abund, rG_abund, # only gall species associated with significant pairwise gall-ptoid interactions
+         vLG.height.mean) %>% # only gall species that varied in size among genotypes
+  mutate(vLG_density = vLG_abund/shootEst.no18,
+         rG_density = rG_abund/shootEst.no18)
+
+df.vLG.size.sub <- na.omit(df) # will use for RDA when I include gall size in the analysis, because not all trees had vLG galls on them, hence the reduced dataset.
+
+## RDA of insect food web
+rda_full <- rda(df.vLG.size.sub[ ,interaxns_noPont]/df.vLG.size.sub$shootEst.no18 ~ vLG_density*vLG.height.mean + rG_density, data = df.vLG.size.sub) # much higher explanatory power without sqrt transforming the data.
+RsquareAdj(rda_full)
+summary(rda_full)
+plot(rda_full, display = c("sp","bp"))
+vif.cca(rda_full)
+anova(rda_full, by = "margin") # why aren't the lower level main effects included?
+
[…]

### 2015-03-06 adonis.table.function.R

@@ -1,0 +1,14 @@
+adonis.table <- function(adonis.object, data.frame){
+  x <- adonis.object
+  mean.sample.size <- mean(table(data.frame$Genotype))
+  genetic.variance <- (x$aov.tab$MeanSqs[1] - x$aov.tab$MeanSqs[2])/mean.sample.size
+  heritability <- genetic.variance/(genetic.variance + x$aov.tab$MeanSqs[2])
+  
+  table.out <- data.frame(df = x$aov.tab$Df[1:2],
+                          SS = x$aov.tab$SumsOfSqs[1:2],
+                          MS = x$aov.tab$MeanSqs[1:2],
+                          Heritability = c(heritability,""),
+                          P = c(x$aov.tab$'Pr(>F)'[1],""))
+  table.out
+}
+

### 2015-03-06 gall_community_size_plot.R

@@ -1,0 +1,74 @@
+require(ggplot2)
+require(gridExtra)
+require(dplyr)
+
+source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
+
+## create theme for figures
+theme_galls <- theme_classic() +
+  theme(axis.title.x = element_text(size = 16, vjust = 0.1), 
+        axis.title.y = element_text(size = 16, vjust = 1),
+        axis.text = element_text(size = 12),
+        legend.title = element_blank(),
+        legend.position = c(1,1),
+        legend.justification = c(1,1))
+
+## Gall community composition figure
+
+# calculate mean abundance of each gall species for each genotype
+gall.comm.df <- tree_level_interaxn_all_plants_traits_size %>%
+  select(Genotype,
+         "Iteomyia salicisverruca" = vLG_abund,
+         "Rabdophaga salicisbrassicoides" = rG_abund,
+         "Cecidomyiid sp. A" = aSG_abund,
+         "Rabdophaga salicisbattatus" = SG_abund) %>%
+  group_by(Genotype) %>%
+  summarise_each(funs(mean.narm = mean(., na.rm = TRUE)))
+levels(gall.comm.df$Genotype)[1] <- "C" # changed Genotype * to C for plotting aesthetics
+gall.comm.df$Genotype <- factor(gall.comm.df$Genotype, levels = LETTERS[1:26])  # reorder for plotting A->Z
+
+gall.comm.melt <- melt(gall.comm.df) # melt for plotting in ggplot
+
+# create figure
+comm.plot <- ggplot(gall.comm.melt, aes(x = Genotype, y = value, fill = variable)) + 
+  geom_bar(stat = "identity", color = "black") +
+  scale_fill_grey(start = 0, end = 1) +
+  scale_y_continuous(expand = c(0,0), limits = c(0,20)) +
+  ylab("Gall density (#/branch)") +
+  xlab("") +
+  theme_galls +
+  annotate("text", x = 1, y = 19, label = "(A)")
+
+## Variation in Iteomyia gall size figure
+gall.size.df <- tree_level_interaxn_all_plants_traits_size 
+levels(gall.size.df$Genotype)[1] <- "C" # changed Genotype * to C for plotting aesthetics
+gall.size.df$Genotype <- factor(gall.size.df$Genotype, levels = LETTERS[1:26]) # reorder for plotting A->Z
+
[…]

### 2015-04-06 ms_willow_network_supplement.Rmd
lots of additions.

@@
 library(ggplot2) # for plotting
-theme_set(theme_classic()) # customize ggplot for prettier default graphs
+theme_set(theme_bw()) # customize ggplot for prettier default graphs
 library(gridExtra) # for grid.arrange
@@
 ```{r gall abund across genotypes, echo = FALSE}
-vLG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = vLG_abund)) + 
-  geom_boxplot() + ylab("Iteomyia salicisverruca density (#/branch)")
-rG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = rG_abund)) + 
-  geom_boxplot() + ylab("Rabdophaga salicisbrassicoides density (#/branch)")
-aSG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = aSG_abund)) + 
-  geom_boxplot() + ylab("Cecidomyid sp. A density (#/branch)")
-SG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = SG_abund)) + 
-  geom_boxplot() + ylab("Rabdophaga salicisbattatus density (#/branch)")
-
-grid.arrange(vLG.abund.plot, rG.abund.plot, aSG.abund.plot, SG.abund.plot)
+gall.plot.df <- as.data.frame(full.df) %>% # melt doesn't like 'tbl' class
+  select(Genotype, 
+         Iteomyia_salicisverruca = vLG_abund,
+         Cecidomyiid_sp.A = aSG_abund,
+         Rabdophaga_salicisbrassicoides = rG_abund, 
+         Rabdophaga_salicisbattatus = SG_abund) %>%
+  melt()
+
+ggplot(gall.plot.df, aes(x = Genotype, y = value, color = variable)) + 
+  geom_boxplot() + facet_wrap( ~ variable, nrow = 2) + theme_bw() + ylab("Gall density (#/branch)")
 ```
@@
 
-N <- 50
+N <- 1000
 link.div.list <- list()
@@
 } # takes a couple minutes with 50 reps
-
+#library(vegan)
 link.div.df <- ldply(link.div.list) %>% tbl_df()
@@
 
+# get predicted weighted linkage density for each genotype
+predict.single.geno <- as.data.frame(round(predict.manyglm(manyglm.full, data.frame(Genotype = levels(full.df$Genotype)), type = "response"),1))
+predict.single.geno$Number.of.Genotypes <- rep(1, 26)
+predict.single.geno$diversity <- diversity(select(predict.single.geno, -Number.of.Genotypes))
+predict.single.geno$repID <- 25001:(25001+25)
+
+link.div.df <- rbind.data.frame(link.div.df, predict.single.geno)
+
 boxplot(diversity ~ factor(Number.of.Genotypes), link.div.df, ylab = "Interaction Diversity", xlab = "Number of Genotypes")
 
-link.melt <- melt(select(link.div.df, -diversity), id.vars = c("repID", "Number.of.Genotypes"))
+link.melt <- as.data.frame(link.div.df) %>%
+  select(-diversity) %>%
+  melt(id.vars = c("repID", "Number.of.Genotypes"))
 
@@
 
-library(bipartite)
+single.web.list <- cast(filter(link.melt.df, Number.of.Genotypes == 1), gall ~ parasitoid | repID, mean)
+
+#library(bipartite)
 web.list.measures <- c()
@@
 }
+
+single.web.list.measures <- c()
+single.genotype.number <- c()
+for(i in 1:dim(web.list)){
+  genotype.number[i] <- names(single.web.list[[i]][1])
+  tmp.web <- single.web.list[[i]][[1]]
+  rownames(tmp.web) <- tmp.web$gall
+  tmp.web <- as.matrix.data.frame(tmp.web[ ,-1])
+  web.list.measures[i] <- networklevel(tmp.web, index = "linkage density", weighted = TRUE)
+}
+
 web.measures.df <- cbind.data.frame(Number.of.Genotypes = as.numeric(genotype.number),
@@
 boxplot(Weighted.Linkage.Density ~ factor(Number.of.Genotypes), web.measures.df, ylab = "Weighted Linkage Density", xlab = "Number of Genotypes")
+
+write.csv(web.measures.df, "web.measures.df.csv")
+
+ggplot(web.measures.df, 
+       aes(x = Number.of.Genotypes, y = Weighted.Linkage.Density)) + 
+  geom_jitter(shape = 1, color = "grey", position = position_jitter(width = 0.15, height = NULL)) + 
+  stat_summary(fun.y = mean, geom = "point", color = "steelblue", size = 8) + 
+  xlab("Genetic variation (# of genotypes)") + ylab("Weighted linkage density") +
+  scale_x_continuous(limits = c(0,26), breaks = 1:26) +
+  theme_bw() + 
+  theme(axis.text = element_text(size = 14),
+        axis.title.x = element_text(size = 16, vjust = 0.1),
+        axis.title.y = element_text(size = 16, vjust = 1),
+        panel.grid = element_blank())
 
@@
 
-library(mgcv)
+#library(mgcv)
+#library(visreg)
 web.sum.lm <- lm(log(median.link.density) ~ log(Number.of.Genotypes), data = web.measures.summary)
@@
 visreg(web.sum.lm) # not quite there
+```
+
+```{r complementarity}
+net.pred.geno <- as.data.frame(round(predict.manyglm(manyglm.full, 
+                                       newdata = data.frame(Genotype = levels(full.df$Genotype)), 
+                                       type = "response"), 1)) %>% # round to avoid pseudoaccuracy 
+  mutate(Genotype = levels(full.df$Genotype)) %>%
+  melt(id.vars = "Genotype")
+
+net.split <- colsplit(net.pred.geno$variable, pattern = "_", names = c("gall","parasitoid"))
+net.pred.geno.df <- cbind(net.pred.geno, net.split)
+
+net.pred.full.df <- net.pred.geno.df %>%
+  group_by(variable, gall, parasitoid) %>%
+  summarise(value = mean(value))
+
+net.list <- cast(net.pred.geno.df, gall ~ parasitoid | Genotype, mean)
+net.list.noU <- net.list[-21]
+
+net.full <- cast(net.pred.full.df, gall ~ parasitoid, mean)
+rownames(net.full) <- net.full$gall
+net.full <- as.matrix.data.frame(net.full[ ,-1]) # remove gall variable
+net.full.measure <- networklevel(net.full, index = c("links per species", "linkage density", "Shannon diversity"), weighted = TRUE) # note that the weighted linkage density here is higher than from my simulation with the 26 genotypes...
+
+library(bipartite)
+nets <- list()
+net.list.measures <- c()
+genotypes <- c()
+for(i in 1:dim(net.list.noU)){
+  genotypes[i] <- names(net.list.noU)[i]
+  tmp.web <- net.list.noU[[i]]
+  rownames(tmp.web) <- tmp.web$gall
+  nets[[i]] <- as.matrix.data.frame(tmp.web[ ,-1])
+  net.list.measures[i] <- networklevel(nets[[i]], index = "linkage density", weighted = TRUE)
+}
+mean(net.list.measures)
+genotypes
+net.list.measures.df <- cbind.data.frame(genotypes, net.list.measures)
+networklevel(nets[[25]], index = c("links per species", "linkage density", "Shannon diversity"), weighted = TRUE)
+
+qplot(net.list.measures.df, x = factor(1), y = net.list.measures, fill = genotypes, , color = "black", geom = "bar", stat = "identity")
+sum(net.list.measures.df$net.list.measures)
+
+ph <- matrix(c(1,1,1), nrow = 1)
+networklevel(ph, index = c("links per species", "linkage density", "Shannon diversity", "vulnerability", "generality"), weighted = TRUE)
+mean(c(3,1,3,1))
 ```
@@
 full.mvabund <- mvabund(full.df[ ,interaxns_noPont]) # create mvabund object for easy analysis
-meanvar.plot(full.mvabund ~ full.df$Genotype) # easy to see that the data do not have a poisson distribution, therefore, a negative binomial may be a good fit.
+full.meanvar <- meanvar.plot(full.mvabund ~ full.df$Genotype, table = TRUE) # easy to see that the data do not have a poisson distribution, therefore, a negative binomial may be a good fit.
 poisson_curve(from = 0.1, to = 3.0, color = "blue")
@@
 neg.binomial_curve(from = 0.1, to = 3.0, theta.negbin = 0.7, color = "black") # negative binomial appears to provide the best fit to this relationship
+
+no.relationships <- which(names(colSums(full.mvabund)) %in% c("rG_Lestodip", "rG_Mesopol", "SG_Platy", "vLG_Eulo"))
+sum(colSums(full.mvabund)[no.relationships])/sum(colSums(full.mvabund)) # less than 13% of the links in the network had relationships to gall abundance and size (see results below)
+```
+
+Analysis of dissimilarity in interaction composition
+```{r}
+links <- full.df[ ,interaxns_noPont] 
+link_abund <- rowSums(links)
+
+links.dist <- cbind.data.frame(Genotype = full.df$Genotype, links) %>%
+  filter(link_abund > 0, Genotype %in% c("*","A","B","D","E","F","I","K","L","Q","S","T","V","W","X","Y","Z"))
+
+links.geno.dist <- meandist(vegdist(links.dist[ ,-1], "bray"), links.dist$Genotype)
+summary(links.geno.dist) # on average, gall communities were 57% dissimilar from each other
+plot(links.geno.dist, cluster = "ward")
 ```
@@
 anova.full <- anova.manyglm(manyglm.full, p.uni = "unadjusted") # Takes about 1 min and 30 sec to run.
-anova.full # vLG_Platy, vLG_Tory, and vLG_Mesopol are driving the community response. rG_Tory is marginally significant and vLG_Eulo is close to marginal as well.
+anova.full # vLG_Platy, vLG_Tory, and vLG_Mesopol are driving the community response. rG_Tory is marginally significant and vLG_Eulo is close to marginal as well. aSG_Tory is significant as well.
+```
+
+```{r}
+full.mean <- cbind.data.frame(Genotype = full.df$Genotype, full.mvabund) %>%
+  group_by(Genotype) %>%
+  summarise_each(funs(mean.narm = mean(., na.rm = TRUE)))
+
+# calculate average dissimilarity in link community composition among willow genotypes using average link abundance data.
+require(vegan)
+full.mean.df.noU <- full.mean %>% filter(Genotype != "U") # genotype U never had any interactions
+full.mean.bray <- vegdist(select(full.mean.df.noU, -Genotype), method = "bray")
+mean(full.mean.bray) # on average, link composition was 77% dissimilar from each other.
+sd(full.mean.bray)
+range(full.mean.bray)
+
+plot(hclust(full.mean.bray, method = "average"), labels = full.mean.df.noU$Genotype)
+#NMDS <- metaMDS(comm = select(full.mean.df.noU, -Genotype), distance = "bray")
+#plot(NMDS, type = "t")
+
+require(tidyr)
+beta_partition_lists <- as.data.frame(full.mean.df.noU) %>%
+  gather(key = Genotype, value = link_abund, aSG_Tory:vLG_Tory) %>% # why don't variable and value columns change?
+  #mutate(variable = as.character(variable)) %>%
+  separate(variable, into = c("Gall","Parasitoid"), sep = "_") %>%
+  filter(value > 0) %>%
+  cast(Gall ~ Parasitoid | Genotype, sum) # note that sum function doesn't do anything. I'm just reshaping the dataframe into multiple lists.
+beta_partition_lists <- beta_partition_lists[-21]
+
+for(i in 1:length(beta_partition_lists)) {
+  rownames(beta_partition_lists[[i]]) <- as.character(beta_partition_lists[[i]]$Gall) # change row names to gall specie names
+  beta_partition_lists[[i]] <- as.data.frame(select(beta_partition_lists[[i]], -Gall)) # remove gall.sp as a column of data.
+Turning this into a matrix is important for easier analysis with betalink package
+  beta_partition_lists[[i]] <- as.matrix(beta_partition_lists[[i]])  # need to be converted into matrices for "as.table" function to work for quantitative data.
+}
+
+### Beta-diversity of interaction network analysis
+source('~/Documents/betalink/R/betalink.R')
+source('~/Documents/betalink/R/vec2data.frame.R')
+source('~/Documents/betalink/R/measures.r')
+source('~/Documents/betalink/R/betalink.b.r')
+source('~/Documents/betalink/R/betalink.dist.r')
+source('~/Documents/betalink/R/betalink.q.R')
+
+# Betalink Quantitative
+betalink_quantitative <- betalink.dist(beta_partition_lists, bf = "bray", triangular = T) #
+betalink_qualitative <- betalink.dist(beta_partition_lists, bf = B01, triangular = T)
+
+mean(betalink_quantitative$WN) # 77% dissimilarity with bray
+mean(betalink_qualitative$WN) # 63% dissimilarity with whittaker
+
+mean(betalink_quantitative$OS) # 33% with bray
+mean(betalink_qualitative$OS) # 6% with whittaker
+
+mean(betalink_quantitative$ST) # 44% with bray
+mean(betalink_qualitative$ST) # 57% with whittaker
+
+mean(betalink_quantitative$contrib) # species turnover contributes to 53% of the dissimilarity in interaction networks, according to bray-curtis
+mean(betalink_qualitative$contrib) # species turnover contributes to 86% of the dissimilarity in interaction networks, according to whittaker index on qualitative data.
 ```
@@
          aSG_abund:vLG_abund, # gall counts. Exclusing SG_abund because it didn't vary among genotypes
+         Genotype, # keeping for later analysis
          vLG.height.mean) %>% # Iteomyia gall size
@@
 net.trait <- mvabund(full.predictors[ ,interaxns_noPont])
-net.trait.sub <- mvabund(full.predictors[ ,c("vLG_Platy","vLG_Mesopol","vLG_Tory","rG_Tory")])
+
+interactions <- full.predictors[ ,interaxns_noPont]
+interactions[interactions > 0] <- 1
+
+net.trait.sub <- mvabund(interactions)
 # even though manyglm doesn't appear to test interaction terms appropriately (sequential vs.
 dropping different terms), there is not interactive effect of vLG height and vLG abundance in univariate negative binomial models for any vLG_parasitoid interaction, so I believe this final model is appropriate for the data.
+net.mvabund.pres <- manyglm(net.trait.sub ~ log.vLG_abund + vLG.height.mean + log.1.rG_abund + log.1.aSG_abund, data = full.predictors, family = "binomial")
+plot(net.mvabund.pres, which = 1:3)
+anova.qual <- anova(net.mvabund.pres, p.uni = "unadjusted")
+anova.qual
+
+summary(glm(cbind(rG_Tory, rG_Platy + rG_Eulo + rG_Lestodip + rG_Mesopol) ~ vLG.height.mean, data = full.predictors, family = "binomial"))
+visreg(glm(cbind(vLG_Tory, vLG_Platy + vLG_Eulo + vLG_Mymarid + vLG_Mesopol) ~ vLG.height.mean + vLG_abund, data = full.predictors, family = "binomial"), scale = "response")
+
 net.mvabund.1.full <- manyglm(net.trait ~ log.vLG_abund*vLG.height.mean + 
@@
 plot(net.mvabund.2, which = 1:3) # residuals look pretty good.
+
+plot(MASS::glm.nb(vLG_Platy ~ vLG_abund + vLG.height.mean, full.predictors))
 ```
@@
 Therefore, we retain all P-values < 0.10 (during at least one run) for the coefficient summary below.
+
+anova.net.null <- anova.manyglm(net.mvabund.2, net.mvabund.6.null) # takes about 30 sec to run
+anova.net.null
 
@@
                                  "log.vLG_abund_vLG_Mymarid",
-                                 "log.vLG.height.mean_vLG_Eulo", # marginal
-                                 "log.vLG.height.mean_rG_Tory", "log.1.rG_abund_rG_Tory",
-                                 "log.vLG_abund_vLG_Mesopol", "log.vLG.height.mean_vLG_Mesopol", 
-                                 "log.vLG.height.mean_vLG_Platy", "log.vLG_abund_vLG_Platy",
-                                 "log.vLG.height.mean_vLG_Tory", "log.vLG_abund_vLG_Tory")) %>% # vLG.height.mean is marginal for vLG_Tory
+                                 "vLG.height.mean_vLG_Eulo", # marginal
+                                 "vLG.height.mean_rG_Tory", "log.1.rG_abund_rG_Tory",
+                                 "vLG_abund_vLG_Mesopol", "vLG.height.mean_vLG_Mesopol", 
+                                 "vLG.height.mean_vLG_Platy", "log.vLG_abund_vLG_Platy",
+                                 "vLG.height.mean_vLG_Tory", "log.vLG_abund_vLG_Tory")) %>% # vLG.height.mean is marginal for vLG_Tory
   select(predictor = X1, response = X2, coefficient = value)
 arrange(sig.coef.df.2, predictor, response)
+1 - exp(-0.2157298) # 19% decrease in vLG_Platy interaction with every one unit increase in gall size.
+1 - exp(-0.2697011) # 24% decrease in vLG_Mesopolobous interaction with every one unit increase in gall size.
+```
+
+The results from the link composition analysis suggest that increases in gall abundance lead to increasing food web complexity for those individual nodes, whereas differences in leaf gall size lead to fundamental differences in link composition.
+
+```{r}
+row.count <- dim(full.predictors)[1]
+arrange(full.predictors, vLG.height.mean)$vLG.height.mean
+
+## generate a data frame that varies on Iteomyia abundance but holds all other variables constant at their means.
+predict.log.vLG_abund <- with(full.predictors,
+                              data.frame(log.vLG_abund = seq(from = min(log.vLG_abund),
+                                                             to = max(log.vLG_abund),
+                                                             length.out = row.count),
+                                         vLG.height.mean = rep(mean(vLG.height.mean), row.count),
+                                         log.1.rG_abund = rep(mean(log.1.rG_abund), row.count),
+                                         log.1.aSG_abund = rep(mean(log.1.aSG_abund), row.count)))
+
+## predict link abundance based on the Iteomyia abundance variation.
+log.vLG_abund.predict <- predict(net.mvabund.2, 
+                       newdata = predict.log.vLG_abund, 
+                       type = "response", se.fit = TRUE) 
+colnames(log.vLG_abund.predict$se.fit) <- colnames(log.vLG_abund.predict$fit) # set column names for SE
+log.vLG_abund.fit.df <- cbind.data.frame(log.vLG_abund.predict$fit, 
+                                             predict.log.vLG_abund)
+log.vLG_abund.SE.df <- cbind.data.frame(log.vLG_abund.predict$se.fit,
+                                        predict.log.vLG_abund)
+log.vLG_abund.predict.df <- rbind.data.frame(log.vLG_abund.fit.df,
+                                             log.vLG_abund.SE.df) %>%
+  mutate(type = c(rep("predict", row.count),
+                  rep("SE", row.count))) %>%
+  select(-log.1.aSG_abund, -log.1.rG_abund, -vLG.height.mean) %>%
+  melt(id.vars = c("type","log.vLG_abund")) %>%
+  spread(type, value)
+
+ggplot(data = filter(log.vLG_abund.predict.df, 
+                     variable %in% c("vLG_Platy","vLG_Mesopol","vLG_Tory")),
+       aes(x = exp(log.vLG_abund), y = predict)) + 
+  geom_line(aes(color = variable)) +
+  geom_ribbon(aes(ymax = predict + SE, ymin = predict - SE, fill = variable), alpha = 0.25) +
+  ylab("Link density (#/branch)") + xlab("Leaf gall midge density (#/branch)")
+
+## generate a data frame that varies on Iteomyia diameter but holds all other variables constant at their means.
+predict.vLG.height.mean <- with(full.predictors,
+                              data.frame(vLG.height.mean = seq(from = min(vLG.height.mean),
+                                                             to = max(vLG.height.mean),
+                                                             length.out = row.count),
+                                         log.vLG_abund = rep(mean(log.vLG_abund), row.count),
+                                         log.1.rG_abund = rep(mean(log.1.rG_abund), row.count),
+                                         log.1.aSG_abund = rep(mean(log.1.aSG_abund), row.count)))
+
+## predict link abundance based on the Iteomyia diameter variation.
+vLG.height.mean.predict <- predict(net.mvabund.2, 
+                       newdata = predict.vLG.height.mean, 
+                       type = "response", se.fit = TRUE) 
+colnames(vLG.height.mean.predict$se.fit) <- colnames(vLG.height.mean.predict$fit) # set column names for SE
+vLG.height.mean.fit.df <- cbind.data.frame(vLG.height.mean.predict$fit, 
+                                             predict.vLG.height.mean)
+vLG.height.mean.SE.df <- cbind.data.frame(vLG.height.mean.predict$se.fit,
+                                        predict.vLG.height.mean)
+vLG.height.mean.predict.df <- rbind.data.frame(vLG.height.mean.fit.df,
+                                             vLG.height.mean.SE.df) %>%
+  mutate(type = c(rep("predict", row.count),
+                  rep("SE", row.count))) %>%
+  select(-log.1.aSG_abund, -log.1.rG_abund, -log.vLG_abund) %>%
+  melt(id.vars = c("type","vLG.height.mean")) %>%
+  spread(type, value)
+
+ggplot(data = filter(vLG.height.mean.predict.df, 
+                     variable %in% c("vLG_Platy","vLG_Mesopol","vLG_Tory","rG_Tory")),
+       aes(x = vLG.height.mean, y = predict)) + 
+  geom_line(aes(color = variable)) +
+  geom_ribbon(aes(ymax = predict + SE, ymin = predict - SE, fill = variable), alpha = 0.25) +
+  ylab("Link density (#/branch)") + xlab("Leaf gall diameter (mm)")
+
+## generate a data frame that varies on bud gall density but holds all other variables constant at their means.
+predict.log.1.rG_abund <- with(full.predictors,
+                              data.frame(log.1.rG_abund = seq(from = min(log.1.rG_abund),
+                                                             to = max(log.1.rG_abund),
+                                                             length.out = row.count),
+                                         log.vLG_abund = rep(mean(log.vLG_abund), row.count),
+                                         vLG.height.mean = rep(mean(vLG.height.mean), row.count),
+                                         log.1.aSG_abund = rep(mean(log.1.aSG_abund), row.count)))
+
+## predict link abundance based on the bud gall variation.
+log.1.rG_abund.predict <- predict(net.mvabund.2, 
+                       newdata = predict.log.1.rG_abund, 
+                       type = "response", se.fit = TRUE) 
+colnames(log.1.rG_abund.predict$se.fit) <- colnames(log.1.rG_abund.predict$fit) # set column names for SE
+log.1.rG_abund.fit.df <- cbind.data.frame(log.1.rG_abund.predict$fit, 
+                                             predict.log.1.rG_abund)
+log.1.rG_abund.SE.df <- cbind.data.frame(log.1.rG_abund.predict$se.fit,
+                                        predict.log.1.rG_abund)
+log.1.rG_abund.predict.df <- rbind.data.frame(log.1.rG_abund.fit.df,
+                                             log.1.rG_abund.SE.df) %>%
+  mutate(type = c(rep("predict", row.count),
+                  rep("SE", row.count))) %>%
+  select(-log.1.aSG_abund, -vLG.height.mean, -log.vLG_abund) %>%
+  melt(id.vars = c("type","log.1.rG_abund")) %>%
+  spread(type, value)
+
+ggplot(data = filter(log.1.rG_abund.predict.df, 
+                     variable %in% c("rG_Tory","rG_Eulo")),
+       aes(x = exp(log.1.rG_abund), y = predict)) + 
+  geom_line(aes(color = variable)) +
+  geom_ribbon(aes(ymax = predict + SE, ymin = predict - SE, fill = variable), alpha = 0.25) +
+  ylab("Link density (#/branch)") + xlab("Bud gall density (#/branch)")
+
+## NOTE THAT IT IS N'T WORKING TO HAVE THE RAW DATA BECAUSE THE SEQUENCE DATA OF THE PREDICTOR VARIABLE MAY HAVE SOME OVERLAPPING VALUES IN ITS ORIGINAL FORM.
+
+## generate a data frame that varies on Iteomyia diameter but holds all other variables constant at their means.
+#log.1.aSG_abund.seq <- arrange(full.predictors, log.1.aSG_abund)$log.1.aSG_abund
+#log.1.aSG_abund.OG <- arrange(full.predictors, log.1.aSG_abund)[ ,interaxns_noPont]
+predict.log.1.aSG_abund <- with(full.predictors,
+                              data.frame(log.1.aSG_abund = seq(from = min(log.1.aSG_abund),
+                                                               to = max(log.1.aSG_abund),
+                                                               length.out = row.count),
+                                         log.vLG_abund = rep(mean(log.vLG_abund), row.count),
+                                         vLG.height.mean = rep(mean(vLG.height.mean), row.count),
+                                         log.1.rG_abund = rep(mean(log.1.rG_abund), row.count)))
+
+## predict link abundance based on the Iteomyia diameter variation.
+log.1.aSG_abund.predict <- predict(net.mvabund.2, 
+                       newdata = predict.log.1.aSG_abund, 
+                       type = "response", se.fit = TRUE) 
+colnames(log.1.aSG_abund.predict$se.fit) <- colnames(log.1.aSG_abund.predict$fit) # set column names for SE
+#log.1.aSG_abund.OG.df <- cbind.data.frame(log.1.aSG_abund.OG,
+          #                                predict.log.1.aSG_abund)
+log.1.aSG_abund.fit.df <- cbind.data.frame(log.1.aSG_abund.predict$fit, 
+                                             predict.log.1.aSG_abund)
+log.1.aSG_abund.SE.df <- cbind.data.frame(log.1.aSG_abund.predict$se.fit,
+                                        predict.log.1.aSG_abund)
+log.1.aSG_abund.predict.df <- rbind.data.frame(#log.1.aSG_abund.OG.df,
+                                               log.1.aSG_abund.fit.df,
+                                             log.1.aSG_abund.SE.df) %>%
+  mutate(type = c(#rep("OG", row.count),
+                  rep("predict", row.count),
+                  rep("SE", row.count))) %>%
+  select(-log.1.rG_abund, -vLG.height.mean, -log.vLG_abund) %>%
+  melt(id.vars = c("type","log.1.aSG_abund")) %>%
+  spread(type, value) # was getting an error about duplicate row identifiers, but it seems to be okay..
+
+ggplot(data = filter(log.1.aSG_abund.predict.df, 
+                     variable %in% c("aSG_Tory")),
+       aes(x = exp(log.1.aSG_abund), y = predict)) + 
+  geom_line(aes(color = variable)) +
+  geom_ribbon(aes(ymax = predict + SE, ymin = predict - SE, fill = variable), alpha = 0.25) +
+  ylab("Link density (#/branch)") + xlab("apical-Stem gall density (#/branch)")
+
+# incorrectly plot the data with all of the points...not controlling for any other predictors though which are important...
+test <- dplyr::select(full.predictors, vLG.height.mean, log.vLG_abund, log.1.rG_abund, vLG_Tory, vLG_Platy, vLG_Mesopol, rG_Tory, rG_Eulo)
+test2 <- melt(as.data.frame(test), id.vars = c("vLG.height.mean", "log.vLG_abund", "log.1.rG_abund"))
+test2 <- mutate(test2, value.bin = ifelse(value > 0, 1, 0))
+test2 <- mutate(test2, log.vLG.cut = ifelse(log.vLG_abund < 1.343, 0, 1))
+
+summary(full.predictors$log.vLG_abund)
+
+library(MASS)
+ggplot(filter(test2, variable %in% c("vLG_Tory","vLG_Platy","vLG_Mesopol")), aes(x = vLG.height.mean, y = value, color = factor(log.vLG.cut))) + geom_point() + geom_smooth(method = "glm.nb") + facet_wrap( ~ variable, ncol = 3)
+ggplot(filter(test2, variable %in% c("vLG_Tory","vLG_Platy","vLG_Mesopol","rG_Tory")), aes(x = vLG.height.mean, y = value.bin)) + geom_point() + geom_smooth(method = "glm", family = "binomial") + facet_wrap( ~ variable, ncol = 1)
+
+test3 <- glmer(vLG_Platy ~ vLG.height.mean + log(vLG_abund) + (1|Genotype), full.predictors, family = poisson)
+overdisp_fun(test3) # no overdispersion
+summary(test3)
+plot(test3) # residuals don't look great though
+visreg(test3, scale = "response", partial = TRUE) # why is the vLG_Platy point so high with the vLG_abund?
+It is impossible...but may suggest poor model fit, since these are the partial residual plots (ref: https://github.com/pbreheny/visreg/issues/1).
+Note however, that this issue is much improved when I used a genearlized linear mixed effects models
+```
+
+```{r}
+full.df$vLG_link_rich <- rowSums(with(full.df, cbind(vLG_Platy, vLG_Tory, vLG_Mesopol, vLG_Mymarid)) > 0)
+full.df$rG_link_rich <- rowSums(with(full.df, cbind(rG_Platy, rG_Tory, rG_Mesopol, rG_Lestodip)) > 0)
+full.df$vLG_link_div <- diversity(with(full.df, cbind(vLG_Platy, vLG_Tory, vLG_Mesopol, vLG_Mymarid)))
+full.df.sub <- full.df %>%
+  filter(vLG_abund > 0) %>%
+  mutate(vLG_abund.sub.log = log(vLG_abund))
+ggplot(full.df.sub, aes(x = vLG_abund, y = vLG_link_rich)) + geom_point() + geom_smooth(method = "glm", family = "poisson") + scale_y_continuous(lim = c(0,3.5))
+
+summary(glm(vLG_link_rich ~ log(vLG_abund), filter(full.df, vLG_abund > 0), family = poisson))
+summary(glm(rG_link_rich ~ (rG_abund), filter(full.df, rG_abund > 0), family = poisson), scale = "response")
+exp(0.11037)
+
+rare.df <- full.df %>%
+  mutate(vLG_link_abund = vLG_Platy + vLG_Tory + vLG_Mesopol + vLG_Mymarid) %>%
+  filter(vLG_link_abund > 1) %>%
+  mutate(vLG_link_rarerich = rarefy(cbind(vLG_Platy, vLG_Tory, vLG_Mesopol, vLG_Mymarid), 2)-1)
+ggplot(rare.df, aes(x = vLG_abund, y = vLG_link_rarerich)) + geom_point() + geom_smooth(method = "lm")
+summary(lm(vLG_link_rarerich ~ vLG_abund, rare.df))
+
+anova(glm(vLG_link_rich ~ Genotype, full.df, family = poisson), test = "Chi")
+vLG_link_rich.abund <- glm(vLG_link_rich ~ vLG_abund, filter(full.df, vLG_abund > 0), family = poisson)
+summary(vLG_link_rich.abund)
+dfun(vLG_link_rich.abund)
+visreg(vLG_link_rich.abund, scale = "response")
+plot(vLG_link_rich ~ vLG_abund, filter(full.df, vLG_abund > 0))
+
+plus.links <- rowSums(full.df[ ,interaxns_noPont])
+rarecurve(filter(full.df[,interaxns_noPont], plus.links > 0))
+visreg(glm(link_richness ~ log(link_abund), filter(full.df, link_abund > 0), family = poisson), scale = "response")
 ```
@@
 ## vLG parasitism
-vLG.ptized.df <- full.df %>%
-  filter(vLG_abund > 0)
-table(vLG.ptized.df$Genotype)
+vLG.ptized.df <- full.predictors %>%
+  mutate(vLG_parasitized = vLG_Tory + vLG_Eulo + vLG_Platy + vLG_Mesopol + vLG_Mymarid)
+table(vLG.ptized.df$Genotype) # data is unbalanced, but this doesn't qualitatively affect the outcome of the glm.
 vLG.ptized.3plus.genos <- c("*","B","D","E","F","H","I","K","L","O","S","V","X","Y","Z")
@@
                   family = 'binomial')
-anova(glm.binomial.disp(vLG.ptized.bin), test = "LR") # just significant, when I account for overdispersion using this method...
+summary(vLG.ptized.bin)
+dfun(vLG.ptized.bin)
+anova(vLG.ptized.bin, test = "LR")
+require(dispmod)
+anova(glm.binomial.disp(vLG.ptized.bin), test = "LR") # still significant after accounting for overdispersion
+visreg(vLG.ptized.bin, scale = "response")
+
+vLG.ptized.pred <- glm(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~ vLG.height.mean,
+                  data = vLG.ptized.df,
+                  family = 'binomial')
+summary(vLG.ptized.pred)
+1-exp(coef(vLG.ptized.pred)[2]) # 25% reduction in the oddes of vLG being parasitized with every one unit increase in gall size.
+dfun(vLG.ptized.pred) # overdispersion
+anova(glm.binomial.disp(vLG.ptized.pred), test = "LR")
+require(visreg)
+visreg(vLG.ptized.pred, scale = "response")
 
 vLG.ptized.bin.glmer <- glmer(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~ (1|Genotype),
-                  data = filter(vLG.ptized.df, Genotype %in% vLG.ptized.3plus.genos), 
+                  data = vLG.ptized.df, # not removing Genotypes with less than 2 replicates
                   family = 'binomial')
 overdisp_fun(vLG.ptized.bin.glmer)
+summary(vLG.ptized.bin.glmer)
 

### 2015-04-06 Rscripts/food web complexity simulation.R

@@ -1,0 +1,115 @@
+source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
+
+require(mvabund)
+require(bipartite)
+
+df <- as.data.frame(tree_level_interaxn_all_plants_traits_size) #%>%
+  filter(Genotype != "U") # never collected any galls and thus gall-ptoid interactions so I removed it from the dataset. 
+genotypes.noU <- levels(df$Genotype)[-21]
+
+full.link.mvabund   <- mvabund(select(df, 
+                                      willow_vLG = vLG_abund,
+                                      willow_rG = rG_abund,
+                                      willow_aSG = aSG_abund,
+                                      willow_SG = SG_abund,
+                                      vLG_Platy, vLG_Mesopol, vLG_Tory, vLG_Eulo, vLG_Mymarid,
+                                      rG_Platy, rG_Mesopol, rG_Tory, rG_Eulo, rG_Lestodip,
+                                      SG_Platy, aSG_Tory))
+full.link.manyglm <- manyglm(full.link.mvabund ~ Genotype, data = df, family = "negative.binomial")
+
+
+predict.allcombos <- function(manyglm.object,
+                              factor.levels,
+                              factor.index, # for extracting column names
+                              predict.type = "response"){
+  
+  # load required libraries
+  require(plyr)
+  require(dplyr)
+  
+  predict.list <- list()
+  samples.vector <- c()
+  for(i in 2:length(factor.levels)){
+    samples <- sort(as.character(sample(factor.levels, i)))
+    tmp.data <- data.frame(samples)
+    colnames(tmp.data) <- names(manyglm.object$data)[factor.index]
+    predict.tmp <- round(predict(manyglm.object, newdata = tmp.data, type = predict.type), 1) # round to prevent pseudoaccuracy for each genotype
+    predict.list[[i]] <- colMeans(predict.tmp) # take mean across genotypes
+    samples.vector[i] <- paste(samples, collapse = " ") # track genotypes used in analysis
+  }
+  # create data frame from list
+  predict.df <- ldply(predict.list) %>%
+    mutate(Number.of.Genotypes = 2:length(factor.levels),
+           Genotype.combo = samples.vector[-1])
+}
+
+N <- 100
+
+full.link.list <- list()
+for(i in 1:N){
+  full.link.list[[i]] <- predict.allcombos(full.link.manyglm, 
[…]

### 2015-04-06 Rscripts/network denddrograms.R

@@ -4,23 +4,52 @@ source('~/Documents/Genotype_Networks/Rscripts/network_management_genotype_level
 library(ggplot2)
 library(ggdendro)
 
-geno.level.net <- genotype_level_interaxn_traits_size[ ,interaxns_noPont]
+geno.level.net <- genotype_level_interaxn_traits_size[-21,interaxns_noPont] # remove Genotype U and focus on interactions
+genotype.labels <- genotype_level_interaxn_traits_size$Genotype[-21]
+levels(genotype.labels)[1] <- "C" # change for plotting aesthetics
+rownames(geno.level.net) <- genotype.labels
 
-geno.members <- table(tree_level_interaxn_all_plants_traits_size$Genotype)
+hc.net <- hclust(d = vegdist(geno.level.net, method = "bray")*100, # multiplied by 100 to interpret as percent dissimilarity
+                 method = "ave")
 
-plot(hclust(d = dist(geno.level.net)^2, method = "cen", members = geno.members), 
-     labels = genotype_level_interaxn_traits_size$Genotype, hang = -1)
+## Currently the best dendrogram for plotting the networks and their relationship to each other.
+dhc.net <- as.dendrogram(hc.net, hang = 0.05)
+#require(ape)
+#plot(as.phylo(hc.net), label.offset = 0.01)
+#plot(dhc.net, main = "", ylab = "Bray-Curtis Dissimilarity", xlab = "", ylim = c(0,100),  leaflab = "perpendicular")
+plot(dhc.net, horiz = TRUE, xlab = "Percent dissimilarity", xlim = c(100,0), dLeaf = -1)
 
-plot(hclust(d = dist(geno.level.net), method = "ave", members = geno.members), 
-     labels = genotype_level_interaxn_traits_size$Genotype, hang = -1)
 
-t <- hclust(d = dist(geno.level.net), method = "ward", members = geno.members)#, 
+## everything below this is old
+dhc.net.df <- dendro_data(dhc.net)
+
+label.positions <- which(segment(dhc.net.df)$x %in% 1:25)
+dhc.net.df$segments$label.positions <- ""
+dhc.net.df$segments$label.positions[label.positions] <- as.character(dhc.net.df$labels$label)
+
+ggplot(data = segment(dhc.net.df), aes(x = x, y = y, xend = xend, yend = yend)) +
+  geom_segment()  +
[…]

### 2015-04-06 Rscripts/network_management_tree_level.R

@@ -3,8 +3,9 @@
 ## Email: barbour@zoology.ubc.ca
 
 ## upload packages for data manipulation. order of libraries is important. always load dplyr last.
-library(reshape2)
 library(reshape)
+library(reshape2)
+
 library(plyr)
 library(dplyr)
 
@@ -162,7 +163,7 @@ tree_level_gall.count.cast <- dcast(tree_level_gall.size.df, Genotype + plant.po
 tree_level_gall.size.count.join <- left_join(tree_level_gall.size.cast, tree_level_gall.count.cast)
 
 ## merge interaction and tree-trait data
-tree_level_interaxn_all_plants_traits_size <- join_all(list(tree_level_interaxn_all_plants, tree_level_traits, tree_level_gall.size.count.join), by = "plant.position") %>%
+tree_level_interaxn_all_plants_traits_size <- join_all(list(tree_level_interaxn_all_plants, tree_level_traits, tree_level_gall.size.count.join)) %>%
   tbl_df()
 
 write.csv(tree_level_interaxn_all_plants_traits_size, '~/Documents/Genotype_Networks/data/tree_level_interaxn_all_plants_traits_size.csv')

### 2015-04-06 Rscripts/simulation plots.R

@@ -1,0 +1,73 @@
+## load required libraries
+require(visreg)
+require(ggplot2)
+require(grid)
+#require(png)
+
+## upload results from simulation
+all.measures <- read.csv("~/Documents/Genotype_Networks/data/food web complexity simulation.csv")
+dim(all.measures)[1] # 2221 unique simulations. 2425 simulations originally run
+
+## upload png of food web composition ordination
+#ord.png <- readPNG("~/Documents/Genotype_Networks/figures/full.link.composition.png")
+#ord.rast <- rasterGrob(ord.png, interpolate = TRUE)
+
+## plot the relationship between genetic variation and total food web complexity (weighted linkage density)
+ggplot(all.measures, 
+       aes(x = Number.of.Genotypes, y = total_complexity)) + 
+  geom_jitter(shape = 1, color = "grey", position = position_jitter(width = 0.15, height = NULL)) + 
+  stat_summary(fun.y = mean, geom = "point", color = "steelblue", size = 8) + 
+  xlab("Genetic variation (no. of genotypes)") + ylab("Food web complexity index") +
+  scale_x_continuous(limits = c(0,25), breaks = 1:25) +
+  scale_y_continuous(limits = c(1,2.4), breaks = seq(1,2.5, by = 0.25)) +
+  theme_bw() + 
+  theme(axis.text = element_text(size = 14),
+        axis.title.x = element_text(size = 16, vjust = 0.1),
+        axis.title.y = element_text(size = 16, vjust = 1),
+        panel.grid = element_blank()) #+
+  #annotation_custom(ord.rast, xmin = 16, xmax = 26, ymin = 1, ymax = 2)
+
+
+## plot the relationship between genetic variation and willow-gall complexity (weighted linkage density)
+ggplot(all.measures, 
+       aes(x = Number.of.Genotypes, y = link.density.plant_gall)) + 
+  geom_jitter(shape = 1, color = "grey", position = position_jitter(width = 0.15, height = NULL)) + 
+  stat_summary(fun.y = mean, geom = "point", color = "steelblue", size = 8) + 
+  xlab("Genetic variation (no. of genotypes)") + ylab("Food web complexity index") +
+  scale_x_continuous(limits = c(0,25), breaks = 1:25) +
+  scale_y_continuous(limits = c(1,2.5), breaks = seq(1,2.5, by = 0.25)) +
+  theme_bw() + 
+  theme(axis.text = element_text(size = 14),
[…]

### 2015-04-06 gall_community_size_plot.R

@@ -66,6 +66,16 @@ mean(gall.comm.bray) # on average, gall communities were 52% dissimilar from eac
 sd(gall.comm.bray)
 range(gall.comm.bray)
 
+# compare to average dissimilarity estimates using dissimilarity index at tree level.
+galls <- full.df %>%
+  select(Genotype, vLG_abund, rG_abund, aSG_abund, SG_abund) %>%
+  mutate(total_abund = vLG_abund + rG_abund + aSG_abund + SG_abund) %>%
+  filter(total_abund > 0)
+
+galls.geno.dist <- meandist(vegdist(galls[ ,-1], "bray"), galls$Genotype)
+summary(galls.geno.dist) # on average, gall communities were 57% dissimilar from each other
+#plot(galls.geno.dist, cluster = "average")
+
 # 
 gall.size.summary <- gall.size.df %>%
   group_by(Genotype) %>%

### 2015-04-06 link_composition_plots.R

@@ -1,0 +1,354 @@
+source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
+source('~/Documents/miscellaneous_R/ggplot_themes.R')
+source('~/Documents/ggnet/bipartite_plot_info.R')
+
+require(ggplot2)
+require(gridExtra)
+require(dplyr)
+require(tidyr)
+
+## create theme for figures
+theme_links <- theme_bw() +
+  theme(axis.title.x = element_text(size = 16, vjust = 0.1), 
+        axis.title.y = element_text(size = 16, vjust = 1),
+        axis.text = element_text(size = 14),
+        strip.text = element_text(size = 14),
+        panel.grid = element_blank(),
+        legend.position = "none")
+point.size.range <- c(1,8)
+label.x.pos <- 1.5
+label.y.galls <- 23
+label.y.gallsize <- 14
+label.y.ptoids <- 9.5
+label.y.ptism <- 1
+ABCD.allgalls <- data.frame(x = rep(label.x.pos,4), y = rep(label.y.galls,4), 
+                            variable = c("Leaf gall", "apical-Stem gall", 
+                                         "Bud gall", "mid-Stem gall"), 
+                            labels = c("(A)","(B)","(C)","(D)"))
+AB.domgalls <- data.frame(x = rep(label.x.pos,2), y = rep(label.y.galls,2), 
+                          variable = c("Leaf gall","Bud gall"), 
+                          labels = c("(A)","(B)"))
+C.gallsize <- data.frame(x = label.x.pos, y = label.y.gallsize, labels = "(C)")
+ABC.domptoids <- data.frame(x = rep(label.x.pos,3), y = rep(label.y.ptoids,3), 
+                            Parasitoid = c("Platygaster", "Mesopolobus", "Torymus"), 
+                            labels = c("(A)","(B)","(C)"))
+D.ptism <- data.frame(x = label.x.pos, y = label.y.ptism, labels = "(D)")
+A.linkabund <- data.frame(x = 0.5, y = 10, labels = "(A)")
+B.ptism <- data.frame(x = 4.25, y = label.y.ptism, labels = "(B)")
+#point.size <- 6
+line.widths <- 3 # for link plots
+## create color-blind friendly palette with grey (taken from http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/)
[…]

### 2015-04-06 ms_willow_network_supplement.Rmd
created figure for simulation and ordination complementarity

@@
 interaxns_noPont # all gall-parasitoid interactions used for analysis
+```
+
+Qualitative differences in gall community composition?
+```{r vegan galls}
+gall.comm <- full.df %>% select(Genotype, vLG_abund, rG_abund, aSG_abund, SG_abund)
+gall.comm.sub <- filter(gall.comm, rowSums(gall.comm[ ,-1]) > 0, 
+                        Genotype != "J", Genotype != "N", Genotype != "U")
+table(gall.comm.sub$Genotype)
+
+gall.comm.raupcrick <- raupcrick(gall.comm.sub[ ,-1])
+
+adonis(gall.comm.raupcrick ~ Genotype, gall.comm.sub)
+
 ```

### 2015-04-06 Rscripts/simulation plots.R

@@ -2,31 +2,49 @@
 require(visreg)
 require(ggplot2)
 require(grid)
+require(dplyr)
+require(gridBase)
 #require(png)
 
 ## upload results from simulation
 all.measures <- read.csv("~/Documents/Genotype_Networks/data/food web complexity simulation.csv")
 dim(all.measures)[1] # 2221 unique simulations. 2425 simulations originally run
 
+single.complexity.max <- max(filter(all.measures, Number.of.Genotypes == 1)$total_complexity, na.rm = TRUE)
+
 ## upload png of food web composition ordination
 #ord.png <- readPNG("~/Documents/Genotype_Networks/figures/full.link.composition.png")
 #ord.rast <- rasterGrob(ord.png, interpolate = TRUE)
 
+# setup multipanel plot for combining ggplot and base graphics (ordination from "link_composition_plots")
+par(mfrow = c(1,2))
+plot.new()
+vps <- baseViewports()
+pushViewport(vps$figure)
+vp1 <- plotViewport(margins = c(2.4, 0.75, 3.1, 0) - 1)#c(1.8, 1, 0, 1))
+
+A.total <- data.frame(x = 0.5, y = 2.4, labels = "(A)")
+
 ## plot the relationship between genetic variation and total food web complexity (weighted linkage density)
-ggplot(all.measures, 
+total <- ggplot(all.measures, 
        aes(x = Number.of.Genotypes, y = total_complexity)) + 
+  geom_hline(yintercept = single.complexity.max, linetype = "dashed") +
   geom_jitter(shape = 1, color = "grey", position = position_jitter(width = 0.15, height = NULL)) + 
-  stat_summary(fun.y = mean, geom = "point", color = "steelblue", size = 8) + 
-  xlab("Genetic variation (no. of genotypes)") + ylab("Food web complexity index") +
+  stat_summary(fun.y = mean, geom = "point", color = "steelblue", size = 5) + 
+  xlab("Genotypic diversity (no. of genotypes)") + ylab("Food web complexity index") +
   scale_x_continuous(limits = c(0,25), breaks = 1:25) +
   scale_y_continuous(limits = c(1,2.4), breaks = seq(1,2.5, by = 0.25)) +
+  geom_text(data = A.total, aes(x = x, y = y, label = labels), 
+            inherit.aes = FALSE, size = 6) +
   theme_bw() + 
-  theme(axis.text = element_text(size = 14),
+  theme(axis.text.y = element_text(size = 12),
[…]

### 2015-04-06 link_composition_plots.R

@@ -330,14 +330,14 @@ rownames(centroids.cap.geno) <- levels(full.links.df$Genotype)[-c(10,14,21)] # r
 
 ## generate plot. Coded out the manner in which I saved the plot
 
-png(file="~/Documents/Genotype_Networks/figures/full.link.composition.png",width=8, height=8, units = "in", res = 300)
+#png(file="~/Documents/Genotype_Networks/figures/full.link.composition.png",width=8, height=8, units = "in", res = 300)
 plot.new()
-plot.window( xlim=c(-1.75,1.5), ylim = c(-1.5,2.5), asp = 1) 
+plot.window( xlim=c(-1.75,1.5), ylim = c(-1.5,2.5), asp = 1, cex.axis = 0.8) 
 abline(h=0, lty="dotted")
 abline(v=0, lty="dotted")
 axis(side=1, tck = -0.015, padj=-0.5)
 axis(side=2, tck= -0.015, las=1)
-title(xlab="CAP 1", ylab="CAP 2")
+title(xlab="CAP 1", ylab="CAP 2", line = 1.75)
 box()
 
 # fill in the plot
@@ -347,8 +347,9 @@ ordiellipse(cap.geno, groups = full.links.df.sub$Genotype,
             border = NA)
 #ordiellipse(cap.geno, groups = full.links.df.sub$Genotype, kind = "se", show.groups="X", col = "grey", draw = "polygon") # code for highlighting specific groups
 text(x = centroids.cap.geno[ ,1], centroids.cap.geno[ ,2], labels = row.names(centroids.cap.geno))
+text(x = -1.5, y = 3, labels = "(B)")
 
-dev.off() # turn off pdf device.
+#dev.off() # turn off pdf device.
 
 
 

### 2015-04-18 ms_willow_network_supplement.Rmd
made supplement and updated main manuscript

--- 2015-03-06 ms_willow_network_supplement.Rmd
@@
 
-The code below replicates the analyses for the qualitative and quantitative network data.
-
-```{r prepare, echo=FALSE}
+```{r prepare, include = FALSE, echo=FALSE}
 ## source in required datasets and functions. 
@@
 source('~/Documents/miscellaneous_R/model_diagnostic_functions.R')
-source('~/Documents/miscellaneous_R/jensen_magnitude_function.R')
+#source('~/Documents/miscellaneous_R/jensen_magnitude_function.R')
 
 ## load required libraries
-#library(vegan) # for adonis analysis
-#library(knitr) # for making tables
+library(vegan) # for adonis analysis
+library(knitr) # for making tables
 library(ggplot2) # for plotting
-theme_set(theme_classic()) # customize ggplot for prettier default graphs
+theme_set(theme_bw()) # customize ggplot for prettier default graphs
 library(gridExtra) # for grid.arrange
@@
 library(HH) # for variance inflation factor analysis (vif) in multiple regression
+library(visreg) # visualize regression output
 
@@
 
-```{r mv abund galls}
+Testing for whether gall size varies among willow genotypes. We were unable to use the mvabund framework for this analysis, so we conducted separate linear models to test for these effects.
+
+```{r gall size plots, echo=FALSE}
+vLG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = vLG.height.mean)) + 
+  geom_boxplot() + ylab("Iteomyia salicisverruca diameter (mm)")
+rG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = rG.height.mean)) + 
+  geom_boxplot() + ylab("Rabdophaga salicisbrassicoides diameter (mm)")
+aSG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = aSG.height.mean)) + 
+  geom_boxplot() + ylab("Cecidomyid sp. A diameter (mm)")
+SG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = SG.height.mean)) + 
+  geom_boxplot() + ylab("Rabdophaga salicisbattatus diameter (mm)")
+
+grid.arrange(vLG.height.mean.plot, rG.height.mean.plot, aSG.height.mean.plot, SG.height.mean.plot)
+```
+
+There were too few data points to reliably test for differences in mid-Stem gall diameter (Rabdophaga salicisbattatus), but we used weighted linear models to test for differences in gall size among willow genotypes.
+We weighted models by the number of galls collected from each plant replicate, because a higher number would reflect a more accurate measure of mean gall size for that particular willow.
+
+```{r gall size models}
+vLG.size.lm <- lm(log(vLG.height.mean) ~ Genotype, data = full.df, weights = vLG.gall.count)
+rG.size.lm <- lm(log(rG.height.mean) ~ Genotype, data = full.df, weights = rG.gall.count)
+aSG.size.lm <- lm(log(aSG.height.mean) ~ Genotype, data = full.df, weights = aSG.gall.count)
+```
+
+```{r gall size results summary, echo=FALSE}
+anova(vLG.size.lm)
+anova(rG.size.lm)
+anova(aSG.size.lm)
+```
+
+We used multivariate GLMs to examine whether gall abundances varied among willow genotypes.
+
+Boxplots of gall abundance among the different genotypes.
+```{r gall abund across genotypes, echo = FALSE}
+gall.plot.df <- as.data.frame(full.df) %>% # melt doesn't like 'tbl' class
+  select(Genotype, 
+         Iteomyia_salicisverruca = vLG_abund,
+         Cecidomyiid_sp.A = aSG_abund,
+         Rabdophaga_salicisbrassicoides = rG_abund, 
+         Rabdophaga_salicisbattatus = SG_abund) %>%
+  melt()
+
+ggplot(gall.plot.df, aes(x = Genotype, y = value, color = variable)) + 
+  geom_boxplot() + facet_wrap( ~ variable, nrow = 2) + theme_bw() + ylab("Gall density (no. per branch)")
+```
+
+Assessing which error distribution is best to model the data. Negative binomial appear to provide the best fit.
+```{r gall abundance negative binomial assumption, echo=FALSE}
 gall.mvabund <- mvabund(full.df[ ,c("vLG_abund","rG_abund","aSG_abund","SG_abund")])
@@
 quasipoisson_curve(from = 0.1, to = 10, quasi.scalar = 2)
-neg.binomial_curve(from = 0.1, to = 10, theta.negbin = 0.7) # negative binomial provides the best fit to the data.
-```
+neg.binomial_curve(from = 0.1, to = 10, theta.negbin = 0.7) # negative binomial
 
-Assessed negative binomial assumptions
-```{r gall abund negative binomial assumptions}
+## specify the multivariate GLM
 manyglm.gall <- manyglm(gall.mvabund ~ Genotype,
@@
                         family = "negative.binomial")
+
+## examine residual plots of the model.
 plot(manyglm.gall, which = 1:3) # residuals look really good.
@@
 
-Tested whether genotype affected gall community composition
-```{r gall abund test}
-anova.gall <- anova.manyglm(manyglm.gall, p.uni = "adjusted")
-anova.gall # vLG_abund, rG_abund and aSG_abund are driving the community response.
+Tested whether genotype affected gall community composition.
+It appears that leaf and bud galls are the major determinants of variation in gall abundance, with apical-Stem galls contributing to a lesser extent.
+Remember that p-values may slightly differ from those reported in the manuscript since this uses a permutation.
+```{r gall abund test, eval=FALSE}
+anova.gall <- anova.manyglm(manyglm.gall, p.uni = "unadjusted")
+anova.gall 
 ```
 
-Boxplots of gall abundance among the different genotypes.
-```{r gall abund across genotypes, echo = FALSE}
-vLG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = vLG_abund)) + 
-  geom_boxplot() + ylab("Iteomyia salicisverruca density (#/branch)")
-rG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = rG_abund)) + 
-  geom_boxplot() + ylab("Rabdophaga salicisbrassicoides density (#/branch)")
-aSG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = aSG_abund)) + 
-  geom_boxplot() + ylab("Cecidomyid sp. A density (#/branch)")
-SG.abund.plot <- ggplot(data = full.df, aes(x = Genotype, y = SG_abund)) + 
-  geom_boxplot() + ylab("Rabdophaga salicisbattatus density (#/branch)")
-
-grid.arrange(vLG.abund.plot, rG.abund.plot, aSG.abund.plot, SG.abund.plot)
+```{r test gall linear models}
+gall.lins <- manylm(log(gall.mvabund+1) ~ Genotype, data = full.df)
+summary(gall.lins)
+anova(gall.lins, p.uni = "unadjusted") # same qualitative results as GLM
 ```
 
-```{r mvabund gall-trait associations}
+Test for quantitative differences  in gall community composition.
+```{r vegan galls, echo=FALSE}
+gall.comm <- full.df %>% select(Genotype, vLG_abund, rG_abund, aSG_abund, SG_abund)
+gall.comm.sub <- filter(gall.comm, rowSums(gall.comm[ ,-1]) > 0, 
+                        Genotype != "J", Genotype != "N", Genotype != "U")
+table(gall.comm.sub$Genotype) # restricted to Genotypes with greater than 2 replicates for composition analyses.
+
+bray.gall.comm.sub <- vegdist(gall.comm.sub[ ,-1], method = "bray")
+
+adonis(bray.gall.comm.sub ~ Genotype, data = gall.comm.sub)
+anova(betadisper(bray.gall.comm.sub, group = gall.comm.sub$Genotype)) # no differences in dispersion among willow genotypes.
+summary(meandist(bray.gall.comm.sub, grouping = gall.comm.sub$Genotype))
+```
+
+Determine which willow traits are resulting in variation in gall abundances.
+```{r mvabund gall-trait associations, echo=FALSE}
 # selects galls and traits for analysis
@@
 
-```{r stepAIC galls-traits}
-## use stepAIC to identify traits that reduce AIC the most and remove them in a stepwise fashion from analysis.
+Starting with the most complex model, we progressively dropped predictors from the model that resulted in the lowest AIC for the model.
+We continued this in stepwise fashion until we arrived at the null model (no predictors).
+Using AIC and likelihood ratio tests, we found that four of these models provided a better fit to the data than the null model.
+For clarity, we only interpret model 11 as this one had the second to lowest AIC value, yet likelihood ratio tests suggested that it provided a significantly better fit than the alternative models.
+```{r stepAIC galls-traits, echo=FALSE, results='asis'}
+## drop variables that reduce AIC the most and remove them in a stepwise fashion from analysis.
 drop1(mvabund.galls.traits)
@@
                    sum(AIC(mod14.null))))
-arrange(AIC.models.galls.traits, AIC)
+kable(arrange(AIC.models.galls.traits, AIC)[1:5,]) # display all of the models that had lower AICs than the null model (including the null model).
 
-## Likelihood ratio tests suggest that all of the models with lower AIC than the null model provide a better fit to the data than the null.
+## Likelihood ratio tests suggest that all of the models with a lower AIC than the null model provide a better fit to the data than the null.
 anova.manyglm(mod14.null, mod13)
@@
 
+## Likelihood ratio tests suggest that model 11 provides a better fit to the data than the two other closest models
+anova.manyglm(mod13, mod11)
+anova.manyglm(mod12, mod11)
+anova.manyglm(mod10, mod11)
+anova.manyglm(mod14.null, mod13, mod12, mod11, mod10)
+```
+
+```{r identify gall-trait associations, eval=FALSE}
 # identify gall-trait associations.
@@
 
-```{r visreg univariate gall-trait associations}
-library(visreg)
+```{r gall trait lms}
+gall.trait.lms <- manylm(galls_galls.traits ~ C_N_imputed + flavanonOLES.PC1 + log_size, 
+                                data = traits_galls.traits)
+summary(gall.trait.lms)
+anova(gall.trait.lms, p.uni = "unadjusted") # qualitatively the same results as GLM.
+
+best.r.sq(galls_galls.traits ~ traits_galls.traits$C_N_imputed + traits_galls.traits$water_content + traits_galls.traits$sal_tannin.PC1 + traits_galls.traits$cinn.PC1 + traits_galls.traits$cinn.PC2 + traits_galls.traits$flavonOLES.PC1 + traits_galls.traits$flavonOLES.PC2 + traits_galls.traits$flavanonOLES.PC1 + traits_galls.traits$log_size + traits_galls.traits$log_trichomes + traits_galls.traits$Height_resid + traits_galls.traits$Density_resid + traits_galls.traits$SLA_resid) # results are pretty close to GLM, retains flavanonOLES.PC1, C:N, but has cinn.PC1 and log.trichomes before log.size.
+Note that these variables were included in the GLM AIC, so it isn't too surprising.
+Not a big concern overall though.
+```
+
+```{r visreg univariate gall-trait associations, eval=FALSE}
+
 # include model with all of the traits, but only examine one of the variables
@@
 
-```{r}
+```{r gall size trait associations, echo=FALSE}
 ## gall size
@@
 summary(size10)
-summary(size12)
+summary(size12) # final model
 summary(size9)
+
+summary(lm(vLG.height.mean ~ luteolin.der2__A320nm, full.df, weights = full.df$vLG.gall.count))
+
+anova(size14.null, size13, size12, size11, size10, size9, size8, size7)
 
@@
 
-```{r gall abund predict}
-gall.predict <- predict(manyglm.gall, 
-                        newdata = data.frame(Genotype = levels(full.df$Genotype)), 
-                        type = "response")
-gall.predict.df <- cbind.data.frame(gall.predict, Genotype = levels(full.df$Genotype))
-gall.predict.melt <- melt(gall.predict.df)
-
-ggplot(gall.predict.melt, aes(x = Genotype, y = value, fill = variable)) + geom_bar(stat = "identity")
-
-colMeans(gall.predict)
-library(vegan)
-diversity(colMeans(gall.predict))
-
-gall.predict.X.R <- predict(manyglm.gall, 
-                            newdata = data.frame(Genotype = c("X","R")),
-                            type = "response")
-colMeans(gall.predict.X.R)
-diversity(colMeans(gall.predict.X.R))
-
-sample(levels(full.df$Genotype), 5)
-
-names(manyglm.gall$data)
-predict.allcombos <- function(manyglm.object, factor.index, predict.type = "response"){
-  require(plyr)
-  require(dplyr)
-  factor.levels <- levels(manyglm.object$data[ ,factor.index])
-
-  predict.list <- list()
-  for(i in 2:length(factor.levels)){
-    tmp.data <- data.frame(sample(factor.levels, i))
-    colnames(tmp.data) <- names(manyglm.object$data)[factor.index]
-    predict.tmp <- predict(manyglm.object, newdata = tmp.data, type = predict.type)
-    predict.list[[i]] <- colMeans(predict.tmp)
-  }
-  predict.df <- ldply(predict.list) %>%
-    mutate(Number.of.Genotypes = 2:length(factor.levels))
-}
-
-N <- 50
-gall.div.list <- list()
-for(i in 1:N){
-  gall.div.list[[i]] <- round(predict.allcombos(manyglm.gall, factor.index = 2), 1) # rounding to avoid pseudoaccuracy
-} # takes a couple minutes with 50 reps
-
-gall.div.df <- ldply(gall.div.list) %>% tbl_df()
-gall.div.df$diversity <- diversity(select(gall.div.df, vLG_abund:SG_abund))
-
-boxplot(diversity ~ factor(Number.of.Genotypes), gall.div.df, ylab = "Gall diversity", xlab = "Number of Genotypes")
-
-N <- 50
-link.div.list <- list()
-for(i in 1:N){
-  link.div.list[[i]] <- round(predict.allcombos(manyglm.full, factor.index = 2), 1)
-} # takes a couple minutes with 50 reps
-
-link.div.df <- ldply(link.div.list) %>% tbl_df()
-link.div.df$diversity <- diversity(select(link.div.df, aSG_Tory:vLG_Tory))
-link.div.df$repID <- 1:dim(link.div.df)[1]
-
-boxplot(diversity ~ factor(Number.of.Genotypes), link.div.df, ylab = "Interaction Diversity", xlab = "Number of Genotypes")
-
-link.melt <- melt(select(link.div.df, -diversity), id.vars = c("repID", "Number.of.Genotypes"))
-
-link.split <- colsplit(link.melt$variable, "_", names = c("gall","parasitoid"))
-link.melt.df <- cbind(link.melt, link.split)
-
-web.list <- cast(link.melt.df, gall ~ parasitoid | repID + Number.of.Genotypes, mean)
-
-library(bipartite)
-web.list.measures <- c()
-genotype.number <- c()
-for(i in 1:dim(web.list)){
-  genotype.number[i] <- names(web.list[[i]][1])
-  tmp.web <- web.list[[i]][[1]]
-  rownames(tmp.web) <- tmp.web$gall
-  tmp.web <- as.matrix.data.frame(tmp.web[ ,-1])
-  web.list.measures[i] <- networklevel(tmp.web, index = "linkage density", weighted = TRUE)
-}
-web.measures.df <- cbind.data.frame(Number.of.Genotypes = as.numeric(genotype.number),
-                                    Weighted.Linkage.Density = web.list.measures)
-
-boxplot(Weighted.Linkage.Density ~ factor(Number.of.Genotypes), web.measures.df, ylab = "Weighted Linkage Density", xlab = "Number of Genotypes")
-
-web.measures.summary <- web.measures.df %>%
-  group_by(Number.of.Genotypes) %>%
-  summarise(median.link.density = median(Weighted.Linkage.Density))
-with(web.measures.summary, max(median.link.density)/min(median.link.density))
-
-library(mgcv)
-web.sum.lm <- lm(log(median.link.density) ~ log(Number.of.Genotypes), data = web.measures.summary)
-summary(web.sum.lm) # Need to understand interpretation here.
-visreg(web.sum.lm) # not quite there
-```
-
-We were unable to use the mvabund framework for this analysis, so we conducted separate linear models to test for these effects.
-```{r gall size}
-plot(vLG.height.mean ~ Genotype, full.df)
-vLG.size.lm <- lm(log(vLG.height.mean) ~ Genotype, filter(full.df, Genotype != "J"))
-#plot(vLG.size.lm)
-summary(vLG.size.lm)
-
-vLG.size.lmer <- lmer(log(vLG.height.mean) ~ (1|Genotype), full.df)
-heritability(vLG.size.lmer) # significant. Note though, that I originally valuate gall size using nested random effect models to account for individual measurements I had for each gall.
-
-plot(rG.height.mean ~ Genotype, full.df)
-rG.size.lm <- lm(log(rG.height.mean) ~ Genotype, full.df)
-#plot(rG.size.lm)
-summary(rG.size.lm)
-
-plot(aSG.height.mean ~ Genotype, full.df) # too few data points to reliably evaluate
-aSG.size.lm <- lm(log(aSG.height.mean) ~ Genotype, full.df)
-#plot(aSG.size.lm)
-summary(aSG.size.lm)
-
-plot(SG.height.mean ~ Genotype, full.df) # too few data points to reliably evaluate
-SG.size.lm <- lm(log(SG.height.mean) ~ Genotype, full.df)
-plot(SG.size.lm)
-summary(SG.size.lm)
-```
-
-```{r gall size variation among willow genotypes}
-vLG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = vLG.height.mean)) + 
-  geom_boxplot() + ylab("Iteomyia salicisverruca diameter (mm)")
-rG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = rG.height.mean)) + 
-  geom_boxplot() + ylab("Rabdophaga salicisbrassicoides diameter (mm)")
-aSG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = aSG.height.mean)) + 
-  geom_boxplot() + ylab("Cecidomyid sp. A diameter (mm)")
-SG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = SG.height.mean)) + 
-  geom_boxplot() + ylab("Rabdophaga salicisbattatus diameter (mm)")
-
-grid.arrange(vLG.height.mean.plot, rG.height.mean.plot, aSG.height.mean.plot, SG.height.mean.plot)
-```
-
 Here, we evaluate the assumptions of mvabund and see which error distribution is appropriate.
@@
 It is easy to see that the negative binomial model (black line) provides the best fit to this data, suggesting that we should specify this as the error distribution in our model
-```{r mv abund interactions explore}
+```{r mv abund interactions explore, echo=FALSE}
 full.mvabund <- mvabund(full.df[ ,interaxns_noPont]) # create mvabund object for easy analysis
-meanvar.plot(full.mvabund ~ full.df$Genotype) # easy to see that the data do not have a poisson distribution, therefore, a negative binomial may be a good fit.
+full.meanvar <- meanvar.plot(full.mvabund ~ full.df$Genotype, table = TRUE) # easy to see that the data do not have a poisson distribution, therefore, a negative binomial may be a good fit.
 poisson_curve(from = 0.1, to = 3.0, color = "blue")
@@
 neg.binomial_curve(from = 0.1, to = 3.0, theta.negbin = 0.7, color = "black") # negative binomial appears to provide the best fit to this relationship
-```
 
-```{r genotype level}
-geno.level.interactions.df <- full.df %>%
-  select(Genotype, aSG_Tory:rG_Platy, rG_Tory, SG_Platy, vLG_Eulo:vLG_Tory, aSG_abund:vLG_abund, vLG.height.mean) %>%
-  group_by(Genotype) %>%
-  summarise_each(funs(mean.na.rm = mean(., na.rm = TRUE))) %>%
-  #na.omit() %>%
-  mutate(log.vLG_abund = log(vLG_abund),
-         log.vLG.height.mean = log(vLG.height.mean),
-         log.1.rG_abund = log(rG_abund+1),
-         log.1.aSG_abund = log(aSG_abund+1))
-
-library(psych)
-corr.test(geno.level.interactions.df[ ,c("vLG_abund","rG_abund","aSG_abund","vLG.height.mean")])
-
-car::scatterplotMatrix(select(geno.level.interactions.df, vLG_abund, rG_abund, aSG_abund, vLG.height.mean))
-corr.test(geno.level.interactions.df[ ,c("log.vLG_abund","log.1.rG_abund","log.1.aSG_abund","log.vLG.height.mean")])
-
-# correlations are rather weak
-corr.test(select(geno.level.interactions.df, aSG_Tory:vLG_Tory))
-corr.test(select(geno.level.interactions.df, vLG_Platy, vLG_Mesopol, vLG_Tory, rG_Tory))
-car::scatterplotMatrix(select(geno.level.interactions.df, vLG_Platy, vLG_Mesopol, vLG_Tory, rG_Tory))
-
-plot(rG_Eulo ~ vLG_Mymarid, geno.level.interactions.df)
-
-plot(log.1.aSG_abund ~ log.vLG_abund, geno.level.interactions.df)
-
-geno.level.mvabund <- mvabund(geno.level.interactions.df[ ,interaxns_noPont])
-
-geno.level.manyglm <- manylm(geno.level.mvabund ~ vLG_abund*vLG.height.mean + rG_abund + aSG_abund, data = geno.level.interactions.df)
-
-plot(geno.level.manyglm, which = 1:3)
-anova.manylm(geno.level.manyglm, p.uni = "unadjusted")
-
-hist(geno.level.interactions.df$vLG_Platy)
-plot(vLG_Mesopol ~ vLG.height.mean, geno.level.interactions.df)
-vLG_Platy.lm <- lm(vLG_Mesopol ~ vLG_abund+vLG.height.mean, geno.level.interactions.df)
-summary(vLG_Platy.lm)
-plot(vLG_Platy.lm)
-AIC(vLG_Platy.lm)
-library(visreg)
-visreg(vLG_Platy.lm, scale = "response")
+#no.relationships <- which(names(colSums(full.mvabund)) %in% c("rG_Lestodip", "rG_Mesopol", "SG_Platy", "vLG_Eulo"))
+#sum(colSums(full.mvabund)[no.relationships])/sum(colSums(full.mvabund)) # less than 13% of the links in the network had relationships to gall abundance and size (see results below)
 ```
@@
 
-```{r}
-link.predict <- predict(manyglm.full, 
-                        newdata = data.frame(Genotype = levels(full.df$Genotype)), 
-                        type = "response")
-link.predict.df <- cbind.data.frame(link.predict, Genotype = levels(full.df$Genotype))
-library(vegan)
-mean(vegdist(select(link.predict.df, -Genotype), method = "horn"))
-sd(vegdist(select(link.predict.df, -Genotype), method = "horn"))
-hist(vegdist(select(link.predict.df, -Genotype), method = "horn"))
-
-link.predict.melt <- melt(link.predict.df)
-
-ggplot(link.predict.melt, aes(x = Genotype, y = value, fill = variable)) + geom_bar(stat = "identity")
-```
-
 Given that a negative binomial distribution seems to provide a good fit to the data, we tested whether the composition of gall-parasitoid interaction varied among willow genotypes.
@@
 Moreover, differences in community composition are driven primarily by 3 interactions: vLG_Platy, vLG_Tory, and vLG_Mesopol.
-```{r}
+```{r, eval=FALSE}
 anova.full <- anova.manyglm(manyglm.full, p.uni = "unadjusted") # Takes about 1 min and 30 sec to run.
-anova.full # vLG_Platy, vLG_Tory, and vLG_Mesopol are driving the community response. rG_Tory is marginally significant and vLG_Eulo is close to marginal as well.
+anova.full # vLG_Platy, vLG_Tory, and vLG_Mesopol are driving the community response. rG_Tory is marginally significant and vLG_Eulo is close to marginal as well. aSG_Tory is significant as well.
+```
+
+```{r test linear models link abundance}
+link.lins <- manylm(log(full.mvabund+1) ~ Genotype, data = full.df)
+summary(link.lins)
+anova(link.lins, p.uni = "unadjusted") # qualitatively same results as GLM
+```
+
+```{r composition analysis of gall-parasitoid links, echo=FALSE}
+link.comm <- cbind.data.frame(Genotype = full.df$Genotype, full.df[ ,interaxns_noPont])
+
+## filtered the data so that all willows have at least one trophic interaction (necessary for quantifying dissimilarity) and that the Genotypes included have 3 or more replicates
+link.comm.sub <- filter(link.comm, rowSums(link.comm[ ,-1]) > 0, # must have at least one trophic interaction
+                        Genotype %in% c("*","B", "D", "I", "K", "L", 
+                                        "Q", "S", "V", "W", "X", "Y", "Z")) 
+table(link.comm.sub$Genotype)
+
+bray.link.comm.sub <- vegdist(link.comm.sub[ ,-1], method = "bray")
+
+adonis(bray.link.comm.sub ~ Genotype, data = link.comm.sub)
+anova(betadisper(bray.link.comm.sub, group = link.comm.sub$Genotype)) # no differences in dispersion among willow genotypes.
+summary(meandist(bray.link.comm.sub, grouping = link.comm.sub$Genotype))
 ```
@@
 First we created a dataset that contained complete observations of the network and predictor variables
-```{r mv abund interaction predictors}
+```{r mv abund interaction predictors, echo=FALSE}
 full.predictors <- full.df %>%
@@
 dim(full.predictors)[1] # 81 data points
+
+net.trait <- mvabund(full.predictors[ ,interaxns_noPont])
 ```
@@
 We then looked for evidence of variance inflation among the predictor variables. But found little evidence for it.
-```{r vif}
-vif(full.predictors[ ,c("log.1.aSG_abund","log.1.rG_abund","log.vLG_abund","log.vLG.height.mean")]) # still little evidence of variance inflation among predictor variables.
+CURRENTLY NOT WORKING
+```{r vif, echo=FALSE}
+vif(xx = as.data.frame(full.predictors[ ,c("log.1.aSG_abund","log.1.rG_abund","log.vLG_abund","vLG.height.mean")])) # Little evidence of variance inflation among predictor variables.
 ```
@@
 Using AIC, we identified 3 equivalenet models (difference in AIC < 2.1 amongst models).
-```{r AIC model assessment}
-net.trait <- mvabund(full.predictors[ ,interaxns_noPont])
-net.trait.sub <- mvabund(full.predictors[ ,c("vLG_Platy","vLG_Mesopol","vLG_Tory","rG_Tory")])
-# even though manyglm doesn't appear to test interaction terms appropriately (sequential vs.
-dropping different terms), there is not interactive effect of vLG height and vLG abundance in univariate negative binomial models for any vLG_parasitoid interaction, so I believe this final model is appropriate for the data.
+```{r AIC model assessment, echo=FALSE}
+
 net.mvabund.1.full <- manyglm(net.trait ~ log.vLG_abund*vLG.height.mean + 
@@
                        family = "negative.binomial")
-MASS::stepAIC(net.mvabund.1.full, direction = "both")
 
@@
 We then examined the which processes underlied variation in the gall-parasitoid interactions.
-```{r}
+```{r, eval=FALSE}
 anova.net <- anova.manyglm(net.mvabund.2, p.uni = "unadjusted") # takes about 2.5 min to run.
@@
 Therefore, we retain all P-values < 0.10 (during at least one run) for the coefficient summary below.
+
+t(coef(net.mvabund.2))[,-1][c(5,4,6,2,3,1,7),c(2,1,3,4)]
+
+anova.net.null <- anova.manyglm(net.mvabund.2, net.mvabund.6.null) # takes about 30 sec to run
+anova.net.null
 
@@
                                  "log.vLG_abund_vLG_Mymarid",
-                                 "log.vLG.height.mean_vLG_Eulo", # marginal
-                                 "log.vLG.height.mean_rG_Tory", "log.1.rG_abund_rG_Tory",
-                                 "log.vLG_abund_vLG_Mesopol", "log.vLG.height.mean_vLG_Mesopol", 
-                                 "log.vLG.height.mean_vLG_Platy", "log.vLG_abund_vLG_Platy",
-                                 "log.vLG.height.mean_vLG_Tory", "log.vLG_abund_vLG_Tory")) %>% # vLG.height.mean is marginal for vLG_Tory
+                                 "vLG.height.mean_vLG_Eulo", # marginal
+                                 "vLG.height.mean_rG_Tory", "log.1.rG_abund_rG_Tory",
+                                 "vLG_abund_vLG_Mesopol", "vLG.height.mean_vLG_Mesopol", 
+                                 "vLG.height.mean_vLG_Platy", "log.vLG_abund_vLG_Platy",
+                                 "vLG.height.mean_vLG_Tory", "log.vLG_abund_vLG_Tory")) %>% # vLG.height.mean is marginal for vLG_Tory
   select(predictor = X1, response = X2, coefficient = value)
 arrange(sig.coef.df.2, predictor, response)
-```
-
-```{r Interaction predictions}
-# calculate the mean values, on the original scale, of all predictor variables
-no.var.data <- with(full.predictors,
-                    data.frame(vLG_abund = mean(vLG_abund),
-                               rG_abund = mean(rG_abund),
-                               aSG_abund = mean(aSG_abund),
-                               vLG.height.mean = mean(vLG.height.mean)))
-
-# http://www.ats.ucla.edu/stat/mult_pkg/faq/general/log_transformed_regression.htm Use this web page for interpreting log and non-transformed coefficients in regression.
-library(visreg)
-par(mfrow = c(1,2))
-vLG_Platy.glm <- MASS::glm.nb(vLG_Platy ~ log(vLG_abund) + vLG.height.mean, full.predictors); summary(vLG_Platy.glm); jensen_magnitude(vLG_Platy.glm, no.var.data)
-visreg(vLG_Platy.glm, scale = "response")
-#plot(vLG_Platy.glm)
-
-vLG_Mesopol.glm <- MASS::glm.nb(vLG_Mesopol ~ log(vLG_abund) + vLG.height.mean, full.predictors); summary(vLG_Mesopol.glm); jensen_magnitude(vLG_Mesopol.glm, no.var.data)
-#plot(vLG_Mesopol.glm)
-visreg(vLG_Mesopol.glm, scale = "response")
-
-vLG_Tory.glm <- MASS::glm.nb(vLG_Tory ~ vLG_abund+vLG.height.mean, full.predictors); summary(vLG_Tory.glm); jensen_magnitude(vLG_Tory.glm, no.var.data) # gall height is marginally significant, but the AIC is slightly better than the model with only vLG_abund.
-visreg(vLG_Tory.glm, scale = "response")
-
-rG_Tory.glm <- glm(rG_Tory ~ rG_abund + vLG.height.mean, full.predictors, family = quasipoisson); summary(rG_Tory.glm); jensen_magnitude(rG_Tory.glm, no.var.data) # underdispersed, which is why I used quasipoisson.
-#plot(rG_Tory.glm)
-visreg(rG_Tory.glm, scale = "response")
-
-rG_Platy.glm <- glm(rG_Platy ~ rG_abund, full.predictors, family = quasipoisson); summary(rG_Platy.glm); jensen_magnitude(rG_Platy.glm, no.var.data)  # underdispersed, which is why I used quasipoisson.
-
-rG_Eulo.glm <- glm(rG_Eulo ~ rG_abund, full.predictors, family = quasipoisson); summary(rG_Eulo.glm); jensen_magnitude(rG_Eulo.glm, no.var.data) # underdispersed, which is why I used quasipoisson
-
-aSG_Tory.glm <- glm(aSG_Tory ~ aSG_abund, full.predictors, family = quasipoisson); summary(aSG_Tory.glm); jensen_magnitude(aSG_Tory.glm, no.var.data) # underdispersed, so I used quasipoisson.
-
-rG_Lestodip.glm <- glm(rG_Lestodip ~ log(rG_abund + 1), full.predictors, family = quasipoisson)
-summary(rG_Lestodip.glm); jensen_magnitude(rG_Lestodip.glm, no.var.data) # rG_abund is marginally significant.
-
-## intercept only models.
-vLG_Eulo.glm <- glm(vLG_Eulo ~ 1, full.predictors, family = poisson); summary(vLG_Eulo.glm); jensen_magnitude(vLG_Eulo.glm, no.var.data) # not a lot of support for including gall height as a predictor, althought it was marginally significant in multivariate model, so I dropped it from the model
-
-SG_Platy.glm <- glm(SG_Platy ~ 1, full.predictors, family = poisson); summary(SG_Platy.glm); jensen_magnitude(SG_Platy.glm, no.var.data) # no evidence of vLG_abund being a clear predictor, so we dropped it from the model
-
-rG_Mesopol.glm <- glm(rG_Mesopol ~ 1, full.predictors, family = poisson); summary(rG_Mesopol.glm); jensen_magnitude(rG_Mesopol.glm, no.var.data)
-
-vLG_Mymarid.glm <- glm(vLG_Mymarid ~ 1, full.predictors, family = poisson); summary(vLG_Mymarid.glm); jensen_magnitude(vLG_Mymarid.glm, no.var.data) # vLG_abund is a significant predictor, but the visreg predictions look so wonky that I don't really trust it, so I just modelled it with an intercept
-```
-
-```{r visreg plots}
-# Platygaster interactions
-layout(matrix(c(1, 2, 3, 4), nrow = 2, ncol = 2, byrow = FALSE))
-visreg(vLG_Platy.glm, scale = "response", ask = F)
-plot(x = range(full.predictors$vLG_abund), y = c(0, 1), type = "n", ylab = "SG_Platy", xlab = "vLG_abund"); abline(h = exp(coef(SG_Platy.glm)), col = "steelblue", lwd = 3) # intercept only models
-visreg(rG_Platy.glm, scale = "response", ask = F)
-
-# Mesopolobus interactions
-layout(matrix(c(1, 2, 3, 0), nrow = 2, ncol = 2, byrow = FALSE))
-visreg(vLG_Mesopol.glm, scale = "response", ask = F)
-plot(x = range(full.predictors$rG_abund), y = c(0, 1), type = "n", ylab = "rG_Mesopol", xlab = "rG_abund"); abline(h = exp(coef(rG_Mesopol.glm)), col = "steelblue", lwd = 3) # intercept only models
-
-# Torymus interactions
-layout(matrix(c(1, 2, 3, 4, 5, 0), nrow = 2, ncol = 3, byrow = FALSE))
-visreg(vLG_Tory.glm, scale = "response", ask = F)
-visreg(rG_Tory.glm, scale = "response", ask = F)
-visreg(aSG_Tory.glm, scale = "response", ask = F)
-
-# Eulophid interactions and other less abundant ones
-layout(matrix(c(1, 2, 3, 4), nrow = 2, ncol = 2, byrow = FALSE))
-plot(x = range(full.predictors$vLG_abund), y = c(0, 1), type = "n", ylab = "vLG_Eulo", xlab = "vLG_abund"); abline(h = exp(coef(vLG_Eulo.glm)), col = "steelblue", lwd = 3)
-visreg(rG_Eulo.glm, scale = "response", ask = F)
-plot(x = range(full.predictors$vLG_abund), y = c(0, 1), type = "n", ylab = "vLG_Mymarid", xlab = "vLG_abund"); abline(h = exp(coef(vLG_Mymarid.glm)), col = "steelblue", lwd = 3)
-visreg(rG_Lestodip.glm, scale = "response", ask = F)
-
-par(mfrow = c(1,1))
-```
-
-```{r table of jensen effects}
-jensen.df <- rbind_all(list(jensen_magnitude(vLG_Platy.glm, no.var.data),
-               jensen_magnitude(vLG_Mesopol.glm, no.var.data),
-               jensen_magnitude(vLG_Tory.glm, no.var.data),
-               jensen_magnitude(vLG_Eulo.glm, no.var.data),
-               jensen_magnitude(vLG_Mymarid.glm, no.var.data),
-               jensen_magnitude(rG_Tory.glm, no.var.data),
-               jensen_magnitude(rG_Mesopol.glm, no.var.data),
-               jensen_magnitude(rG_Platy.glm, no.var.data),
-               jensen_magnitude(rG_Mesopol.glm, no.var.data),
-               jensen_magnitude(rG_Lestodip.glm, no.var.data),
-               jensen_magnitude(aSG_Tory.glm, no.var.data),
-               jensen_magnitude(SG_Platy.glm, no.var.data)))
-jensen.df <- mutate(jensen.df, Interactions = c("vLG_Platy","vLG_Mesopol","vLG_Tory","vLG_Eulo","vLG_Mymarid","rG_Tory","rG_Mesopol","rG_Platy","rG_Mesopol","rG_Lestodip","aSG_Tory","SG_Platy"))
-plot(log(Magnitude.Difference) ~ log(mean.no.var), filter(jensen.df, Magnitude.Difference != "1"))
-visreg(lm(log(Magnitude.Difference) ~ log(mean.no.var), filter(jensen.df, Magnitude.Difference != "1")), scale = "response", xtrans = log, xlab = "log(mean.no.var)")
-
-interactions.split <- colsplit(jensen.df$Interactions, "_", names = c("gall","parasitoid"))
-jensen.df <- cbind(jensen.df, interactions.split)
-
-var.web <- cast(jensen.df, gall ~ parasitoid, mean, value = "mean.with.var")
-rownames(var.web) <- var.web$gall
-var.web <- as.matrix.data.frame(var.web[ ,-1])
-
-no.var.web <- cast(jensen.df, gall ~ parasitoid, mean, value = "mean.no.var")
-rownames(no.var.web) <- no.var.web$gall
-no.var.web <- as.matrix.data.frame(no.var.web[ ,-1])
-
-library(bipartite)
-plotweb(var.web, method = "normal")
-plotweb(no.var.web, method = "normal")
-```
-
-```{r remove weak interactions}
-colSums(full.predictors[ ,interaxns_noPont]) # remove rG_Lestodip, rG_Mesopol, vLG_Mymarid, and SG_Platy.
-SG_Platy had 9 interactions, but all of these interactions came from a single, multichambered gall.
-
-net.trait.noweak <- mvabund(full.predictors[ ,c("aSG_Tory", "rG_Eulo", "rG_Tory", "vLG_Eulo", "vLG_Mesopol","vLG_Platy", "vLG_Tory")])
-
-net.mvabund.2.noweak <- manyglm(net.trait.noweak ~ log.vLG_abund + log.vLG.height.mean + 
-                         log.1.rG_abund + log.1.aSG_abund,
-                       data = full.predictors,
-                       family = "negative.binomial")
+1 - exp(-0.2157298) # 19% decrease in vLG_Platy interaction with every one unit increase in gall size.
+1 - exp(-0.2697011) # 24% decrease in vLG_Mesopolobous interaction with every one unit increase in gall size.
 ```
@@
 ```{r}
-## Identify the effect of phenotypic variation (gall density and gall size) on food web structure
-
-## create data with the same mean, but NO VARIANCE in predictor variables. For some reason, the predict.manyglm needs a dataset that is the same dimension as the original one.
-row.count <- dim(full.predictors)[1]
-no.var.data <- with(full.predictors, 
-                  data.frame(log.vLG_abund = rep(log(mean(vLG_abund)), row.count),
-                             log.1.rG_abund = rep(log(mean(rG_abund)+1), row.count),
-                             log.1.aSG_abund = rep(log(mean(aSG_abund)+1), row.count),
-                             log.vLG.height.mean = rep(log(mean(vLG.height.mean)), row.count)))
-# increase mean by one standard deviation. With this type of code, I should be able to make predictions for how evolution will affect the food web.
-no.var.data.vLG.size.Increase <- with(full.predictors, 
-                  data.frame(log.vLG_abund = rep(log(mean(vLG_abund)), row.count),
-                             log.1.rG_abund = rep(log(mean(rG_abund)+1), row.count),
-                             log.1.aSG_abund = rep(log(mean(aSG_abund)+1), row.count),
-                             log.vLG.height.mean = rep(log(mean(vLG.height.mean)+sd(vLG.height.mean)), row.count)))
-
-jensen.net.mvabund.2 <- jensen_magnitude_manyglm(net.mvabund.2, original.data = full.predictors,
-                                                 comparison.data = no.var.data)
-jensen.net.mvabund.2.noweak <- jensen_magnitude_manyglm(net.mvabund.2.noweak, original.data = full.predictors, comparison.data = no.var.data)
-jensen.net.mvabund.2.meanIncrease <- jensen_magnitude_manyglm(net.mvabund.2, original.data = no.var.data, comparison.data = no.var.data.vLG.size.Increase)
-
-# used a wilcoxon signed rank test because the differences were not normally distributed
-with(jensen.net.mvabund.2, hist(var.predict - no.var.predict, breaks = seq(0,2,0.01)))
-with(jensen.net.mvabund.2, wilcox.test(var.predict, no.var.predict, 
-                                       paired = TRUE, conf.int = TRUE))
-magnitude.lm <- lm(log(Percent.Difference) ~ log(no.var.predict), jensen.net.mvabund.2)
-summary(magnitude.lm); visreg(magnitude.lm, scale = "response", xtrans = log)
-
-# I get the same qualitative results even after removing the weakest interactions from the dataset
-with(jensen.net.mvabund.2.noweak, wilcox.test(var.predict, no.var.predict, 
-                                              paired = TRUE, conf.int = TRUE)) 
-magnitude.lm.noweak <- lm(log(Percent.Difference) ~ log(no.var.predict), 
-                          jensen.net.mvabund.2.noweak)
-summary(magnitude.lm.noweak); visreg(magnitude.lm.noweak, scale = "response", xtrans = log)
+net.trait.lms <- manylm(net.trait ~ log.vLG_abund + vLG.height.mean + 
+                         log.1.rG_abund + log.1.aSG_abund,
+                       data = full.predictors)
+summary(net.trait.lms) # non significant effect of gall height. It was the least important variable though in the GLM as well with mostly marginally significant effects.
 ```
 
-```{r effects of variation on food web structure}
-library(bipartite)
-jensen.to.use <- jensen.net.mvabund.2.noweak
-interactions.split <- colsplit(jensen.to.use$response.variables, 
-                               "_", names = c("gall","parasitoid"))
-jensen.to.use <- cbind(jensen.to.use, interactions.split)
+The results from the link composition analysis suggest that increases in gall abundance lead to increasing food web complexity for those individual nodes, whereas differences in leaf gall size lead to fundamental differences in link composition.
 
-var.web <- cast(jensen.to.use, gall ~ parasitoid, mean, value = "var.predict")
-rownames(var.web) <- var.web$gall
-var.web <- as.matrix.data.frame(var.web[ ,-1])
+Then we examined whether the proportion of galls parasitized varied among willow genotypes
+```{r gall parasitism rates}
+vLG.ptized.glm <- glm(vLG_parasitized/vLG_abund ~ Genotype, data = full.df, 
+                      weights = vLG_abund, family = "binomial")
+vLG.Platy.glm <- glm(vLG_Platy/vLG_abund ~ Genotype, data = full.df, 
+                      weights = vLG_abund, family = "binomial")
+vLG.Mesopol.glm <- glm(vLG_Mesopol/vLG_abund ~ Genotype, data = full.df, 
+                      weights = vLG_abund, family = "binomial")
+vLG.Tory.glm <- glm(vLG_Tory/vLG_abund ~ Genotype, data = full.df, 
+                      weights = vLG_abund, family = "binomial")
+vLG.Eulo.glm <- glm(vLG_Eulo/vLG_abund ~ Genotype, data = full.df, 
+                      weights = vLG_abund, family = "binomial")
+vLG.Mymar.glm <- glm(vLG_Mymarid/vLG_abund ~ Genotype, data = full.df, 
+                      weights = vLG_abund, family = "binomial")
 
-no.var.web <- cast(jensen.to.use, gall ~ parasitoid, mean, value = "no.var.predict")
-rownames(no.var.web) <- no.var.web$gall
-no.var.web <- as.matrix.data.frame(no.var.web[ ,-1])
+rG.ptized.glm <- glm(rG_parasitized/rG_abund ~ Genotype, data = full.df, 
+                     weights = rG_abund, family = "binomial")
 
-par(mfrow = c(2,1))
-plotweb(var.web, method = "normal")
-plotweb(no.var.web, method = "normal")
-
-cbind(networklevel(var.web, 
-                   index = c("linkage density", "interaction evenness")),
-      networklevel(no.var.web,
-                   index = c("linkage density", "interaction evenness")))
+aSG.ptized.glm <- glm(aSG_parasitized/aSG_abund ~ Genotype, data = full.df,
+                      weights = aSG_abund, family = "binomial")
 ```
 
-```{r}
-vLG.ptized.glm <- glm(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~ vLG.height.mean, full.df, family = binomial)
-summary(vLG.ptized.glm)
-visreg(vLG.ptized.glm, scale = "response")
-vLG.ptized.no.var.data <- with(full.predictors, data.frame(vLG.height.mean = mean(vLG.height.mean)))
-jensen_magnitude(vLG.ptized.glm, no.var.data = vLG.ptized.no.var.data)
-#plot(vLG.ptized.glm)
-
-vLG.Platy.glm <- glm(cbind(vLG_Platy, vLG_abund - vLG_Platy) ~ vLG.height.mean, full.df, family = binomial)
-summary(vLG.Platy.glm)
-visreg(vLG.Platy.glm, scale = "response")
-vLG.Platy.no.var.data <- with(full.predictors, data.frame(vLG.height.mean = mean(vLG.height.mean)))
-jensen_magnitude(vLG.Platy.glm, no.var.data = vLG.Platy.no.var.data)
-
-vLG.Tory.glm <- glm(cbind(vLG_Tory, vLG_abund - vLG_Tory) ~ vLG.height.mean, full.df, family = binomial)
-summary(vLG.Tory.glm)
-visreg(vLG.Tory.glm, scale = "response")
-vLG.Tory.no.var.data <- with(full.predictors, data.frame(vLG.height.mean = mean(vLG.height.mean)))
-jensen_magnitude(vLG.Tory.glm, no.var.data = vLG.Tory.no.var.data)
-
-#plot(vLG_Platy ~ Genotype, full.df)
-#plot(vLG_Tory ~ Genotype, full.df)
-
-vLG.ptized.glmer <- glmer(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~ log(vLG.height.mean) + (1|Genotype), full.df, family = binomial)
-summary(vLG.ptized.glmer)
-visreg(vLG.ptized.glmer, scale = "response")
-mean(predict(vLG.ptized.glmer, type = "response"))
-#mean(predict(vLG.ptized.glmer, newdata = vLG.ptized.no.var.data, type = "response"))
-
-vLG_Platy.glmer <- glmer(vLG_Platy ~ log(vLG_abund) + log(vLG.height.mean) + (1|Genotype), full.df, family = poisson)
-summary(vLG_Platy.glmer)
-overdisp_fun(vLG_Platy.glmer) # no overdispersion when I model genotype as a randome effect.
-visreg(vLG_Platy.glmer, scale = "response")
+Summary of results from parasitism models
+```{r gall parasitism likelihood ratio tests, echo=FALSE}
+anova(vLG.ptized.glm, test = "LR")
+anova(vLG.Platy.glm, test = "LR")
+anova(vLG.Mesopol.glm, test = "LR")
+anova(vLG.Tory.glm, test = "LR")
+anova(vLG.Eulo.glm, test = "LR")
+anova(vLG.Mymar.glm, test = "LR")
+anova(rG.ptized.glm, test = "LR")
+anova(aSG.ptized.glm, test = "LR")
 ```
 
+Further explore the factors determining gall parasitism rates among willow genotypes. These were the best models as determined by AIC.
 ```{r}
-## Examine whether the proportion of galls parasitized varies among willow genotypes
+vLG_total.ptism <- glm(vLG_parasitized/vLG_abund ~ vLG.height.mean, 
+                       data = full.df, weights = vLG_abund, family = "binomial")
+summary(vLG_total.ptism)
 
-## vLG parasitism
-vLG.ptized.df <- full.df %>%
-  filter(vLG_abund > 0)
-table(vLG.ptized.df$Genotype)
-vLG.ptized.3plus.genos <- c("*","B","D","E","F","H","I","K","L","O","S","V","X","Y","Z")
-colSums(vLG.ptized.df[ ,c("vLG_parasitized","vLG_abund","vLG_Tory","vLG_Eulo","vLG_Platy",
-                         "vLG_Mesopol","vLG_Mymarid")])
+vLG_Platy.ptism <- glm(vLG_Platy/vLG_abund ~ vLG.height.mean*vLG_abund, 
+                       data = full.df, weights = vLG_abund, family = "binomial")
+summary(vLG_Platy.ptism)
 
-# interesting, only marginally significant with quasibinomial
-vLG.ptized.bin <- glm(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~ Genotype,
-                  data = filter(vLG.ptized.df, Genotype %in% vLG.ptized.3plus.genos), 
-                  family = 'binomial')
-anova(glm.binomial.disp(vLG.ptized.bin), test = "LR") # just significant, when I account for overdispersion using this method...
+vLG_Mesopol.ptism <- glm(vLG_Mesopol/vLG_abund ~ vLG.height.mean*vLG_abund, 
+                       data = full.df, weights = vLG_abund, family = "binomial")
+summary(vLG_Mesopol.ptism)
 
-vLG.ptized.bin.glmer <- glmer(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~ (1|Genotype),
-                  data = filter(vLG.ptized.df, Genotype %in% vLG.ptized.3plus.genos), 
-                  family = 'binomial')
-overdisp_fun(vLG.ptized.bin.glmer)
-
-test.2 <- rpt.binomGLMM.add(y = with(vLG.ptized.df, cbind(vLG_parasitized, vLG_abund - vLG_parasitized)), groups = vLG.ptized.df$Genotype) 
-test.2
-library(emdbook)
-library(MASS)
-vLG.ptized.df <- mutate(vLG.ptized.df, 
-                        plant.position = factor(plant.position))
-library(lme4)
-sc.abund <- scale(vLG.ptized.df$vLG_abund)
-sc.height <- scale(vLG.ptized.df$vLG.height.mean)
-test.4 <- glmer(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~  vLG.height.mean + (1|Genotype),
-                  data = vLG.ptized.df,
-                  family = 'binomial')
-summary(test.4)
-AIC(test.4)
-confint(test.4, method = "profile") # variance > 0!
-overdisp_fun(test.4) # no clear evidence of overdispersion
-ggQQ_ranef(ranef(test.4)$Genotype$'(Intercept)') # and the residuals look great!
-visreg(test.4, scale = "response")
-R.link  <- .6411 / (.6411 +pi^2 /3) # heritability is reasonable as well
-R.link.low  <- .1153673 / (.1153673 +pi^2 /3) 
-R.link.high  <- 1.44 / (1.44 +pi^2 /3) 
-profile(test.4, which = 1)
-summary(test.4)
-test.3 <- rpt.binomGLMM.multi(y = with(vLG.ptized.df, cbind(vLG_parasitized, vLG_abund - vLG_parasitized)), groups = vLG.ptized.df$Genotype)
-
-library(RLRsim)
-
-#plot(vLG.ptized.bin)
-summary(vLG.ptized.bin)
-dfun(vLG.ptized.bin)
-anova(vLG.ptized.bin, test = "LR") # still significant after retaining Genotypes with 3 plus replicates.
-
-vLG_ptized.bin.pred <- gam(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~ s(vLG.height.mean),
-                  data = full.df, 
-                  family = 'binomial')
-summary(vLG_ptized.bin.pred)
-AIC(vLG_ptized.bin.pred)
-visreg(vLG_ptized.bin.pred, scale = "response")
-#jensen.single(vLG_ptized.bin.pred)
-
-# vLG_Platy test
-vLG_Platy.bin <- glm(cbind(vLG_Platy, vLG_abund - vLG_Platy) ~ Genotype,
-                  data = filter(full.df, Genotype %in% vLG.ptized.3plus.genos), 
-                  family = 'quasibinomial')
-plot(vLG_Platy.bin)
-summary(vLG_Platy.bin)
-anova(vLG_Platy.bin, test = "F")
-
-library(dispmod)
-vLG_Platy.bin.pred <- glm(cbind(vLG_Platy, vLG_abund - vLG_Platy) ~ vLG.height.mean,
-                  data = full.predictors, 
-                  family = 'binomial')
-dfun(vLG_Platy.bin.pred)
-summary(vLG_Platy.bin.pred)
-glm.binomial.disp(vLG_Platy.bin.pred)
-AIC(vLG_Platy.bin.pred)
-plot(vLG_Platy.bin.pred)
-visreg(vLG_Platy.bin.pred, scale = "response")
-jensen.single(vLG_Platy.bin.pred)
-glm.binomial.disp(vLG_Platy.bin.pred)
-qAIC(vLG_Platy.bin.pred, dispersion = dfun(vLG_Platy.bin.pred))
-
-dfun(vLG_Platy.bin.pred)
-
-vLG_Platy.bin.pred.glmer <- glmer(cbind(vLG_Platy, vLG_abund - vLG_Platy) ~ vLG.height.mean + (1|Genotype),
-                  data = full.df, 
-                  family = 'binomial')
-summary(vLG_Platy.bin.pred.glmer)
-overdisp_fun(vLG_Platy.bin.pred.glmer)
-AIC(vLG_Platy.bin.pred.glmer)
-plot(vLG_Platy.bin.pred.glmer)
-ggQQ_ranef(ranef(vLG_Platy.bin.pred.glmer)$Genotype$'(Intercept)') # not too bad.
-
-# AIC gam interaction 207.6398, AIC = 204 with main effect model in gam. 213 with main model
-
-## Identify the effect of phenotypic variation (gall density and gall size) on food web structure
-
-jensen.single(final.vLG_Platy)
-jensen.single(vLG_Mesopol.bin.pred)
-jensen.single(vLG_Tory.bin.pred)
-
-# vLG_Mesopol test
-vLG_Mesopol.bin <- glm(cbind(vLG_Mesopol, vLG_abund - vLG_Mesopol) ~ Genotype,
-                  data = filter(full.df, Genotype %in% vLG.ptized.3plus.genos), 
-                  family = 'binomial')
-plot(vLG_Mesopol.bin)
-summary(vLG_Mesopol.bin)
-dfun(vLG_Mesopol.bin)
-anova(vLG_Mesopol.bin, test = "LR")
-
-vLG_Mesopol.bin.pred <- gam(cbind(vLG_Mesopol, vLG_abund - vLG_Mesopol) ~ vLG.height.mean,
-                  data = full.df, 
-                  family = 'quasibinomial')
-summary(vLG_Mesopol.bin.pred)
-visreg(vLG_Mesopol.bin.pred, xvar = "vLG.height.mean", scale = "response")
-dfun(vLG_Mesopol.bin.pred)
-
-jensen.single(vLG_Mesopol.bin.pred)
-
-# vLG_Tory test
-vLG_Tory.bin <- glm(cbind(vLG_Tory, vLG_abund - vLG_Tory) ~ Genotype,
-                  data = filter(full.df, Genotype %in% vLG.ptized.3plus.genos), 
-                  family = 'binomial')
-plot(vLG_Tory.bin)
-summary(vLG_Tory.bin)
-anova(vLG_Tory.bin, test = "LR")
-
-vLG_Tory.bin.pred <- glm(cbind(vLG_Tory, vLG_abund - vLG_Tory) ~ vLG_abund + vLG.height.mean,
-                  data = full.df, 
-                  family = 'binomial')
-summary(vLG_Tory.bin.pred)
-AIC(vLG_Tory.bin.pred) # 136.1538 is the lowest score. Main effect model
-visreg(vLG_Tory.bin.pred, scale = "response")
-
-# vLG_Eulo test
-vLG_Eulo.bin <- glm(cbind(vLG_Eulo, vLG_abund - vLG_Eulo) ~ Genotype,
-                  data = filter(full.df, Genotype %in% vLG.ptized.3plus.genos), 
-                  family = 'binomial')
-#plot(vLG_Eulo.bin)
-summary(vLG_Eulo.bin)
-anova(vLG_Eulo.bin, test = "LR") # not significant
-
-# not testing vLG_Mymarid because it is in such low abundance
-predict.df <- data.frame(Genotype = vLG.ptized.3plus.genos)
-vLG.genotype.predictions <- cbind.data.frame(predict.df, 
-                                         vLG.Proportion.Ptized = predict(vLG.ptized.bin,
-                                                                               newdata = predict.df,
-                                                                               type = "response"))
-summary(vLG.genotype.predictions)
-
-ggplot(filter(vLG.ptized.df, Genotype %in% vLG.ptized.3plus.genos),
-       aes(x = Genotype, y = vLG_parasitized/vLG_abund)) +
-  geom_point(aes(size = vLG_abund), color = "grey", shape = 1,
-             position = position_jitter(height = 0, width = 0.1)) + 
-  scale_size(range = c(3,15)) + 
-  geom_point(data = vLG.genotype.predictions, 
-             aes(x = Genotype, y = vLG.Proportion.Ptized),
-             color = "steelblue", size = 30, shape = "--") +
-  theme_classic() +
-  ylab("Proportion of galls parasitized")
-range(filter(vLG.ptized.df, Genotype %in% vLG.ptized.3plus.genos)$vLG_abund)
-
-## rG parasitism
-rG.ptized.df <- full.df %>%
-  filter(rG_abund > 0)
-table(rG.ptized.df$Genotype)
-rG.ptized.3plus.genos <- c("*","A","D", "F", "G", "H","I","K","O","Q", "R", "S","X", "Y", "Z") 
-colSums(rG.ptized.df[ ,c("rG_parasitized","rG_abund","rG_Tory","rG_Eulo","rG_Platy",
-                         "rG_Mesopol","rG_Lestodip")]) # too few rG_eulo, rG_Platy, rG_Mesopol, and rG_Lestodip to test for differences in parasitism among genotypes.
-
-rG.ptized.bin <- glm(cbind(rG_parasitized, rG_abund - rG_parasitized) ~ Genotype,
-                  data = filter(full.df, Genotype %in% rG.ptized.3plus.genos), 
-                  family = 'binomial')
-#plot(rG.ptized.bin) # residuals not great
-summary(rG.ptized.bin)
-anova(rG.ptized.bin, test = "LR") # marginally significant after retaining Genotypes with 3 plus replicates.
-
-# rG_Tory test
-rG_Tory.bin <- glm(cbind(rG_Tory, rG_abund - rG_Tory) ~ Genotype,
-                  data = filter(full.df, Genotype %in% rG.ptized.3plus.genos), 
-                  family = 'binomial')
-plot(rG_Tory.bin)
-summary(rG_Tory.bin)
-anova(rG_Tory.bin, test = "LR") # not significant
-
-rG_Eulo.bin <- glm(cbind(rG_Eulo, rG_abund - rG_Eulo) ~ Genotype,
-                  data = filter(full.df, Genotype %in% rG.ptized.3plus.genos), 
-                  family = 'binomial')
-plot(rG_Eulo.bin)
-summary(rG_Eulo.bin)
-anova(rG_Eulo.bin, test = "LR") # not significant
-
-predict.df <- data.frame(Genotype = rG.ptized.3plus.genos)
-rG.genotype.predictions <- cbind.data.frame(predict.df, 
-                                         rG.Proportion.Ptized = predict(rG.ptized.bin,
-                                                                               newdata = predict.df,
-                                                                               type = "response"))
-summary(rG.genotype.predictions)
-
-ggplot(filter(rG.ptized.df, Genotype %in% rG.ptized.3plus.genos),
-       aes(x = Genotype, y = rG_parasitized/rG_abund)) +
-  geom_point(aes(size = rG_abund), color = "grey", shape = 1,
-             position = position_jitter(height = 0, width = 0.1)) + 
-  scale_size(range = c(3,15)) + 
-  geom_point(data = rG.genotype.predictions, 
-             aes(x = Genotype, y = rG.Proportion.Ptized),
-             color = "steelblue", size = 30, shape = "--") +
-  theme_classic() +
-  ylab("Proportion of galls parasitized")
-range(filter(rG.ptized.df, Genotype %in% rG.ptized.3plus.genos)$rG_abund)
-
-## aSG parasitism
-aSG.ptized.df <- full.df %>%
-  filter(aSG_abund > 0)
-table(aSG.ptized.df$Genotype)
-aSG.ptized.3plus.genos <- c("G","I","K","Q")
-
-aSG.ptized.bin <- glm(cbind(aSG_parasitized, aSG_abund - aSG_parasitized) ~ Genotype,
-                  data = filter(full.df, Genotype %in% aSG.ptized.3plus.genos), 
-                  family = 'binomial')
-plot(aSG.ptized.bin) # residuals not great
-summary(aSG.ptized.bin)
-anova(aSG.ptized.bin, test = "LR") # marginally significant after retaining Genotypes with 3 plus replicates.
-predict.df <- data.frame(Genotype = aSG.ptized.3plus.genos)
-aSG.genotype.predictions <- cbind.data.frame(predict.df, 
-                                         aSG.Proportion.Ptized = predict(aSG.ptized.bin,
-                                                                               newdata = predict.df,
-                                                                               type = "response"))
-summary(aSG.genotype.predictions)
-
-ggplot(filter(aSG.ptized.df, Genotype %in% aSG.ptized.3plus.genos),
-       aes(x = Genotype, y = aSG_parasitized/aSG_abund)) +
-  geom_point(aes(size = aSG_abund), color = "grey", shape = 1,
-             position = position_jitter(height = 0, width = 0.1)) + 
-  scale_size(range = c(3,15)) + 
-  geom_point(data = aSG.genotype.predictions, 
-             aes(x = Genotype, y = aSG.Proportion.Ptized),
-             color = "steelblue", size = 30, shape = "--") +
-  theme_classic() +
-  ylab("Proportion of galls parasitized")
-range(filter(aSG.ptized.df, Genotype %in% aSG.ptized.3plus.genos)$aSG_abund)
-
-## SG parasitism
-SG.ptized.df <- full.df %>%
-  filter(SG_abund > 0)
-table(SG.ptized.df$Genotype) # no Genotypes with 3+ replicates, so I'm unable to evaluate whether SG parasitism defers among willow genotypes.
-
-##
-uni.check.bin <- glm(vLG_Platy/vLG_abund ~ vLG_abund*vLG.height.mean, full.predictors,
-                     family = binomial,
-                     weights = vLG_abund)
-AIC(uni.check.bin)
-summary(uni.check.bin)
-visreg(uni.check.bin, xvar = "vLG_abund", by = "vLG.height.mean", scale = "response")
-
-# NEED TO THINK ABOUT HOW I SCALE THIS UP TO DIFFERENT NUMBERS OF SHOOTS SAMPLED.
-
-#### messing around with different ideas.
-## Create mvabund model based on probability of observing an interaction
-prob.df <- full.predictors %>%
-  mutate(succ.vLG_Platy.succ = vLG_Platy,
-         fail.vLG_Platy.fail = vLG_abund - vLG_Platy,
-         succ.vLG_Tory.succ = vLG_Tory,
-         fail.vLG_Tory.fail = vLG_abund - vLG_Tory,
-         succ.vLG_Mesopol.succ = vLG_Mesopol,
-         fail.vLG_Mesopol.fail = vLG_abund - vLG_Mesopol,
-         succ.vLG_Eulo.succ = vLG_Eulo,
-         fail.vLG_Eulo.fail = vLG_abund - vLG_Eulo,
-         succ.rG_Platy.succ = rG_Platy,
-         fail.rG_Platy.fail = rG_abund - rG_Platy,
-         succ.rG_Tory.succ = rG_Tory,
-         fail.rG_Tory.fail = rG_abund - rG_Tory,
-         succ.rG_Mesopol.succ = rG_Mesopol,
-         fail.rG_Mesopol.fail = rG_abund - rG_Mesopol,
-         succ.rG_Eulo.succ = rG_Eulo,
-         fail.rG_Eulo.fail = rG_abund - rG_Eulo,
-         succ.SG_Platy.succ = SG_Platy,
-         fail.SG_Platy.fail = SG_abund - SG_Platy,
-         succ.aSG_Tory.succ = aSG_Tory,
-         fail.aSG_Tory.fail = aSG_abund - aSG_Tory) %>%
-  select(succ.vLG_Platy.succ:fail.aSG_Tory.fail) %>%
-  mvabund()
-
-vLG.interaction.df <- mvabund(select(full.predictors, vLG_Eulo, vLG_Platy, vLG_Mymarid, vLG_Tory, vLG_Mesopol))
-vLG.interaction.df[vLG.interaction.df > 0] <- 1
-
-test <- manyglm(vLG.interaction.df ~ vLG_density*vLG.height.mean, data = full.predictors, family = "binomial")
-plot(test)
-anova(test, p.uni = "adjusted")
-length(prob.df[prob.df > 1] > 0)
-
-test <- MASS::glm.nb(vLG_Platy ~ offset(log(vLG_abund)) + vLG.height.mean, data = full.predictors) #
-summary(test)
-coef(test)
-AIC(test)
-#exp(sum(-4.5 + .34*5 -0.25*8))
-exp(sum(coef(test)[c(1:3)]*c(1,0.02,8))) # 8 mm gall at a density of 11 per 100 shoots. Would be 6.11 parasitoids per 100 shoots.
-predict(test, type = "response")*10
-visreg(test, scale = "response")
+vLG_Tory.ptism <- glm(vLG_Tory/vLG_abund ~ vLG.height.mean + vLG_abund, 
+                       data = full.df, weights = vLG_abund, family = "binomial")
+summary(vLG_Tory.ptism)
 ```
 
-```{r, functions, echo=FALSE}
-kable.adonis <- function(adonis.object){
-  require(knitr)
-  kable(adonis.object$aov.tab[ ,-5], # remove R2 calculation
-        col.names = c("df","SS","MS","F","P"), 
-        digits = c(0,2,2,2,3))
-}
-
-heritability.adonis <- function(adonis.object, data){
-  gen.row <- which(rownames(adonis.object$aov.tab) == "Genotype")
-  res.row <- which(rownames(adonis.object$aov.tab) == "Residuals")
-  gen.MS <- adonis.object$aov.tab$MeanSqs[gen.row]
-  res.MS <- adonis.object$aov.tab$MeanSqs[res.row]
-  mean.sample.size <- mean(table(data$Genotype))
-  gen.variance <- (gen.MS - res.MS)/mean.sample.size
-  heritability <- gen.variance/(gen.variance + res.MS) # note that residual MS is equal to the residual variance
-  print(heritability)
-}
-
-heritability <- function(merMod.object){
-  require(lme4)
-  require(RLRsim)
-
-  var.comps <- as.data.frame(VarCorr(merMod.object))
-  genetic.var <- var.comps$vcov[1]
-  residual.var <- var.comps$vcov[2]
-  heritability <- genetic.var/(genetic.var + residual.var)
-  restricted.likelihood.ratio.test <- exactRLRT(merMod.object)
-  data.frame(#Model = formula(merMod.object),
-             Heritability = round(heritability, 2),
-             RLRT_statistic = round(restricted.likelihood.ratio.test$statistic, 2),
-             P = round(restricted.likelihood.ratio.test$p.value, 3))
-}
-
-jensen.single <- function(model.object, predict.type = "response"){
-  model <- model.object
-
-  var.predict <- predict(model.object, type = predict.type)
-
-  predictor.names <- names(model.object$model)[-1] # omits response variable
-  data <- data.frame(model.object$model[ ,predictor.names])
-
-  no.var.data <- list()
-  for(i in 1:length(data)){
-    no.var.data[[i]] <- mean(data[ ,i])
-  }
-  no.var.data <- as.data.frame(no.var.data)
-  names(no.var.data) <- predictor.names
-
-  no.var.predict <- predict(model.object, newdata = no.var.data, type = predict.type)
-
-  data.frame(mean.with.var = mean(var.predict, na.rm = TRUE),
-             mean.no.var = mean(no.var.predict, na.rm = TRUE))
-}
-
-jensen.single <- function(model.object, no.var.data, predict.type = "response"){
-  model <- model.object
-
-  var.predict <- predict(model.object, type = predict.type)
-
-  predictor.names <- names(model.object$model)[-1] # omits response variable
-  data <- data.frame(model.object$model[ ,predictor.names])
-
-  #no.var.data <- list()
-  #for(i in 1:length(data)){
-   # no.var.data[[i]] <- mean(data[ ,i])
-  #}
-  #no.var.data <- as.data.frame(no.var.data)
-  #names(no.var.data) <- predictor.names
-
-  no.var.predict <- predict(model.object, newdata = no.var.data, type = predict.type)
-
-  data.frame(mean.with.var = mean(var.predict, na.rm = TRUE),
-             mean.no.var = mean(no.var.predict, na.rm = TRUE))
-}
-
+```{r summary of models, echo=FALSE}
+car::Anova(vLG_total.ptism)
+car::Anova(vLG_Platy.ptism)
+car::Anova(vLG_Mesopol.ptism)
+car::Anova(vLG_Tory.ptism)
 ```
 
-```{r}
-library(psych)
-gall.df.cors <- full.df %>%
-  mutate(vLG_density = log(vLG_abund+1) - log(shootEst.no18),
-         vLG_density2 = log(vLG_abund/shootEst.no18+0.01),
-         vLG_density3 = sqrt(vLG_abund/shootEst.no18),
-         rG_density = log(vLG_abund+1) - log(shootEst.no18),
-         rG_density2 = log(rG_abund/shootEst.no18+0.01),
-         rG_density3 = sqrt(rG_abund/shootEst.no18)) %>%
-  select(vLG.height.mean, vLG_density:rG_density3)
-trait.df.cors <- full.df %>%
-  select(Total_Area:flavanonOLES.PC1)
-
-gall.trait.cors <- corr.test(x = trait.df.cors, y = gall.df.cors)
-round(gall.trait.cors$r,2)
-# further examine all correlations greater than 0.20
-# vLG.height.mean: salicortin__A270nm, luteolin.glucoside.5.glu.__A320nm, quercetyin.der1, luteolin.der1, luteolin der2, termulacin, Carbon, sal_tannin.PC1, flavonoOLES.PC1
-# vLG_density: Total_Area, Height, D_mean_smoothed, ampelopsin.der__A320nm, myricitrin__A320nm, eriodictyol.7.glucoside__A270nm, luteolin.der1__A320nm, C_N_imputed (sal_tannin.PC1?
-a bit low).
-Note though the Nitrogen and C_N_ratio are not good predictors for the "un_imputed" data.
-# rG_density: total_area, height, D_mean_smoothed, ampelopsin.der__A320nm, eriodictyol.7.glucoside__A270nm, luteolin.der1__A320nm, C_N_imputed
-
-library(car)
-# need to work on code below.
-scatterplotMatrix(cbind.data.frame(gall.df.cors$vLG.height.mean, 
-                                   trait.df.cors[ ,"salicortin__A270nm", 
-                                                 "luteolin.glucoside.5.glu.__A320nm",
-                                                 "quercetin.der1__A320nm", 
-                                                 "luteolin.der1__A320nm", 
-                                                 "luteolin der2__A320nm", 
-                                                 "termulacin__A320nm",
-                                                 "Carbon", "sal_tannin.PC1", 
-                                                 "flavonoOLES.PC1"]))
+The odds of a leaf gall being parasitized decreased by 25% with every one mm increase in gall size.
+```{r, eval=FALSE}
+1-exp(coef(vLG_total.ptism)[2]) # 25% reduction in the odds of vLG being parasitized with every one unit increase in gall size.
 ```
-
-```{r}
-interaction_abund <- rowSums(full.df[ ,interaxns_noPont])
-interaction_richness <- rowSums(full.df[ ,interaxns_noPont] > 0)
-ggplot(full.df, aes(x = Genotype, y = interaction_abund)) +
-  geom_boxplot()
-
-ggplot(full.df, aes(x = Genotype, y = interaction_abund/shootEst.no18)) +
-  geom_boxplot() # G, H, M, N, O, P, R, U all have 75% of their observations at zero for interaction density
-
-ggplot(full.df, aes(x = Genotype, y = interaction_richness)) +
-  geom_boxplot() # G, H, M, N, O, P, R, U all have 75% of their observations at zero for interaction richness
-
-non.zero.genos <- c("*","A","B","D","E","F","I","J","K","L","Q","S","T","V","W","X","Y","Z")
-
-df.non.zero.genos <- filter(full.df, Genotype %in% non.zero.genos)
-hist(log(rowSums(df.non.zero.genos[ ,interaxns_noPont]) +1))
-interaction_abund.non.zero.genos <- rowSums(df.non.zero.genos[ ,interaxns_noPont])
-library(lme4)
-interaction.abund.lmer <- lmer(log(interaction_abund+1) ~ offset(log(shootEst.no18)) + (1|Genotype), full.df)
-heritability(interaction.abund.lmer) # ~0.28 with non.zero.genos.
-0.33 with full dataset, but the residuals look pretty good actually.
-Actually, setting an offset and log(x+1) transforming the response variable seems to do a pretty good job of making the random effect normally distributed.
-ggQQ_ranef(ranef(interaction.abund.lmer)$Genotype$'(Intercept)')
-hist(ranef(interaction.abund.lmer)$Genotype$'(Intercept)')
-```
-
-Table 1: Quantitative full data set with euclidean distance
-```{r, results='asis', echo=FALSE}
-full.adonis <- adonis(log(full.df.quant+1) ~ log(shootEst.no18) + Genotype, 
-                      data = full.df, method = "euclidean")
-heritability.adonis(full.adonis, full.df)
-kable.adonis(full.adonis)
-```
-
-Table S2: Qualitative full data set with euclidean distance
-```{r, results='asis', echo=FALSE}
-full.adonis.qual <- adonis(full.df.qual ~ log(shootEst.no18) + Genotype, 
-                      data = full.df, method = "euclidean")
-heritability.adonis(full.adonis.qual, full.df)
-kable.adonis(full.adonis.qual)
-```
-
-Table S3: Qualitative full data set with Raup-crick null model
-```{r, results='asis', echo=FALSE}
-rc.dist.full <- raupcrick(full.df.qual) # calculate raupcrick dissimilarity for teasing apart differences in alpha vs. beta-diversity
-rc.dist.full.adonis <- adonis(rc.dist.full ~ log(shootEst.no18) + Genotype, 
-                      data = full.df)
-heritability.adonis(rc.dist.full.adonis, full.df)
-kable.adonis(rc.dist.full.adonis)
-```
-
-Table S3: Qualitative sub data set with Raup-crick null model
-```{r, results='asis', echo=FALSE}
-rc.dist.sub <- raupcrick(sub.df.qual) # calculate raupcrick dissimilarity for teasing apart differences in alpha vs. beta-diversity
-rc.dist.sub.adonis <- adonis(rc.dist.sub ~ log(shootEst.no18) + Genotype, 
-                      data = sub.df)
-heritability.adonis(rc.dist.sub.adonis, sub.df)
-kable.adonis(rc.dist.sub.adonis)
-```
-
-Table S2: Quantitative data subset, euclidean
-```{r, results='asis', echo=FALSE}
-sub.adonis.euclid <- adonis(log(sub.df.quant+1) ~ log(shootEst.no18) + Genotype, 
-                            data = sub.df, method = "euclidean")
-kable.adonis(sub.adonis.euclid)
-```
-
-Table S3: Quantitative data subset, with bray-curtis dissimilarity
-```{r, results='asis', echo=FALSE}
-sub.adonis.bray <- adonis(sub.df.quant ~ log(shootEst.no18) + Genotype, 
-                          data = sub.df, method = "bray")
-kable.adonis(sub.adonis.bray)
-```
-
-Table S3: Qualitative data subset, with Jaccard dissimilarity
-```{r, results='asis', echo=FALSE}
-sub.adonis.jac <- adonis(sub.df.qual ~ log(shootEst.no18) + Genotype, 
-                          data = sub.df, method = "jaccard")
-heritability.adonis(adonis.object = sub.adonis.jac, data = sub.df)
-kable.adonis(adonis.object = sub.adonis.jac)
-```
-
-Table S4: Qualitative data subset, with euclidean dissimilarity
-```{r, results='asis', echo=FALSE}
-sub.adonis.euc <- adonis(sub.df.qual ~ log(shootEst.no18) + Genotype, 
-                          data = sub.df, method = "euclidean")
-kable.adonis(adonis.object = sub.adonis.euc)
-```
-
-Heritability of focal gall-parasitoid guild interactions
-```{r gall-ptoid heritability, echo=FALSE}
-## Variation among genotypes
-gall.ptoid.variation <- cbind(Genotype = full.df$Genotype, full.df[ ,c("vLG_egg","vLG_ecto","rG_ecto")]/full.df$shootEst.no18) %>%
-  group_by(Genotype) %>%
-  select(vLG_egg, vLG_ecto, rG_ecto) %>%
-  summarise_each(funs(mean))
-
-max(filter(gall.ptoid.variation, vLG_egg > 0)$vLG_egg)/min(filter(gall.ptoid.variation, vLG_egg > 0)$vLG_egg) # 35-fold variation in vLG_egg parasitism among genotypes that hosted this interaction
-
-max(filter(gall.ptoid.variation, vLG_ecto > 0)$vLG_ecto)/min(filter(gall.ptoid.variation, vLG_ecto > 0)$vLG_ecto) # 20-fold variation in vLG_ecto parasitism among genotypes that hosted this interaction
-
-max(filter(gall.ptoid.variation, rG_ecto > 0)$rG_ecto)/min(filter(gall.ptoid.variation, rG_ecto > 0)$rG_ecto) # 9-fold variation in vLG_ecto parasitism among genotypes that hosted this interaction
-
-## Iteomyia (vLG) - Egg parasitoid 
-vLG_egg.lmer.full <- lmer(log(vLG_egg + 1) ~ log(shootEst.no18) + (1|Genotype), 
-                          data = full.df)
-heritability(vLG_egg.lmer.full)
-ggQQ_ranef(ranef(vLG_egg.lmer.full)$Genotype$"(Intercept)") # non-normal random effect
-
-vLG_egg.lmer.sub <- lmer(log(vLG_egg + 1) ~ log(shootEst.no18) + (1|Genotype), 
-                         data = sub.df)
-heritability(vLG_egg.lmer.sub)
-ggQQ_ranef(ranef(vLG_egg.lmer.sub)$Genotype$"(Intercept)") # normal random effect
-
-## Iteomyia (vLG) - Larval parasitoid (ecto)
-vLG_ecto.lmer.full <- lmer(log(vLG_ecto + 1) ~ log(shootEst.no18) + (1|Genotype), 
-                          data = full.df)
-heritability(vLG_ecto.lmer.full)
-ggQQ_ranef(ranef(vLG_ecto.lmer.full)$Genotype$"(Intercept)") # close to normal random effect
-
-vLG_ecto.lmer.sub <- lmer(log(vLG_ecto + 1) ~ log(shootEst.no18) + (1|Genotype), 
-                         data = sub.df)
-heritability(vLG_ecto.lmer.sub) # marginally significant
-ggQQ_ranef(ranef(vLG_ecto.lmer.sub)$Genotype$"(Intercept)") # normal random effect
-
-## Rabdophaga salicisbrassicoides (rG) - Larval parasitoid (ecto)
-rG_ecto.lmer.full <- lmer(log(rG_ecto + 1) ~ log(shootEst.no18) + (1|Genotype), 
-                          data = full.df)
-heritability(rG_ecto.lmer.full)
-ggQQ_ranef(ranef(rG_ecto.lmer.full)$Genotype$"(Intercept)") # non-normal random effect
-
-rG_ecto.lmer.sub <- lmer(log(rG_ecto + 1) ~ log(shootEst.no18) + (1|Genotype), 
-                         data = sub.df)
-heritability(rG_ecto.lmer.sub)
-ggQQ_ranef(ranef(rG_ecto.lmer.sub)$Genotype$"(Intercept)") # close to normal random effect
-```
-
-Gall density and gall size variation among willow genotypes
-```{r}
-gall.density.variation <- cbind(Genotype = full.df$Genotype, full.df[ ,c("vLG_abund","rG_abund")]/full.df$shootEst.no18) %>%
-  group_by(Genotype) %>%
-  select(vLG_abund, rG_abund) %>%
-  summarise_each(funs(mean))
-
-max(filter(gall.density.variation, vLG_abund > 0)$vLG_abund)/min(filter(gall.density.variation, vLG_abund > 0)$vLG_abund) # 67-fold variation in Iteomyia among willow genotypes
-
-max(filter(gall.density.variation, rG_abund > 0)$rG_abund)/min(filter(gall.density.variation, rG_abund > 0)$rG_abund) # 62-fold variation in Rabdophaga-bud among willow genotypes.
-
-vLG.lmer.full <- lmer(log(vLG_abund + 1) ~ log(shootEst.no18) + (1|Genotype),
-                      data = full.df)
-heritability(merMod.object = vLG.lmer.full)
-
-rG.lmer.full <- lmer(log(rG_abund + 1) ~ log(shootEst.no18) + (1|Genotype),
-                      data = full.df)
-heritability(merMod.object = rG.lmer.full)
-
-## vLG size. To test this, I took advantage of the full gall data set and used a nested random effect model.
-vLG.size.df <- filter(gall.size.df, gall.sp == "vLG")
-table(vLG.size.df$Genotype) # large heterogeneity in sample sizes
-table(vLG.size.df$plant.position) # large heterogeneity in sample sizes
-
-vLG.size.lmer <- lmer(gall.height ~ 1 + (1 | Genotype) + (1 |plant.position:Genotype),
-                      data = vLG.size.df)
-vLG.size.lmer.gen <- update(vLG.size.lmer, .~. - (1 | plant.position:Genotype)) # only Genotype as random effect
-vLG.size.lmer.pp <- update(vLG.size.lmer, .~. - (1 | Genotype)) # only plant.position as random effect
-
-vLG.size.lmer.sum <- summary(vLG.size.lmer)
-summary(vLG.size.lmer.gen)
-summary(vLG.size.lmer.pp)
-
-gen.var <- vLG.size.lmer.sum$varcor$Genotype[1] # Variance due to Genotype
-pp.var <- vLG.size.lmer.sum$varcor$'plant.position:Genotype'[1] # variance due to plant position
-res.var <- 4.1133 # residual variance - don't know how to extract this from summary object.
-
-H2.vLG.size <- gen.var/(gen.var + pp.var + res.var) # 0.15
-exactRLRT(m = vLG.size.lmer.gen, mA = vLG.size.lmer, m0 = vLG.size.lmer.pp) # this model tests whether
-
-## rG size. To test this, I took advantage of the full gall data set and used a nested random effect model.
-rG.size.df <- filter(gall.size.df, gall.sp == "rG")
-table(rG.size.df$Genotype) # large heterogeneity in sample sizes
-table(rG.size.df$plant.position) # large heterogeneity in sample sizes
-
-rG.size.lmer <- lmer(gall.height ~ 1 + (1 | Genotype) + (1 |plant.position:Genotype),
-                      data = rG.size.df)
-rG.size.lmer.gen <- update(rG.size.lmer, .~. - (1 | plant.position:Genotype)) # only Genotype as random effect
-rG.size.lmer.pp <- update(rG.size.lmer, .~. - (1 | Genotype)) # only plant.position as random effect
-
-rG.size.lmer.sum <- summary(rG.size.lmer)
-summary(rG.size.lmer.gen)
-summary(rG.size.lmer.pp)
-
-gen.var <- rG.size.lmer.sum$varcor$Genotype[1] # Variance due to Genotype
-pp.var <- rG.size.lmer.sum$varcor$'plant.position:Genotype'[1] # variance due to plant position
-res.var <- 1.80518 # residual variance - don't know how to extract this from summary object.
-
-H2.rG.size <- gen.var/(gen.var + pp.var + res.var) # 0.0418
-exactRLRT(m = rG.size.lmer.gen, mA = rG.size.lmer, m0 = rG.size.lmer.pp) # this model tests whether 
-```
-
-RDA analysis of community response to Iteomyia density and size as well as R. salicisbrassicoides density
-```{r RDA, echo=FALSE}
-full.rda.df <- full.df %>%
-  filter(vLG.height.mean > 0) %>% 
-  mutate(vLG_density = vLG_abund/shootEst.no18,
-         rG_density = rG_abund/shootEst.no18)
-full.rda.df.interaxn <- full.rda.df[ ,interaxns_noPont]
-full.rda.df.interaxn.qual <- ifelse(full.rda.df[ ,interaxns_noPont] > 0, 1, 0)
-
-full.rda.df.interaxn.log1 <- log(full.rda.df.interaxn+1)
-full.rda.df.interaxn.log1.offset <- full.rda.df.interaxn.log1 - log(full.rda.df$shootEst.no18) # this is the same as applying an "offset" to all of the data.
-In other words, to interpret it in terms of density
-
-rda.full <- rda(full.rda.df.interaxn.log1.offset ~  
-                  vLG_density*vLG.height.mean + rG_density, data = full.rda.df)
-RsquareAdj(rda.full)
-anova(rda.full)
-anova(rda.full, by = "margin")
-
-rda.traits.df <- na.omit(data.frame(full.rda.df.interaxn, 
-                                  shootEst.no18 = full.rda.df$shootEst.no18,
-                                  C_N_imputed = full.rda.df$C_N_imputed, 
-                                  Height = full.rda.df$Height, 
-                                  sal_tannin.PC1 = full.rda.df$sal_tannin.PC1, 
-                                  flavonOLES.PC1 = full.rda.df$flavonOLES.PC1))
-rda.full.traits <- rda(log(rda.traits.df[ ,interaxns_noPont]+1) ~ 
-                         Condition(log(shootEst.no18)) + 
-                         sal_tannin.PC1, data = rda.traits.df)
-RsquareAdj(rda.full.traits)
-anova(rda.full.traits)
-anova(rda.full.traits, by = "margin")
-vif.cca(rda.full.traits)
-plot(rda.full.traits)
-
-rda.full.qual <- rda(full.rda.df.interaxn.qual ~  
-                  vLG_density + vLG.height.mean + rG_density, data = full.rda.df)
-RsquareAdj(rda.full.qual)
-anova(rda.full.qual)
-anova(rda.full.qual, by = "margin")
-
-pos.interactions <- which(rowSums(full.rda.df.interaxn.qual) > 0)
-rda.full.qual.jac <- capscale(full.rda.df.interaxn.qual[pos.interactions, ] ~  
-                  vLG_density+vLG.height.mean + rG_density, 
-                  data = full.rda.df[pos.interactions, ], distance = "jaccard")
-RsquareAdj(rda.full.qual.jac)
-plot(rda.full.qual.jac)
-anova(rda.full.qual.jac)
-anova(rda.full.qual.jac, by = "margin")
-vif.cca(rda.full.qual.jac)
-
-## vLG-egg parasitoid response
-vLG_egg.lm <- lm(log(vLG_egg + 1) ~ offset(log(shootEst.no18)) + 
-                   vLG_density*vLG.height.mean,
-                 data = full.rda.df)
-summary(vLG_egg.lm)
-plot(vLG_egg.lm)
-library(visreg)
-visreg(vLG_egg.lm, xvar = "vLG_density", by = "vLG.height.mean")
-
-## vLG-ecto parasitoid response
-vLG_ecto.lm <- lm(log(vLG_ecto+1) ~ offset(log(shootEst.no18)) + vLG_density + vLG.height.mean,
-                 data = full.rda.df)
-summary(vLG_ecto.lm)
-plot(vLG_ecto.lm)
-
-## rG-ecto parasitoid response
-rG_ecto.lm <- lm(log(rG_ecto+1) ~ offset(log(shootEst.no18)) + rG_density,
-                 data = full.rda.df)
-summary(rG_ecto.lm)
-
-```
-
-```{r functional responses}
-library(ggplot2)
-full.rda.df.small.galls <- filter(full.rda.df, vLG.height.mean < 8)
-ggplot(full.rda.df.small.galls, aes(x = vLG_density, y = vLG_egg/vLG_abund)) +
-  geom_point(aes(size = vLG_abund)) +
-  stat_smooth(method = "glm", family = binomial, aes(weight = full.rda.df.small.galls$vLG_abund))
-
-vLG_egg.bin <- glm(cbind(vLG_egg, vLG_abund - vLG_egg) ~ vLG_density*vLG.height.mean, 
-                   data = full.rda.df, family = binomial)
-summary(vLG_egg.bin)
-anova(update(vLG_egg.bin, .~. - vLG_density*vLG.height.mean), vLG_egg.bin, test = "Chi")
-
-plot(vLG_egg.bin)
-visreg(vLG_egg.bin, xvar = "vLG_density", by = "vLG.height.mean", scale = "response")
-visreg2d(vLG_egg.bin, x = "vLG.height.mean", y = "vLG_density", scale = "response")
-
-vLG_ecto.bin <- glm(cbind(vLG_ecto, vLG_abund - vLG_ecto) ~ vLG_density + vLG.height.mean,
-                    data = full.rda.df, family = binomial)
-summary(vLG_ecto.bin)
-anova(update(vLG_ecto.bin, .~. - vLG_density - vLG.height.mean), vLG_ecto.bin, test = "Chi")
-plot(vLG_ecto.bin)
-visreg(vLG_ecto.bin, scale = "response")
-visreg2d(vLG_ecto.bin, x = "vLG.height.mean", y = "vLG_density", scale = "response")
-
-rG_ecto.bin <- glm(cbind(rG_ecto, rG_abund - rG_ecto) ~ I(rG_abund/shootEst.no18),
-                    data = full.df, family = binomial)
-summary(rG_ecto.bin)
-anova(update(rG_ecto.bin, .~. - I(rG_abund/shootEst.no18)), rG_ecto.bin, test = "Chi")
-plot(rG_ecto.bin)
-```
-
-```{r}
-# try mantel test
-data.for.mantel <- na.omit(select(full.df, plant.position, shootEst.no18,
-                                  aSG_Tory:rG_Platy, rG_Tory, SG_Platy, vLG_Eulo:vLG_Tory, # all interactions
-                                  Total_Area:HCH.tremulacin.__A220nm, water_content,
-                                  specific_leaf_area, C_N_imputed))
-traits.stand <- decostand(select(data.for.mantel, Total_Area:C_N_imputed), method = "standardize")
-traits.dist <- vegdist(traits.stand, method = "euclidean")
-network.stand <- (select(data.for.mantel, aSG_Tory:vLG_Tory)/data.for.mantel$shootEst.no18)
-network.dist <- vegdist(network.stand, method = "euclidean")
-
-network.for.dist <- mantel(xdis = traits.dist, ydis = network.dist)
-
-```
-
-Gall density mechanisms
-```{r}
-traits <- full.df %>%
-  select(plant.position,
-         Total_Area, Height, Density, # architecture traits
-         Trichome.No., specific_leaf_area, water_content, C_N_imputed, # other leaf quality traits
-         sal_tannin.PC1, cinn.PC1, cinn.PC2, # salicylates/tannins and phenolic acids
-         flavonOLES.PC1, flavonOLES.PC2, flavanonOLES.PC1) %>% # flavonoids
-  mutate(log_size = log(Total_Area),
-         log_trichomes = log(Trichome.No.+1))
-traits <- na.omit(traits)
-height_resid <- residuals(lm(Height ~ log_size, traits))
-density_resid <- residuals(lm(Density ~ log_size, traits))
-sla_resid <- residuals(lm(specific_leaf_area ~ water_content, traits))
-
-traits.df <- cbind.data.frame(select(traits, plant.position, water_content:log_trichomes), 
-                              height_resid, density_resid, sla_resid)
-
-trait.names <- colnames(traits.df)[-1]
-
-gall.density.size.df <- full.df %>% 
-  select(vLG_abund, rG_abund, shootEst.no18, plant.position, vLG.height.mean) %>%
-  mutate(vLG_density = vLG_abund/shootEst.no18,
-         rG_density = rG_abund/shootEst.no18) %>%
-  select(plant.position:rG_density)
-
-gall.density.size.traits.df <- left_join(gall.density.size.df, traits.df)
-vLG.df <- na.omit(select(gall.density.size.traits.df, vLG_density, water_content:sla_resid))
-vLG.size.df <- na.omit(select(gall.density.size.traits.df, vLG.height.mean, water_content:sla_resid))
-rG.df <- na.omit(select(gall.density.size.traits.df, rG_density, water_content:sla_resid))
-#focal.df <- na.omit(gall.density.size.traits.df)
-#focal.df.traits <- focal.df[ ,trait.names]
-
-# Iteomyia salicisverruca (vLG) abundance model
-vLG.dens.glm <- glm(vLG_abund > 0 ~ offset(log(shootEst.no18)) + C_N_imputed + Total_Area + sal_tannin.PC1, full.df, family = binomial)
-summary(vLG.dens.glm)
-plot(vLG.dens.glm)
-
-vLG.dens.lm <- lm(log(vLG_abund) ~ offset(log(shootEst.no18)) + C_N_imputed + Total_Area + sal_tannin.PC1, filter(full.df, vLG_abund > 0))
-summary(vLG.dens.lm)
-plot(vLG.dens.lm)
-visreg(vLG.dens.lm)
-
-plot(log(vLG_abund/shootEst.no18) ~ sal_tannin.PC1, filter(full.df, vLG_abund > 0))
-library(MuMIn)
-
-dredge(vLG.full)
-vLG.null <- lm(vLG.df$vLG_density ~ 1, vLG.df[ ,trait.names])
-vLG.full <- lm(vLG.df$vLG_density ~ ., vLG.df[ ,trait.names])
-summary(vLG.full) # adj. R2 = 0.03 DOESN'T PASS THE TEST!!!
-vLG.step <- step(vLG.null, scope = formula(vLG.full), direction = "forward")
-vLG.update <- update(vLG.null, .~. + sal_tannin.PC1 + log_size + height_resid) 
-summary(vLG.update)
-#plot(vLG.update)
-
-vLG.size.null <- lm(vLG.size.df$vLG.height.mean ~ 1, vLG.size.df[ ,trait.names])
-vLG.size.full <- lm(vLG.size.df$vLG.height.mean ~ ., vLG.size.df[ ,trait.names])
-summary(vLG.size.full) # adj. R2 = 0.03 DOESN'T PASS THE TEST!!!
-vLG.size.step <- step(vLG.size.null, scope = formula(vLG.size.full), direction = "forward")
-vLG.size.update <- update(vLG.size.null, .~. + flavonOLES.PC1) 
-summary(vLG.size.update)
-
-rG.null <- lm(rG.df$rG_density ~ 1, rG.df[ ,trait.names])
-rG.full <- lm(rG.df$rG_density ~ ., rG.df[ ,trait.names])
-summary(rG.full) # adj. R2 = 0.12 DOESN'T PASS THE TEST!!!
-rG.step <- step(rG.null, scope = formula(rG.full), direction = "forward")
-rG.update <- update(rG.null, .~. + log_size  + height_resid + 
-    density_resid) 
-summary(rG.update)
-
-```

### 2015-04-18 Rscripts/morgue_ms_willow_network_supplement.R

@@ -1,3 +1,1641 @@
+```{r}
+require(visreg)
+abund <- which(full.df$vLG_abund > 0)
+summary(full.df$vLG_abund[abund])
+cut.vLG_abund <- cut(x = full.df$vLG_abund, breaks = c(1, 4, 22))
+ptism.glm <- glm(vLG_parasitized/vLG_abund ~ vLG.height.mean+vLG_abund, data = full.df, 
+                 weights = vLG_abund, family = "binomial")
+summary(ptism.glm)
+test <- visreg(ptism.glm, xvar = "vLG.height.mean", by = "vLG_abund", scale = "response")
+```
+
+```{r}
+## SG parasitism
+SG.ptized.glm <- glm(SG_parasitized/SG_abund ~ Genotype, data = full.df, weights = SG_abund, family = "binomial")
+summary(SG.ptized.glm)
+anova(SG.ptized.glm, test = "LR")
+
+## vLG parasitism
+vLG.ptized.df <- full.predictors %>%
+  mutate(vLG_parasitized = vLG_Tory + vLG_Eulo + vLG_Platy + vLG_Mesopol + vLG_Mymarid)
+table(vLG.ptized.df$Genotype) # data is unbalanced, but this doesn't qualitatively affect the outcome of the glm.
+vLG.ptized.3plus.genos <- c("*","B","D","E","F","H","I","K","L","O","S","V","X","Y","Z")
+colSums(vLG.ptized.df[ ,c("vLG_parasitized","vLG_abund","vLG_Tory","vLG_Eulo","vLG_Platy",
+                          "vLG_Mesopol","vLG_Mymarid")])
+
+# interesting, only marginally significant with quasibinomial
+vLG.ptized.bin <- glm(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~ Genotype,
+                      data = filter(vLG.ptized.df, Genotype %in% vLG.ptized.3plus.genos), 
+                      family = 'binomial')
+summary(vLG.ptized.bin)
+dfun(vLG.ptized.bin)
+anova(vLG.ptized.bin, test = "LR")
+require(dispmod)
+anova(glm.binomial.disp(vLG.ptized.bin), test = "LR") # still significant after accounting for overdispersion
+visreg(vLG.ptized.bin, scale = "response")
+
+vLG.ptized.pred <- glm(cbind(vLG_parasitized, vLG_abund - vLG_parasitized) ~ vLG.height.mean,
+                       data = vLG.ptized.df,
+                       family = 'binomial')
[…]

### 2015-04-18 Rscripts/simulation plots.R

@@ -79,6 +79,9 @@ web.measures.summary <- all.measures %>%
   group_by(Number.of.Genotypes) %>%
   summarise(mean.complexity = mean(total_complexity, na.rm = TRUE))
 with(web.measures.summary, max(mean.complexity)/min(mean.complexity))
+with(web.measures.summary, max(mean.complexity)/single.complexity.max) # 8% higher due to complimentarity effects
+
+max(all.measures$total_complexity, na.rm = TRUE)/single.complexity.max # up to 15% higher due to complimentarity effects.
 
 ## linear model with just the mean data. 
 web.sum.lm <- lm(log(mean.complexity) ~ log(Number.of.Genotypes), data = web.measures.summary)

### 2015-04-18 link_composition_plots.R

@@ -35,6 +35,8 @@ ABC.domptoids <- data.frame(x = rep(label.x.pos,3), y = rep(label.y.ptoids,3),
 D.ptism <- data.frame(x = label.x.pos, y = label.y.ptism, labels = "(D)")
 A.linkabund <- data.frame(x = 0.5, y = 10, labels = "(A)")
 B.ptism <- data.frame(x = 4.25, y = label.y.ptism, labels = "(B)")
+AB.ptism <- data.frame(x = rep(4.25,2), y = rep(1,2), 
+                       labels = c("(A)","(B)"), cut.vLG_abund = c("Low leaf gall abundance (1-4 per branch)", "High leaf gall abundance (5 - 22 per branch)"))
 #point.size <- 6
 line.widths <- 3 # for link plots
 ## create color-blind friendly palette with grey (taken from http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/)
@@ -152,6 +154,12 @@ link.df$Parasitoid <- factor(link.df$Parasitoid, levels = c("Platygaster",
 ## order genotypes by mean leaf gall abundance
 link.df$Genotype <- factor(link.df$Genotype, levels = vLG.df.sum$Genotype)
 
+## link summary
+link.summary <- link.df %>% group_by(Genotype, Gall, Parasitoid) %>% summarise_each(funs(mean)) %>% arrange(Gall)
+max(filter(link.summary, Gall == "Leaf gall", Parasitoid == "Platygaster")$value) # 3.7
+max(filter(link.summary,  Gall == "Leaf gall", Parasitoid == "Mesopolobus")$value) # 8
+max(filter(link.summary,  Gall == "Leaf gall", Parasitoid == "Torymus")$value) # 1.4
+
 ## link composition plot for all possible interactions. For supplementary materials.
 link.composition.plot <- ggplot(link.df, aes(x = Genotype, y = value, fill = Parasitoid)) +
   facet_grid(Parasitoid ~ Gall) + 
@@ -222,17 +230,18 @@ attack.df <- as.data.frame(tree_level_interaxn_all_plants_traits_size) %>%
   select(vLG.height.mean, vLG_abund, vLG_Platy, vLG_Mesopol, vLG_Tory) %>%
   gather(vLG.height.mean, vLG_abund)
 
-platy <- glm(value/vLG_abund ~ vLG.height.mean, data = filter(attack.df, variable == "vLG_Platy"), family = 'binomial', weights = vLG_abund)
[…]

### 2015-04-20 ms_willow_network_supplement.Rmd
first complete draft of network manuscript and supplementary info

@@
 summary(meandist(bray.gall.comm.sub, grouping = gall.comm.sub$Genotype))
+```
+
+```{r}
+gall.corrs.df <- full.df %>% select(Genotype, vLG.height.mean, vLG.gall.count, vLG_abund, rG_abund, aSG_abund)
+gall.corrs.pheno <- select(gall.corrs.df, -Genotype, -vLG.gall.count)
+library(psych)
+library(car)
+scatterplotMatrix(gall.corrs.pheno)
+corr.test(gall.corrs.pheno, use = "pairwise", method = "pearson")
+
+gall.corrs.geno <- gall.corrs.df %>%
+  group_by(Genotype) %>%
+  summarise(leaf_gall_size = weighted.mean(vLG.height.mean, vLG.gall.count, na.rm = TRUE),
+            leaf_gall_abund = mean(vLG_abund, na.rm = TRUE),
+            bud_gall_abund = mean(rG_abund, na.rm = TRUE),
+            apical_stem_gall_abund = mean(aSG_abund, na.rm = TRUE)) %>%
+  select(leaf_gall_size, leaf_gall_abund, bud_gall_abund, apical_stem_gall_abund)
+scatterplotMatrix(gall.corrs.geno)
+corr.test(gall.corrs.geno)
 ```

### 2015-04-20 Rscripts/simulation plots.R

@@ -9,6 +9,7 @@ require(gridBase)
 ## upload results from simulation
 all.measures <- read.csv("~/Documents/Genotype_Networks/data/food web complexity simulation.csv")
 dim(all.measures)[1] # 2221 unique simulations. 2425 simulations originally run
+table(all.measures$Number.of.Genotypes)
 
 single.complexity.max <- max(filter(all.measures, Number.of.Genotypes == 1)$total_complexity, na.rm = TRUE)
 
@@ -88,7 +89,19 @@ web.sum.lm <- lm(log(mean.complexity) ~ log(Number.of.Genotypes), data = web.mea
 summary(web.sum.lm) # Need to understand interpretation here.
 visreg(web.sum.lm) # not quite there
 
+## non-linear model with just the mean data
+require(mgcv)
+web.sum.gam <- gam(mean.complexity ~ s(Number.of.Genotypes), data = web.measures.summary)
+summary(web.sum.gam)
+visreg(web.sum.gam)
+
+
 ## linear model with all of the data. 
 web.lm <- lm(log(total_complexity) ~ log(Number.of.Genotypes), data = all.measures)
 summary(web.lm)
 visreg(web.lm)
+
+## non-linear model with all of the data
+web.gam <- gam(total_complexity ~ s(Number.of.Genotypes), data = all.measures)
+summary(web.gam)
+visreg(web.gam)

### 2015-06-30 Rscripts/complexity sim Gravel.R
manuscript before incorporating final coauthor comments
except I believe this has Jordi and Miguel’s comments

updated cover letter based on Greg and Jordi's comments

update link_composition_plots with reduced fig sizes

focused on fig 3 and 4

updated link_compostion_plots by resizing Fig_5

trying new figures sizes to fit in PNAS

updated text of ms with minor edits after reading paper copy in detail

updated figure legends with minor edits

updated dimensions for PNAS.

still need to convert to .tif and then try formatting.

converted .png to .tif in terminal

updated fig. 6 with lower res

@@ -1,0 +1,242 @@
+# In order to get an expectation of how my data is behaving, I could randomly shuffle everything and see what patterns I observe. Alternatively, I could create different scenarios to see what is going on (ref notebook). Let's try shuffling things up first...
+
+# PERHAPS BOOTSTRAPPING IS THE WAY TO GO. NEED TO THINK ABOUT THE ASSUMPTIONS BEHIND THIS A BIT MORE.
+# bootstrapping relies on one crucial assumption: the data accurately represents the true population...
+# add a calculation of functional diversity and see if this has a stronger relationship with food-web complexity.
+
+# Gravel simulation
+
+source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
+
+#require(mvabund)
+require(bipartite)
+require(tidyr)
+require(FD)
+
+df <- tree_level_interaxn_all_plants_traits_size %>%
+  tbl_df() %>%
+  filter(Genotype != "U") %>% # never collected any galls and thus gall-ptoid interactions so I removed it from the dataset. 
+  select(Genotype, 
+         willow_vLG = vLG_abund,
+         willow_rG = rG_abund,
+         willow_aSG = aSG_abund,
+         willow_SG = SG_abund,
+         vLG_Platy, vLG_Mesopol, vLG_Tory, vLG_Eulo, vLG_Mymarid,
+         rG_Platy, rG_Mesopol, rG_Tory, rG_Eulo, rG_Lestodip,
+         SG_Platy, aSG_Tory)#, #%>%#,
+         #Total_Area:HCH.tremulacin.__A220nm,
+         #water_content:N_imputed)# %>%
+ # group_by(Genotype) # this will ensure that the sample_n function subsamples each genotype
+
+geno.avgs <- df %>%
+  group_by(Genotype) %>%
+  summarise_each(funs(mean.narm = mean(., na.rm = TRUE), n())) 
+geno.avgs$avg.interaction <- rowMeans(geno.avgs[ ,2:17])
+geno.avgs$interaction.rich <- rowSums(geno.avgs[ ,2:17]>0)
+
+plot(geno.avgs$willow_vLG_n, geno.avgs$interaction.rich)
+cor.test(geno.avgs$avg.interaction, geno.avgs$willow_vLG_n)
+
+#df.new <- geno.avgs
+df.new <- select(df, Genotype, willow_vLG, willow_rG, willow_aSG,
[…]

### 2015-06-30 Rscripts/simulation plots.R

@@ -3,49 +3,80 @@ require(visreg)
 require(ggplot2)
 require(grid)
 require(dplyr)
-require(gridBase)
-#require(png)
+require(piecewiseSEM)
+require(semPlot)
 
 ## upload results from simulation
-all.measures <- read.csv("~/Documents/Genotype_Networks/data/food web complexity simulation.csv")
-dim(all.measures)[1] # 2221 unique simulations. 2425 simulations originally run
-table(all.measures$Number.of.Genotypes)
+all.measures <- read.csv("~/Documents/Genotype_Networks/data/food web complexity simulation Gravel.csv")
+#all.measures <- read.csv("~/Documents/Genotype_Networks/data/food web complexity simulation.csv")
+#dim(all.measures)[1] # 2221 unique simulations. 2425 simulations originally run
+#table(all.measures$Number.of.Genotypes)
 
-single.complexity.max <- max(filter(all.measures, Number.of.Genotypes == 1)$total_complexity, na.rm = TRUE)
+#single.complexity.max <- max(filter(all.measures, Number.of.Genotypes == 1)$total_complexity, na.rm = TRUE)
 
-## upload png of food web composition ordination
-#ord.png <- readPNG("~/Documents/Genotype_Networks/figures/full.link.composition.png")
-#ord.rast <- rasterGrob(ord.png, interpolate = TRUE)
+## plot updated Gravel simulation
+#all.measures.4 <- filter(all.measures, reps.sampled == 4)
+all.measures.NAzero <- all.measures
+all.measures.NAzero[is.na(all.measures.NAzero)] <- 0
+
+all.measures.NAone <- all.measures
+all.measures.NAone[is.na(all.measures.NAone)] <- 1
+
+# check plant sampling number
+ggplot(all.measures,#,filter( plants_sampled %in% c(3,4,6,9,9,10,12)),#.NAone,
+       aes(x = genotype.richness, y = total_complexity, group = reps.sampled, color = reps.sampled)) + # genotype.richness
+  #facet_wrap(~reps.sampled, nrow = 2) +
+  geom_point(color = "grey", shape = 1) +
+  geom_smooth(size = 2)
+
+ggplot(all.measures,#,filter( plants_sampled %in% c(3,4,6,9,9,10,12)),#.NAone,
+       aes(x = plants_sampled, y = total_complexity, group = genotype.richness, color = genotype.richness)) + # genotype.richness
+  #facet_wrap(~reps.sampled, nrow = 2) +
[…]

### 2015-06-30 link_composition_plots.R

@@ -1,64 +1,76 @@
 source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
 source('~/Documents/miscellaneous_R/ggplot_themes.R')
 source('~/Documents/ggnet/bipartite_plot_info.R')
+source('~/Documents/ggnet/tripartite_plot_info.R')
 
 require(ggplot2)
 require(gridExtra)
 require(dplyr)
 require(tidyr)
+require(vegan)
 
-## create theme for figures
+## for an unknown reason, ggsave needs to be modified for me to save my arrangeGrob object ----
+ggsave <- ggplot2::ggsave; body(ggsave) <- body(ggplot2::ggsave)[-2]
+
+## create theme for figures ----
 theme_links <- theme_bw() +
-  theme(axis.title.x = element_text(size = 16, vjust = 0.1), 
-        axis.title.y = element_text(size = 16, vjust = 1),
-        axis.text = element_text(size = 14),
-        strip.text = element_text(size = 14),
+  theme(axis.title.x = element_text(size = 11, vjust = -0.25), # size = 12
+        axis.title.y = element_text(size = 11, vjust = 1.25), # size = 12
+        axis.text.y = element_text(size = 9),#10
+        axis.text.x = element_text(size = 8), #12
+        strip.text = element_text(size = 9), #9
         panel.grid = element_blank(),
         legend.position = "none")
-point.size.range <- c(1,8)
-label.x.pos <- 1.5
-label.y.galls <- 23
-label.y.gallsize <- 14
-label.y.ptoids <- 9.5
-label.y.ptism <- 1
+point.size.range <- c(1,5) #c(1,5)
+point.size.range.ptism <- c(1,3.5)#c(0.5,2)
+label.x.pos <- 1.5 #2
+label.y.galls <- 23 # 21
+label.y.gallsize <- 13.75
+label.y.ptoids <- 9.5 #9.75
+label.y.ptism <- 1.05 #0.975
+label.size <- 3 # 3
 ABCD.allgalls <- data.frame(x = rep(label.x.pos,4), y = rep(label.y.galls,4), 
-                            variable = c("Leaf gall", "apical-Stem gall", 
-                                         "Bud gall", "mid-Stem gall"), 
+                            variable = c("Leaf gall", "Apical-stem gall", 
+                                         "Bud gall", "Mid-stem gall"), 
                             labels = c("(A)","(B)","(C)","(D)"))
[…]

### 2016-01-25 rptR/R/pqlglmm.pois.model.R
removed PNAS digital guidelines and rptR package that I never used for the final analysis

@@ -1,18 +1,0 @@
-pqlglmm.pois.model <- function(y, groups, link, returnR=TRUE) {
-  mod     <-  glmmPQL(y ~ 1,random=~1|groups,  family=quasipoisson(link=eval(link)), verbose=FALSE) 
-  VarComp <- nlme::VarCorr(mod)
-  beta0   <- as.numeric(mod$coefficients$fixed)
-  omega   <- (as.numeric(VarComp[2,1]))
-  var.a   <- (as.numeric(VarComp[1,1]))
-  if (link=="log") {
-    R.link  <- var.a/(var.a + omega*log(1/exp(beta0)+1))
-    EY 		<- exp(beta0+var.a/2)
-    R.org 	<- EY*(exp(var.a)-1)/(EY*(exp(var.a)-1)+omega) 
-  }
-  if (link=="sqrt") {
-    R.link  <- var.a/(var.a + omega*0.25)
-    R.org 	<- NA 
-  }	
-  if(returnR) return(list(R.link=R.link, R.org=R.org))
-  else return(list(beta0=beta0, omega=omega, var.a=var.a))
-}

### 2016-01-25 rptR/R/pqlglmm.pois.model.offset.R

@@ -1,18 +1,0 @@
-pqlglmm.pois.model.offset <- function(y, offset.trans, groups, link, returnR=TRUE) {
-  mod     <-  glmmPQL(y ~ 1 + offset(offset.trans),random=~1|groups,  family=quasipoisson(link=eval(link)), verbose=FALSE, control = lmeControl(opt = "optim")) 
-  VarComp <- nlme::VarCorr(mod)
-  beta0   <- as.numeric(mod$coefficients$fixed)
-  omega   <- (as.numeric(VarComp[2,1]))
-  var.a   <- (as.numeric(VarComp[1,1]))
-  if (link=="log") {
-    R.link  <- var.a/(var.a + omega*log(1/exp(beta0)+1))
-    EY   	<- exp(beta0+var.a/2)
-    R.org 	<- EY*(exp(var.a)-1)/(EY*(exp(var.a)-1)+omega) 
-  }
-  if (link=="sqrt") {
-    R.link  <- var.a/(var.a + omega*0.25)
-    R.org 	<- NA 
-  }	
-  if(returnR) return(list(R.link=R.link, R.org=R.org))
-  else return(list(beta0=beta0, omega=omega, var.a=var.a))
-}

### 2016-01-25 rptR/R/print.rpt.R

@@ -1,74 +1,0 @@
-print.rpt <- function(x, ...) {
-	if(x$datatype=="Gaussian" & length(x$P)==1 & length(x$R)==1) {
-		cat("\n", "Repeatability calculation using the ", x$method, " method", "\n\n",
-			"R  = ", round(x$R,3), "\n",
-			"SE = ", round(x$se,3), "\n",
-			"CI = [", round(x$CI.R[1],3), ", ", round(x$CI.R[2],3), "]", "\n",
-			"P  = ", signif(x$P, 3), "\n\n", 
-			sep="")  		
-	} 
-	if(x$datatype=="Gaussian" & length(x$P)>1 & length(x$R)==1) {
-		cat("\n", "Repeatability calculation using the ", x$method, " method", "\n\n",
-			"R  = ", round(x$R,3), "\n",
-			"SE = ", round(x$se,3), "\n",
-			"CI = [", round(x$CI.R[1],3), ", ", round(x$CI.R[2],3), "]", "\n",
-			"P  = ", signif(x$P[1], 3), " [", attr(x$P, "names")[1], "]", "\n", 
-			"     ", signif(x$P[2], 3), " [", attr(x$P, "names")[2], "]", "\n\n", 
-			sep="")
-	} 
-	if(x$datatype=="Gaussian" & length(x$P)==1 & length(x$R)>1) {
-		print("h3")
-		cat("\n", "Repeatability calculation using the ", x$method, " method", "\n\n")
-		for(i in 1: length(x$R)) {
-			cat("Repeatability for ", names(x$R)[i], "\n",
-			"R  = ", round(x$R[i],3), "\n",
-			"SE = ", round(x$se[i],3), "\n",
-			"CI = [", round(x$CI.R[i,1],3), ", ", round(x$CI.R[i,2],3), "]", "\n",
-			"P  = ", signif(x$P[i], 3), "\n\n", 
-			sep="")
-		}
-	}
-	if(x$datatype=="Gaussian" & length(x$P)>1 & length(x$R)>1) {
-		cat("\n", "Repeatability calculation using the ", x$method, " method", "\n\n")
-		for(i in 1: length(x$R)) {
-			cat("Repeatability for ", names(x$R)[i], "\n",
-			"R  = ", round(x$R[i],3), "\n",
-			"SE = ", round(x$se[i],3), "\n",
[…]

### 2016-01-25 rptR/R/rpt.R

@@ -1,36 +1,0 @@
-rpt <- function(y, groups, 
-			   datatype=c("Gaussian", "binomial", "proportion", "count"),  
-			   method=c("corr", "ANOVA", "REML", "MCMC", "GLMM.add", "GLMM.multi"),  
-			   link=c("logit", "probit", "log", "sqrt"),
-			   CI=0.95, nboot=1000, npermut=1000) {
-	if(datatype=="Gaussian") {
-		if(length(method)>1) {
-			warning("Linear mixed model fitted by REML used by default. Change using argument 'method', if required ('corr', 'ANOVA', 'REML' and 'MCMC' allowed for Gaussian data).")
-			method<-"REML" 
-		}
-		if (method=="REML")  return(rpt.remlLMM(y, groups, CI=CI, nboot=nboot, npermut=npermut))
-		if (method=="MCMC")  return(rpt.mcmcLMM(y, groups, CI=CI))
-		if (method=="ANOVA") return(rpt.aov(y, groups, CI=CI, npermut=npermut))	
-		if (method=="corr")  return(rpt.corr(y, groups, CI=CI, nboot=nboot, npermut=npermut)) 
-	}
-	if(datatype=="binomial" | datatype=="proportion") {
-		if(length(method)>1) {
-			warning("Generalised linear mixed model with multiplicative overdispersion fitted by PQL used by default. Change using argument 'method', if required ('GLMM.add' and 'GLMM.multi' allowed for Binomial data).")
-			method<-"GLMM.multi" 
-		}
-		if (method=="GLMM.multi") return(rpt.binomGLMM.multi(y, groups, link, CI=CI, nboot=nboot, npermut=npermut))
-		if (method=="GLMM.add") return(rpt.binomGLMM.add(y, groups, CI=CI))
-	}
-	if(datatype=="count") {
-		if(length(method)>1) {
-			warning("Generalised linear mixed model with multiplicative overdispersion fitted by PQL used by default. Change using argument 'method', if required ('GLMM.add' and 'GLMM.multi' allowed for count data).")
-			method<-"GLMM.multi"
-		}
-		if(length(link)>1) {
-			link="log"
-			warning("Log link will be used.")
-		}
-		if (method=="GLMM.multi") return(rpt.poisGLMM.multi(y, groups, link, CI=CI, nboot=nboot, npermut=npermut)) 
-		if (method=="GLMM.add")   return(rpt.poisGLMM.add(y, groups, CI=CI)) 
-	} 
-}

### 2016-01-25 rptR/R/rpt.adj.R

@@ -1,33 +1,0 @@
-rpt.adj <- function(formula, grname, data,
-			   datatype=c("Gaussian", "binomial", "proportion", "count"),  
-			   method=c("corr", "ANOVA", "REML", "MCMC", "GLMM.add", "GLMM.multi"),  
-			   link=c("logit", "probit", "log", "sqrt"),
-			   CI=0.95, nboot=1000, npermut=1000) {
-	if(datatype=="Gaussian") {
-		if(length(method)>1) {
-			warning("Linear mixed model fitted by REML used by default. Change using argument 'method', if required ('corr', 'ANOVA', 'REML' and 'MCMC' allowed for Gaussian data).")
-			method<-"REML" 
-		}
-		if (method=="REML")  return(rpt.remlLMM.adj(formula, grname, data, CI=CI, nboot=nboot, npermut=npermut))
-		if (method=="MCMC")  warning("Not jet implemented") # return(rpt.mcmcLMM.adj(formula, grname, data, CI=CI))
-		if (method=="ANOVA") warning("Not jet implemented") # return(rpt.aov(y, groups, CI=CI, npermut=npermut))	
-		if (method=="corr")  warning("Not jet implemented") # return(rpt.corr(y, groups, CI=CI, nboot=nboot, npermut=npermut)) 
-	}
-	if(datatype=="binomial" | datatype=="proportion") {
-		if(length(method)>1) {
-			warning("Generalised linear mixed model with multiplicative overdispersion fitted by PQL used by default. Change using argument 'method', if required ('GLMM.add' and 'GLMM.multi' allowed for Binomial data).")
-			method<-"GLMM.multi" 
-		}
-		if (method=="GLMM.multi") warning("Not jet implemented") # return(rpt.binomGLMM.multi(y, groups, link, CI=CI, nboot=nboot, npermut=npermut))
-		if (method=="GLMM.add") warning("Not jet implemented") # return(rpt.binomGLMM.add(y, groups, CI=CI))
-	}
-	if(datatype=="count") {
-		if(length(method)>1) {
-			warning("Generalised linear mixed model with multiplicative overdispersion fitted by PQL used by default. Change using argument 'method', if required ('GLMM.add' and 'GLMM.multi' allowed for count data).")
-			method<-"GLMM.multi"
-		}
-		if(length(link)>1) link="log"
[…]

### 2016-01-25 rptR/R/rpt.aov.R

@@ -1,42 +1,0 @@
-rpt.aov <- function(y, groups, CI=0.95, npermut=1000) {	
-	# initial checks
-    if(length(y) != length(groups)) stop("y and group are not of equal length")
-	if(npermut < 1) npermut <- 1
-	# preparation
-	groups <- factor(groups)
-    k  <- length(levels(groups))
-	N  <- length(y)
-    n0 <- 1/(k-1)*(N - sum(table(groups)^2)/N)
-	# functions: point estimates of R
-	R.pe <- function(y, groups, n0) {
-		gm   <- mean(y)
-		MSa  <- sum(tapply(y, groups, function(x) (mean(x)-gm)^2 * length(x)))   / (k-1)
-		MSw  <- sum(tapply(y, groups, function(x) sum((x - mean(x))^2)))  / (N-k)
-		R    <- ((MSa-MSw)/n0 ) / ( (MSa-MSw)/n0 + MSw)
-		return(R) 
-	}
-	# point estimation according to equations 4 and 5
-	R        <- R.pe(y, groups, n0)
-	# confidence interval estimation according to equation 6 and 7
-	se       <- sqrt(( 2*(N-1)*(1-R)^2 * (1 + (n0-1)*R)^2) / (n0^2 * (N-k)*(k-1)))
-	CI.R     <- R + c(1,-1)* qt((1-CI)/2,k-1)*se
-	# significance test from ANOVA
-	P.aov    <- anova(lm(y ~ groups))[5][1,1]
-	# significance test by permutation
-	permut <- function(y, groups, N, n0) {
-		sampy <- sample(y, N)
-		return(R.pe(sampy, groups, n0))
-	}
-	if(npermut > 1) {
-		R.permut <- c(R, replicate(npermut-1, permut(y, groups, N, n0), simplify=TRUE))
-		P.permut <- sum(R.permut >= R)/npermut
-	}
-	else {
-		R.permut <- R
-		P.permut <- NA
-	}
-	# return of results
-	res <- list(call=match.call(), datatype="Gaussian", method="ANOVA", CI=CI, R=R, se=se, CI.R=CI.R, P=c(P.aov=P.aov, P.permut=P.permut), R.permut=R.permut) 
-	class(res) <- "rpt"
-	return(res) 
-}

### 2016-01-25 rptR/R/rpt.binomGLMM.add.R

@@ -1,43 +1,0 @@
-rpt.binomGLMM.add <- function(y, groups, CI=0.95, prior=NULL, verbose=FALSE, ...) {
-	# initial checks
-	if (is.null(dim(y))) 
-		y      <- cbind(y, 1-y)
-	if(any(is.na(y))) warning("missing values in y")
-	# preparation	
-	n          <- rowSums(y)
-	groups     <- factor(groups)
-	# model fitting
-	if(all(n==1)) {
-		if(is.null(prior)) prior=list(R=list(V=1,fix=1),G=list(G1=list(V=1,nu=1,alpha.mu=0,alpha.V=25^2)))
-		mod    <- MCMCglmm(m ~ 1, random= ~ groups, data=data.frame(m=y[,1],nm=y[,2],groups=groups), prior=prior, family="categorical", verbose=verbose, ...) 
-	}
-	else {
-		if(is.null(prior)) prior=list(R=list(V=1e-10,nu=-1),G=list(G1=list(V=1,nu=1,alpha.mu=0,alpha.V=25^2)))
-		mod    <- MCMCglmm(cbind(m, nm) ~ 1, random= ~ groups, data=data.frame(m=y[,1],nm=y[,2],groups=groups), prior=prior, family="multinomial2", verbose=verbose, ...)
-	}
-	# ezxtraction of posterior distributions
-	var.a      <- mod$VCV[,"groups"]
-	var.e      <- mod$VCV[,"units"]
-	postR.link <- var.a / (var.a + var.e + pi^2 /3)
-	beta0      <- mod$Sol
-	P          <- exp(beta0) / (1+ exp(beta0))
-	postR.org  <- (var.a * P*P / (1+exp(beta0))^2 ) / ((var.a + var.e)* P*P / (1+exp(beta0))^2 +  P*(1-P))
-	# point estimate on link and original scale
-	R.link     <- posterior.mode( postR.link )
-	R.org      <- posterior.mode( postR.org )
-	# credibility interval estimation from paterior distribution
-	CI.link    <- coda::HPDinterval(postR.link, CI)[1,]
-	CI.org     <- coda::HPDinterval(postR.org, CI)[1,]
-	se.link    <- sd(postR.link)
-	se.org     <- sd(postR.org)
-	# P value equivilents that are not appropriate
-	P.link     <- NA
-	P.org      <- NA
-	# return of results
[…]

### 2016-01-25 rptR/R/rpt.binomGLMM.multi.R

@@ -1,105 +1,0 @@
-rpt.binomGLMM.multi <- function(y, groups, link=c("logit", "probit"), CI=0.95, nboot=1000, npermut=1000) {
-	# initial checks
-	if (is.null(dim(y))) 
-		y <- cbind(y, 1-y)
-	if (nrow(y) != length(groups)) 
-		stop("y and group have to be of equal length")
-	if(nboot < 0) 	nboot <- 0
-	if(npermut < 1) npermut <- 1
-	if(length(link) > 1)
-		link   <- link[1]
-	if(link != "logit" & link != "probit") 
-		stop("inappropriate link (has to be 'logit' or 'probit')")
-	if(any(is.na(y))) {
-		warning("missing values in y are removed")
-		groups <- groups[-rowSums(is.na(y)) > 0]
-		y      <- y[-rowSums(is.na(y)) > 0,]
-	}
-	# preparation
-	groups <- factor(groups)
-	n <- rowSums(y)
-	N <- nrow(y)
-	k <- length(levels(groups))
-	# functions
-	pqlglmm.binom.model <- function(y, groups, n, link, returnR=TRUE) {
-		if(all(n==1)) mod <- glmmPQL(y ~ 1, random=~1|groups,  family=binomial(link=eval(link)), verbose=FALSE)
-		else          mod <- glmmPQL(y ~ 1, random=~1|groups,  family=quasibinomial(link=eval(link)), verbose=FALSE)	
-		VarComp  <- nlme::VarCorr(mod)
-		beta0    <- as.numeric(mod$coefficients$fixed)
-		if(all(n==1)) omega <- 1
-			else      omega <- (as.numeric(VarComp[2,1]))
-		var.a  <- (as.numeric(VarComp[1,1]))
-		if (link=="logit") {
-			R.link  <- var.a / (var.a+omega*pi^2 /3)
-			P 		<- exp(beta0) / (1+ exp(beta0))
-			R.org   <- (var.a * P*P / (1+exp(beta0))^2 ) / (var.a * P*P / (1+exp(beta0))^2 + omega * P*(1-P)) 
-		}
-		if (link=="probit") {
-			R.link  <- var.a / (var.a+omega)
-			R.org   <- NA 
-		}
-		if(returnR) return(list(R.link=R.link, R.org=R.org))
-		else return(list(beta0=beta0, omega=omega, var.a=var.a)) 
-	}
-	# point estimate
[…]

### 2016-01-25 rptR/R/rpt.corr.R

@@ -1,55 +1,0 @@
-rpt.corr <- function(y, groups, CI=0.95, nboot=1000, npermut=1000) {
-	# initial checks
-	if(length(y)!= length(groups)) 
-		stop("y and group have to be of equal length")
-	if(nboot < 0)	nboot <- 0
-	if(npermut < 1) npermut <- 1
-	if(any(is.na(y))) 
-		stop("missing values in y ")
-	if(any(is.na(groups)))
-		stop("missing values in groups ")
-	if(!all(table(groups)==2))     
-		stop("not exactly two data points per group")
-	# preparation
-	sortix <- sort.int(as.numeric(groups),index.return=TRUE)$ix
-	y1     <- y[sortix][seq(1,length(y),by=2)]
-	y2     <- y[sortix][seq(2,length(y),by=2)]
-	k      <- length(unique(groups))
-	# functions: point estimates of R
-	R.pe  <- function(y1, y2, k) {
-		y <- c(y1, y2)
-		R <- 1/(k-1)*sum((y1-mean(y))*(y2-mean(y))) / var(y)
-		return(R) 
-	}	
-	# point estimation according to equations 4 and 5
-	R      <- R.pe(y1, y2, k)
-	# confidence interval estimation according to equation 6 and 7
-	bootstr <- function(y1, y2, k) {
-		samp<- sample(1:k, k, replace=TRUE)
-		R.pe(y1[samp], y2[samp], k)
-	}
-	if(nboot > 0)
-		R.boot <- replicate(nboot, bootstr(y1, y2, k), simplify=TRUE) 
-	else
-		R.boot <- NA
-	CI.R   <- quantile(R.boot, c((1-CI)/2, 1-(1-CI)/2), na.rm=TRUE)
-	se     <- sd(R.boot, na.rm=TRUE)
-	# significance test by permutation
-	permut   <- function(y1, y2, k) {
-		samp <- sample(1:k, k)
-		R.pe(y1[samp], y2, k)
-	}
-	if(npermut > 1) {
-		R.permut <- c(R, replicate(npermut-1, permut(y1, y2, k), simplify=TRUE))
-		P.permut <- sum(R.permut >= R)/npermut
-	}
-	else {
-		R.permut <- R
-		P.permut <- NA
-	}
-	# return of results
-	res <- list(call=match.call(), datatype="Gaussian", method="corr", CI=CI, 
[…]

### 2016-01-25 rptR/R/rpt.mcmcLMM.R

@@ -1,23 +1,0 @@
-rpt.mcmcLMM <- function(y, groups, CI=0.95, prior=NULL, verbose=FALSE, ...){
-	# initial checks
-	if(length(y)!= length(groups)) stop("y and group are of unequal length")
-	# preparation
-	groups <- factor(groups)
-	if(is.null(prior)) prior <- list(R=list(V=1,n=10e-2), G=list(G1=list(V=1,n=10e-2)) )
- 	# point estimation according to model 8 and equation 9
-	mod   <- MCMCglmm(y ~ 1, random=~groups, family="gaussian", data=data.frame(y=y,groups=groups), prior=prior, verbose=verbose, ...)
-	var.a <- mod$VCV[,"groups"]
-	var.e <- mod$VCV[,"units"]
-	postR <- var.a / (var.a + var.e)
-	# point estimate
-	R     <- posterior.mode( postR )
-	# credibility interval estimation from paterior distribution
-	CI.R    <- coda::HPDinterval(postR,CI)[1,]
-	se 	    <- sd(postR)
-	# 'significance test'
-	P 	  <- NA
-	res = list(call=match.call(), datatype="Gaussian", method="LMM.MCMC", CI=CI, 
-				R=R, CI.R=CI.R, se=se, P=P, R.post=postR ) 
-	class(res) <- "rpt"
-	return(res) 
-}

### 2016-01-25 rptR/R/rpt.poisGLMM.add.R

@@ -1,34 +1,0 @@
-rpt.poisGLMM.add <- function(y, groups, CI=0.95, prior=NULL, verbose=FALSE, ...) {
-	# initial checks
-	if(any(is.na(y))) warning("missing values in y")
-	# preparation
-	groups     <- factor(groups)
-	# model fitting
-	if(is.null(prior)) prior=list(R=list(V=1e-10,nu=-1),G=list(G1=list(V=1,nu=1,alpha.mu=0,alpha.V=25^2)))
-	mod        <- MCMCglmm(y ~ 1, random=~groups, family="poisson", data=data.frame(y=y,groups=groups), prior=prior, verbose=verbose, ...)
-	# ezxtraction of posterior distributions
-	var.a      <- mod$VCV[,"groups"]
-	var.e      <- mod$VCV[,"units"]
-	beta0      <- mod$Sol
-	postR.link <- var.a/(var.a + var.e+log(1/exp(beta0)+1))
-	EY 		   <- exp(beta0+(var.e+var.a)/2)
-	postR.org  <- EY*(exp(var.a)-1)/(EY*(exp(var.e+var.a)-1)+1)
-	# point estimate on link and original scale
-	R.link     <- posterior.mode( postR.link )
-	R.org      <- posterior.mode( postR.org )
-	# credibility interval estimation from paterior distribution
-	CI.link    <- coda::HPDinterval(postR.link, CI)[1,]
-	CI.org     <- coda::HPDinterval(postR.org, CI)[1,]
-	se.link    <- sd(postR.link)
-	se.org     <- sd(postR.org)
-	# P value equivilents that are not appropriate
-	P.link     <- NA
-	P.org      <- NA
-	# return of results
-	res 	   <- list(call=match.call(), datatype="count", method="MCMC", CI=CI,
-			      R.link=R.link, se.link=se.link, CI.link=CI.link, P.link=P.link, 
-				  R.org = R.org, se.org=se.org, CI.org=CI.org, P.org=P.org,
-				  R.post=list(R.link=as.vector(postR.link), R.org=as.vector(postR.org)) )
-	class(res) <- "rpt"
-	return(res) 
-}

### 2016-01-25 rptR/R/rpt.poisGLMM.multi.R

@@ -1,89 +1,0 @@
-rpt.poisGLMM.multi = function(y, groups, link=c("log", "sqrt"), CI=0.95, nboot=1000, npermut=1000) {
-	# initial checks
-	if(length(y) != length(groups)) 
-		stop("y and group hav to be of equal length")
-	if(nboot < 0)   nboot <- 0
-	if(npermut < 1) npermut <- 1
-	if(length(link) > 1)
-		link   <- link[1]
-	if(link != "log" &  link != "sqrt") 
-		stop("inappropriate link (has to be 'log' or 'sqrt')")
-	if(any(is.na(y))) {
-		warning("missing values in y are removed")
-		groups <- groups[!is.na(y)]
-		y      <- y[!is.na(y)]
-	}
-	# preparation
-	groups <- factor(groups)
-	N <- length(y)
-	k <- length(levels(groups))
-	# functions
-	pqlglmm.pois.model <- function(y, groups, link, returnR=TRUE) {
-		mod     <-  glmmPQL(y ~ 1,random=~1|groups,  family=quasipoisson(link=eval(link)), verbose=FALSE) 
-		VarComp <- nlme::VarCorr(mod)
-		beta0   <- as.numeric(mod$coefficients$fixed)
-		omega   <- (as.numeric(VarComp[2,1]))
-		var.a   <- (as.numeric(VarComp[1,1]))
-		if (link=="log") {
-			R.link  <- var.a/(var.a + omega*log(1/exp(beta0)+1))
-			EY 		<- exp(beta0+var.a/2)
-			R.org 	<- EY*(exp(var.a)-1)/(EY*(exp(var.a)-1)+omega) 
-		}
-		if (link=="sqrt") {
-			R.link  <- var.a/(var.a + omega*0.25)
-			R.org 	<- NA 
-		}	
-		if(returnR) return(list(R.link=R.link, R.org=R.org))
-		else return(list(beta0=beta0, omega=omega, var.a=var.a))
-	}
-	# point estimation according to model 17 equations 18-20
-	R   <- pqlglmm.pois.model(y, groups, link)
-	# confidence interval estimation by parametric bootstrapping
-	bootstr <- function(y, groups, k, N, beta0, var.a, omega, link) {
-		groupMeans <- rnorm(k, 0, sqrt(var.a))
-		if(link=="log")  mu <- exp(beta0 + groupMeans[groups])
[…]

### 2016-01-25 rptR/R/rpt.poisGLMM.multi.offset.MAB.R

@@ -1,93 +1,0 @@
-# Matt Barbour has added an offset parameter to allow for estimation of density.
-
-rpt.poisGLMM.multi.offset = function(y, offset.trans, groups, link=c("log", "sqrt"), CI=0.95, nboot=1000, npermut=1000) {
-  # initial checks
-  if(length(y) != length(groups)) 
-    stop("y and group hav to be of equal length")
-  if(nboot < 0)   nboot <- 0
-  if(npermut < 1) npermut <- 1
-  if(length(link) > 1)
-    link   <- link[1]
-  if(link != "log" &  link != "sqrt") 
-    stop("inappropriate link (has to be 'log' or 'sqrt')")
-  if(any(is.na(y))) {
-    warning("missing values in y are removed")
-    groups <- groups[!is.na(y)]
-    y      <- y[!is.na(y)]
-  }
-  # preparation
-  groups <- factor(groups)
-  N <- length(y)
-  k <- length(levels(groups))
-  # functions
-  pqlglmm.pois.model <- function(y, offset.trans, groups, link, returnR=TRUE) {
-    mod     <-  glmmPQL(y ~ 1 + offset(offset.trans),random=~1|groups,  family=quasipoisson(link=eval(link)), verbose=FALSE) 
-    VarComp <- nlme::VarCorr(mod)
-    beta0   <- as.numeric(mod$coefficients$fixed)
-    omega   <- (as.numeric(VarComp[2,1]))
-    var.a   <- (as.numeric(VarComp[1,1]))
-    if (link=="log") {
-      R.link  <- var.a/(var.a + omega*log(1/exp(beta0)+1))
-      EY 		<- exp(beta0+var.a/2)
-      R.org 	<- EY*(exp(var.a)-1)/(EY*(exp(var.a)-1)+omega) 
-    }
-    if (link=="sqrt") {
-      R.link  <- var.a/(var.a + omega*0.25)
-      R.org 	<- NA 
-    }	
-    if(returnR) return(list(R.link=R.link, R.org=R.org))
-    else return(list(beta0=beta0, omega=omega, var.a=var.a))
-  }
-  # point estimation according to model 17 equations 18-20
-  R   <- pqlglmm.pois.model(y, offset.trans, groups, link)
-  
-  #browser()
-  # confidence interval estimation by parametric bootstrapping
[…]

### 2016-01-25 rptR/R/rpt.remlLMM.R

@@ -1,76 +1,0 @@
-rpt.remlLMM <- function(y, groups, CI=0.95, nboot=1000, npermut=1000) {
-	# initial checks
-	if(length(y)!= length(groups)) stop("y and group are of unequal length")
-	if(nboot < 0) 	nboot <- 0
-	if(npermut < 1) npermut <- 1
-	if(any(is.na(y))) {
-		warning("missing values in y are removed")
-		groups <- groups[!is.na(y)]
-		groups <- factor(groups)
-		y      <- y[!is.na(y)] 
-	}
-	# preparation
-	groups <- factor(groups)
-    k <- length(unique(groups))
-	N <- length(y)
-	# functions: point estimates of R
-	R.pe <- function(y, groups) {
-        varComps <- try(nlme::VarCorr(lme(y ~ 1, random = ~1 | groups)), silent=TRUE)
-		if(class(varComps)=="try-error") {
-			warning("Convergence problems in lme (most likely during bootstrap or permutation). R is set to NA for this iteration")
-			R = NA
-		}
-		else {
-			var.a <- as.numeric(varComps[1, 1])
-			var.e <- as.numeric(varComps[2, 1])
-			R <- var.a/(var.a + var.e)
-		}
-		return(R) 
-	}
-	# point estimation according to model 8 and equation 9
-	R <- R.pe(y, groups)
-	# confidence interval estimation by parametric bootstrapping
-	bootstr <- function(y, groups, k, N, beta0, var.a, var.e) {
-		y.boot <- beta0 + rnorm(k, 0, sqrt(var.a))[groups] + rnorm(N, 0, sqrt(var.e))
-		R.pe(y.boot, groups) 
-	}
-	if(nboot > 0) {
-		mod <- try(lme(y ~ 1, random = ~1 | groups), silent=TRUE)
-		if(class(mod)=="try-error") {
-			warning("Convergence problems in lme. Model is refitted.")
-			mod <- lme(y ~ 1, random = ~1 | groups)                
-		}
-		beta0    <- as.numeric(summary(mod)$tTable[,1])
-		varComps <- nlme::VarCorr(mod)
-		var.a    <- as.numeric(varComps[1,1])
-		var.e    <- as.numeric(varComps[2,1])	
[…]

### 2016-01-25 rptR/R/rpt.remlLMM.adj.R

@@ -1,72 +1,0 @@
-rpt.remlLMM.adj = function(formula, grname, data, CI=0.95, nboot=1000, npermut=1000) {
-	mod         <- lmer(formula, data=data)
-	if(nboot < 0) 	nboot <- 0
-	if(npermut < 1) npermut <- 1
-	e1 = environment()
-	# point estimates of R
-	R.pe <- function(formula, data, grname, peYN=FALSE) {
-		mod.fnc = lmer(formula, data)
-		varComps <- lme4::VarCorr(mod.fnc)
-		if(peYN & any(varComps==0) & nboot > 0) {
-			assign("nboot", 0, envir=e1)
-			warning("(One of) the point estimate(s) for the repeatability was exactly zero; parametric bootstrapping has been skipped.")
-		}
-		var.a    <- as.numeric(varComps[grname])
-		var.p    <- sum(as.numeric(varComps)) + attr(varComps, "sc")^2
-		#var.e    <- as.numeric(attr(varComps, "sc")^2)
-		R        <- var.a / var.p
-		return(R) 
-	}
-	R <- R.pe(formula, data, grname, peYN=TRUE)
-	names(R) = grname
-	# confidence interval estimation by parametric bootstrapping
-	bootstr <- function(mod, formula, data, grname) {
-		mod.sim <- arm::sim(mod, n.sim=2)   # for some reason it is not possible to set n.sim=1
-		y <- mod@X %*% matrix(mod.sim@fixef[1,]) + t(as.matrix(mod@Zt)) %*% matrix(unlist(mod.sim@ranef)[seq(1, length(unlist(mod.sim@ranef)), by=2)])
-		y <- y + rnorm(length(y), 0, attr(lme4::VarCorr(mod), "sc"))
-		data[,names(mod@frame)[1]] = as.vector(y)
-		R.pe(formula, data, grname)
-	}
-	if(nboot > 0) R.boot   <- replicate(nboot, bootstr(mod, formula, data, grname), simplify=TRUE)
-		else R.boot <- matrix(rep(NA, length(grname)), nrow=length(grname))
-	if(length(grname) == 1) {
-		CI.R     <- quantile(R.boot, c((1-CI)/2,1-(1-CI)/2), na.rm=TRUE)
-		se <- sd(R.boot)
-		names(se) = grname }
-	else {
[…]

### 2016-01-25 rptR/R/rptR.functions.R

@@ -1,73 +1,0 @@
-#### Functions taken from rptR package
-
-# general count model function
-pqlglmm.pois.model <- function(y, # response variable
-                               groups, # grouping variable
-                               link, # link function
-                               returnR=TRUE) {
-  mod     <-  glmmPQL(y ~ 1,random=~1|groups,  family=quasipoisson(link=eval(link)), verbose=FALSE) 
-  VarComp <- nlme::VarCorr(mod)
-  beta0   <- as.numeric(mod$coefficients$fixed)
-  omega   <- (as.numeric(VarComp[2,1]))
-  var.a   <- (as.numeric(VarComp[1,1]))
-  if (link=="log") {
-    R.link  <- var.a/(var.a + omega*log(1/exp(beta0)+1))
-    EY   	<- exp(beta0+var.a/2)
-    R.org 	<- EY*(exp(var.a)-1)/(EY*(exp(var.a)-1)+omega) 
-  }
-  if (link=="sqrt") {
-    R.link  <- var.a/(var.a + omega*0.25)
-    R.org 	<- NA 
-  }	
-  if(returnR) return(list(R.link=R.link, R.org=R.org))
-  else return(list(beta0=beta0, omega=omega, var.a=var.a))
-}
-
-# general count model with offset option. Most parameters same as for general count model
-pqlglmm.pois.model.offset <- function(y, 
-                                      offset.trans, # variable to use for offsetting that has already been transformed according to the link function used.
-                                      groups, 
-                                      link, 
-                                      returnR=TRUE) {
-  mod     <-  glmmPQL(y ~ 1 + offset(offset.trans),random=~1|groups,  family=quasipoisson(link=eval(link)), verbose=FALSE) 
-  VarComp <- nlme::VarCorr(mod)
-  beta0   <- as.numeric(mod$coefficients$fixed)
-  omega   <- (as.numeric(VarComp[2,1]))
-  var.a   <- (as.numeric(VarComp[1,1]))
-  if (link=="log") {
-    R.link  <- var.a/(var.a + omega*log(1/exp(beta0)+1))
-    EY     <- exp(beta0+var.a/2)
-    R.org 	<- EY*(exp(var.a)-1)/(EY*(exp(var.a)-1)+omega) 
-  }
-  if (link=="sqrt") {
-    R.link  <- var.a/(var.a + omega*0.25)
-    R.org 	<- NA 
-  }	
[…]

### 2016-01-25 morgue/ms_willow_network.Rmd
moved old R scripts to morgue

---
title: "Plant genetic variation drives dissimilarity in an insect food web"
author: "Matthew A. Barbour"
date: "January 19, 2015"
output: word_document
---

### Abstract
The structure of species interaction networks is an emergent property of interactions between individual organisms. While we know this network structure can have a profound influence on community dynamics, we still have a limited understanding of how different genotypes and phenotypes of individual organisms contribute to overall network structure. We address this knowledge gap by using a common garden experiment to examine how genetic and phenotypic variation within a dominant plant species influences an insect herbivore-parasitoid interaction network. We found that plant genetic variation was an important driver of dissimilarity in the herbivore-parasitoid network. Specifically, we found that herbivore suscpetibility to attack from particular parasitoid species and guilds depended on willow genotype. This variation in herbivore susceptibility was determined by both the density and size of herbivores, which in turn was affected by leaf quality and plant architecture traits. Taken together, our results indicate that genetic variation can play a key role in structuring ecological networks. Furthermore, our results highlight the potential for microevolutionary processes to shape both the structure and dynamics of ecological networks.

### Introduction  
Ecological networks provide a map of who interacts with whom in diverse communities. By accurately quantifying their structure, we can begin to understand the processes that organize these networks (CITE) as well as how network structure affects the persistence of ecological communities (CITE). Typically though, studies quantifying the structure of ecological networks have been resolved to the species level (Ings et al., 2009). These species-level networks implicitly assume that each species consists of a homogenous population of individuals, all of which interact with individuals of different species in the same way. Yet, virtually all populations are heterogenous mixtures of individuals and this phenotypic variation can have large effects on the structure and dynamics of ecological networks (CITE INDIVIDUAL STUDIES AND TREE REVIEWS). Consequently, if we continue to ignore how individual differences scale up to affect species interactions, we risk overlooking fundamental drivers of how ecological networks are organized (CITE INDIVIDUAL STUDIES HERE).

Genetic variation is a key driver of individual differences and there are a growing number of examples demonstrating its cascading effects on multitrophic interactions (Antonovics 1992; Whitham et al. 2003, 2012). For example, genetic variation within plant species can influence the density and size of herbivores, which in turn alters interactions with higher trophic levels (Bailey et al., 2006; Johnson 2008; Abdala-Roberts and Mooney 2014). Similarly, genetic and phenotypic variation within predators has been found to mediate the strength of trophic cascades (Post et al. 2008; Harmon et al., 2009; Bassar et al., 2010). While it is clear that microevolutionary processes within a single species can influence pairwise species interactions and community composition at different trophic levels, these studies have neglected to quantify changes in the organization of the ecological network.

Understanding how genetic and phenotypic variation influences the organization of ecological networks is important, as this enhances our ability to predict the eco-evolutionary dynamics of multitrophic communities (CITE JORDI and Carlos Melian). For example, if trait variation in a plant species influences their susceptibility to different herbivore species, and this trait variation is heritable, then we have a mechanism by which genetic variation drives dissimilarity in the herbivore community (Agrawal 2005; Barbour et al. 2015). Morever, if the genetic and phenotypic variation in the plant is large enough, then the resulting dissimilarity in the herbivore community may influence the functional and numerical response of upper trophic levels, providing a mechanism for how genetic variation could drive dissimilarity in an ecological network. From here, we can begin to predict how trait evolution will affect the eco-evolutionary dynamics of networks (Guimeras et al., 20011; Nuismer et al., 2012). Despite this promising avenue for future research, we are currently lacking empirical tests of whether genetic variation drives dissimilarity in the structure of ecological networks (but see Bukovinsky et al. 2008).

Here, we test the hypothesis that genetic and phenotypic variation drives dissimilarity in ecological networks using a subset of the insect food web associated with the coastal willow *Salix hookeriana*. Specifically, this insect food web consisted of four species of gall midges (Family: Cecidomyiidae) and the six species of natural enemies that attack them (Fig. 1A). This plant-insect food web was ideal for testing our hypothesis for two reasons. Firstly, *S. hookeriana* displays considerable genetic and phenotypic variation that results in variation in the density of galling insects (Barbour et al. 2015 in press). *A priori*, we expect that variation in gall density would affect both the functional and numerical response of their natural enemies, thereby providing a mechanistic link for how plant genetic variation could affect this food web. Second, the closed nature of their galls restricts their parasitoid community to a small number of species that can successfully locate and parasitize galls. Consequently, it is easy to identify and quantify the sources of mortality, making this food web amenable to building quantitative interaction networks (van Veen et al. 2006).

If genetic and phenotypic variation within species drives dissimilarity in ecological networks, we would predict the following three things. Firstly, we predict that the structure of insect food webs will vary among willow genotypes. Secondly, we predict that variation in the insect food web will be explained by the functional and numerical responses of the natural enemies to genetic variation in gall density and gall size. Finally, we predict that variation in gall density and gall size will result from genetic variation in leaf quality and plant architectural traits. Taken together, our study seeks to understand how genetic variation shapes the structure of ecological networks. And since genetic variation provides the raw material for evolution, establishing a genetic basis to the organization of ecological networks will begin to inform the eco-evolutionary dynamics of multitrophic communities.

### Methods

## Common Garden
To isolate the effects of genetic variation within Salix hookeriana (hereafter ‘willow’) on the insect food web, we used a common garden experiment consisting of 26 different genotypes of S. hookeriana (13 males; 13 females), located at Humboldt Bay National Wildlife Refuge (HBNWR) (40°40'53"N, 124°12'4"W) near Loleta, California, USA. Willow genotypes were collected from a single population of willows growing around Humboldt Bay. This common garden was planted in February 2009 with 25 clonal replicates (i.e., stem cuttings) of each willow genotype in a completely randomized design in two hectares of a former cattle pasture at HBNWR. Willows in our garden begin flowering in February and reach their peak growth in early August. During this study, willows had reached 2-4 m in height. Further details on the genotyping and planting of the common garden are available in Barbour et al. (2015, in press).

## Dissimilarity of insect food web
To build quantitative insect food webs, we collected galls from about 5 randomly chosen replicates of each genotype in September 2012 (N = 146 trees, range = 4-9 trees per genotype). For each replicate willow, we collected all galls occurring on one randomly selected basal branch. For each branch, we estimated the number of shoots based on an allometric equation using the stem diameter of the sampled basal branch (mean ± SD shoot count = 280 ± 124; details in supplementary materials). We then used these shoot estimates to quantify the density of gall-parasitoid interactions on each replicate willow. To quantify gall-parasitoid interactions, we placed collected galls into 30 mL plastic transport vials (loosely capped at the end), which we maintained at room temperature in the lab for four months. We then opened galls under a dissecting scope and determined whether the gall survived or was parasitized, and if parasitized, the identity of the parasitoid species. We omitted from analyses those galls for which we could not reliably determine the cause of mortality.

**FRAMEWORK FOR MEASURING NETWORK BETA-DIVERSITY** We measured the dissimilarity of insect food webs among willows by analyzing both qualitative and quantitaive differences in the composition of their links (i.e., species interactions) (Poisot et al. 2012). These differences in link composition can then be further partitioned into species turnover and interaction components. Specifically, differences due to species turnover arise from the gain/loss of species (either gall or parasitoid) altering link composition, whereas differences due to interactions occur when species switch with whom they are interacting, despite having the same species composition. To measure qualitative differences, we transformed the quantitative network into presence/absence data and calculated the pairwise Euclidean distances between sites. We used Euclidean distance because we considered the joint absence of the same interaction between sites as meaningful. We note though that we obtain the same qualitative results when we use other common dissimilarity indices for presence/absence (e.g. Jaccard, Sorenson, Whittaker; details in supplementary info). 
Given that the dissimilarity in network structure between two sites will be positively correlated with differences in species richness between sites, we used a probabilistic null model to tease apart the contribution of differences in species richnes the Raup-Crick null model ('raupcrick' function in *vegan* package in R).*tighten up the language here*

To measure quantitative differences, we used Euclidean distance. As the with presence/absence data, other common dissimilarity indices gave qualitatively the same results (e.g. Bray-Curtis, Horn-Morisita; details in supplementary info). To get a better understanding of the processes driving the quantitative dissimilarity among genotypes, Since we were also interested in understanding which 
FOR QUANTITATIVE DIFFERENCES, WE ALSO USED EUCLIDEAN DISTANCE. AS WITH THE QUALITATIVE DATA, OTHER DISSIMILARITY INDICES GAVE QUALITATIVELY THE SAME RESULTS, WHICH WE REPORT IN THE SUPPLEMENTARY INFO.

**ANALYZING NETWORK BETA-DIVERSITY** To test whether willow genetic variation drove beta-diversity of gall-parasitoid networks, we used permutational ANOVA (PERMANOVA, 'adonis' function with *vegan* package in R) on pairwise Euclidean distances between link composition, with willow genotype modelled as a random effect. We then partitioned the relative contribution of species turnover and interaction components (with betalink package in R) to understand the processes driving network beta-diversity among genotypes. To estimate the proportion of variance in network dissimilarity explained by willow genotype, we calculated broad-sense heritability (*H^2*) using the equation: H2 = VG / VP, where VG is the total genotypic variance among clones, and VP is the total phenotypic variance, calculated as the sum of the residual (Vr) and genetic variance (Lynch & Walsh 1998). Following Lynch and Walsh (1998), we estimated Vg and Vp from out PERMANOVA table. Broad-sense heritability values range between 0-1, where values close to zero indicate low heritability (i.e., gall resistance is strongly influenced by the environment), and values close to 1 indicate high heritability (i.e., gall resistance is strongly controlled by underlying genetic variation).

To get a better understanding of the processes driving the quantiative dissimilarity in networks, we analyzed whether the dominant gall-parasitoid interactions varied among willow genotypes. To do this, we analyzed random effect models, where willow genotype was specified as a random effect and the density of a particular gall-parasitoid interaction was the response variable. Since we only had one response variable in each model, we used restricted maximum likelihood (REML) to estimate the variance due to willow genotype (VG) and unexplained residuals (Vr), which we used for calculating the broad-sense heritability of each interaction. We then performed restricted likelihood-ratio tests to examine whether willow genotype explained a significant proportion of the variance in each gall-parasitoid interaction ('RLRTsim' function in *RLRTsim* package in R).

To get at a more interpretable understanding of how species turnover could affect galling insects, we also divided up the natural enemy community into three different guilds. Egg parasitoids, larval parasitoids, and predators. These guilds may impose different functional constraints on galling insects. In particular, gall size. For example, for Cecidomyiid midges, larva initiate gall development, therefore we would not expect any characteristics of the gall to impose a functional constraint (e.g. gall size) on the likelihood of parasitoid attack, and is therefore causing selection on this stage of their life history. In contrast, larval parasitoids attack after the gall is formed and therefore, we would expect them to be influenced by the size of the gall and potentiall constrained from being able to oviposit in large galls. In our system, one of the species is a predatory midge which presumably attacks the larval stage as well, but giving their distant evolutionary relationship to parasitoid wasps, may be constrained by other factors, which is why we retained it as a separate guild. Therefore, we also analyzed differences in link composition when parasitoid species were collapsed into different guilds.

## Genetic variation in gall density and size affects functional and numerical responses of natural enemies
To test whether gall density (count per 100 shoots) and gall size (measured as the maximum diameter perpendicular to direction of plant tissue growth, to the nearest 0.01 mm) varied among willow genotypes, we used our gall collections from each replicate willow tree (N = 146). As with pairwise gall-parasitoid interactions, we analyzed separate random effect models for the density of each gall species. Since we had individual-level measurements for gall size, we analyzed a nested random effect model, where willow replicates were nested within willow genotype. We then used the variance due to willow replicate as another source of phenotypic variance (i.e. Vp = Vg + Vwillow rep + Vres) for our estimates of broad-sense heritability. For both gall density and gall size, we restricted our analyses to gall species that were associated with genetic variation in gall-parasitoid interactions. Gall density and gall size were transformed as needed to improve normality and reduce heteroscedasticity of model residuals.*check on whether reducing heteroscadesiticty is meaningful here*

To test whether gall density and gall size influenced the numerical response of natural enemies, we used linear regression. Specifically, we modelled gall-parasitoid density as our response variable with gall density, mean gall size (tree-level), and their interaction, as our predictor variables. To test whether gall density and gall size influence the functional response of natural enemies, we used generalized linear mixed-effect models (GLMMs, 'glmer' function in *lme4* package in R). GLMMS were appropriate because they enabled us to account for the hierarchical structure of our data and model the probability of observing a gall-parasitoid interaction as our response variable (family = binomial). Specifically, we modelled individual-level gall size, tree-level gall density, and their interaction, as fixed effects with each willow replicate as a random effect. We then used parametric bootstrapping to test the effect of gall density, gall size, and their interaction ('pbkrtest' function in *pbkrtest* package in R). We always started with the most complex model and removed non-significant predictors (P > 0.05) until we identified the most parsimonious model. *for numerical response, decide whether it's useful to model this at the genotype-level*

## Genetic variation in plant traits influences gall density and size
To identify the plant traits influencing variation in gall density and gall size, we first measured 40 different traits associated with variation in leaf quality (36 traits) and plant architecture (4 traits). Details on how these willow traits were sampled and quantified are given in Barbour et al. (2015, in press), but we summarize which traits were sampled here. Leaf quality traits included: phenolic chemistry (7 classes of compounds, 31 individual metabolites), trichome density, specific leaf area (SLA), water content, and percent Carbon and Nitrogen (converted to C:N). Plant architecture traits included: plant size, fractal dimension (index of architectural complexity), height, and foliage density. Each of these 40 traits exhibited significant broad-sense heritable variation among willow genotypes (mean leaf quality H2 = 0.72; mean architecture H2 = 0.27; range of H2 = 0.15 - 0.97; Barbour et al., 2015 in press). We then used multiple linear regression with forward model selection to identify the key plant traits mediating variation in gall density ang gall size. First though, we mitigated multicollinearity using three different methods: principal components analysis (phenolic chemistry only), residual and sequential regression (Graham 2003), and omitting highly correlated traits (details in Barbour et al. 2015, in press). This resulted in 12 predictor variables (DETAILS). We then used the forward model selection approach advocated by Blanchet et al. (2008), which prevents inclusion of spurious variables (i.e., inflated Type 1 error) and overestimation of explained variance (i.e., R2).

### Results

## Genetic variation drives dissimilarity in insect food web
In concordance with our prediction, we found that willow genotype was a major driver of both qualitative and quantitative dissimilarity in the insect food web (Fig. 1b). The majority of the dissimilarity in the insect food web was due to species turnover. In other words, gall species varied in their susceptibility to different parasitoid species depending on willow genotype (Fig. 1b). This dissimilarity in the insect food web was due in part to differences in link richness among willow genotypes. Importantly though, we observed dissimilarity in the insect food web among genotypes after accounting for differences in link richness, suggesting that willow genotype imposes a systematic ecological filter on different gall-parasitoid interactions.

Interestingly, we found willow genotype explained a higher percentage of the variance when collapsed parasitoids into different functional groups, suggesting that this was a better characterization of these interactions. Quantitatively, we found that variation larval parasitoids interaction with Iteomyia and Rabdophaga (bud) and egg parasitoids with Iteomyia were the primary drivers of variation in the quantitative data (Fig. 2a). Specifically, the 
*need amount of variation* Specifically, the frequency of parasitism from the egg, endoparasitoid Platygaster on Iteomyia varied 34.9-fold among willow genotypes (H2 = 0.31, RLRT = 21.61, P < 0.001), while parasitism from the larval, ectoparasitoids Mesopolobus (H2 = 0.11, RLRT = 3.77, P = 0.024) and Torymus (H2 = 0.25, RLRT = 14.75, P < 0.001) varied 10.5- and 5.7-fold among willow genotypes, respectively.

## Genetic variation in gall density and size affects functional and numerical responses of natural enemies
We found that willows displayed heritable variation in resistance to galling insects in terms of both the density and size of galls. Specifically, the density for four of the five gall species varied between 22.8- and 70.2-fold among willow genotypes (Fig. 2A; range of H2 = 0.12 - 0.36), but was the most pronounced for the most common gall former, Iteomyia. Iteomyia was also the only gall species that varied in size (2.3-fold) among willow genotypes (Fig. 2B; H2 = 0.13, RLRT = 3.68, P = 0.022).

Heritable variation in resistance to galling insects was responsible for the observed dissimilarity in the insect food web. Specifically, the interaction between Iteomyia density and size as well as Rabdophaga (bud) density explained 35% of the variance in the insect food web (F(4,76) = 10.30, P = 0.005). In particular, this effect was driven by the numerical response of egg parasitoids to Iteomyia density on trees with small galls (< 9 mm). Larval parasitoids also had a positive numerical response to Iteomyia density, but were negatively affected by increasing gall size. As expected, larval parasitoids attacking Rabdophaga (bud) galls also had a positive numerical response to Rabdophaga (bud) gall density.

When we examined the functional responses of the different parasitoid guilds, we revealed a potentially important behavior that may be mediating the dynamics of these interactions.

The probability of an egg parasitoid attacking Iteomyia was highest on small galls (< 8 mm) (Fig. 3a). For these small galls, the probability of an egg parasitoid attacking Iteomyia increased 3-fold with increasing gall density (Fig. 3a). Similar to the egg parasitoids, larval parasitoids also had higher attack rates on small galls; however, the probability of a larval parasitoid attacking Iteomyia actually decreased with increasing gall density (Fig. 3a). In contrast to parasitism on Iteomyia, the probability of larval parasitoids attacking Rabdophaga (bud) was independent of both gall density and gall size.

In terms of natural enemy functional responses, we found that both gall density and gall size were important predictors of the probability of parasitoid attack on Iteomyia, but neither of these were important in explaining the probability of attack for the Rabdophaga bud galler. Specifically, the probability of Platygaster parasitizing Iteomyia decreased 18-fold over the range of gall sizes (Fig. 3A). For small-to-intermediate size galls though, the probability of Platygaster parasitizing Iteomyia increased 4.5-fold over the range of gall densities (Fig. 3C; 𝛘2 = 20.61, P < 0.001). In contrast to Platygaster, the probability of Mesopolobus parasitizing Iteomyia was highest for intermediate sized galls (Fig. 3B) and decreased 2.8-fold over the range of gall densities (Fig. 3D; 𝛘2 = 18.10, P = 0.003). The probability of Torymus parasitizing Iteomyia was not influenced by gall size, but similar to Mesopolobus, parasitism decreased 19-fold over the range of gall densities (Fig. 3E; 𝛘2 = 10.82, P = 0.001).

##Genetic variation in plant traits influences gall density and size
We found that variation in both the density and size of galls was explained by both leaf quality and plant architecture traits. For example, the density of both Iteomyia and R. salicisbrassicoides galls was higher on shorter willows with higher leaf C:N (Iteomyia R2 = 0.17, F2,119 = 12.14, P < 0.001; Rabdophaga (bud): R2 = 0.15, F2,120 = 10.97, P < 0.001). Interestingly, the size of Iteomyia galls was not influenced by either willow height or leaf C:N. Instead, gall size was larger on willows with higher concentrations of salicylates and flavones in their leaves (R2 = 0.14, F2,75 = 5.88, P = 0.004). *put it all together? see if these traits explain any of the variation in gall-parasitoid interactions?*

### Discussion

### Acknowledgements

### References

### Figure Legends

### Supplementary Information

```{r}

```

![Figure 1](figures_network_ms_portrait.001.png)

![Figure 2]()

![Figure 3](figures_network_ms_portrait.002.png)

### 2016-01-25 adonis.table.function.R
moved adonis.table.function to miscellaneous_R repository

@@ -1,14 +1,0 @@
-adonis.table <- function(adonis.object, data.frame){
-  x <- adonis.object
-  mean.sample.size <- mean(table(data.frame$Genotype))
-  genetic.variance <- (x$aov.tab$MeanSqs[1] - x$aov.tab$MeanSqs[2])/mean.sample.size
-  heritability <- genetic.variance/(genetic.variance + x$aov.tab$MeanSqs[2])
-  
-  table.out <- data.frame(df = x$aov.tab$Df[1:2],
-                          SS = x$aov.tab$SumsOfSqs[1:2],
-                          MS = x$aov.tab$MeanSqs[1:2],
-                          Heritability = c(heritability,""),
-                          P = c(x$aov.tab$'Pr(>F)'[1],""))
-  table.out
-}
-

### 2016-01-25 Rscripts/best_QuaBiMo_function.R
moved to misc R repository

@@ -1,18 +1,0 @@
-best_QuaBiMo <- function(web, QuaBiMo_reps = 100) {
-  require('bipartite')
-  
-  observed_web_repeats <- list()
-  for(i in 1:QuaBiMo_reps){ 
-    observed_web_repeats[[i]] <- web
-  }
-  
-  QuaBiMo_repeats <- sapply(observed_web_repeats, computeModules)
-  best_observed_value <- max(sapply(QuaBiMo_repeats, function(x) x@likelihood)) # takes the maximum out of a user specified number of replications, since this is an optimization algorithm and we want to avoid reaching a local optimum.
-  
-  best_reps <- which.max(sapply(QuaBiMo_repeats, function(x) x@likelihood)) # identify best partition of modularity algorithm
-  
-  output <- list(best_observed_value = best_observed_value,
-                 best_reps = best_reps,
-                 best_module_info = QuaBiMo_repeats[best_reps])
-}
-

### 2016-01-25 Rscripts/matrix_highlights.R

@@ -1,61 +1,0 @@
-# matrix highlighter function for plotweb2_highlights
-
-matrix_highlighter <- function(web, high.names = "", low.names = "") {
-  
-  web <- as.matrix(web)
-  
-  grab.colnames <- which(colnames(web) %in% high.names)
-  grab.rownames <- which(rownames(web) %in% low.names)
-  
-  web.sub <- web
-  web.sub[ ,-grab.colnames] <- 0
-  web.sub[-grab.rownames, ] <- 0
-  
-  out <- web.sub
-}
-
-
-matrix_highlighter_2 <- function(web, interaction.names = "", replace.values = NULL) {
-  require(dplyr)
-  
-  if(!is.null(replace.values)){
-    replace.df <- data.frame(interaction.names = interaction.names, replace.values = replace.values)
-  }
-  
-  web <- data.frame(web)
-  web$low.nodes <- factor(rownames(web))
-  
-  web.melt <- melt(web)
-  web.melt <- mutate(web.melt, interactions = paste(low.nodes, variable, sep = "_"))
-  
-  web.melt.sub <- web.melt
-  web.melt.sub[-which(web.melt.sub$interactions %in% interaction.names), 3] <- 0
-  
-  if(!is.null(replace.values)){
-    interaction.order <- web.melt.sub$interactions[which(web.melt.sub$interactions %in% interaction.names)]
-    replace.df <- mutate(replace.df, interaction.order = factor(interaction.order, ordered = TRUE))
-    test <- match(replace.df$interaction.names, replace.df$interaction.order)
-    #replace.df.sort <- sort_df(replace.df, "interaction.order")
-    test2 <- web.melt.sub[which(web.melt.sub$interactions %in% interaction.names), ]
-    test3 <- replace.df$replace.values[test]
-    
-    web.melt.sub <- web.melt.sub %>% 
-      left_join(replace.df %>% select(interactions = interaction.names, replace.values),
-                missing = 0) %>%
-      mutate(value = ifelse(is.na(replace.values),yes = 0,no = replace.values))
-  }
-    #web.melt.sub[which(web.melt.sub$interactions %in% interaction.names), 3] <- replace.df$replace.values[test]
-  #}
-
-  web.melt.sub.matrix <- dcast(web.melt.sub, low.nodes ~ variable, value.var = "value")
[…]

### 2016-01-25 Rscripts/plotweb2_highlights.R

@@ -1,624 +1,0 @@
-
-plotweb2_highlights <- function (web, 
-                                 web2, 
-                                 highlight.web = NULL,
-                                 highlight.web.high = NULL,
-                                 method = "cca", 
-                                 empty = FALSE, # when FALSE, method changes to "normal"
-                                 labsize = 1, 
-                                 ybig = 1, 
-                                 y_width = 0.1,
-                                 y_width_scale = 0,
-                                 spacing = 0.05, 
-                                 arrow = "no", 
-                                 col.interaction = "grey80", 
-                                 bord.col.interaction = "black",
-                                 col.pred = "grey10", 
-                                 col.prey = "grey10",
-                                 lab.space = 1, 
-                                 lablength = NULL, # buggy if not set to null
-                                 sequence = NULL, 
-                                 low.abun = NULL, 
-                                 low.abun.col = "green", 
-                                 high.abun = NULL, 
-                                 high.abun.col = "red", 
-                                 method2 = "cca", 
-                                 empty2 = TRUE, 
-                                 spacing2 = 0.05, 
-                                 arrow2 = "no", 
-                                 col.interaction2 = "grey80", 
-                                 bord.col.interaction2 = "black",
-                                 col.pred2 = "grey10", 
-                                 col.prey2 = NA, 
-                                 lablength2 = NULL, # buggy if not set to null
-                                 sequence.pred2 = NULL, 
-                                 low.abun2 = NULL, 
-                                 low.abun.col2 = "white", 
-                                 bord.low.abun.col2 = "black",
[…]

### 2016-01-25 Rscripts/analysis_empirical_data.R
transferring code from ms_willow_network_supplement
This code will contain a clean and explained version of the analyses
used in the manuscript

@@ -1,0 +1,559 @@
+##
+## Author: Matt Barbour
+
+## source in required datasets and functions ----
+source('~/Documents/Genotype_Networks/Rscripts/network_management_tree_level.R')
+source('~/Documents/miscellaneous_R/model_diagnostic_functions.R')
+#source('~/Documents/miscellaneous_R/jensen_magnitude_function.R')
+
+## load required libraries ----
+library(vegan) # for adonis analysis
+#library(knitr) # for making tables
+#library(ggplot2) # for plotting
+#theme_set(theme_bw()) # customize ggplot for prettier default graphs
+#library(gridExtra) # for grid.arrange
+library(mvabund) # multivariate analysis
+library(HH) # for variance inflation factor analysis (vif) in multiple regression
+library(visreg) # visualize regression output
+
+## create datasets for analysis ----
+full.df <- tree_level_interaxn_all_plants_traits_size 
+interaxns_noPont # all gall-parasitoid interactions used for analysis
+
+#Testing for whether gall size varies among willow genotypes. We were unable to use the mvabund framework for this analysis, so we conducted separate linear models to test for these effects.
+
+## gall size plots
+vLG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = vLG.height.mean)) + 
+  geom_boxplot() + ylab("Iteomyia salicisverruca diameter (mm)")
+rG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = rG.height.mean)) + 
+  geom_boxplot() + ylab("Rabdophaga salicisbrassicoides diameter (mm)")
+aSG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = aSG.height.mean)) + 
+  geom_boxplot() + ylab("Cecidomyid sp. A diameter (mm)")
+SG.height.mean.plot <- ggplot(data = full.df, aes(x = Genotype, y = SG.height.mean)) + 
+  geom_boxplot() + ylab("Rabdophaga salicisbattatus diameter (mm)")
+
+grid.arrange(vLG.height.mean.plot, rG.height.mean.plot, aSG.height.mean.plot, SG.height.mean.plot)
+
+
[…]

### 2016-01-25 Rscripts/fxns_mvabund_diagnostics.R
added separate file for mvabund diagnostics

@@ -1,0 +1,15 @@
+# Curve plots that are meant to be plotted on top of meanvar.plots in the 'mvabund' package or any other mean-variance plots. 
+poisson_curve <- function(from, to, add = TRUE, color = "blue", line.width = 3, line.type = 1){
+  curve(expr = 1*x, # poisson assumes that mean = variance
+        from = from, to = to, add = add, col = color, lwd = line.width, lty = line.type)
+}
+
+quasipoisson_curve <- function(from, to, quasi.scalar, add = TRUE, color = "red", line.width = 3, line.type = 1){
+  curve(expr = quasi.scalar*x, # quaispoisson assumes that variance scales linearly with mean according to a constant.
+        from = from, to = to, add = add, col = color, lwd = line.width, lty = line.type)
+}
+
+neg.binomial_curve <- function(from, to, theta.negbin, add = TRUE, color = "black", line.width = 3, line.type = 1){
+  curve(expr = x + theta.negbin*x^2, # negative binomial assumes this mean-variance relationship (quadratic)
+        from = from, to = to, add = add, col = color, lwd = line.width, lty = line.type)
+}
